### 2013-10-18 gammaRays.tex
Initial commit

\documentclass{article}

\usepackage{commath}
\usepackage{fullpage}

\allowdisplaybreaks

\DeclareMathOperator*{\argmin}{argmin}

\author{Maxim Piskunov}
\title{Spectral Lags in Very High Energy Emission of Gamma Ray Bursts}

\begin{document}
\maketitle

\section{Introduction}

\section{Observations}

\section{Model}

\subsection{Assumptions}

To understand the observed spectral lag, lets explore geometry of the jet. First of all, lets make some assumptions:
\begin{enumerate}
\item{Time $t = 0$, a spherical shell of plasma is emitted. The center is called the central engine.}
\item{The shell points propagate with a constant velocity $v \sim 1$, so at the time $t$ the radius of the shall is $v t$.}
\item{Each point of the shell is an isotropic radiator in its reference frame.}
\item{
	The radiation intensity is a function of the radiator position and the radiation frequency:
	\begin{equation}
		\eta\left(r,\theta,\omega\right) = 
		\frac{\eta_0}{1 + \left(\frac{r}{r_0}\right)^n}
		\exp\left(
			-\left(\frac{\theta}{\theta_0}\right)^2
			\left(\frac{\omega}{\omega_0}\right)^{-2k}
		\right)
		\left(\frac{\omega}{\omega_0}\right)^\alpha
	\end{equation}
	$\eta$ is a number of particles emitted per volume per frequency. It is a function of the distance $r$ from the central engine, of the off-axis angle $\theta$, and of the radiation frequency $\omega$.
}
\end{enumerate}

The burst is fully specified by the following set of parameters:
\begin{itemize}
\item{$v$, the velocity of the shell, $\gamma = \frac{1}{\sqrt{1-v^2}} \gg 1$;}
\item{$\eta_0$, which defines the intensity scale;}
\item{$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
\item{$n$, which determines the sharpness of the jet end, $n > 3$;}
\item{$\omega_0$, a characteristic radiation frequency;}
\item{$\theta_0$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;}
\item{$k$, which determines how much the opening angle changes with frequency, $k < 0$;}
\item{$\alpha$, the bare spectral index.}
\end{itemize}

Now we have all we need to calculate the observed light curves, and then the stretching factors.

\subsection{Photon observation time}

Lets begin by calculating a time at which some particular photon is observed. This time is a function of the radiator location $\left(r, \theta, \phi\right)$, as well as the observer location $\left(d, \chi, 0\right)$. (Here we choose coordinates so that the rotation angle of the observer is $0$.) Lets assume for now that the observer is far enough to resolve the jet geometry, yet close enough so that the expansion of space is negligible. This assumption allows us to ignore the effects of cosmology for this calculation.

The observation time is a sum of two terms: the time interval from $t = 0$ to the photon emission (the plasma time), and the time interval between the emission and the observation (the photon time):
\begin{equation*}
t\left(r, \theta, \phi, d, \chi\right) = t_\text{plasma}\left(r\right) + t_\text{photon}\left(r, \theta, \phi, d, \chi\right)
\end{equation*}
$t_\text{plasma}$ is easy to compute since plasma moves with uniform velocity:
\begin{equation*}
t_\text{plasma}\left(r\right) = \frac{r}{v}
\end{equation*}
$t_\text{photon}$ is a distance between the radiator and the observer:
\begin{align*}
t_\text{photon}\left(r, \theta, \phi, d, \chi\right) &= \sqrt{
	\left( d \cos\chi - r \cos\theta \right)^2 +
	\left( d \sin\chi - r \sin\theta \cos\phi \right)^2 +
	\left( r \sin\theta \sin\phi \right)^2
} \\
&= d\sqrt{
	\left( \cos\chi - \frac{r}{d} \cos\theta \right)^2 +
	\left( \sin\chi - \frac{r}{d} \sin\theta \cos\phi \right)^2 +
	\left(\frac{r}{d} \sin\theta \sin\phi \right)^2
} \\
&\sim d\sqrt{
	\cos^2{\chi} - 2\frac{r}{d} \cos\theta \cos\chi + \sin^2\chi - 2\frac{r}{d} \sin\theta \cos\phi \sin\chi
} \\
&\sim d\left(
	1 - 
	\frac{r}{d} \left( \cos\theta \cos\chi + \sin\theta \cos\phi \sin\chi \right)
\right) \\
&= d - r\left( \cos\theta \cos\chi + \sin\theta \cos\phi \sin\chi \right)
\end{align*}
Combining these expressions together, we get:
\begin{equation*}
t\left(r, \theta, \phi, d, \chi\right) = d + r\left( \frac{1}{v} - \cos\theta \cos\chi - \sin\theta \cos\phi \sin\chi \right)
\end{equation*}
Since no photon can reach the observer faster than the speed of light, and the photons emitted at $t = 0$ reach the observer at $t = d$, this time $t = d$ is the start of the observer's light curve. So we can define a more convenient time origin:
\begin{equation*}
\tau \left(r, \theta, \phi, \chi\right) = t\left(r, \theta, \phi, d, \chi\right) - d = r\left( \frac{1}{v} - \cos\theta \cos\chi - \sin\theta \cos\phi \sin\chi \right)
\end{equation*}
Finally, we can use the spherical law of cosines to write this in terms of the great-circle distance $\sigma\left( \theta, \phi, \chi \right)$ between the points $\left( \theta, \phi \right)$ and $\left( \chi, 0 \right)$:
\begin{equation}
\tau \left(r, \theta, \phi, \chi\right) = r\left( \frac{1}{v} - \cos\sigma\left( \theta, \phi, \chi \right) \right)
\end{equation}

Note that $\tau$ doesn't depend on $d$ anymore. But this is only true if the scale factor did not change since emission to observation. For a distant observer we should account for the change in the scale factor, which will stretch the distances between photons:
\begin{equation}
\tau \left(r, \theta, \phi, z, \chi \right) = r\left( \frac{1}{v} - \cos\sigma\left( \theta, \phi, \chi \right) \right) \left( 1 + z \right)
\end{equation}
Here $z$ is the redshift of the burst from the point of view of the observer.

\subsection{Cosmology}
Before calculating light curves and stretching factors, we will need some equations from cosmology.

First of all, the metric of the expanding Universe:
\begin{equation}
\dif s^2 = -\dif t^2 + a^2\left(t\right) \dif r^2 + a^2\left(t\right) r^2 \dif \Omega_2
\end{equation}
We define $a\left(t_\text{obs}\right) = 1$, where $t_\text{obs}$ is the observation time.

We need to understand how the scale factor changes with time. For that we assume that the energy content of the Universe consists of matter $\Omega_m$ and vacuum energy $\Omega_\Lambda$ only, so that the Friedmann equation takes the following form:
\begin{align*}
\left( \frac{\dot{a}\left(t\right)}{a\left(t\right)} \right)^2 &= \Omega_m H_\text{obs}^2 \frac{1}{a^3\left(t\right)} + H_\text{obs}^2 \Omega_\Lambda \\
\dot{a}\left(t\right) &= a\left(t\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda} \\
\dif t &= \frac{\dif a}{a\left(t\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}}
\end{align*}
Here $H_\text{obs} = H\left(t_\text{obs}\right) = \dot{a}\left( t_\text{obs} \right)$ is the Hubble parameter at the observation time.

We also need to know the areas of two spheres: the photon sphere, the sphere over which the photons emitted in a particular burst are spread; and the bursts sphere, the sphere over which the bursts at a particular redshift are distributed.

Lets begin with the photon sphere. This sphere has the origin at $r = 0$ and $t = 0$, at the central engine of a particular burst. Taking $\dif s = 0$ and $\dif \Omega_2 = 0$ in the equation for metric, we get:
\begin{equation*}
\dif r = \frac{\dif t}{a\left(t\right)}
\end{equation*}
And now we integrate over time to find the observer's position:
\begin{align*}
r\left(t_\text{obs}\right) &= \int_0^{t_\text{obs}}\frac{\dif t}{a\left(t\right)} \\
&= \int_{a\left(0\right)}^1\frac{\dif a}{a^2\left(t\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}} \\
&= \frac{\, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{a^3\left(0\right) \Omega_\Lambda }\right)-a\left(0\right) \, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{\Omega_\Lambda }\right)}{a\left(0\right) H_\text{obs} \sqrt{\Omega_\Lambda }}
\end{align*}
Here $_2F_1\left(a, b; c; z\right) = \frac{\Gamma\left(c\right)}{\Gamma\left(b\right)\Gamma\left(c-b\right)}\int_0^1 \frac{t^{b-1}\left(1-t\right)^{c-b-1}}{\left(1-t z\right)^a} \dif t$ is a hypergeometric function. Here $a\left(0\right) = \frac{1}{1+z}$, so, finally,
\begin{equation}
r\left(z\right) = \frac{\left(1+z\right)\, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{\Omega_\Lambda}\left(1+z\right)^3\right) - \, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{\Omega_\Lambda }\right)}{H_\text{obs} \sqrt{\Omega_\Lambda }}
\end{equation}
The area of the photon sphere is then:
\begin{equation}
A_\text{ph}\left(z\right) = 4 \pi a^2\left(t_\text{obs}\right) r^2\left(z\right) = 4 \pi r^2\left(z\right)
\end{equation}

The second sphere is the bursts sphere. It has its origin at the observer's position. We define a new primed set of coordinates, so that $r' = 0$ and $t' = 0$ at the observation event. We can relate the primed coordinates of the burst to the observer's unprimed coordinates:
\begin{align*}
t'_\text{burst} &= -t_\text{obs} \\
r'\left(t'_\text{burst}\right) &= r'\left(-t_\text{obs}\right) = \int_{-t_\text{obs}}^0\frac{\dif t'}{a'\left(t'\right)} = \int_0^{t_\text{obs}}\frac{\dif t}{a\left(t\right)} = r\left(t_\text{obs}\right)
\end{align*}
We see that the radius of the bursts sphere is the same as the radius of the photon sphere. Finally, the area is:
\begin{equation}
A_\text{b}\left(z\right) = 4 \pi a'^2\left( -t_\text{obs} \right) r^2\left(z\right) = 4 \pi a^2\left(0\right) r^2\left(z\right) = \frac{4 \pi r^2\left(z\right)}{\left(1+z\right)^2}
\end{equation}

We will need to know one more thing about the bursts sphere -- the volume of the infinitesimal shell surrounding it. For that we can again use the metric:
\begin{align}
\dif V\left(z\right) &= - A_\text{b}\left(z\right) a\left(0\right) \dif r \nonumber\\
&= - A_\text{b}\left(z\right) a\left(0\right) \frac{\dif a}{a^2\left(0\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}} \nonumber\\
&= A_\text{b}\left(z\right) \frac{\frac{1}{\left(1+z\right)^2}\dif z \left(1+z\right)}{H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}} \nonumber\\
&= A_\text{b}\left(z\right) \frac{1}{\left(1+z\right)} \frac{\dif z}{H_\text{obs} \sqrt{\Omega_m\left(1+z\right)^3 + \Omega_\Lambda}} \nonumber\\
&= \frac{4\pi r^2\left(z\right)}{\left(1+z\right)^3} \frac{\dif z}{H_\text{obs} \sqrt{\Omega_m\left(1+z\right)^3 + \Omega_\Lambda}}
\end{align}

\subsection{Light curve}
We now know enough to compute the observable quantity - the number of observed photons $p$. For that we need to integrate the radiation intensity $\eta$ over four things. Two burst-related things: radiator positions and frequencies at emission. Two observer-related things: observation time and frequencies at observation. We relate the first two and the second two with the delta functions:
\begin{align}
p\left( z, \chi; \tau_1, \tau_2; \omega_1, \omega_2 \right) &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)} \int_0^\infty \dif r \int_0^{\pi} r \dif \theta \int_0^{2\pi} r \sin{\theta} \dif \phi \int_0^\infty \dif \omega' \int_{\tau_1}^{\tau_2} \dif \tau \int_{\omega_1}^{\omega_2} \dif \omega \, \eta\left( r, \theta, \omega' \right) \nonumber\\
&\qquad {} \times \underbrace{\frac{1}{\gamma^2\left( 1 - v \cos\sigma\left(\theta, \phi, \chi\right) \right)^2}}_{\text{aberration}} \underbrace{\frac{1}{\gamma\left( 1 - v \cos\sigma\left(\theta, \phi, \chi\right) \right)}}_{\text{time dilation}} \nonumber\\
&\qquad {} \times \delta\left( \frac{\omega'}{\underbrace{\gamma \left( 1 - v\cos\sigma\left(\theta, \phi, \chi\right) \right)}_{\text{relativistic shift}} \underbrace{\left(1+z\right)}_{\text{cosmological shift}}} - \omega \right) \nonumber\\
&\qquad {} \times \delta\left( \tau - r\left( \frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right) \right) \left(1+z\right) \right)
\end{align}
Here $A_\text{det}$ is the effective area of the detector. We have taken account for the four relativistic effects, which affect intensities and frequencies of the radiators: relativistic aberration; time dilation of the radiators relative to the observer; relativistic blue/redshift; and cosmological redshift.

We can use the delta functions to do the integrals over $\omega'$ and $r$. For that lets transform the first one:
\begin{align*}
\delta\left( \frac{\omega'}{\gamma \left( 1 - v\cos\sigma \right)\left(1+z\right)} - \omega \right) &= \delta\left( \frac{1}{\gamma\left( 1-v\cos\sigma \right)\left(1+z\right)} \left( \omega' - \gamma\left( 1 - v\cos\sigma \right) \left(1+z\right) \omega \right) \right) \\
&= \gamma\left( 1 - v\cos\sigma \right) \left(1+z\right) \delta\left( \omega' - \gamma\left( 1 - v\cos\sigma \right) \left(1+z\right) \omega \right)
\end{align*}
And the second one:
\begin{align*}
\delta\left(\tau - r\left( \frac{1}{v} - \cos\sigma \right)\left(1+z\right)\right) &= \delta\left( \left( \frac{1}{v} - \cos\sigma \right) \left( 1+z \right) \left( r - \frac{\tau}{\left(\frac{1}{v} - \cos\sigma\right)\left(1+z\right)} \right) \right) \\
&= \frac{1}{\left(\frac{1}{v} - \cos\sigma\right)\left(1+z\right)} \delta\left( r - \frac{\tau}{\left(\frac{1}{v} - \cos\sigma\right)\left(1+z\right)} \right)
\end{align*}
After the transformations the expression for $p$ takes the following form:
\begin{align}
p\left( z, \chi; \tau_1, \tau_2; \omega_1, \omega_2 \right) &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)} \int_0^{\pi} \dif \theta \int_0^{2\pi} \dif \phi \int_{\tau_1}^{\tau_2} \dif \tau \int_{\omega_1}^{\omega_2} \dif \omega \nonumber\\
&\qquad {} \times \eta\left( \frac{\tau}{\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)\left(1+z\right)}, \theta, \gamma\left( 1 - v\cos\sigma\left(\theta, \phi, \chi\right) \right)\left(1+z\right)\omega  \right) \nonumber\\
&\qquad {} \times \frac{\tau^2 \sin\theta}{\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)^3 \left(1+z\right)^2 \gamma^2 \left(1-v\cos\sigma\left(\theta, \phi, \chi\right)\right)^2} \nonumber\\
&= \frac{A_\text{det}}{A_\text{ph}\left(z\right)} \frac{1}{v^2 \gamma^2 \left(1+z\right)^2} \int_{\tau_1}^{\tau_2} \dif \tau \, \tau^2 \int_0^{\frac{\pi}{2}} \dif \theta \int_0^{2\pi} \dif \phi \int_{\omega_1}^{\omega_2} \dif \omega \frac{\sin\theta}{\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)^5} \nonumber\\
&\qquad {} \times \eta\left( \frac{\tau}{\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)\left(1+z\right)}, \theta, \gamma\left( 1 - v\cos\sigma\left(\theta, \phi, \chi\right) \right)\left(1+z\right)\omega  \right)
\end{align}

Furthermore, we can do the integrals over $\omega$ and $\tau$, if we write $\eta$ explicitly:
\begin{align}
p\left( z, \chi; \tau_1, \tau_2; \omega_1, \omega_2 \right) &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)} \frac{1}{v^2 \gamma^2 \left(1+z\right)^2} \int_0^{\pi} \dif \theta \int_0^{2\pi} \dif \phi \int_{\omega_1}^{\omega_2} \dif \omega \int_{\tau_1}^{\tau_2} \dif \tau \, \tau^2 \frac{\sin\theta}{\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)^5} \nonumber\\
&\qquad {} \times \frac{\eta_0}{1 + \left(\frac{\tau}{r_0\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)\left(1+z\right)}\right)^n} \nonumber\\
&\qquad {} \times \exp\left(
	-\left(\frac{\theta}{\theta_0}\right)^2
	\left(\frac{\omega\gamma\left( 1 - v\cos\sigma\left(\theta, \phi, \chi\right) \right)\left(1+z\right)}{\omega_0}\right)^{-2k}
\right) \nonumber\\
&\qquad {} \times \left(\frac{\omega\gamma\left( 1 - v\cos\sigma\left(\theta, \phi, \chi\right) \right)\left(1+z\right)}{\omega_0}\right)^\alpha \nonumber\\
&= \frac{A_\text{det}}{A_\text{ph}\left(z\right)} \frac{\eta_0}{\left(v\gamma\left(1+z\right)\right)^{2-\alpha}} \int_0^{\pi} \dif \theta \int_0^{2\pi} \dif \phi \frac{\sin\theta}{\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)^{5-\alpha}} \nonumber\\
&\qquad {} \times \int_{\omega_1}^{\omega_2} \dif \omega \exp\left(
	-\left(\frac{\theta}{\theta_0}\right)^2
	\left(\frac{\omega\gamma\left( 1 - v\cos\sigma\left(\theta, \phi, \chi\right) \right)\left(1+z\right)}{\omega_0}\right)^{-2k}
\right) \left(\frac{\omega}{\omega_0}\right)^\alpha \nonumber\\
&\qquad {} \times \int_{\tau_1}^{\tau_2} \dif \tau \, \tau^2 \frac{1}{1 + \left(\frac{\tau}{r_0\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)\left(1+z\right)}\right)^n} \nonumber\\
&= \frac{A_\text{det}}{A_\text{ph}\left(z\right)}
\frac{\eta_0}{\left(v\gamma\left(1+z\right)\right)^{2-\alpha}}
\int_0^{\pi} \dif \theta \int_0^{2\pi} \dif \phi \frac{\sin\theta}{\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)^{5-\alpha}} \nonumber\\
&\qquad {} \times \left( I\left(z,\chi,\omega_2; \theta,\phi\right) - I\left(z,\chi,\omega_1; \theta,\phi\right) \right) \left( J\left(z,\chi,\tau_2; \theta,\phi\right) - J\left( z,\chi,\tau_1; \theta,\phi \right) \right)
\end{align}
Here $I$ and $J$ are the indefinite integrals over $\omega$ and $\tau$:
\begin{align}
I\left(z,\chi,\omega;\theta,\phi\right) &= \frac{
	\omega
	\left(\frac{\omega}{\omega_0}\right)^{\alpha }
	E_{\frac{\alpha +1}{2 k}+1}\left(
		\left(\frac{\theta}{\theta_0}\right)^2
		\left(\frac{\omega}{\omega_0}\right)^{-2k}
		\left(
			\gamma \left(1-v \cos \sigma\left(\theta ,\phi ,\chi \right) \right) \left(1+z\right)
		\right)^{-2 k}
	\right)
}{
	2 k
} \\
J\left(z,\chi,\tau;\theta,\phi\right) &= \frac{\tau^3}{3} \,
_2F_1\left(
	1,
	\frac{3}{n};
	\frac{n+3}{n};
	-\left(\frac{\tau}{
		r_0 \left( \frac{1}{v} - \cos\sigma\left( \theta,\phi,\chi \right) \right) \left(1+z\right)
	}\right)^n
\right)
\end{align}
where $E_n$ function $E_n\left(x\right) = \int_1^\infty \frac{e^{-x t} \dif t}{t^n}$.

The remaining integrals over $\theta$ and $\chi$ are hard to do symbolically, so we compute them numerically. To optimize this computation we can use the assumption of small $\theta$ and $\chi$, so that:
\begin{align*}
\sin\theta &\sim \theta \\
\cos\sigma\left(\theta, \phi, \chi \right) &= \cos\theta \cos\chi + \sin\theta \sin\chi \cos\phi \sim 1 - \frac{\theta^2}{2} - \frac{\chi^2}{2} + \theta\chi\cos\phi
\end{align*}
Using that, and the fact that this integral is an even function of $\theta$, we arrive to the optimized expressions for $p$, $I$ and $J$:
\begin{align}
p\left( z, \chi; \tau_1, \tau_2; \omega_1, \omega_2 \right) &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)}
\frac{2\eta_0}{\left(v\gamma\left(1+z\right)\right)^{2-\alpha}}
\int_0^{\infty} \dif \theta \int_0^{\pi} \dif \phi \frac{\theta}{\left(\frac{1}{v} - 1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi\right)^{5-\alpha}}\nonumber\\
&\qquad {} \times \left( I\left(z,\chi,\omega_2; \theta,\phi\right) - I\left(z,\chi,\omega_1; \theta,\phi\right) \right) \left( J\left(z,\chi,\tau_2; \theta,\phi\right) - J\left( z,\chi,\tau_1; \theta,\phi \right) \right) \\
I\left(z,\chi,\omega;\theta,\phi\right) &= \frac{
	\omega
	\left(\frac{\omega}{\omega_0}\right)^{\alpha }
	E_{\frac{\alpha +1}{2 k}+1}\left(
		\left(\frac{\theta}{\theta_0}\right)^2
		\left(\frac{\omega}{\omega_0}\right)^{-2 k}
		\left(
			v \gamma \left(1+z\right) \left(\frac{1}{v} - 1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi \right)
		\right)^{-2 k}
	\right)
}{
	2 k
} \\
J\left(z,\chi,\tau;\theta,\phi\right) &= \frac{\tau^3}{3} \,
_2F_1\left(
	1,
	\frac{3}{n};
	\frac{n+3}{n};
	-\left(\frac{\tau}{
		r_0 \left( \frac{1}{v} -  1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi \right) \left(1+z\right)
	}\right)^n
\right)
\end{align}

Finally, we will need to compute limits of $p$ for $\omega_2 \rightarrow \infty$ and $\tau_2 \rightarrow \infty$. It requires us to know the limit of $I$ for $\omega \rightarrow \infty$ (we assumed that $k < 0$):
\begin{equation}
I\left(z,\chi,\infty;\theta,\phi\right) = \lim_{\omega \rightarrow \infty} I\left(z,\chi,\omega;\theta,\phi\right) = 0
\end{equation}
and the limit of $J$ for $\tau \rightarrow \infty$ ($n > 3$ by assumption):
\begin{equation}
J\left(z,\chi,\infty;\theta,\phi\right) = \lim_{\tau \rightarrow \infty} J\left(z,\chi,\tau;\theta,\phi\right) = \left(r_0 \left( \frac{1}{v} -  1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi \right) \left(1+z\right)\right)^3 \frac{\pi}{n\sin\frac{3\pi}{n}}
\end{equation}

Now, when we know $p$, we can compute many different things with it. Examples include:
\begin{itemize}
\item{the total number of particles observed in a given energy range, $p_\infty\left(z,\chi; \omega_1,\omega_2\right) = p\left( z,\chi; 0,\infty; \omega_1,\omega_2 \right)$;}
\item{the fraction of photons observed during a given time interval, $\Phi\left(z,\chi; \tau_1,\tau_2; \omega_1,\omega_2\right) = \frac{p\left( z,\chi; \tau_1,\tau_2; \omega_1,\omega_2\right)}{p_\infty\left( z,\chi; \omega_1,\omega_2 \right)}$;}
\item{the duration of the burst $T_f\left( z,\chi;\omega_1,\omega_2 \right)$, that is the time by which the fraction $f$ of photons is observed. We can compute it by solving the following equation for $T_f$: $p\left( z,\chi; 0,T_f; \omega_1,\omega_2\right) = f p_\infty\left( z,\chi; \omega_1,\omega_2 \right)$;}
\item{the stretching factor, which we will discuss in the next section.}
\end{itemize}

\subsection{Stretching factor}
The stretching factor for a continuous light curve is defined exactly like the stretching factor for a discrete one. It is the value $\kappa$ which makes the KS-distance minimal:
\begin{equation}
\kappa\left(z,\chi; \omega_1, \omega_2, \omega_3\right) = \argmin_\kappa \max_\tau\left| \Phi\left(z,\chi; 0,\tau; \omega_1,\omega_2\right) - \Phi\left( z,\chi; 0,\kappa \tau; \omega_2,\omega_3 \right) \right|
\end{equation}

But the maximum of an absolute value cannot be differentiated, so it is a complicated thing to compute numerically. Lets instead rewrite the expression for $\kappa$ in terms of a positive and a negative KS-distances:
\begin{align*}
D_+\left(z,\chi; \kappa; \omega_1, \omega_2, \omega_3\right) &= \max_\tau\left( \Phi\left(z,\chi; 0,\tau; \omega_1,\omega_2\right) - \Phi\left( z,\chi; 0,\kappa \tau; \omega_2,\omega_3 \right) \right) \\
D_-\left(z,\chi; \kappa; \omega_1, \omega_2, \omega_3\right) &= \min_\tau\left( \Phi\left(z,\chi; 0,\tau; \omega_1,\omega_2\right) - \Phi\left( z,\chi; 0,\kappa \tau; \omega_2,\omega_3 \right) \right) \\
\kappa\left(z,\chi; \omega_1, \omega_2, \omega_3\right) &= \argmin_\kappa \max\left(D_+\left(z,\chi; \kappa; \omega_1, \omega_2, \omega_3\right), -D_-\left(z,\chi; \kappa; \omega_1, \omega_2, \omega_3\right)\right)
\end{align*}

Take note that $\Phi$ monotonously increases with $\tau$. It implies that $D_+$ and $D_-$ monotonously decrease with $\kappa$, so as $D_+ + D_-$. So there is a single value of $\kappa$, for which
\begin{equation}
D_+\left(z,\chi; \kappa; \omega_1, \omega_2, \omega_3\right) = -D_-\left(z,\chi; \kappa; \omega_1, \omega_2, \omega_3\right)
\end{equation}
And this value of $\kappa$ also makes the $\max\left({D_+, -D_-}\right)$ minimal, since $D_+$ monotonously increase, and $-D_-$ monotonously decrease with $\kappa$.

So we have now a simplified way to compute $\kappa$ by solving an equation instead of computing the minimum. And $\Phi$ can be differentiated, which makes it easier to compute maximums and minimums of their differences.

Being able to compute the stretching factor, we could now compare our model predictions with observations given the position of an observer. However, we don't know the observer's off-axis angle $\chi$, so we cannot check the stretching factor prediction directly. Instead, we will focus on a series of tests, which will ensure that our model doesn't contradict existing observations. These tests will be discussed in the following sections.

\subsection{Total energy}
\subsection{Spectrum}
\subsection{Distribution}
\subsection{Case study}

\section{Summary}

%\bibliographystyle{plain}
%\bibliography{refs}

\end{document}

### 2013-10-18 gammaRays.tex
Revised Stretching factor section

@@
 \subsection{Stretching factor}
-The stretching factor for a continuous light curve is defined exactly like the stretching factor for a discrete one. It is the value $\kappa$ which makes the KS-distance minimal:
+The stretching factor for a continuous light curve is defined exactly like the stretching factor for a discrete one. It is the value of $\kappa$ which makes the KS-distance minimal:
 \begin{equation}
@@
 
-But the maximum of an absolute value cannot be differentiated, so it is a complicated thing to compute numerically.
-Lets instead rewrite the expression for $\kappa$ in terms of a positive and a negative KS-distances:
+The maximum of an absolute value cannot be differentiated, so the computation of $\kappa$ by the given definition is complicated.
+Lets instead rewrite this expression in terms of a positive and a negative KS-distances:
 \begin{align*}
@@
 \end{equation}
-And this value of $\kappa$ also makes the $\max\left({D_+, -D_-}\right)$ minimal, since $D_+$ monotonously increase, and $-D_-$ monotonously decrease with $\kappa$.
+And this value of $\kappa$ also makes the $\max\left({D_+, -D_-}\right)$ minimal, since $D_+$ monotonously decrease, and $-D_-$ monotonously increase with $\kappa$.
 
-So we have now a simplified way to compute $\kappa$ by solving an equation instead of computing the minimum.
+So we have now a simpler way to compute $\kappa$ by solving an equation instead of computing the minimum.
 And $\Phi$ can be differentiated, which makes it easier to compute maximums and minimums of their differences.
@@
 Being able to compute the stretching factor, we could now compare our model predictions with observations given the position of an observer.
-However, we don't know the observer's off-axis angle $\chi$, so we cannot check the stretching factor prediction directly.
+We don't know the observer's off-axis angle $\chi$, however, so we cannot check the stretching factor prediction directly.
 Instead, we will focus on a series of tests, which will ensure that our model doesn't contradict existing observations.

### 2013-10-20 gammaRays.tex
Fixed definition of \eta

@@
 	\end{equation}
-	$\eta$ is a number of particles emitted per volume per frequency.
+	$\eta$ is a number of particles emitted per volume per solid angle per frequency.
 It is a function of the distance $r$ from the central engine, of the off-axis angle $\theta$, and of the radiation frequency $\omega$.
@@
 \subsection{Total energy}
+
+The first test is to compute the total energy radiated from the burst, and to ensure that it doesn't exceed the mass of the star from which the burst originated.
+
+However, the largest contribution to the total energy in our model is due to unobservable low-energy photons.
+Moreover, the total energy integral diverges on low energies as we will see later.
+This divergence is understandable, since our model is only valid only for large energy photons.
+Still, we have to ensure that the total energy of radiated high energy photons is low enough.
+
+To do that, we have to modify the expression for the radiation intensity, to cut off all the radiators with frequencies outside the $\left(\omega_1, \omega_2\right)$ interval:
+\begin{equation}
+	\eta_{\omega_1}^{\omega_2}\left(r,\theta,\omega\right) = \begin{cases}
+		0 & \text{if } \omega < \omega_1 \\
+		\eta\left(r,\theta,\omega\right) & \text{if } \omega_1 < \omega < \omega_2 \\
+		0 & \text{if } \omega_2 < \omega
+	\end{cases}
+\end{equation}
+
 \subsection{Spectrum}

### 2013-10-20 gammaRays.tex
Written Total Energy section up to a divergent integral

--- 2013-10-18 gammaRays.tex
@@
 \item{$k$, which determines how much the opening angle changes with frequency, $k < 0$;}
-\item{$\alpha$, the bare spectral index.}
+\item{$\alpha$, the bare spectral index, $\alpha < -2k - 1$}
 \end{itemize}
@@
 \subsection{Total energy}
+
+The first test is to compute the total energy radiated from the burst, and to ensure that it doesn't exceed the mass of the star from which the burst originated.
+
+To compute the total energy, we need to multiply the radiation intensity by frequency, and integrate it over frequencies $\omega$, volume $\left(r,\theta,\phi\right)$, and observer positions $\left(\sigma, \xi\right)$.
+We assumed $\theta \ll 1$, and have taken into account the same relativistic effects as we did in the observed particle count computation:
+\begin{align}
+E &= \int_0^\infty \dif \omega \int_0^\infty \dif r \int_0^\infty r \dif \theta \int_0^{2\pi} r \, \theta \dif \phi \int_0^\infty \dif \sigma \int_0^{2\pi} \sin\sigma \dif \xi \nonumber\\
+&\quad {} \times \eta\left( r,\theta,\omega \right) \underbrace{\frac{1}{\gamma\left(1-v\cos\sigma \right)}}_{\text{time dilation}} \underbrace{\frac{1}{\gamma^2\left(1-v\cos\sigma \right)^2}}_{\text{aberration}} \underbrace{\frac{\omega}{\gamma\left(1-v\cos\sigma \right)}}_{\text{relativistic shift}} \\
+&= \frac{4\pi^2}{\gamma^4} \int_0^\infty \dif \omega \, \omega \int_0^\infty \dif r \, r^2 \int_0^\infty \dif \theta \, \theta \, \eta\left( r,\theta,\omega \right) \int_0^\infty \dif \sigma \frac{\sin\sigma}{\left(1-v\cos\sigma \right)^4} \nonumber
+\end{align}
+An integral over $\sigma$ is computable analytically:
+\begin{equation*}
+\int_0^\infty \dif \sigma \frac{\sin\sigma}{\left(1-v\cos\sigma \right)^4} = \frac{2\left(3+v^2\right)}{3\left(1-v^2\right)^3} = \frac{2}{3}\gamma^6 \left(4 - \frac{1}{\gamma^2}\right)
+\end{equation*}
+Substituting the result to the integral for $E$ we get:
+\begin{equation}
+E = \frac{32\pi^2}{3}\left(\gamma^2 - \frac{1}{4}\right) \int_0^\infty \dif \omega \, \omega \int_0^\infty \dif r \, r^2 \int_0^\infty \dif \theta \, \theta \, \eta\left( r,\theta,\omega \right)
+\end{equation}
+Now we can use the expression for $\eta$ to do the remaining integrals:
+\begin{align*}
+E = \frac{32\pi^2 \eta_0}{3}\left(\gamma^2 - \frac{1}{4}\right) \int_0^\infty \dif \omega \, \omega \left(\frac{\omega}{\omega_0}\right)^\alpha \int_0^\infty \dif r \, \frac{r^2}{1 + \left(\frac{r}{r_0}\right)^n} \int_0^\infty \dif \theta \, \theta \, \exp\left(-\left(\frac{\theta}{\theta_0}\right)^2\left(\frac{\omega}{\omega_0}\right)^{-2k}\right)
+\end{align*}
+The integrals over $r$ and $\theta$ can be done symbolically:
+\begin{equation*}
+\int_0^\infty \dif r \, \frac{r^2}{1 + \left(\frac{r}{r_0}\right)^n} = \frac{\pi}{n\sin\left(\frac{3\pi}{n}\right)}r_0^3
+\end{equation*}
+\begin{equation*}
+\int_0^\infty \dif \theta \, \theta \, \exp\left(-\left(\frac{\theta}{\theta_0}\right)^2\left(\frac{\omega}{\omega_0}\right)^{-2k}\right) = \frac{1}{2}\theta_0^2 \left(\frac{\omega}{\omega_0}\right)^{2k}
+\end{equation*}
+Now we have only one integral left:
+\begin{equation}
+E = \frac{16\pi^3}{3 n\sin\left(\frac{3\pi}{n}\right)} \left(\gamma^2 - \frac{1}{4}\right) \, \eta_0 \, r_0^3 \, \theta_0^2 \int_0^\infty \dif \omega \, \omega \left(\frac{\omega}{\omega_0}\right)^{2k+\alpha}
+\end{equation}
+
 \subsection{Spectrum}

### 2013-10-20 gammaRays.tex
Drafted Total Energy section

@@
 \end{equation*}
-Substituting the result to the integral for $E$ we get:
+Substituting this result into the expression for $E$ we get:
 \begin{equation}
@@
 \end{equation*}
-Now we have only one integral left:
+Now we only have one integral left:
 \begin{equation}
 E = \frac{16\pi^3}{3 n\sin\left(\frac{3\pi}{n}\right)} \left(\gamma^2 - \frac{1}{4}\right) \, \eta_0 \, r_0^3 \, \theta_0^2 \int_0^\infty \dif \omega \, \omega \left(\frac{\omega}{\omega_0}\right)^{2k+\alpha}
+\end{equation}
+
+Note, however, that since $\alpha < -2k-1$ this integral diverges for $\omega \rightarrow 0$.
+But our model is not expected to describe low-energy radiation of the burst, so we should not integrate over low-frequency radiators.
+Nevertheless, we should ensure that the total energy of the high-energy radiation is not too large.
+To compute this energy we will only integrate over those radiators which emission we can observe, that is over the radiators with frequencies greater than $\frac{\omega_1}{\gamma}$, where $\omega_1$ is the smallest observable photon frequency.
+Now we can compute the last integral and get the final expression for $E$:
+\begin{align}
+E\left(\omega_1\right)
+&= \frac{16\pi^3}{3 n\sin\left(\frac{3\pi}{n}\right)} \left(\gamma^2 - \frac{1}{4}\right) \, \eta_0 \, r_0^3 \, \theta_0^2 \int_{\frac{\omega_1}{\gamma}}^\infty \dif \omega \, \omega \left(\frac{\omega}{\omega_0}\right)^{2k+\alpha} \nonumber\\
+&= \frac{16\pi^3}{3 n\sin\left(\frac{3\pi}{n}\right)} \left(\gamma^2 - \frac{1}{4}\right) \, \eta_0 \, r_0^3 \, \theta_0^2 \frac{\omega_1^2 \left(\frac{\omega_1}{\gamma \omega_0}\right)^{2k+\alpha}}{\gamma ^2 (-2k-\alpha-2)} \nonumber\\
+&= \frac{16\pi^3}{3 n\left(-2k-\alpha-2\right)\sin\left(\frac{3\pi}{n}\right)} \gamma^{-2k-\alpha} \left(1 - \frac{1}{4\gamma^2}\right) \, \eta_0 \, r_0^3 \, \theta_0^2 \frac{\omega_0^{-2k-\alpha}}{\omega_1^{-2k-\alpha-2}}
+\end{align}
+
+This energy should not be larger than a typical mass $M_s$ of a massive star. So, finally, we arrive at the first constraint for the burst parameters:
+\begin{equation}
+\frac{16\pi^3}{3 n\left(-2k-\alpha-2\right)\sin\left(\frac{3\pi}{n}\right)} \gamma^{-2k-\alpha} \left(1 - \frac{1}{4\gamma^2}\right) \, \eta_0 \, r_0^3 \, \theta_0^2 \frac{\omega_0^{-2k-\alpha}}{\omega_1^{-2k-\alpha-2}} < M_s
 \end{equation}

### 2013-11-21 gammaRays.tex
Small style improvements

@@
 \item{The shell points propagate with a constant velocity $v \sim 1$, so at the time $t$ the radius of the shall is $v t$.}
-\item{Each point of the shell is an isotropic radiator in its reference frame.}
+\item{Each point of the shell is an isotropic radiator in its rest frame.}
 \item{
@@
 
-Lets begin by calculating a time at which some particular photon is observed.
+Lets begin by computing a time at which some particular photon is observed.
 This time is a function of the radiator location $\left(r, \theta, \phi\right)$, as well as the observer location $\left(d, \chi, 0\right)$.
@@
 
-We also need to know the areas of two spheres: the photon sphere, the sphere over which the photons emitted in a particular burst are spread; and the bursts sphere, the sphere over which the bursts at a particular redshift are distributed.
+We also need to know the areas of two spheres: the photon sphere, a sphere over which photons emitted in a particular burst are spread; and the bursts sphere, a sphere over which bursts at a particular redshift are distributed.
 
@@
 \end{align*}
-Using that, and the fact that this integral is an even function of $\theta$, we arrive to the optimized expressions for $p$, $I$ and $J$:
+Using that, and the observation that this integral is an even function of $\theta$, we arrive to the optimized expressions for $p$, $I$ and $J$:
 \begin{align}
@@
 
-\subsection{Spectrum}
 \subsection{Distribution}
-\subsection{Case study}
+\subsection{Burst fraction with observable high-energy photons}
 

### 2013-11-21 gammaRays.tex
drafted Distribution section (without pictures)

@@
 
-\subsection{Distribution}
-\subsection{Burst fraction with observable high-energy photons}
+\subsection{Distribution of stretching factors}
+
+The second test is to calculate the stretching factors distribution of the observable bursts, and to compare it with observations.
+
+The computation of the exact and precise distribution is technically hard (because stretching factors neither change monotonously with redshift, nor with off-axis angle), and computationally intensive.
+So we will not compute the precise distribution.
+
+Instead, we will use the Monte-Carlo method to produce a large representative sample of stretching factors.
+Then, we will calculate an empirical CDF of this sample, and compare it with observations using the KS-test.
+Since we can compute much larger sample than that of observations, we will not lose much precision due to this simplification.
+
+We still have one ingredient missing though -- the evolution of the bursts density.
+We assume that the bursts density is roughly proportional to the stars density, and the stars density is roughly proportional to the matter density, which changes as $\left(1+z\right)^3$.
+It is clear, however, that since no stars existed at very small redshifts, the burst density should decline there, so we add an exponential cutoff to it:
+\begin{equation}
+\rho = \rho_0 \left(1+z\right)^3 \exp\left(-\frac{z}{z_c}\right)
+\end{equation}
+Here $\rho_0$ is a normalization factor, and $z_c$ is a redshift scale, after which the density is cut off.
+
+Now when we have all the ingredients, we can describe the algorithm for computing a sample of stretching factors:
+\begin{enumerate}
+\item{
+	Select a redshift $z$. The distribution of redshifts is determined by the bursts count in a given volume. Its CDF is:
+	\begin{equation}
+	\Phi_z\left(z\right) = \frac{\int_{z_\text{min}}^z \rho\left(z'\right) \dif V\left(z'\right)}{\int_{z_\text{max}}^{z_\text{max}} \rho\left(z'\right) \dif V\left(z'\right)}
+	\end{equation}
+	Here $\left(z_\text{min}, z_\text{max}\right)$ is the range from which we are selecting redshifts.
+$z_\text{min}$ should be some small number so it is unlikely to detect a burst with $z < z_\text{min}$.
+And $z_\text{max}$ is the largest redshift from which a burst can be observed:
+	\begin{equation}
+	p_\infty\left(z_\text{max}, 0., \omega_2, \omega_3\right) = p_\text{min}
+	\end{equation}
+	Here $p_\text{min}$ is the smallest number of particles required to claim an observation.
+
+	Now we can select a properly distributed random redshift by solving an equation for $z$:
+	\begin{equation}
+	\Phi_z\left(z\right) = x
+	\end{equation}
+	where $x$ is a uniformly distributed random number in the range $0$ to $1$.
+}
+\item{
+	Select an off-axis angle $\chi$. All directions for the burst jets are equally probable. So the CDF of the off-axis angles distribution is:
+	\begin{equation}
+	\Phi_\chi\left(\chi\right) = \frac{\int_0^\chi \sin\chi' \dif\chi'}{\int_0^{\chi_\text{max}} \sin\chi' \dif\chi'} \approx \frac{\int_0^\chi \chi' \dif\chi'}{\int_0^{\chi_\text{max}} \chi' \dif\chi'} = \left(\frac{\chi}{\chi_\text{max}}\right)^2
+	\end{equation}
+	Here $\chi_\text{max}$ is the largest off-axis angle with which a burst at redshift $z_\text{min}$ can be observed:
+	\begin{equation}
+	p_\infty\left(z_\text{min}, \chi_\text{max}, \omega_2, \omega_3\right) = p_\text{min}
+	\end{equation}
+
+	Now as with $z$ distribution we can find the random $\chi$ by solving the equation:
+	\begin{align*}
+	\Phi_\chi\left(\chi\right) = \left(\frac{\chi}{\chi_\text{max}}\right)^2 = y \\
+	\chi = \chi_\text{max}\sqrt{y}
+	\end{align*}
+	where $y$ is a uniformly distributed random number in the range $0$ to $1$.
+
+	Note that $p\left(z,\chi\right)$ decreases with $z$, as well as with $\chi$, so all observable burst configurations can be generated except for the unlikely $z<z_\text{min}$ case.
+}
+\item{If the burst is observable: $p_\infty\left(z,\chi,\omega_2,\omega_3\right) > p_\text{min}$, add $\kappa\left(z,\chi\right)$ to the sample.}
+\item{Repeat from the step $1$.}
+\end{enumerate}
+
+For our set of parameters, this algorithm produced a sample, empirical CDF of which you can see on fig.
+The error margins of the observed stretching factors are too large to use the KS-test on them.
+It appears, however, that our model doesn't contradict existing observations.
+
+\subsection{Very high energy bursts fraction}
 

### 2013-11-22 gammaRays.tex
Rewritten description of the distribution computation algorithm

--- 2013-11-21 gammaRays.tex
@@
 
-\subsection{Distribution}
-\subsection{Burst fraction with observable high-energy photons}
+\subsection{Distribution of stretching factors}
+
+The second test is to calculate the distribution of the stretching factors of the observable bursts, and to compare it with observations.
+
+The computation of the exact and precise distribution is technically hard (because stretching factors neither change monotonously with redshift, nor with off-axis angle), and computationally intensive.
+So we will not compute the precise distribution.
+
+Instead, we will use the Monte-Carlo method to produce a large representative sample of stretching factors.
+Then, we will calculate an empirical CDF of this sample, which can be compared with observations using the KS-test.
+Since we can compute much larger sample than that of observations, we will not lose much precision due to this simplification.
+
+We still have one ingredient missing though -- the evolution of the bursts density.
+We assume that the density is roughly proportional to the stars density, and the stars density is roughly proportional to the matter density, which changes with $z$ as $\left(1+z\right)^3$.
+It is clear, however, that since no stars existed at very small redshifts, the burst density should decline there, so we add an exponential cutoff to it:
+\begin{equation}
+\rho = \rho_0 \left(1+z\right)^3 \exp\left(-\frac{z}{z_c}\right)
+\end{equation}
+Here $\rho_0$ is a normalization factor, and $z_c$ is a redshift scale, after which the density is cut off.
+
+Now when we have all the ingredients, we can compute the sample with the following steps:
+\begin{enumerate}
+\item{
+	First of all we need to define the range from which to select redshifts and off-axis angles.
+We should include all observable jet positions in this range.
+However, to avoid dropping too many points corresponding to invisible bursts, we should keep the range as small as possible.
+We also want to make the region rectangular to be able to select redshifts and angles separately.
+
+	Lets start by selecting a range for redshifts.
+Ideally, we want this range to start with $z=0$.
+Then, however, there will be visible jets for all possible off-axis angles, which will make our angles range too large.
+Note also, that since the bursts count increases with redshift as $z^3$, the probability to observe a low-redshift burst is very small.
+So instead of selecting the smallest redshift to be $0$, we select it to be a small number $z_\text{min}$.
+
+	$z_\text{max}$ is defined by the farthest jet, which can be observed.
+Since the burst observability $p_\infty\left(z,\chi,\omega_2,\omega_3\right)$ decreases with both redshift and $\chi$, we can find the maximum redshift by solving the following equation:
+	\begin{equation}
+	p_\infty\left(z_\text{max},0,\omega_2,\omega_3\right) = p_\text{min}
+	\end{equation}
+	where $p_\text{min}$ is the minimum number of particles required to claim an observation.
+}
+\item{
+	The range for off-axis angles is easier to compute since nothing prohibits us from selecting the smallest angle to be $0$.
+
+	The observability declines with $z$ and $\chi$, so the observable burst with the largest $\chi$ should be located at the redshift $z_\text{min}$.
+We can find this angle by solving the similar equation, as we did for the redshifts:
+	\begin{equation}
+	p_\infty\left(z_\text{min},\chi_\text{max},\omega_2,\omega_3\right) = p_\text{min}
+	\end{equation}
+}
+\item{
+	Now we are in a position to select a particular properly distributed random redshift. The CDF of the distribution of redshifts is a ratio of redshift counts in different space volumes:
+	\begin{equation}
+	\Phi_z\left(z\right) = \frac{\int_{z_\text{min}}^z \rho\left(z'\right) \dif V\left(z'\right)}{\int_{z_\text{max}}^{z_\text{max}} \rho\left(z'\right) \dif V\left(z'\right)}
+	\end{equation}
+
+	To generate a redshift, we should uniformly select a value of $\Phi_z$, and solve the corresponding equation for $z$:
+	\begin{equation}
+	\Phi_z\left(z\right) = x
+	\end{equation}
+	where $x$ is a random variable uniformly distributed in the range $0$ to $1$.
+}
+\item{
+	An off-axis angle can be selected in a similar way. The CDF of the angles distribution is a ratio of spherical areas:
+	\begin{equation}
+	\Phi_\chi\left(\chi\right) = \frac{\int_0^\chi \sin\chi' \dif\chi'}{\int_0^{\chi_\text{max}} \sin\chi' \dif\chi'} \approx \frac{\int_0^\chi \chi' \dif\chi'}{\int_0^{\chi_\text{max}} \chi' \dif\chi'} = \left(\frac{\chi}{\chi_\text{max}}\right)^2
+	\end{equation}
+
+	As with redshifts, we get a properly distributed $\chi$ by solving the equation:
+	\begin{align}
+	\Phi_\chi\left(\chi\right) = \left(\frac{\chi}{\chi_\text{max}}\right)^2 = y \\
+	\chi = \chi_\text{max}\sqrt{y}
+	\end{align}
+	Here $y$ is an another random variable uniformly distributed in the range $0$ to $1$.
+}
+\item{Now we should check, if the burst in a selected position can be observed: $p_\infty\left(z,\chi,\omega_2,\omega_3\right) > p_\text{min}$.
+If it is, add $\kappa\left(z,\chi\right)$ to the sample.
+If not, repeat from the step $3$.}
+\item{If the sample is not as big as we want yet, repeat from the step $3$.}
+\end{enumerate}
+
+For our set of parameters, this algorithm produced a sample, empirical CDF of which you can see on fig.
+However, the error margins of the observed stretching factors are too large to use the KS-test on them.
+Nevertheless, It appears that our model doesn't contradict existing observations.
+
+\subsection{Very high energy bursts fraction}
 

### 2013-11-23 gammaRays.tex
Partially drafted Very high energy bursts fraction section

@@
 
+Our final test is to calculate how much of the observable bursts can be observed in the very high energy range -- GeV and above.
+
+To do so, we need to compute the number of bursts visible in a given energy range, which is the integral over space and jet directions:
+\begin{align}
+b\left(\omega_1,\omega_2\right) &= \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} 2\pi \sin\chi \dif\chi \nonumber\\
+&\approx 2\pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} \chi \dif\chi = \pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_2\right)
+\end{align}
+Here $z_\text{max}\left(\omega_1,\omega_2\right)$ and $\chi_\text{max}\left(z,\omega_1,\omega_2\right)$ are the same values we used in the previous sections: the maximum redshift from which a burst can be observed, and the maximum off-axis angle with which the burst at redshift $z$ can be observed.
+
+The fraction in question is simply the ratio of these:
+\begin{equation}
+f\left(\omega_1,\omega_2,\omega_3\right) = \frac{b\left(\omega_2,\omega_3\right)}{b\left(\omega_1,\omega_3\right)} = \frac{\int_0^{z_\text{max}\left(\omega_2,\omega_3\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_2,\omega_3\right)}{\int_0^{z_\text{max}\left(\omega_1,\omega_3\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_3\right)}
+\end{equation}
+
+The ratio for our set of parameters is $f_m \approx 0.072$. The ratio computed from observations is $f_o = ...$. The binary distribution gives us the probability for it to happen:
+\begin{equation}
+...
+\end{equation}
+which means that the model fits observations in this test.
+
 \section{Summary}

### 2013-11-23 gammaRays.tex
Small phrasing changes in VHE bursts fraction section

@@
 
-Our final test is to calculate how much of the observable bursts can be observed in the very high energy range -- GeV and above.
+Our final test is to calculate the fraction of the observable bursts which can be observed at the energies of GeV and above.
 
@@
 \end{align}
-Here $z_\text{max}\left(\omega_1,\omega_2\right)$ and $\chi_\text{max}\left(z,\omega_1,\omega_2\right)$ are the same values we used in the previous sections: the maximum redshift from which a burst can be observed, and the maximum off-axis angle with which the burst at redshift $z$ can be observed.
+Here $z_\text{max}\left(\omega_1,\omega_2\right)$ and $\chi_\text{max}\left(z,\omega_1,\omega_2\right)$ are the same values we used in the previous section: the maximum redshift from which a burst can be observed, and the maximum off-axis angle with which the burst at redshift $z$ can be observed.
 

### 2013-11-24 gammaRays.tex
Added kappaHistogram figures

@@
 \usepackage{fullpage}
+\usepackage{graphicx}
+\usepackage{subcaption}
 
@@
 
-For our set of parameters, this algorithm produced a sample, empirical CDF of which you can see on fig.
+\begin{figure}[kappaDistribution]
+        \centering
+        \begin{subfigure}[kappaDistributionHistogram]{0.4\textwidth}
+                \includegraphics[width=\textwidth]{kappaDistributionHistogram}
+                \label{fig:kappaDistributionHistogram}
+        \end{subfigure}
+        \begin{subfigure}[kappaDistributionCDF]{0.4\textwidth}
+                \includegraphics[width=\textwidth]{kappaDistributionCDF}
+                \label{fig:kappaDistributionCDF}
+        \end{subfigure}
+        \caption{The histogram and the CDF of the stretching factors sample produced by our model.
+The sample contains $200$ points.
+You can see the stretching factors computed from the observations below the $\kappa$ axis (not yet implemented).}
+        \label{fig:kappaDistribution}
+\end{figure}
+
+For our set of parameters, this algorithm produced a sample you can see on the figure \ref{fig:kappaDistribution}.
 However, the error margins of the observed stretching factors are too large to use the KS-test on them.

### 2013-11-25 gammaRays.tex
computed kappa sample to 250 points
added observations to distirbution plots

--- 2013-11-23 gammaRays.tex
@@
 
-For our set of parameters, this algorithm produced a sample, empirical CDF of which you can see on fig.
+\begin{figure}[kappaDistribution]
+        \centering
+        \begin{subfigure}[kappaDistributionHistogram]{0.49\textwidth}
+                \includegraphics[width=\textwidth]{kappaDistributionHistogram}
+                \label{fig:kappaDistributionHistogram}
+        \end{subfigure}
+        \begin{subfigure}[kappaDistributionCDF]{0.49\textwidth}
+                \includegraphics[width=\textwidth]{kappaDistributionCDF}
+                \label{fig:kappaDistributionCDF}
+        \end{subfigure}
+        \caption{The histogram and the CDF of the stretching factors sample produced by our model (blue).
+The sample contains $250$ points.
+The red curve is an empirical CDF of the observed sample (3 points).}
+        \label{fig:kappaDistribution}
+\end{figure}
+
+Our set of parameters led to a sample you can see on the figure \ref{fig:kappaDistribution}.
 However, the error margins of the observed stretching factors are too large to use the KS-test on them.

### 2013-11-25 gammaRays.tex
computed observations in the fraction test

@@
 However, the error margins of the observed stretching factors are too large to use the KS-test on them.
-Nevertheless, It appears that our model doesn't contradict existing observations.
+Nevertheless, it appears that our model doesn't contradict existing observations.
 
@@
 
-Our final test is to calculate the fraction of the observable bursts which can be observed at the energies of GeV and above.
+Out final test is the comparison of answers to this question: given the bursts which were observed in a high energy range $\left(\omega_1,\omega_3\right)$, which fraction of them can also be observed in a very high energy range $\left(\omega_2,\omega_3\right)$?
 
-To do so, we need to compute the number of bursts visible in a given energy range, which is the integral over space and jet directions:
+To calculate this value, we need to compute the number of bursts visible in a given energy range, which is the integral over space and jet directions:
 \begin{align}
@@
 \end{align}
-Here $z_\text{max}\left(\omega_1,\omega_2\right)$ and $\chi_\text{max}\left(z,\omega_1,\omega_2\right)$ are the same values we used in the previous section: the maximum redshift from which a burst can be observed, and the maximum off-axis angle with which the burst at redshift $z$ can be observed.
+Here $z_\text{max}\left(\omega_1,\omega_2\right)$ and $\chi_\text{max}\left(z,\omega_1,\omega_2\right)$ are the same values we used in the previous section: the maximum redshift from which a burst can be observed in a given energy range, and the maximum off-axis angle with which the burst at redshift $z$ can be observed.
 
-The fraction in question is simply the ratio of these:
+The fraction we want to compute is the ratio of these integrals:
 \begin{equation}
@@
 
-The ratio for our set of parameters is $f_m \approx 0.072$. The ratio computed from observations is $f_o = ...$. The binary distribution gives us the probability for it to happen:
+The ratio for our set of parameters is $f_m = 0.072$.
+The LAT burst catalog \ref{LATbc} contains $n = 35$ bursts, $3$ of which were observed in the energy range ${1\text{GeV}, \infty}$.
+So the ratio is $f_o = \frac{3}{35} \approx 0.086$.
+The binary distribution gives us the probability for it to happen:
 \begin{equation}
-...
+\text{Pr}\left(k_o = k\right) = \frac{n}{k} f^k \left(1-f\right)^{n-k} \approx 0.223585
 \end{equation}
-which means that the model fits observations in this test.
+which means that the model doesn't contradict observations in this test.
 

### 2013-12-01 gammaRays.tex
Added reference to LAT catalog
Computed kappa sample to 300 points
Drafted an abstract

@@
 \usepackage{subcaption}
+\usepackage{units}
 
@@
 \maketitle
+
+\begin{abstract}
+Numerous observations confirm that the high energy $(> \unit[100]{MeV})$ emission of gamma ray bursts is delayed with respect to the low energy emission.
+However, the difference of light curves in various high energy bands has not been studied properly.
+
+In this paper we consider all the bursts observed by Fermi-LAT since 2008 August 4 to 2011 August 1, for which at least $10$ events with energies $\unit[1]{GeV}$ or higher were observed.
+There are $3$ of them: GRB080916C, GRB090902B, and GRB090926A.
+We study their light curves in two bands, $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.
+The Kolmogorov-Smirnov test is used to check whether the light curves for these two bands are the same.
+No significant difference was found for GRB080916C and GRB090902B.
+However, we observed with statistical significance of $3.4 \sigma$, that the higher energy light curve of GRB090926A is stretched with respect to the lower-energy one.
+
+We suggest a simple geometrical model to explain this result.
+The main assumption is the jet opening angle dependence on radiation energy -- the most energetic photons are emitted near the axis of the jet.
+To test this model, we compute the total energy of the burst, and confirm that it is below the constraint.
+We also compute the fraction of observable bursts in $(\unit[100]{MeV}, \unit[1]{GeV})$ band, which can also be observed in higher energies.
+This fraction matches the observations.
+Finally, we predict the distribution of observable stretching factors, which may be tested in the future when more observational data will be available.
+\end{abstract}
 
@@
         \caption{The histogram and the CDF of the stretching factors sample produced by our model (blue).
-The sample contains $250$ points.
+The sample contains $300$ points.
 The red curve is an empirical CDF of the observed sample (3 points).}
@@
 The ratio for our set of parameters is $f_m = 0.072$.
-The LAT burst catalog \ref{LATbc} contains $n = 35$ bursts, $3$ of which were observed in the energy range ${1\text{GeV}, \infty}$.
+The LAT burst catalog \cite{Ackermann:2013zfa} contains $n = 35$ bursts, $3$ of which were observed in the energy range $\left(\unit[1]{GeV}, \infty\right)$.
 So the ratio is $f_o = \frac{3}{35} \approx 0.086$.
@@
 \begin{equation}
-\text{Pr}\left(k_o = k\right) = \frac{n}{k} f^k \left(1-f\right)^{n-k} \approx 0.223585
+\text{Pr}\left(f_o = \frac{k}{n}\right) = \binom{n}{k} f_m^k \left(1-f_m\right)^{n-k} \approx 0.22
 \end{equation}
@@
 
-%\bibliographystyle{plain}
-%\bibliography{refs}
+\bibliographystyle{alpha}
+\bibliography{gammaRays}
 

### 2013-12-08 Paper/gammaRays.tex
Added GRObservations c++ code

--- 2013-12-01 gammaRays.tex
@@
 
-\author{Maxim Piskunov}
-\title{Spectral Lags in Very High Energy Emission of Gamma Ray Bursts}
+\author{Maxim S. Piskunov and Grigory I. Rubtsov}
+\title{Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model}
 
@@
 \section{Introduction}
+
+Gamma ray bursts are the most energetic explosions in the Universe, therefore they are of considerable interest for particle physics.
+Recently developed instruments, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{SwiftReference} allowed for extensive study of gamma ray bursts \cite{Vianello:2013ela, Gehrels:2013xd}.
+Many discoveries were made: for example, it was found that the energy emitted by GRBs in gamma rays remains relatively constant for different bursts \cite{Bloom:2003wy}.
+Temporal variations of GRB spectra were studied.
+The spectral lags were observed between different low-energy bands \cite{Yi:2005ht}, as well as between low-energy and high energy radiation \cite{}.
+However, recent observations of luminous bursts by Fermi allowed to study temporal variations even in high energy spectra.
+In this study we used KS-test to search for spectral lags between two high energy bands: $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.
 
@@
 \begin{itemize}
-\item{$v$, the velocity of the shell, $\gamma = \frac{1}{\sqrt{1-v^2}} \gg 1$;}
-\item{$\eta_0$, which defines the intensity scale;}
+\item{$v$, the velocity of the shell, $\gamma = \frac{1}{\sqrt{1-v^2}} \gg 1$.
+
+	If $\gamma$ is small enough, high photon densities would make the jet plasma optically thick to pair production.
+This in turn would suppress emitted flux, and this is not observed.
+From this argument the lower bound of $\gamma \ge 100$ was derived \cite{something}.
+The limited total energy of the burst provides an upper bound: $\gamma \le \gamma_\text{max}$ (see section 3.6).
+Finally, the model dependent computations of $\gamma$ give the values ranging from $50$ for low-energy bursts to $1000$ for high-energy ones \cite{Ghirlanda:2011ux,Ghisellini:2012nc}.
+$\gamma$ may vary from burst to burst, and it should decrease during the burst due to the dissipation of energy.
+As a first approximation we select {\boldmath$\gamma = 100$} for this study.
+}
+\item{$\eta_0$, which defines the luminosity scale.
+
+	$\eta_0$ controls the number of photons we see from the bursts, and so can be roughly estimated from the number of luminous bursts we observe.
+We can use this and the measured burst redshifts to estimate both $\eta_0$ and the bursts density $\rho_0$, which will be discussed later.
+For this study we choose {\boldmath$\eta_0 = 1 \times 10^{45}$}.
+}
 \item{$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
@@
         \caption{The histogram and the CDF of the stretching factors sample produced by our model (blue).
-The sample contains $300$ points.
+The sample contains $600$ points.
 The red curve is an empirical CDF of the observed sample (3 points).}
@@
 
-\section{Summary}
+\subsection{Parameter values}
+
+\subsection{Results}
+
+\section{Discussion}
 
 \bibliographystyle{alpha}
-\bibliography{gammaRays}
+\bibliography{/Users/maxitg/Documents/bibliography.bib}
 

### 2013-12-08 GRObservations/GRObservations/GRBurst.cpp

@@ -1,0 +1,231 @@
+//
+//  GRBurst.cpp
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 24.03.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#include <sys/stat.h>
+#include <sys/errno.h>
+
+#include <math.h>
+
+#include <fstream>
+#include <sstream>
+
+#include "GRBurst.h"
+
+#define TIME_EXTENTION_FACTOR 1.3
+#define PHOTONS_QUALITY       0.95
+#define LENGTHENING_MIN       0.1
+#define LENGTHENING_MAX       10.
+#define LENGTHENING_STEP      1.001
+
+void GRBurst::init() {    
+    double duration = (endOffset - startOffset) * TIME_EXTENTION_FACTOR;
+    double center = time + (startOffset + endOffset) / 2.;
+    double startTime = center - duration/2.;
+    double endTime = center + duration/2.;
+        
+    query.startTime = startTime;
+    query.endTime = endTime;
+    query.location = location;
+        
+    query.init();
+    
+    //duration = duration * 100;
+    duration = 86400;
+    endTime = startTime;
+    startTime = startTime - duration;
+    
+    backgroundQuery.startTime = startTime;
+    backgroundQuery.endTime = endTime;
+    backgroundQuery.location = location;
+    
+    backgroundQuery.init();
+    
+    error = GRBurstErrorNotDownloaded;
+}
+
+void GRBurst::download() {
+    if (query.error == GRFermiLATDataServerQueryErrorNotDownloaded) query.download();
+    if (query.error != GRFermiLATDataServerQueryErrorNotProcessed && query.error != GRFermiLATDataServerQueryErrorNotRead && !query.error == GRFermiLATDataServerQueryErrorOk) {
+        error = GRBurstErrorQueryError;
+        errorDescription = query.errorDescription;
+        return;
+    }
+    
+    if (backgroundQuery.error == GRFermiLATDataServerQueryErrorNotDownloaded) backgroundQuery.download();
+    if (backgroundQuery.error != GRFermiLATDataServerQueryErrorNotProcessed && backgroundQuery.error != GRFermiLATDataServerQueryErrorNotRead && backgroundQuery.error != GRFermiLATDataServerQueryErrorOk) {
+        error = GRBurstErrorBackgroundQueryError;
+        errorDescription = backgroundQuery.errorDescription;
+        return;
+    }
+
[…]

### 2013-12-08 GRObservations/GRObservations/GRBurst.h

@@ -1,0 +1,83 @@
+//
+//  GRBurst.h
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 24.03.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#ifndef __Gamma_Rays__GRBurst__
+#define __Gamma_Rays__GRBurst__
+
+#include <stdio.h>
+
+#include <iostream>
+#include <vector>
+#include <string>
+#include <algorithm>
+
+#include "GRFermiLATDataServerQuery.h"
+#include "GRDistribution.h"
+
+using namespace std;
+
+enum GRBurstError {
+    GRBurstErrorOk,
+    GRBurstErrorNotDownloaded,
+    GRBurstErrorNotProcessed,
+    GRBurstErrorNotRead,
+    
+    GRBurstErrorQueryError,
+    GRBurstErrorBackgroundQueryError,
+    
+    GRBurstErrorMkdirFailed
+    };
+
+class GRBurst {
+public:
+    GRBurstError error;
+    string errorDescription;
+    
+    string name;
+    double time;
+    GRLocation location;
+    
+    double startOffset;
+    double endOffset;
+        
+public:
+    GRFermiLATDataServerQuery backgroundQuery;
+    GRFermiLATDataServerQuery query;
+    
+    vector <GRFermiLATPhoton> mevPhotons;
+    vector <GRFermiLATPhoton> mevBackgroundPhotons;
+    
+    vector <GRFermiLATPhoton> gevPhotons;
+    vector <GRFermiLATPhoton> gevBackgroundPhotons;
+    
+public:
+    GRDistribution mevDistribution;
+    GRDistribution gevDistribution;
+    
+public:
+    double minLengthening[5];
+    double maxLengthening[5];
+    vector <float> lengtheningValues;
+    vector <float> lengtheningProbabilities;
+    
+    double trivialProbability;
+    
+private:
+    void calculateBackground();
+        
+public:
+    void init();
+    void download();
+    void process();
+    void read();
+    void evaluate();
+    
+    void clear();
+};
+
+#endif /* defined(__Gamma_Rays__GRBurst__) */

### 2013-12-08 GRObservations/GRObservations/GRDistribution.cpp

@@ -1,0 +1,189 @@
+//
+//  GRDistribution.cpp
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 07.04.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#include <math.h>
+#include <cmath>
+
+#include <fstream>
+
+#include <glpk.h>
+
+#include "GRDistribution.h"
+
+vector <GRDistributionCDFPoint> GRDistribution::cdf() {
+    vector <GRDistributionCDFPoint> result;
+    result.reserve(2*values.size());
+    for (int i = 0; i < values.size(); i++) {
+        GRDistributionCDFPoint point;
+        point.value = values[i];
+        point.probability = (double)i/values.size();
+        result.push_back(point);
+        point.value = values[i];
+        point.probability = (double)(i+1)/values.size();
+        result.push_back(point);
+    }
+    return result;
+}
+
+double GRDistribution::kolmogorovSmirnovProbability(double distance, int n1, int n2) {
+    double n = (double)n1*n2/(n1+n2);
+    double z = (sqrt(n) + 0.12 + 0.11/sqrt(n)) * distance;
+    
+    if (z == 0.) return 1.;
+    if (z < 1.18) {
+        double y = exp(-1.23370055013616983/pow(z,2.));
+        return 1. - 2.25675833419102515*sqrt(-log(y))*(y + pow(y,9) + pow(y,25) + pow(y,49));
+    } else {
+        double x = exp(-2.*pow(z,2.));
+        return 2.*(x - pow(x,4) + pow(x,9));
+    }
+}
+
+double GRDistribution::kolmogorovSmirnovDistance(double probability, int n1, int n2) {
+    double y,logy,yp,x,xp,f,ff,u,t;
+    double z;
+    if (probability == 1.) z = 0.;
+    else if (probability > 0.3) {
+        f = -0.392699081698724155*pow(1.-probability, 2.);
+        
+        const double ooe = 0.367879441171442322;
+        double tp,up,to=0.;
+        if (f < -0.2) up = log(ooe-sqrt(2*ooe*(f+ooe)));
+        else up = -10.;
+        do {
+            up += (tp=(log(f/up)-up)*(up/(1.+up)));
+            if (tp < 1.e-8 && abs(tp+to)<0.01*abs(tp)) break;
+            to = tp;
[…]

### 2013-12-08 GRObservations/GRObservations/GRDistribution.h

@@ -1,0 +1,39 @@
+//
+//  GRDistribution.h
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 07.04.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#ifndef __Gamma_Rays__GRDistribution__
+#define __Gamma_Rays__GRDistribution__
+
+#include <iostream>
+#include <vector>
+#include <algorithm>
+
+using namespace std;
+
+struct GRDistributionCDFPoint {
+    double value;
+    double probability;
+};
+
+class GRDistribution {
+public:
+    vector <double> values;
+    double estimatedLinearComponent;
+    double start;
+    double end;
+    
+private:
+    double kolmogorovSmirnovProbability(double distance, int n1, int n2);
+    double kolmogorovSmirnovDistance(double probability, int n1, int n2);
+public:
+    vector <GRDistributionCDFPoint> cdf();
+    float kolmogorovSmirnovTest(GRDistribution distribution, double lengthening = 1., bool plot = false, ostream &mev = cout, ostream &gev = cout);
+    pair<double, double> lengtheningLimits(GRDistribution distribution, float probability, bool *success);
+};
+
+#endif /* defined(__Gamma_Rays__GRDistribution__) */

### 2013-12-08 GRObservations/GRObservations/GRExposureMap.cpp

@@ -1,0 +1,75 @@
+//
+//  GRExposureMap.cpp
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 28.04.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#include "GRExposureMap.h"
+
+int GRExposureMap::findEnergy(float energy) {
+    if (energy < energies[0]) return -1;
+    else {
+        int lowerEnergy = 0;
+        int upperEnergy = (int)energies.size()-1;
+        while (upperEnergy - lowerEnergy > 1) {
+            if (energies[(upperEnergy + lowerEnergy)/2] > energy) {
+                upperEnergy = (upperEnergy + lowerEnergy)/2;
+            } else {
+                lowerEnergy = (upperEnergy + lowerEnergy)/2;
+            }
+        }
+        return lowerEnergy;
+    }
+}
+
+float GRExposureMap::exposure(float energy, GRLocation location) {
+    int lowerIndex = findEnergy(energy);
+    int upperIndex;
+    if (lowerIndex == -1) {
+        lowerIndex = 0;
+        upperIndex = 1;
+    } else if (lowerIndex == (int)energies.size()-1) {
+        lowerIndex = (int)energies.size()-2;
+        upperIndex = (int)energies.size()-1;
+    } else {
+        upperIndex = lowerIndex+1;
+    }
+    
+    double lowerSpread = exposure(lowerIndex, location);
+    double upperSpread = exposure(upperIndex, location);
+    double lowerEnergy = energies[lowerIndex];
+    double upperEnergy = energies[upperIndex];
+    
+    if (lowerEnergy == upperEnergy) return lowerSpread;
+    else return lowerSpread + (upperSpread - lowerSpread) / (upperEnergy - lowerEnergy) * (energy - lowerEnergy);
+}
+
+float GRExposureMap::exposure(int energyIndex, GRLocation location) {
+    int raLow = location.ra;
+    int raHigh = raLow+1;
+    if (raHigh == 360) {
+        raLow = 358;
+        raHigh = 359;
+    }
+    
+    int decLow = location.dec + 90.;
+    int decHigh = decLow+1;
+    if (decHigh == 180) {
+        decLow = 178;
+        decHigh = 179;
+    }
+    
+    double rl = raLow;
+    double rh = raHigh;
+    double dl = decLow - 90.;
+    double dh = decHigh - 90.;
+    
[…]

### 2013-12-08 GRObservations/GRObservations/GRExposureMap.h

@@ -1,0 +1,31 @@
+//
+//  GRExposureMap.h
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 28.04.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#ifndef __Gamma_Rays__GRExposureMap__
+#define __Gamma_Rays__GRExposureMap__
+
+#include <iostream>
+#include <vector>
+
+#include "GRLocation.h"
+
+using namespace std;
+
+class GRExposureMap {
+public:
+    vector <float> energies;
+    vector <vector <vector <float> > > exposures; // [ra][dec]
+    
+private:
+    int findEnergy(float energy);
+public:
+    float exposure(int energyIndex, GRLocation location);
+    float exposure(float energy, GRLocation location);
+};
+
+#endif /* defined(__Gamma_Rays__GRExposureMap__) */

### 2013-12-08 GRObservations/GRObservations/GRFermiLATDataServerQuery.cpp

@@ -1,0 +1,670 @@
+//
+//  GRFermiLATDataServerQuery.cpp
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 23.04.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#include <sstream>
+#include <fstream>
+using namespace std;
+
+#include <sys/stat.h>
+#include <sys/errno.h>
+#include <unistd.h>
+
+#ifdef __APPLE__
+#include <CommonCrypto/CommonDigest.h>
+#include <CoreFoundation/CoreFoundation.h>
+#include <Security/Security.h>
+#include <Security/SecTransform.h>
+#include <Security/SecEncodeTransform.h>
+#else
+#include <openssl/sha.h>
+#include <openssl/bio.h>
+#include <openssl/evp.h>
+#endif
+
+#include <curl/curl.h>
+
+#include <fitsio.h>
+
+#include "GRFermiLATDataServerQuery.h"
+
+void GRFermiLATDataServerQuery::calculateHash() {
+    hash.clear();
+    
+    double parameters[4];
+    parameters[0] = startTime;
+    parameters[1] = endTime;
+    parameters[2] = location.ra;
+    parameters[3] = location.dec;
+    int parametersSize = 4*sizeof(double);
+    
+    int digestLength;
+    
+#ifdef __APPLE__
+    
+    digestLength = CC_SHA256_DIGEST_LENGTH;
+    CC_SHA256_CTX context;
+    unsigned char md[CC_SHA256_DIGEST_LENGTH];
+    CC_SHA256_Init(&context);
+    CC_SHA256_Update(&context, (unsigned char*)parameters, parametersSize);
+    CC_SHA256_Final(md, &context);
+    
+#else
+    
+    digestLength = SHA256_DIGEST_LENGTH;
+    SHA256_CTX context;
+    unsigned char md[SHA256_DIGEST_LENGTH];
+    SHA256_Init(&context);
+    SHA256_Update(&context, (unsigned char*)parameters, parametersSize);
+    SHA256_Final(md, &context);
+    
+#endif
+    
+    hash.reserve(digestLength*2);
+    for (int i = 0; i < digestLength; i++) {
+        unsigned char currentChar = md[i];
+        int symbols[2] = {currentChar/16, currentChar%16};
+        for (int i = 0; i < 2; i++) {
[…]

### 2013-12-08 GRObservations/GRObservations/GRFermiLATDataServerQuery.h

@@ -1,0 +1,99 @@
+//
+//  GRFermiLATDataServerQuery.h
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 23.04.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#ifndef __Gamma_Rays__GRFermiLATDataServerQuery__
+#define __Gamma_Rays__GRFermiLATDataServerQuery__
+
+#include <iostream>
+#include <vector>
+
+#include "GRFermiLATPhoton.h"
+#include "GRLocation.h"
+#include "GRPsf.h"
+#include "GRExposureMap.h"
+
+using namespace std;
+
+enum GRFermiLATDataServerQueryError {
+    GRFermiLATDataServerQueryErrorOk,
+    GRFermiLATDataServerQueryErrorNotDownloaded,
+    GRFermiLATDataServerQueryErrorNotProcessed,
+    GRFermiLATDataServerQueryErrorNotRead,
+    // 4
+    GRFermiLATDataServerQueryErrorMkdir,
+    GRFermiLATDataServerQueryErrorFileOpen,
+    GRFermiLATDataServerQueryErrorSymlink,
+    // 7
+    GRFermiLATDataServerQueryErrorCurlInit,
+    GRFermiLATDataServerQueryErrorCurlPerform,
+    // 9
+    GRFermiLATDataServerQueryErrorFermiDataServerTooEarly,
+    GRFermiLATDataServerQueryErrorFermiDataServerUnknown,
+    // 11
+    GRFermiLATDataServerQueryErrorFermiDataServerQuiryStateUnknown,
+    // 12
+    GRFermiLATDataServerQueryErrorFermiDataServerEmptyResults,
+    GRFermiLATDataServerQueryErrorFermiDataServerUnknownFile,
+    // 14
+    GRFermiLATDataServerQueryErrorNoEventListFile,
+    GRFermiLATDataServerQueryErrorNoFilteredFile,
+    GRFermiLATDataServerQueryErrorNoSpacecraftFile,
+    GRFermiLATDataServerQueryErrorNoLtCubeFile,
+    // 18
+    GRFermiLATDataServerQueryErrorGtselectFailed,
+    GRFermiLATDataServerQueryErrorGtmktimeFailed,
+    GRFermiLATDataServerQueryErrorGtltcubeFailed,
+    GRFermiLATDataServerQueryErrorGtexpcube2Failed,
[…]

### 2013-12-08 GRObservations/GRObservations/GRFermiLATPhoton.cpp

@@ -1,0 +1,52 @@
+//
+//  GRFermiLATPhoton.cpp
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 24.03.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#include <math.h>
+
+#include <sstream>
+
+#include "GRFermiLATPhoton.h"
+
+string GRFermiLATPhoton::energyDescription() {
+    if (energy == 0.) return "0";
+    
+    ostringstream result;
+    
+    int magnitude = (int)round(log10(energy));
+    
+    if (magnitude < -3 || magnitude >= 12) result << energy * 1e6 << " eV";
+    else if (magnitude < 0) result << energy * 1e3 << " keV";
+    else if (magnitude < 3) result << energy * 1e0 << " MeV";
+    else if (magnitude < 6) result << energy * 1e-3 << " GeV";
+    else if (magnitude < 9) result << energy * 1e-6 << " TeV";
+    else result << energy * 1e-9 << " PeV";
+    
+    return result.str();
+}
+
+string GRFermiLATPhoton::conversionTypeName() {
+    if (conversionType == GRFermiConversionTypeFront) return "front";
+    else return "back";
+}
+
+string GRFermiLATPhoton::eventClassName() {
+    if (eventClass == GRFermiEventClassUltraclean) return "ultraclean";
+    else if (eventClass == GRFermiEventClassClean) return "clean";
+    else if (eventClass == GRFermiEventClassSource) return "source";
+    else return "transient";
+}
+
+bool GRFermiLATPhoton::operator<(GRFermiLATPhoton right) const {
+    return time < right.time;
+}
+
+string GRFermiLATPhoton::description() {
+    ostringstream result;
+    result << "Photon (" << energyDescription() << ", " << eventClassName() << ", " << conversionTypeName() << ") at MET " << (int)time << " from " << location.description();
+    return result.str();
+}

### 2013-12-08 GRObservations/GRObservations/GRFermiLATPhoton.h

@@ -1,0 +1,54 @@
+//
+//  GRFermiLATPhoton.h
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 24.03.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#ifndef __Gamma_Rays__GRFermiLATPhoton__
+#define __Gamma_Rays__GRFermiLATPhoton__
+
+#include <iostream>
+#include <string>
+
+#include "GRLocation.h"
+
+enum GRFermiConversionType {
+    GRFermiConversionTypeFront = 0,
+    GRFermiConversionTypeBack = 1
+    };
+
+enum GRFermiEventClass {
+    GRFermiEventClassTransient = 0,
+    GRFermiEventClassSource = 1,
+    GRFermiEventClassClean = 2,
+    GRFermiEventClassUltraclean = 3
+    };
+
+const int GRFermiConversionTypesCount = 2;
+const int GRFermiEventClassesCount = 4;
+
+const GRFermiConversionType GRFermiConversionTypes[] = {GRFermiConversionTypeFront, GRFermiConversionTypeBack};
+const GRFermiEventClass GRFermiEventClasses[] = {GRFermiEventClassTransient, GRFermiEventClassSource, GRFermiEventClassClean, GRFermiEventClassUltraclean};
+
+class GRFermiLATPhoton {
+public:
+    GRLocation location;
+    float energy;
+    double time;
+    GRFermiConversionType conversionType;
+    GRFermiEventClass eventClass;
+    
+private:
+    string conversionTypeName();
+    string eventClassName();
+    string energyDescription();
+    
+public:
+    GRFermiLATPhoton(double time, GRLocation location, float energy, GRFermiConversionType conversionType, GRFermiEventClass fermiEventClass) : time(time), location(location), energy(energy), conversionType(conversionType), eventClass(fermiEventClass) {};
+    bool operator<(GRFermiLATPhoton right) const;
+    string description();
+};
+
+#endif /* defined(__Gamma_Rays__GRFermiLATPhoton__) */

### 2013-12-08 GRObservations/GRObservations/GRLocation.cpp

@@ -1,0 +1,68 @@
+//
+//  GRCelestialSpherePoint.cpp
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 24.03.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#include <math.h>
+
+#include <sstream>
+
+#include "GRLocation.h"
+
+GRLocation::GRLocation(GRCoordinateSystem system, float angle0, float angle1, float error) : error (error){
+    if (system == GRCoordinateSystemJ2000) {
+        ra = angle0;
+        dec = angle1;
+    } else if (system ==GRCoordinateSystemGalactic) {
+        double l = angle0 * M_PI / 180.;
+        double b = angle1 * M_PI / 180.;
+        
+        double pole_ra = 192.859508 * M_PI / 180.;
+        double pole_dec = 27.128336 * M_PI / 180.;
+        double posangle = (122.932-90.0) * M_PI / 180.;
+        
+        double ra_rad = atan2( (cos(b)*cos(l-posangle)), (sin(b)*cos(pole_dec) - cos(b)*sin(pole_dec)*sin(l-posangle)) ) + pole_ra;
+        double dec_rad = asin( cos(b)*cos(pole_dec)*sin(l-posangle) + sin(b)*sin(pole_dec) );
+        
+        ra = (float)(ra_rad * 180. / M_PI);
+        dec = (float)(dec_rad * 180. / M_PI);
+    }
+}
+
+GRLocation::GRLocation(GRCoordinateSystem system, float angle0, float angle1) {
+    *this = GRLocation(system, angle0, angle1, error);
+}
+
+GRLocation::GRLocation() {
+    ra = 0.;
+    dec = 0.;
+    error = 0.;
+}
+
+float GRLocation::separation(GRLocation location) {
+    double ra1Rad = ra * M_PI / 180.;
+    double dec1Rad = dec * M_PI / 180.;
+    double ra2Rad = location.ra * M_PI / 180.;
+    double dec2Rad = location.dec * M_PI / 180.;
+    double raDiff = ra1Rad - ra2Rad;
+    if (raDiff < 0) raDiff = -raDiff;
+    double decDiff = dec1Rad - dec2Rad;
+    if (decDiff < 0) decDiff = -decDiff;
+    
[…]

### 2013-12-08 GRObservations/GRObservations/GRLocation.h

@@ -1,0 +1,39 @@
+//
+//  GRCelestialSpherePoint.h
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 24.03.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#ifndef __Gamma_Rays__GRCelestialSpherePoint__
+#define __Gamma_Rays__GRCelestialSpherePoint__
+
+#include <iostream>
+#include <string>
+#include <vector>
+
+using namespace std;
+
+enum GRCoordinateSystem {
+    GRCoordinateSystemJ2000 = 0,
+    GRCoordinateSystemGalactic = 1
+    };
+
+class GRLocation {
+public:
+    float ra;
+    float dec;
+    float error;
+    
+public:
+    GRLocation(GRCoordinateSystem system, float ra, float dec, float error);
+    GRLocation(GRCoordinateSystem system, float ra, float dec);
+    GRLocation();
+    double operator==(GRLocation location);
+    float separation(GRLocation location);
+    bool isSeparated(GRLocation location);
+    string description();
+};
+
+#endif /* defined(__Gamma_Rays__GRCelestialSpherePoint__) */

### 2013-12-08 GRObservations/GRObservations/GRPsf.cpp

@@ -1,0 +1,112 @@
+//
+//  GRPsf.cpp
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 31.03.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#include <math.h>
+
+#include <memory>
+#include <sstream>
+
+#include <fitsio.h>
+
+#include "GRPsf.h"
+
+using namespace std;
+
+int GRPsf::findEnergy(float energy) {
+    if (energy < energies[0]) return -1;
+    else {
+        int lowerEnergy = 0;
+        int upperEnergy = (int)energies.size()-1;
+        while (upperEnergy - lowerEnergy > 1) {
+            if (energies[(upperEnergy + lowerEnergy)/2] > energy) {
+                upperEnergy = (upperEnergy + lowerEnergy)/2;
+            } else {
+                lowerEnergy = (upperEnergy + lowerEnergy)/2;
+            }
+        }
+        return lowerEnergy;
+    }
+}
+
+float GRPsf::spread(int energyIndex, float probability) {
+    double currentProbability = 0.;
+    double previousProbability = 0.;
+    int i;
+    for (i = 0; i < angles.size()-1 && currentProbability < probability; i++) {
+        previousProbability = currentProbability;
+        currentProbability += (probabilityDensity[energyIndex][i] + probabilityDensity[energyIndex][i+1]) * M_PI * (cos(angles[i] * M_PI / 180.) - cos(angles[i+1] * M_PI / 180.));
+    }
+    if (i == angles.size()-1 && currentProbability < probability) {
+        i++;
+        previousProbability = currentProbability;
+        currentProbability = 1.;
+    }
+    
+    double lowerAngle = angles[i-1];
+    double upperAngle = (i == angles.size() ? 180. : angles[i]);
+    double lowerProbability = previousProbability;
+    double upperProbability = currentProbability;
+    
+    if (lowerProbability == upperProbability) return lowerAngle;
+    else return lowerAngle + (upperAngle - lowerAngle) / (upperProbability - lowerProbability) * (probability - lowerProbability);
+}
+
+float GRPsf::spread(float energy, float probability) {
+    int lowerIndex = findEnergy(energy);
+    int upperIndex;
+    if (lowerIndex == -1) {
+        lowerIndex = 0;
+        upperIndex = 1;
+    } else if (lowerIndex == (int)energies.size()-1) {
+        lowerIndex = (int)energies.size()-2;
[…]

### 2013-12-08 GRObservations/GRObservations/GRPsf.h

@@ -1,0 +1,34 @@
+//
+//  GRPsf.h
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 31.03.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#ifndef __Gamma_Rays__GRPsf__
+#define __Gamma_Rays__GRPsf__
+
+#include <iostream>
+#include <fstream>
+#include <vector>
+#include <string>
+
+using namespace std;
+
+class GRPsf {
+public:
+    vector <float> energies;
+    vector <float> angles;
+    vector <vector <float> > probabilityDensity;
+    
+private:
+    int findEnergy(float energy);
+public:
+    float spread(float energy, float probability);
+    float spread(int energyIndex, float probability);
+    ostream& writeSpreads(float probability, ostream &stream);
+    string description();
+};
+
+#endif /* defined(__Gamma_Rays__GRPsf__) */

### 2013-12-08 GRObservations/GRObservations/main.cpp

@@ -1,0 +1,123 @@
+//
+//  main.cpp
+//  GRObservations
+//
+//  Created by Maxim Piskunov on 15.04.2013.
+//  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
+//
+
+#include <math.h>
+#include <string>
+#include <iostream>
+
+#include "GRBurst.h"
+
+using namespace std;
+
+double probability(double sigma) {
+    return 1-erf((double)sigma/sqrt(2.));
+}
+
+double sigma(double prob) {
+    if (prob == 0.) return INFINITY;
+    if (prob == 1.) return 0.;
+    
+    double sigmaLow = 0.;
+    double sigmaHigh = 1.;
+    while (probability(sigmaHigh) > prob) sigmaHigh *= 2.;
+    
+    while (sigmaHigh - sigmaLow > 0.01) {
+        if (probability((sigmaLow + sigmaHigh)/2.) > prob) sigmaLow = (sigmaLow + sigmaHigh)/2.;
+        else sigmaHigh = (sigmaLow + sigmaHigh)/2.;
+    }
+    
+    return (sigmaLow + sigmaHigh) / 2.;
+}
+
+int main(int argc, const char * argv[])
+{
+    ofstream log("log");
+    ofstream interestingLog("interestingLog");
+    
+    vector <GRBurst> burstCatalog;
+    
+    if (argc != 2) {
+        cout << "Usage: GRObservations burstsCatalogLocation" << endl;
+        return 1;
+    }
+    
+    ifstream burstsFile(argv[1]);
+    while (!burstsFile.eof()) {
+        string name;
+        double time;
+        float ra;
+        float dec;
+        float error;
+        double startOffset;
+        double endOffset;
+        
+        burstsFile >> name >> time >> ra >> dec >> error >> startOffset >> endOffset;
+        
+        if (name == "") continue;
+        
+        if ((endOffset - startOffset) != 0) {
+            GRBurst burst;
+            burst.name = name;
+            burst.time = time;
+            burst.location.ra = ra;
+            burst.location.dec = dec;
+            burst.location.error = error;
+            burst.startOffset = startOffset;
+            burst.endOffset = endOffset;
+            burstCatalog.push_back(burst);
+            cout << name << " " << time << " " << GRLocation(GRCoordinateSystemJ2000, ra, dec, error).description() << endl;
+        }
+    }
+    
+    for (int i = 0; i < burstCatalog.size(); i++) {
[…]

### 2013-12-09 Paper/gammaRays.tex
Drafted 1/3 of introduction

@@
 
-Gamma ray bursts are the most energetic explosions in the Universe, therefore they are of considerable interest for particle physics.
-Recently developed instruments, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{SwiftReference} allowed for extensive study of gamma ray bursts \cite{Vianello:2013ela, Gehrels:2013xd}.
-Many discoveries were made: for example, it was found that the energy emitted by GRBs in gamma rays remains relatively constant for different bursts \cite{Bloom:2003wy}.
-Temporal variations of GRB spectra were studied.
-The spectral lags were observed between different low-energy bands \cite{Yi:2005ht}, as well as between low-energy and high energy radiation \cite{}.
-However, recent observations of luminous bursts by Fermi allowed to study temporal variations even in high energy spectra.
-In this study we used KS-test to search for spectral lags between two high energy bands: $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.
+Gamma Ray Bursts (GRBs) are among the most energetic events in the Universe, therefore they might provide new knowledge for particle physics.
+Recent observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} allowed for extensive study of these explosions \cite{Vianello:2013ela,Gehrels:2013xd}.
+These studies led to several interesting results.
+For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}.
+Temporal variations of spectra were also studied.
+The spectral lags were found between different low energy bands \cite{Yi:2005ht}.
+The very high energy radiation was discovered to be extended if compared to x-ray emission \cite{Vianello:2013ela,Lange:2013uh}.
+Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.
+In this paper we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
 

### 2013-12-12 Paper/gammaRays.tex
Drafted 2/3 of introduction
The 2nd part was about observations

@@
 \usepackage{units}
+\usepackage{hyperref}
 
@@
 The very high energy radiation was discovered to be extended if compared to x-ray emission \cite{Vianello:2013ela,Lange:2013uh}.
-Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.
+Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we'll call them low and high energy bands).
 In this paper we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
+
+We take the GRB list from the Fermi-LAT catalog \cite{Ackermann:2013zfa}.
+Tables 2 and 4 in \cite{Ackermann:2013zfa} provide the time, durations and locations of the bursts, which we use to download the events data from the LAT Data Server\footnote{\url{http://fermi.gsfc.nasa.gov/cgi-bin/ssc/LAT/LATDataQuery.cgi}}.
+We also download events during a long time interval before the burst, and use the technique introduced in \cite{Rubtsov:2011qq} to estimate background radiation in both energy bands.
+Most of the bursts in the catalog, however, do not have enough events with over $\unit[1]{GeV}$ energies to do desired computations, so we choose only those bursts for study, for which at least $10$ photons were detected in the high energy band.
+This leaves $3$ of them: 080916C \cite{Tajima:2009az}, 090902B \cite{Abdo:2009pg} and 090926A \cite{Bregeon:2011bu}.
+
+For these 3 bursts we compute both high and low energy CDFs.
+We subtract the low energy background estimate from the low energy CDF, and add the high energy background estimate to it.
+This makes the low energy CDF non-monotonous and, rigorously speaking, we cannot use the 2-sample Kolmogorov-Smirnov test on it.
+However, since there are much more high energy events than there are low energy ones, we can think of the low energy CDF as continuous.
+Therefore, this non-monotonicity will be negligible, and will not harm the KS-test results very much.
+
+Finally, we stretch the high energy CDF by different factors, and compare it to the low energy CDF using the KS-test.
+If we require $2\sigma$-significant probability to exclude a particular stretching factor, then the stretching factor of $1$ (which means no stretching) is allowed for GRBs 080916C and 090902B.
+However, for 090926A, the high energy light curve should be stretched with respect to the low energy one.
 
@@
 This in turn would suppress emitted flux, and this is not observed.
-From this argument the lower bound of $\gamma \ge 100$ was derived \cite{something}.
+From this argument the lower bound of $\gamma \ge 100$ was derived \cite{}.
 The limited total energy of the burst provides an upper bound: $\gamma \le \gamma_\text{max}$ (see section 3.6).
@@
 
-\bibliographystyle{alpha}
-\bibliography{/Users/maxitg/Documents/bibliography.bib}
+\bibliographystyle{hep}
+\bibliography{gammaRays.bib}
 

### 2013-12-12 GRObservations/GRObservations/GRBurst.cpp

@@ -50,7 +50,7 @@ void GRBurst::init() {
 
 void GRBurst::download() {
     if (query.error == GRFermiLATDataServerQueryErrorNotDownloaded) query.download();
-    if (query.error != GRFermiLATDataServerQueryErrorNotProcessed && query.error != GRFermiLATDataServerQueryErrorNotRead && !query.error == GRFermiLATDataServerQueryErrorOk) {
+    if (query.error != GRFermiLATDataServerQueryErrorNotProcessed && query.error != GRFermiLATDataServerQueryErrorNotRead && query.error != GRFermiLATDataServerQueryErrorOk) {
         error = GRBurstErrorQueryError;
         errorDescription = query.errorDescription;
         return;

### 2013-12-12 Paper/gammaRays.tex
Added affiliations and model overview graphics

@@
 \usepackage{commath}
+\usepackage{graphicx}
 \usepackage{fullpage}
-\usepackage{graphicx}
 \usepackage{subcaption}
@@
 
-\author{Maxim S. Piskunov and Grigory I. Rubtsov}
 \title{Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model}
+\author{
+	Maxim S. Piskunov$^{1}$\thanks{maxit@ms2.inr.ac.ru} and
+	Grigory I. Rubtsov$^{1}$\thanks{grisha@ms2.inr.ac.ru} \\
+	$^{1}$Institute for Nuclear Research RAS, 117312, Moscow, Russia\\
+}
 
@@
 Recent observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} allowed for extensive study of these explosions \cite{Vianello:2013ela,Gehrels:2013xd}.
-These studies led to several interesting results.
+These and previous studies led to several interesting results.
 For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}.
@@
 However, for 090926A, the high energy light curve should be stretched with respect to the low energy one.
+
+We propose that this result can be explained by the curvature effects (that is the effects of the jet geometry).
+This effects were explored by multiple authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.
+However, these studies were only concerned with x-ray radiation, and, more importantly, they assumed that the distribution of radiation sources is homogeneous throughout the jet.
+On the contrary, we propose a different idea: that the highest energy radiators are concentrated near the axis of the burst, that is the burst opening angle depends on energy.
+
+\begin{figure}[modelOverview]
+        \centering
+        \includegraphics[width=\textwidth]{modelOverview}
+        \caption{Model Overview}
+        \label{fig:modelOverview}
+\end{figure}
 

### 2013-12-22 PosterRussian/GRPosterRussian.tex
Added poster in russian

\documentclass[russian]{beamer}

\usetheme{confposter}

\usepackage[size=a0,orientation=portrait,scale=1.6]{beamerposter}
\usepackage{units}
\usepackage{float}
\usepackage{subfig}
\usepackage{fontspec}
\usepackage{picture}
\usepackage{eso-pic}
\defaultfontfeatures{Ligatures=TeX}

\setmainfont{Times New Roman Cyr}
\setsansfont{Times New Roman Cyr}

\title{Временное растяжение ГэВ излучения гамма-всплесков:\\[0.2ex]Fermi LAT и геометическая модель}
\author{
	{\Large Максим Пискунов$^{*}$ и Григорий Рубцов} \\[0.5ex]
	{\large Институт ядерных исследований РАН} \\[0.4ex]
	{\large МГУ им. М.В. Ломоносова} \\[0.6ex]
	{\normalsize $^{*}$\href{mailto:maxit@ms2.inr.ac.ru}{maxit@ms2.inr.ac.ru}} \\
	\hfill {\normalsize \url{https://github.com/maxitg/GammaRays}}
}

\begin{document}

	\begin{frame}
		\begin{columns}[t]

			\begin{column}[t]{0.4\linewidth}

			   	\begin{block}{Введение}
					Наблюдения показывают, что излучение гамма-всплесков с энергией выше \unit[100]{МэВ} систематически наблюдается позже, чем низкоэнергичное излучение. Различия же кривых блеска в различных диапазонах высокоэнергичного (> \unit[100]{МэВ}) излучения изучены хуже. В данной работе мы изучаем различия кривых блеска в диапазонах (\unit[100]{МэВ}, \unit[1]{ГэВ}) и (\unit[1]{ГэВ}, \unit[300]{ГэВ}).
			   	\end{block}

			   	\begin{block}{Наблюдения}

					\begin{figure}

						\centering
						{\large Сравнение кривых блеска GRB 090926A (тест Колмогорова-Смирнова)}
						\subfloat{\includegraphics[width=\textwidth]{090926A.pdf}} \\
						\subfloat{\includegraphics[width=\textwidth]{090926AStretched.pdf}} \\
						\label{Figures}

					\end{figure}

					\begin{itemize}
						\item{Рассмотрены 3 гамма-всплеска: 080916C, 090902B и 090926A.}
						\item{Кривые блеска 080916C и 090902B в рассматриваемых диапазонах совпадают в пределах $2\sigma$.}
						\item{Излучение 090926A в диапазоне (\unit[1]{ГэВ}, \unit[300]{ГэВ}) растянуто относительно менее энергичного со статистической значимостью $3.1\sigma$.}
					\end{itemize}

			   	\end{block}

			\end{column}

			\begin{column}[t]{0.4\linewidth}

			   	\begin{block}{Модель}

			   		\begin{figure}
						\centering
						\includegraphics[width=\textwidth]{modelExplanation.pdf}
						\label{fig:modelExplanation}
			   		\end{figure}

					Мы предлагаем простую геометрическую модель, объясняющую временное растяжение. Основное предположение -- чем выше энергия излучения, тем ближе к оси джета оно излучается.

					\begin{enumerate}
						\item{В момент времени $t = 0$, центральная машина излучает сферическую ударную волну.}
						\item{Ударная волна распространяется с ультрарелятивистской скоростью.}
						\item{Каждая точка джета -- изотропный излучатель в собственной системе отсчёта.}
						\item{
							Интенсивность -- функция пространственных координат и частоты:
							\begin{equation*}
								\eta\left(r,\theta,\omega\right) = 
								\frac{\eta_0}{1 + \left(r/r_0\right)^n}
								e^{
									-\left(\theta/\theta_0\right)^2
									\left(\omega/\omega_0\right)^{-2k}
								}
								\left(\omega/\omega_0\right)^\alpha
							\end{equation*}
						}
					\end{enumerate}

			   	\end{block}

			   	\begin{block}{Результаты}

			   		Наблюдаемое временное растяжение ГэВ излучения объясняется в рамках геометрической модели:

			   		\begin{figure}
			   			\centering
			   			\includegraphics[width=\textwidth]{kappaDistributionHistogram}
			   			\label{fig:kappaDistributionHistogram}
			   		\end{figure}

			   		Также получены (для выбранных значений параметров):

			   		\begin{itemize}
			   			\item{Полная энергия, излученная в диапазоне $\gtrsim$ \unit[1]{ГэВ}. Результат меньше предельного значения.}
			   			\item{Доля всплесков среди наблюдаемых в диапазоне (\unit[100]{МэВ}, \unit[1]{ГэВ}), которые также можно увидеть в диапазоне (1 ГэВ, 300 ГэВ) $f_m = 0.072$. Наблюдаемое значение $f_o = 0.086$.}
			   		\end{itemize}

			   	\end{block}

			\end{column}

		\end{columns}
	\end{frame}

\end{document}

### 2013-12-22 GRObservations/GRObservations/GRBurst.cpp

@@ -181,9 +181,15 @@ void GRBurst::evaluate() {
         maxLengthening[i] = 0.;
     }
     
+    double bestValue = 0., bestProbability = 0.;
     for (double value = LENGTHENING_MIN; value <= LENGTHENING_MAX; value *= LENGTHENING_STEP) {
+        double probability = gevDistribution.kolmogorovSmirnovTest(mevDistribution, value);
         lengtheningValues.push_back(value);
-        lengtheningProbabilities.push_back(gevDistribution.kolmogorovSmirnovTest(mevDistribution, value));
+        lengtheningProbabilities.push_back(probability);
+        if (probability > bestProbability) {
+            bestProbability = probability;
+            bestValue = value;
+        }
         
         for (int i = 0; i < 5; i++) {
             if (lengtheningProbabilities[lengtheningProbabilities.size()-1] < probabilityValues[i]) continue;
@@ -197,6 +203,10 @@ void GRBurst::evaluate() {
     trivialProbability = gevDistribution.kolmogorovSmirnovTest(mevDistribution, 1., true, mev, gev);
     mev.close();
     gev.close();
+    
+    mev.open((name + "/stretchedMeV").c_str());
+    gevDistribution.kolmogorovSmirnovTest(mevDistribution, bestValue, true, mev, gev);
+    mev.close();
     
     cout << endl;
 }

### 2013-12-22 GRObservations/GRObservations/GRDistribution.cpp

@@ -128,6 +128,7 @@ float GRDistribution::kolmogorovSmirnovTest(GRDistribution distribution, double 
     vector <pair<double, double> > mevPoints;
     vector <pair<double, double> > gevPoints;
     
+    double bestSelfPoint, bestDistributionPoint, bestValue;
     for (int i = 0; i < values.size(); i++) {
         bool selfInProgress = (values[i] >= start) && (values[i] <= end);
         bool distributionInProgress = (values[i] >= distribution.start) && (values[i] <= distribution.end);
@@ -144,7 +145,13 @@ float GRDistribution::kolmogorovSmirnovTest(GRDistribution distribution, double 
         
         selfPoint = (double)(i+1) / values.size();
         if (plot) gevPoints.push_back(make_pair(values[i], selfPoint));
-        maxDistance = max(maxDistance, abs(selfPoint - distributionPoint));
+        
+        if (maxDistance < abs(selfPoint - distributionPoint)) {
+            maxDistance = abs(selfPoint - distributionPoint);
+            bestSelfPoint = selfPoint;
+            bestDistributionPoint = distributionPoint;
+            bestValue = values[i];
+        }
     }
     
     for (int i = 0; i < distribution.values.size(); i++) {
@@ -167,10 +174,16 @@ float GRDistribution::kolmogorovSmirnovTest(GRDistribution distribution, double 
         
         if (plot) mevPoints.push_back(make_pair(distribution.values[i], distributionPoint));
         
-        maxDistance = max(maxDistance, abs(selfPoint - distributionPoint));
+        if (maxDistance < abs(selfPoint - distributionPoint)) {
+            maxDistance = abs(selfPoint - distributionPoint);
+            bestSelfPoint = selfPoint;
+            bestDistributionPoint = distributionPoint;
+            bestValue = distribution.values[i];
+        }
     }
     
     if (plot) {
+        cout << lengthening << ": " << bestValue << " " << bestSelfPoint << " " << bestDistributionPoint << endl;
         sort(mevPoints.begin(), mevPoints.end());
         sort(gevPoints.begin(), gevPoints.end());
         for (int i = 0; i < mevPoints.size(); i++) {

### 2013-12-22 GRObservations/GRObservations/GRFermiLATDataServerQuery.cpp

@@ -80,6 +80,7 @@ void GRFermiLATDataServerQuery::calculateHash() {
 void GRFermiLATDataServerQuery::init() {
     error = GRFermiLATDataServerQueryErrorNotDownloaded;
     calculateHash();
+    cout << this->hash << endl;
     
     ifstream status((hash + "/status").c_str());
     int statusCode = 0;
@@ -353,14 +354,14 @@ void GRFermiLATDataServerQuery::download() {
 }
 
 string GRFermiLATDataServerQuery::irfName(GRFermiEventClass eventClass, GRFermiConversionType conversionType) {
-    string irfName = "P7";
+    string irfName = "P7REP_";
     
     if (eventClass == GRFermiEventClassUltraclean) irfName += "ULTRACLEAN";
     else if (eventClass == GRFermiEventClassClean) irfName += "CLEAN";
     else if (eventClass == GRFermiEventClassSource) irfName += "SOURCE";
     else irfName += "TRANSIENT";
     
-    irfName += (string)"_V6" + (eventClass == GRFermiEventClassSource ? "MC" : "") + "::";
+    irfName += (string)"_V15::";
     
     if (conversionType == GRFermiConversionTypeBack) irfName += "BACK";
     else irfName += "FRONT";
@@ -391,7 +392,7 @@ int GRFermiLATDataServerQuery::gtmktime() {
     ostringstream cmd;
     cmd << fixed << "gtmktime" << " ";
     cmd << "scfile=" << hash << "/spacecraft.fits" << " ";
-    cmd << "filter=" << "\"DATA_QUAL==1 && LAT_CONFIG==1\"" << " ";
+    cmd << "filter=" << "\"DATA_QUAL>0 && LAT_CONFIG==1 && ABS(ROCK_ANGLE)<52\"" << " ";
     cmd << "roicut=" << "no" << " ";
     cmd << "evfile=" << hash << "/filtered.fits" << " ";
     cmd << "outfile=" << hash << "/timed.fits" << " ";

### 2013-12-22 GRObservations/GRObservations/main.cpp

@@ -92,7 +92,7 @@ int main(int argc, const char * argv[])
             else return -1;
         }
         
-        log << burstCatalog[i].name << " " << (burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.estimatedLinearComponent) << endl;
+        log << burstCatalog[i].name << " " << (burstCatalog[i].mevDistribution.values.size() - burstCatalog[i].mevDistribution.estimatedLinearComponent) << " " << (burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.estimatedLinearComponent) << endl;
         
         if ((burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.estimatedLinearComponent) < 10) {
             burstCatalog[i].clear();

### 2014-01-17 Paper/gammaRays.tex
Added better model overview
Drafted arguments for the main assumption of the model

@@
 For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}.
+This suggests that most of the bursts must have similar energetics in their rest frame.
 Temporal variations of spectra were also studied.
@@
 We take the GRB list from the Fermi-LAT catalog \cite{Ackermann:2013zfa}.
-Tables 2 and 4 in \cite{Ackermann:2013zfa} provide the time, durations and locations of the bursts, which we use to download the events data from the LAT Data Server\footnote{\url{http://fermi.gsfc.nasa.gov/cgi-bin/ssc/LAT/LATDataQuery.cgi}}.
+Tables 2 and 4 in \cite{Ackermann:2013zfa} provide the time, durations and locations of the bursts, which we use to download events data from the LAT Data Server\footnote{\url{http://fermi.gsfc.nasa.gov/cgi-bin/ssc/LAT/LATDataQuery.cgi}}.
 We also download events during a long time interval before the burst, and use the technique introduced in \cite{Rubtsov:2011qq} to estimate background radiation in both energy bands.
@@
 
-For these 3 bursts we compute both high and low energy CDFs.
+For these 3 bursts we compute both high and low energy distributions of photon times.
 We subtract the low energy background estimate from the low energy CDF, and add the high energy background estimate to it.
 This makes the low energy CDF non-monotonous and, rigorously speaking, we cannot use the 2-sample Kolmogorov-Smirnov test on it.
-However, since there are much more high energy events than there are low energy ones, we can think of the low energy CDF as continuous.
-Therefore, this non-monotonicity will be negligible, and will not harm the KS-test results very much.
+However, since there is a much greater number of photons on lower energies, we can think of the low energy CDF as continuous.
+Therefore, this non-monotonicity will be negligible, and will not harm the KS-test results much.
 
@@
 We propose that this result can be explained by the curvature effects (that is the effects of the jet geometry).
-This effects were explored by multiple authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.
+These effects were explored by multiple authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.
 However, these studies were only concerned with x-ray radiation, and, more importantly, they assumed that the distribution of radiation sources is homogeneous throughout the jet.
-On the contrary, we propose a different idea: that the highest energy radiators are concentrated near the axis of the burst, that is the burst opening angle depends on energy.
+We propose a contrary idea: that the highest energy radiators are concentrated near the axis of the burst, that is the burst opening angle depends on energy.
+We cannot prove this assumption rigorously, but we have arguments supporting it.
+
+First of all, there are around 750 GRBs detected by GBM, half of which were in the LAT field of view \cite{Vianello:2013ela}.
+However, only about 30 of them were detected by the LAT, and only 3 of them were bright in the high energy band.
+If we extrapolate the uniform jet model to very high energies, it means that these groups of bursts are internally different: some of them produce VHE radiation, and other do not.
+These differences are hard to explain given that burst energetics are similar \cite{Bloom:2003wy}.
+Nevertheless, these differences in burst counts can easily be explained by our model.
+In our model, the opening angle of low energy jet is much larger than the opening angle of a high energy jet.
+Therefore the most common scenario for an observable jet is that the off-axis angle of the observer is smaller than low energy jet opening angle, but larger than the opening angle of a high energy jet.
+Because of that, most of the bursts are only seen at low energies.
+The 3 bursts we study in this paper were seen from the lowest off-axis angles.
+
+Secondly, consider the jet plasma right after emission from the central engine. There, two processes happen simultaneously:
+\begin{enumerate}
+	\item{Plasma particles near the jet boundary change directions, therefore increasing the jet opening angle.}
+	\item{Plasma particles lose energy, therefore decreasing radiation frequency emitted from the jet.}
+\end{enumerate}
+We argue that these two processes are correlated, because they happen due to the same particle interaction processes.
+And since they are correlated, jets with large opening angles will have lower energies, which is our main assumption.
 
@@
         \centering
-        \includegraphics[width=\textwidth]{modelOverview}
-        \caption{Model Overview}
+        \includegraphics[width=0.7\textwidth]{modelOverview}
+        \caption{
+ %       	Model Overview.
+Since the total path (first as a plasma particle, then as a photon) is longer for high energies particles, the high energy emission is stretched compared to low energy emission.
+ 			Here the red cone represents the propagation volume of low energy plasma, while the blue cone stands for high energy radiators.
+The green disk is the observer.
+$\theta_2$ and $\theta_1$ are opening angles of high and low energy jets, $\chi$ is an observer's off-axis angle.
+Every ``particle'' emitted from the central engine should go through two stages before being observer: traveling in plasma to the radiation point, and then traveling to the observer as photons, If $\chi < \theta_1$, then due to the relativistic beaming effect, the majority of observed photons is radiated near the point $R_1$, which lies on the line from the central engine to the observer.
+However, it might be different for the high energy radiation.
+If $\theta_2 < \chi$, then there is no high energy plasma near $R_1$, and ``particles'' will have to move near the jet axis first, and then turn at emission to the observer's direction.
+For each moment of emission, this path is longer, therefore the high energy emission will be stretched with respect to the low energies.
+        }
         \label{fig:modelOverview}
@@
 
+Finally, this assumption is able to explain the non-trivial stretching factor of the GRB 090926A.
+The detailed computations will be described in the following sections, however, you can see why the stretching is happening on the figure \ref{fig:modelOverview}.
+
 \section{Observations}
+
+\subsection{Download and Process}
+\subsection{Linear Component (background)}
+\subsection{KS-test}
+\subsection{Results}
 

### 2014-01-17 GRObservations/GRObservations/GRBurst.cpp

@@ -22,25 +22,42 @@
 #define LENGTHENING_MAX       10.
 #define LENGTHENING_STEP      1.001
 
-void GRBurst::init() {    
-    double duration = (endOffset - startOffset) * TIME_EXTENTION_FACTOR;
-    double center = time + (startOffset + endOffset) / 2.;
-    double startTime = center - duration/2.;
-    double endTime = center + duration/2.;
-        
-    query.startTime = startTime;
-    query.endTime = endTime;
+#define BACKGROUND_DURATION   86400
+
+double GRBurst::middleTime() {
+    return time + (startOffset + endOffset) / 2.;
+}
+
+double GRBurst::duration() {
+    return (endOffset - startOffset) * TIME_EXTENTION_FACTOR;
+}
+
+double GRBurst::startTime() {
+    return middleTime() - duration()/2.;
+}
+
+double GRBurst::endTime() {
+    return middleTime() + duration()/2.;
+}
+
+double GRBurst::backgroundStartTime() {
+    return startTime() - BACKGROUND_DURATION;
+}
+
+double GRBurst::backgroundEndTime() {
+    return startTime();
+}
+
+void GRBurst::init() {
+    
+    query.startTime = startTime();
+    query.endTime = endTime();
     query.location = location;
         
     query.init();
     
-    //duration = duration * 100;
-    duration = 86400;
-    endTime = startTime;
-    startTime = startTime - duration;
-    
-    backgroundQuery.startTime = startTime;
-    backgroundQuery.endTime = endTime;
+    backgroundQuery.startTime = backgroundStartTime();
+    backgroundQuery.endTime = backgroundEndTime();
     backgroundQuery.location = location;
     
     backgroundQuery.init();
@@ -133,7 +150,8 @@ void GRBurst::read() {
     for (int i = 0; i < query.events.size(); i++) {
         GRFermiLATPhoton photon = query.events[i];
         photon.location.error = query.psfs[photon.eventClass][photon.conversionType].spread(photon.energy, PHOTONS_QUALITY);
-        if (photon.eventClass == GRFermiEventClassTransient) continue;
+        //if (photon.eventClass == GRFermiEventClassTransient) continue;
+        if (photon.time < startTime() || photon.time > endTime()) continue;
         if (location.isSeparated(photon.location)) continue;
         
[…]

### 2014-01-17 GRObservations/GRObservations/GRBurst.h

@@ -44,6 +44,8 @@ public:
     
     double startOffset;
     double endOffset;
+    
+    double redshift;
         
 public:
     GRFermiLATDataServerQuery backgroundQuery;
@@ -67,9 +69,22 @@ public:
     
     double trivialProbability;
     
+    string topEnergyDescription (int n);
+    
+    double flux(double minEnergy, double maxEnergy);
+    double backgroundFlux(double minEnergy, double maxEnergy);
+    int photonCount(double minEnergy, double maxEnergy);
+    int backgroundPhotonCount(double minEnergy, double maxEnergy);
+    
 private:
     void calculateBackground();
-        
+    double middleTime();
+    double duration();
+    double startTime();
+    double endTime();
+    double backgroundStartTime();
+    double backgroundEndTime();
+            
 public:
     void init();
     void download();

### 2014-01-17 GRObservations/GRObservations/GRFermiLATPhoton.h

@@ -43,11 +43,11 @@ public:
 private:
     string conversionTypeName();
     string eventClassName();
-    string energyDescription();
     
 public:
     GRFermiLATPhoton(double time, GRLocation location, float energy, GRFermiConversionType conversionType, GRFermiEventClass fermiEventClass) : time(time), location(location), energy(energy), conversionType(conversionType), eventClass(fermiEventClass) {};
     bool operator<(GRFermiLATPhoton right) const;
+    string energyDescription();
     string description();
 };
 

### 2014-01-17 GRObservations/GRObservations/main.cpp

@@ -9,6 +9,7 @@
 #include <math.h>
 #include <string>
 #include <iostream>
+#include <sstream>
 
 #include "GRBurst.h"
 
@@ -55,8 +56,9 @@ int main(int argc, const char * argv[])
         float error;
         double startOffset;
         double endOffset;
+        double redshift;
         
-        burstsFile >> name >> time >> ra >> dec >> error >> startOffset >> endOffset;
+        burstsFile >> name >> time >> ra >> dec >> error >> startOffset >> endOffset >> redshift;
         
         if (name == "") continue;
         
@@ -69,8 +71,9 @@ int main(int argc, const char * argv[])
             burst.location.error = error;
             burst.startOffset = startOffset;
             burst.endOffset = endOffset;
+            burst.redshift = redshift;
             burstCatalog.push_back(burst);
-            cout << name << " " << time << " " << GRLocation(GRCoordinateSystemJ2000, ra, dec, error).description() << endl;
+            cout << name << " " << time << " " << GRLocation(GRCoordinateSystemJ2000, ra, dec, error).description() << " z = " << redshift << endl;
         }
     }
     
@@ -86,7 +89,6 @@ int main(int argc, const char * argv[])
         if (burstCatalog[i].error != GRBurstErrorOk) {
             cout << "error " << burstCatalog[i].error << " " << burstCatalog[i].query.error << " " << burstCatalog[i].backgroundQuery.error << " : " << burstCatalog[i].errorDescription << endl;
             if (burstCatalog[i].query.error == 19) {
-                burstCatalog[i].clear();
                 continue;
             }
             else return -1;
@@ -95,7 +97,6 @@ int main(int argc, const char * argv[])
         log << burstCatalog[i].name << " " << (burstCatalog[i].mevDistribution.values.size() - burstCatalog[i].mevDistribution.estimatedLinearComponent) << " " << (burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.estimatedLinearComponent) << endl;
         
         if ((burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.estimatedLinearComponent) < 10) {
-            burstCatalog[i].clear();
             continue;
         }
         
@@ -112,12 +113,36 @@ int main(int argc, const char * argv[])
[…]

### 2014-01-18 Paper/gammaRays.tex
Introduction draft completed

@@
         \caption{
- %       	Model Overview.
-Since the total path (first as a plasma particle, then as a photon) is longer for high energies particles, the high energy emission is stretched compared to low energy emission.
- 			Here the red cone represents the propagation volume of low energy plasma, while the blue cone stands for high energy radiators.
-The green disk is the observer.
-$\theta_2$ and $\theta_1$ are opening angles of high and low energy jets, $\chi$ is an observer's off-axis angle.
-Every ``particle'' emitted from the central engine should go through two stages before being observer: traveling in plasma to the radiation point, and then traveling to the observer as photons, If $\chi < \theta_1$, then due to the relativistic beaming effect, the majority of observed photons is radiated near the point $R_1$, which lies on the line from the central engine to the observer.
-However, it might be different for the high energy radiation.
-If $\theta_2 < \chi$, then there is no high energy plasma near $R_1$, and ``particles'' will have to move near the jet axis first, and then turn at emission to the observer's direction.
-For each moment of emission, this path is longer, therefore the high energy emission will be stretched with respect to the low energies.
+        	Model Overview.
+Here the red and the blue cones represent volumes through which low and high energy plasma propagates.
+In the case depicted the observer's off-axis angle is smaller than the opening angle of the low energy jet, so most of the observable low energy photons will travel along the straight line from the central engine (due to the relativistic beaming effect).
+In the same time, the observer's off-axis angle is larger that the opening angle of the high energy jet, so high energy radiation will still originate near the center of the jet (because it is the only place where there are high energy radiators).
+The observation time of a photon is a sum of two things: the time interval spent in plasma as a radiator (which is approximately a distance from the central engine to the point of emission); and the time interval from emission to detection (which is the distance from the point of emission to the observer's location).
+Given a position of the shock front, this sum is larger for high energy photons.
+Because of that, high energy emission will be observed later throughout the burst duration, therefore the high energy light curve will be stretched.
         }
@@
 
+\begin{figure}[sampleLightCurve]
+	\centering
+	\includegraphics[width=0.49\textwidth]{sampleLightCurve}
+	\caption{
+		High and low energy light curves produced by the geometrical model.
+Parameter values are the same as discussed in the section \ref{sec:parameters}.
+Cosmological redshift of the burst $z = 1$.
+Observer's off-axis angle $\chi = 0.01$.
+	}
+	\label{fig:sampleLightCurve}
+\end{figure}
+
 Finally, this assumption is able to explain the non-trivial stretching factor of the GRB 090926A.
-The detailed computations will be described in the following sections, however, you can see why the stretching is happening on the figure \ref{fig:modelOverview}.
+The detailed computations will be described in the following sections, but you can see why the stretching is happening on the figure \ref{fig:modelOverview}.
+The computations confirm that this qualitative picture is indeed correct, you can see a sample light curve on the figure \ref{fig:sampleLightCurve}, and predicted distribution of stretching factors on the figure \ref{fig:kappaDistribution}.
 
@@
 \subsection{Parameter values}
+\label{sec:parameters}
 

### 2014-01-21 Paper/gammaRays.tex
Cleared main.cpp

@@
 Gamma Ray Bursts (GRBs) are among the most energetic events in the Universe, therefore they might provide new knowledge for particle physics.
-Recent observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} allowed for extensive study of these explosions \cite{Vianello:2013ela,Gehrels:2013xd}.
+Modern observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} made possible to study these explosions extensively \cite{Vianello:2013ela,Gehrels:2013xd}.
 These and previous studies led to several interesting results.
@@
 \subsection{Download and Process}
+
+Download in progress.
+
 \subsection{Linear Component (background)}

### 2014-01-21 GRObservations/GRObservations/main.cpp

@@ -38,7 +38,6 @@ double sigma(double prob) {
 int main(int argc, const char * argv[])
 {
     ofstream log("log");
-    ofstream interestingLog("interestingLog");
     
     vector <GRBurst> burstCatalog;
     
@@ -56,11 +55,11 @@ int main(int argc, const char * argv[])
         float error;
         double startOffset;
         double endOffset;
-        double redshift;
         
-        burstsFile >> name >> time >> ra >> dec >> error >> startOffset >> endOffset >> redshift;
+        burstsFile >> name >> time >> ra >> dec >> error >> startOffset >> endOffset;
         
         if (name == "") continue;
+        if (name[0] == '#') continue;
         
         if ((endOffset - startOffset) != 0) {
             GRBurst burst;
@@ -71,9 +70,8 @@ int main(int argc, const char * argv[])
             burst.location.error = error;
             burst.startOffset = startOffset;
             burst.endOffset = endOffset;
-            burst.redshift = redshift;
             burstCatalog.push_back(burst);
-            cout << name << " " << time << " " << GRLocation(GRCoordinateSystemJ2000, ra, dec, error).description() << " z = " << redshift << endl;
+            cout << name << " " << time << " " << GRLocation(GRCoordinateSystemJ2000, ra, dec, error).description() << endl;
         }
     }
     
@@ -94,19 +92,23 @@ int main(int argc, const char * argv[])
             else return -1;
         }
         
-        log << burstCatalog[i].name << " " << (burstCatalog[i].mevDistribution.values.size() - burstCatalog[i].mevDistribution.estimatedLinearComponent) << " " << (burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.estimatedLinearComponent) << endl;
+        log << burstCatalog[i].name << ":" << endl;
+        log << "\tLow energy count  = " << burstCatalog[i].mevDistribution.values.size() - burstCatalog[i].mevDistribution.estimatedLinearComponent << endl;
+        log << "\tHigh energy count = " << burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.estimatedLinearComponent << endl;
+        
         
         if ((burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.estimatedLinearComponent) < 10) {
[…]

### 2014-01-21 GRObservations/GRObservations/GRBurst.cpp
Clear GRBurst.cpp and GRBurst.h

@@ -24,40 +24,19 @@
 
 #define BACKGROUND_DURATION   86400
 
-double GRBurst::middleTime() {
-    return time + (startOffset + endOffset) / 2.;
-}
-
-double GRBurst::duration() {
-    return (endOffset - startOffset) * TIME_EXTENTION_FACTOR;
-}
-
-double GRBurst::startTime() {
-    return middleTime() - duration()/2.;
-}
-
-double GRBurst::endTime() {
-    return middleTime() + duration()/2.;
-}
-
-double GRBurst::backgroundStartTime() {
-    return startTime() - BACKGROUND_DURATION;
-}
-
-double GRBurst::backgroundEndTime() {
-    return startTime();
-}
-
 void GRBurst::init() {
     
-    query.startTime = startTime();
-    query.endTime = endTime();
+    double middle = time + (startOffset + endOffset) / 2.;
+    double duration = (endOffset - startOffset) * TIME_EXTENTION_FACTOR;
+    
+    query.startTime = middle - duration/2.;
+    query.endTime = middle + duration/2.;
     query.location = location;
         
     query.init();
     
-    backgroundQuery.startTime = backgroundStartTime();
-    backgroundQuery.endTime = backgroundEndTime();
+    backgroundQuery.startTime = query.startTime - BACKGROUND_DURATION;
+    backgroundQuery.endTime = query.startTime;
     backgroundQuery.location = location;
     
     backgroundQuery.init();
@@ -150,8 +129,6 @@ void GRBurst::read() {
     for (int i = 0; i < query.events.size(); i++) {
         GRFermiLATPhoton photon = query.events[i];
         photon.location.error = query.psfs[photon.eventClass][photon.conversionType].spread(photon.energy, PHOTONS_QUALITY);
-        //if (photon.eventClass == GRFermiEventClassTransient) continue;
-        if (photon.time < startTime() || photon.time > endTime()) continue;
         if (location.isSeparated(photon.location)) continue;
         
         if (photon.energy < 1000.) mevPhotons.push_back(photon);
@@ -163,8 +140,6 @@ void GRBurst::read() {
     for (int i = 0; i < backgroundQuery.events.size(); i++) {
         GRFermiLATPhoton photon = backgroundQuery.events[i];
[…]

### 2014-01-21 GRObservations/GRObservations/GRBurst.h

@@ -69,13 +69,6 @@ public:
     
     double trivialProbability;
     
-    string topEnergyDescription (int n);
-    
-    double flux(double minEnergy, double maxEnergy);
-    double backgroundFlux(double minEnergy, double maxEnergy);
-    int photonCount(double minEnergy, double maxEnergy);
-    int backgroundPhotonCount(double minEnergy, double maxEnergy);
-    
 private:
     void calculateBackground();
     double middleTime();
@@ -91,8 +84,6 @@ public:
     void process();
     void read();
     void evaluate();
-    
-    void clear();
 };
 
 #endif /* defined(__Gamma_Rays__GRBurst__) */

### 2014-01-21 Paper/gammaRays.tex
Revised the Introduction

@@
 The spectral lags were found between different low energy bands \cite{Yi:2005ht}.
-The very high energy radiation was discovered to be extended if compared to x-ray emission \cite{Vianello:2013ela,Lange:2013uh}.
-Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we'll call them low and high energy bands).
-In this paper we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
+The very high energy radiation was discovered to be extended relative to x-ray emission \cite{Vianello:2013ela,Lange:2013uh}.
+Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
+In this study we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
 
 We take the GRB list from the Fermi-LAT catalog \cite{Ackermann:2013zfa}.
-Tables 2 and 4 in \cite{Ackermann:2013zfa} provide the time, durations and locations of the bursts, which we use to download events data from the LAT Data Server\footnote{\url{http://fermi.gsfc.nasa.gov/cgi-bin/ssc/LAT/LATDataQuery.cgi}}.
-We also download events during a long time interval before the burst, and use the technique introduced in \cite{Rubtsov:2011qq} to estimate background radiation in both energy bands.
-Most of the bursts in the catalog, however, do not have enough events with over $\unit[1]{GeV}$ energies to do desired computations, so we choose only those bursts for study, for which at least $10$ photons were detected in the high energy band.
+Tables 2 and 4 in \cite{Ackermann:2013zfa} provide the time, durations and locations of the bursts, which we use to download observational data from the LAT Data Server\footnote{\url{http://fermi.gsfc.nasa.gov/cgi-bin/ssc/LAT/LATDataQuery.cgi}}.
+We also download events during a day before the burst, and use the technique introduced in \cite{Rubtsov:2011qq} to estimate background radiation in both energy bands.
+Most of the bursts in the catalog, however, do not have enough events with over $\unit[1]{GeV}$ energies to do desired computations, so we choose only those bursts, from which at least $10$ photons were detected in the high energy band.
 This leaves $3$ of them: 080916C \cite{Tajima:2009az}, 090902B \cite{Abdo:2009pg} and 090926A \cite{Bregeon:2011bu}.
@@
 If we require $2\sigma$-significant probability to exclude a particular stretching factor, then the stretching factor of $1$ (which means no stretching) is allowed for GRBs 080916C and 090902B.
-However, for 090926A, the high energy light curve should be stretched with respect to the low energy one.
+However, for 090926A, the stretching factor of $1$ is prohibited, so the high energy light curve should be stretched with respect to the low energy one.
 
@@
 These effects were explored by multiple authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.
-However, these studies were only concerned with x-ray radiation, and, more importantly, they assumed that the distribution of radiation sources is homogeneous throughout the jet.
-We propose a contrary idea: that the highest energy radiators are concentrated near the axis of the burst, that is the burst opening angle depends on energy.
-We cannot prove this assumption rigorously, but we have arguments supporting it.
+However, these studies were only concerned with x-ray radiation, and, even more importantly, they assumed that the distribution of radiation sources is homogeneous throughout the jet.
+We propose a contrary idea: that the highest energy radiators are concentrated near the axis of the burst, meaning that the burst opening angle depends on energy.
+We cannot prove this assumption rigorously, but have arguments supporting it.
 
-First of all, there are around 750 GRBs detected by GBM, half of which were in the LAT field of view \cite{Vianello:2013ela}.
+First of all, there are around 750 GRBs detected by GBM, half of which were in the LAT field of view at the moment of observation \cite{Vianello:2013ela}.
 However, only about 30 of them were detected by the LAT, and only 3 of them were bright in the high energy band.
-If we extrapolate the uniform jet model to very high energies, it means that these groups of bursts are internally different: some of them produce VHE radiation, and other do not.
+If we extrapolate the uniform jet model to very high energies, this observation would mean that these groups of bursts are internally different: some of them produce VHE radiation, and other do not.
 These differences are hard to explain given that burst energetics are similar \cite{Bloom:2003wy}.
 Nevertheless, these differences in burst counts can easily be explained by our model.
-In our model, the opening angle of low energy jet is much larger than the opening angle of a high energy jet.
-Therefore the most common scenario for an observable jet is that the off-axis angle of the observer is smaller than low energy jet opening angle, but larger than the opening angle of a high energy jet.
-Because of that, most of the bursts are only seen at low energies.
-The 3 bursts we study in this paper were seen from the lowest off-axis angles.
+In our model, the opening angle of a jet is inversely proportional to the energy of photons it radiates.
+Therefore the most common scenario is that the off-axis angle of the observer is smaller than the low energy jet opening angle, but larger than the opening angle of a high energy jet.
+Because of that, most of the bursts can only be seen at low energies.
+The 3 bursts we study in this paper, were seen, according to our model, from the lowest off-axis angles.
 
-Secondly, consider the jet plasma right after emission from the central engine. There, two processes happen simultaneously:
+Second, consider the plasma right after its ejection from the central engine. Two processes happen there simultaneously:
 \begin{enumerate}
-	\item{Plasma particles near the jet boundary change directions, therefore increasing the jet opening angle.}
-	\item{Plasma particles lose energy, therefore decreasing radiation frequency emitted from the jet.}
+	\item{Particles near the jet boundary change their movement directions, therefore increasing the opening angle of the jet.}
+	\item{Particles lose energy, therefore decreasing radiation frequency emitted by the jet.}
 \end{enumerate}
-We argue that these two processes are correlated, because they happen due to the same particle interaction processes.
-And since they are correlated, jets with large opening angles will have lower energies, which is our main assumption.
+We argue that these two processes are correlated, for they happen due to the same particle interaction processes.
+And since they are correlated, jets with larger opening angles should have lower energies, which is the assumption we are trying to justify.
 
@@
         	Model Overview.
-Here the red and the blue cones represent volumes through which low and high energy plasma propagates.
-In the case depicted the observer's off-axis angle is smaller than the opening angle of the low energy jet, so most of the observable low energy photons will travel along the straight line from the central engine (due to the relativistic beaming effect).
-In the same time, the observer's off-axis angle is larger that the opening angle of the high energy jet, so high energy radiation will still originate near the center of the jet (because it is the only place where there are high energy radiators).
-The observation time of a photon is a sum of two things: the time interval spent in plasma as a radiator (which is approximately a distance from the central engine to the point of emission); and the time interval from emission to detection (which is the distance from the point of emission to the observer's location).
+Here the red and the blue cones represent the volumes through which low and high energy plasma propagates.
+In case depicted the observer's off-axis angle is smaller than the opening angle of the low energy jet, so, due to the relativistic beaming effect, the most of the observable low energy photons will travel along the straight line from the central engine.
+Also, the observer's off-axis angle is larger that the opening angle of the high energy jet, so the high energy radiation will still originate near the center of the jet (because it is the only place where there are high energy radiators).
+The observation time of a photon is a sum of two things: the time interval spent in plasma as a radiator (which approximately equals to the distance from the central engine to the point of emission); and the time interval from emission to detection (which is the distance from the point of emission to the observer's location).
 Given a position of the shock front, this sum is larger for high energy photons.
@@
 Parameter values are the same as discussed in the section \ref{sec:parameters}.
-Cosmological redshift of the burst $z = 1$.
+Redshift $z = 1$.
 Observer's off-axis angle $\chi = 0.01$.
@@
 
-Finally, this assumption is able to explain the non-trivial stretching factor of the GRB 090926A.
-The detailed computations will be described in the following sections, but you can see why the stretching is happening on the figure \ref{fig:modelOverview}.
-The computations confirm that this qualitative picture is indeed correct, you can see a sample light curve on the figure \ref{fig:sampleLightCurve}, and predicted distribution of stretching factors on the figure \ref{fig:kappaDistribution}.
+Finally, our assumption is able to explain the non-trivial stretching factor of the GRB 090926A.
+Detailed computations will be described in the following sections, but you can see the main idea on the figure \ref{fig:modelOverview}.
+The computations confirm that this qualitative picture is indeed correct.
+You can see a sample light curve on the figure \ref{fig:sampleLightCurve}, and predicted distribution of stretching factors on the figure \ref{fig:kappaDistribution}.
 

### 2014-01-22 Paper/gammaRays.tex
Drafted Burst Data Queries

@@
 \usepackage{hyperref}
+\usepackage{siunitx}
 
@@
 
-\subsection{Download and Process}
+\subsection{Burst Data Queries}
 
-Download in progress.
+\begin{table}[bursts]
+	\centering
+	\small
+	\begin{tabular}{ l | S[table-format=9.3] | S[table-format=3.3] | S[table-format=2.3] | S[table-format=1.5] | S[table-format=2.1] | S[table-format=3.1] }
+		\multicolumn{1}{ c |}{\texttt{name}} & $\mathrm{\texttt{time}}$ & \texttt{location.ra} & \texttt{location.dec} & \texttt{location.err} & \texttt{startOffset} & \texttt{endOffset} \\
+		\multicolumn{1}{ c |}{GRB} & $\mathrm{GBM\ Trigger\ Time}$ & $\mathrm{R.A.}$ & $\mathrm{Dec.}$ & $\mathrm{Location\ Error}$ & $T_{05}$ & $T_{95}$ \\
+		\multicolumn{1}{ c |}{Name}& $\mathrm{MET,\ sec}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{deg}$ & $\mathrm{sec}$ & $\mathrm{sec}$ \\
+		\hline
+		080825C	&	241366429.105	&	233.9  	&	 -4.5  	&	0.75   	&	 3.2	&	 29.4	\\
+		080916C	&	243216766.614	&	119.85 	&	-56.64 	&	0.0001 	&	 5.0	&	209.8	\\
+		081006 	&	244996175.173	&	136.32 	&	-62.05 	&	0.52   	&	 0.7	&	115.0	\\
+		081024B	&	246576161.864	&	322.95 	&	 21.2  	&	0.22   	&	 0.1	&	191.0	\\
+		090217 	&	256539404.560	&	204.83 	&	 -8.42 	&	0.35   	&	 6.2	&	 68.0	\\
+		090323 	&	259459364.630	&	190.71 	&	 17.053	&	0.0001 	&	15.9	&	293.9	\\
+		090328 	&	259925808.510	&	 90.67 	&	-41.715	&	0.0002 	&	18.8	&	652.9	\\
+		090510 	&	263607781.971	&	333.55 	&	-26.583	&	0.0004 	&	 0.6	&	 45.6	\\
+		090626 	&	267683530.880	&	170.03 	&	-33.49 	&	0.22   	&	52.2	&	299.9	\\
+		090902B	&	273582310.313	&	264.94 	&	 27.324	&	0.001  	&	 7.7	&	825.0	\\
+		090926A	&	275631628.990	&	353.4  	&	-66.32 	&	0.01   	&	 5.5	&	225.0	\\
+		091003 	&	276237347.585	&	251.52 	&	 36.625	&	0.0005 	&	 3.9	&	452.6	\\
+		091031 	&	278683230.850	&	 71.49 	&	-57.65 	&	0.23   	&	 3.1	&	206.2	\\
+		100116A	&	285370262.240	&	305.01 	&	 14.43 	&	0.17   	&	 3.0	&	141.0	\\
+		100414A	&	292904423.990	&	192.11 	&	  8.693	&	0.0005 	&	17.4	&	288.6	\\
+		110120A	&	317231981.230	&	 61.5  	&	-12.0  	&	0.36   	&	 0.5	&	112.8	\\
+		110428A	&	325675112.410	&	  5.59 	&	 64.849	&	0.00001	&	10.7	&	407.6	\\
+		110721A	&	332916465.760	&	333.2  	&	-38.5  	&	0.20   	&	 0.1	&	239.0	\\
+		110731A	&	333803371.954	&	280.504	&	-28.537	&	0.0001 	&	 3.0	&	 24.1
+	\end{tabular}
+	\caption{Burst data used in our study.
+The first row contains the corresponding \texttt{GRBurst} class variable names.
+The data is taken from the tables 2 and 4 of \cite{Ackermann:2013zfa}.
+\url{https://github.com/maxitg/GammaRays/blob/master/GRObservations/bursts}}
+	\label{tab:bursts}
+\end{table}
 
+Before we start to compute the Kolmogorov-Smirnov probabilities, we need to download the actual observational data from the Fermi LAT data server.
+We take the data about the bursts from the catalog \cite{Ackermann:2013zfa}, which contains all the bright bursts seen by the LAT since 2008 August 4 to 2011 August 1.
+Specifically, we take 4 pieces of information from the tables 2 and 4 there:
+\begin{itemize}
+	\item{GBM trigger time (table 2). It is used as a reference point for the GRB time.}
+	\item{Location of the GRB (table 2). We need burst locations to filter out photons coming from other sources.}
+	\item{Location error (table 2). Location errors are used to improve the accuracy of filtering, that is to avoid losing statistics by filtering too much.}
+	\item{$T_{05}$ -- $T_{95}$ interval of the LAT-detected emission (table 4).
+We need precise intervals for two things.
+First, to understand where is the fixed point of the stretching, $T_{05}$ is used as a good estimate for that.
+Second, the burst interval is used to download events data from the data server.}
+\end{itemize}
+The data obtained is summarized in the table \ref{tab:bursts}.
+
+Finally, to download the data about observed photons, we fill the form at the LAT data server front page with the following values:
+\begin{itemize}
+	\item{{\bf Object name or coordinates}. Here the coordinates are filled from the table \ref{tab:bursts}.}
+	\item{{\bf Coordinate system}. We use J2000 here.}
+	\item{{\bf Search radius (degrees)}. The value of $60$ is used here. Events will be filtered precisely later on by using the point spread functions (PSFs) of the LAT.}
+	\item{
+		{\bf Observation dates}.
+We extend the duration of the burst by $15\%$ relative to the table \ref{tab:bursts} values to both past and future to have a decent safety margin.
+So the time ranges we fill in here are:
+		\begin{align*}
+			\texttt{time} + \texttt{startOffset} &- 0.15\left(\texttt{endOffset}-\texttt{startOffset}\right),\\
+			\texttt{time} + \texttt{endOffset} &+ 0.15\left(\texttt{endOffset}-\texttt{startOffset}\right)
+		\end{align*}
+	}
+	\item{{\bf Time system}. MET is used.}
+	\item{{\bf Energy range (MeV)}. The whole energy range is required for our study. So we fill in $30,\,300000$.}
+	\item{{\bf LAT data type}. Extended, because we need transient photons for better statistics.}
+	\item{{\bf Spacecraft data}. Checked. We need it for both events filtering, and calculation of the PSFs and exposure maps.}
+\end{itemize}
+
+The data we downloaded contains a lot of background and radiation from other sources. We need to filter it before using it.
+
+\subsection{Data Filtering}
+\subsection{Point Spread Functions}
+\subsection{Exposure Maps}
 \subsection{Linear Component (background)}

### 2014-01-23 Paper/gammaRays.tex
Drafted filtering with gtselect and gtmktime

@@
 \usepackage{siunitx}
+\usepackage{listings}
+
+\lstset{columns=fullflexible,basicstyle=\ttfamily}
 
@@
 
-\subsection{Burst Data Queries}
+\subsection{Burst Data Download}
 
@@
 We extend the duration of the burst by $15\%$ relative to the table \ref{tab:bursts} values to both past and future to have a decent safety margin.
-So the time ranges we fill in here are:
+So the time range we fill in here is:
 		\begin{align*}
@@
 
-The data we downloaded contains a lot of background and radiation from other sources. We need to filter it before using it.
+For background estimation purposes, we also download observational data during a day before the burst, specifically in the time range:
+\begin{align*}
+	\texttt{time} + \texttt{startOffset} &- 0.15\left(\texttt{endOffset} - \texttt{startOffset}\right) - \unit[86400]{sec},\\
+	\texttt{time} + \texttt{startOffset} &- 0.15\left(\texttt{endOffset} - \texttt{startOffset}\right)
+\end{align*}
+All other values filled in the form are the same.
 
-\subsection{Data Filtering}
+Before reading the data, we need to apply the basic filtering as described in Fermi guidelines.
+Two tools from the Fermi Science Tools package are used for that task (the same filtering is applied to both burst and background data).
+
+The first one is \texttt{gtselect}.
+It filters the Earth limb emission.
+We run the tool with the following parameters\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/explore_latdata_burst.html}}:
+\begin{lstlisting}
+gtselect infile=[eventFile] outfile=filtered.fits
+	ra=INDEF dec=INDEF rad=180 tmin=INDEF tmax=INDEF emin=100 emax=300000
+	zmax=100 evclass=0 convtype=-1 evtable=EVENTS
+\end{lstlisting}
+Here \texttt{[eventFile]} is the file downloaded in the previous step.
+
+The second one is \texttt{gtmktime}.
+It uses the spacecraft file as well as events file, and filters out time periods when Fermi LAT was not operational.
+Running \texttt{gtmktime} is also required for exposure maps calculation, which we'll need later on.
+Parameters are the following\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/data_preparation.html}}:
+\begin{lstlisting}
+gtmktime scfile=[spacecraft] evfile=filtered.fits outfile=timed.fits
+	filter="DATA_QUAL>0 && LAT_CONFIG==1" roicut=yes
+\end{lstlisting}
+where \texttt{[spacecraft]} is the FITS containing the spacecraft data.
+
+Now, with basic filtering done, we can read events data from \texttt{timed.fits}.
+This file, however, contains observations of a large area of the sky.
+Before using the data obtained from it, we should perform a more elaborate filtering by location using Fermi LAT's point spread function, which will be described in the following section.
+
 \subsection{Point Spread Functions}
 \subsection{Exposure Maps}
-\subsection{Linear Component (background)}
 \subsection{KS-test}

### 2014-01-23 GRObservations/GRObservations/GRFermiLATDataServerQuery.cpp

@@ -392,8 +392,8 @@ int GRFermiLATDataServerQuery::gtmktime() {
     ostringstream cmd;
     cmd << fixed << "gtmktime" << " ";
     cmd << "scfile=" << hash << "/spacecraft.fits" << " ";
-    cmd << "filter=" << "\"DATA_QUAL>0 && LAT_CONFIG==1 && ABS(ROCK_ANGLE)<52\"" << " ";
-    cmd << "roicut=" << "no" << " ";
+    cmd << "filter=" << "\"DATA_QUAL>0 && LAT_CONFIG==1\"" << " ";
+    cmd << "roicut=" << "yes" << " ";
     cmd << "evfile=" << hash << "/filtered.fits" << " ";
     cmd << "outfile=" << hash << "/timed.fits" << " ";
     cmd << "chatter=" << 0 << " ";

### 2014-01-23 Paper/gammaRays.tex
Fixed bursts table format (to prevent overfull warnings)

@@
 	\small
-	\begin{tabular}{ l | S[table-format=9.3] | S[table-format=3.3] | S[table-format=2.3] | S[table-format=1.5] | S[table-format=2.1] | S[table-format=3.1] }
+	\begin{tabular}{ l | S[table-format=9.3] | S[table-format=3.3] | S[table-format=-2.3] | S[table-format=1.5] | S[table-format=2.1] | S[table-format=3.1] }
 		\multicolumn{1}{ c |}{\texttt{name}} & $\mathrm{\texttt{time}}$ & \texttt{location.ra} & \texttt{location.dec} & \texttt{location.err} & \texttt{startOffset} & \texttt{endOffset} \\

### 2014-01-23 Paper/gammaRays.tex
Revised data download

@@
 	\item{Location of the GRB (table 2). We need burst locations to filter out photons coming from other sources.}
-	\item{Location error (table 2). Location errors are used to improve the accuracy of filtering, that is to avoid losing statistics by filtering too much.}
+	\item{Location error (table 2). Location errors are used to improve accuracy of filtering, specifically to avoid losing statistics by filtering too much.}
 	\item{$T_{05}$ -- $T_{95}$ interval of the LAT-detected emission (table 4).
 We need precise intervals for two things.
-First, to understand where is the fixed point of the stretching, $T_{05}$ is used as a good estimate for that.
+First, to understand where is the fixed point of the stretching ($T_{05}$ is used as a proxy for that).
 Second, the burst interval is used to download events data from the data server.}
@@
 	\item{{\bf Object name or coordinates}. Here the coordinates are filled from the table \ref{tab:bursts}.}
-	\item{{\bf Coordinate system}. We use J2000 here.}
-	\item{{\bf Search radius (degrees)}. The value of $60$ is used here. Events will be filtered precisely later on by using the point spread functions (PSFs) of the LAT.}
+	\item{{\bf Coordinate system}. J2000.}
+	\item{{\bf Search radius (degrees)}. $60$. Events will be filtered by location later on by using the point spread functions (PSFs) of the LAT.}
 	\item{
 		{\bf Observation dates}.
-We extend the duration of the burst by $15\%$ relative to the table \ref{tab:bursts} values to both past and future to have a decent safety margin.
-So the time range we fill in here is:
+To have a decent safety margin, we extend the duration of the burst by $15\%$ to both past and future relative to the table \ref{tab:bursts} time ranges.
+So, we fill in the following values:
 		\begin{align*}
@@
 	}
-	\item{{\bf Time system}. MET is used.}
-	\item{{\bf Energy range (MeV)}. The whole energy range is required for our study. So we fill in $30,\,300000$.}
+	\item{{\bf Time system}. MET.}
+	\item{{\bf Energy range (MeV)}. $30,\,300000$. The whole energy range is required for our study.}
 	\item{{\bf LAT data type}. Extended, because we need transient photons for better statistics.}
-	\item{{\bf Spacecraft data}. Checked. We need it for both events filtering, and calculation of the PSFs and exposure maps.}
+	\item{{\bf Spacecraft data}. Checked. It is required for both events filtering, and calculation of the PSFs and exposure maps.}
 \end{itemize}
 
-For background estimation purposes, we also download observational data during a day before the burst, specifically in the time range:
+For background estimation, we also download observational data during a day before the burst. All the values are put the same into the form, except the time range:
 \begin{align*}

### 2014-01-24 Paper/gammaRays.tex
Described PSF calculation

@@
 \author{
-	Maxim S. Piskunov$^{1}$\thanks{maxit@ms2.inr.ac.ru} and
-	Grigory I. Rubtsov$^{1}$\thanks{grisha@ms2.inr.ac.ru} \\
+	Maxim S. Piskunov$^{1}$\thanks{\href{mailto:maxit@ms2.inr.ac.ru}{maxit@ms2.inr.ac.ru}} and
+	Grigory I. Rubtsov$^{1}$\thanks{\href{mailto:grisha@ms2.inr.ac.ru}{grisha@ms2.inr.ac.ru}} \\
 	$^{1}$Institute for Nuclear Research RAS, 117312, Moscow, Russia\\
@@
 
-The first one is \texttt{gtselect}.
+The first one is \texttt{gtselect}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtselect.txt}}.
 It filters the Earth limb emission.
@@
 
-The second one is \texttt{gtmktime}.
+The second one is \texttt{gtmktime}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtmktime.txt}}.
 It uses the spacecraft file as well as events file, and filters out time periods when Fermi LAT was not operational.
@@
 \begin{lstlisting}
-gtmktime scfile=[spacecraft] evfile=filtered.fits outfile=timed.fits
+gtmktime scfile=[spacecraft] sctable=SC_DATA
 	filter="DATA_QUAL>0 && LAT_CONFIG==1" roicut=yes
+	evfile=filtered.fits evtable=EVENTS outfile=timed.fits
+	apply_fiter=yes
 \end{lstlisting}
@@
 \subsection{Point Spread Functions}
+
+\texttt{gtpsf}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtpsf.txt}} tool requires the livetime map to generate the PSFs.
+This map is computed by \texttt{gtltcube}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtltcube.txt}} with the following parameters:
+\begin{lstlisting}
+gtltcube evfile=timed.fits evtable=EVENTS scfile=[spacecraft] sctable=SC_DATA
+	outfile=ltcube.fits
+	dcostheta=0.025 binsz=1 phibins=0 tmin=0 tmax=0 zmax=180 zmin=0
+\end{lstlisting}
+
+Having the maps, we can compute the PSFs:
+\begin{lstlisting}
+gtpsf expcube=ltcube.fits outfile=psf_[IrfName].fits outtable=PSF irfs=[IrfName]
+	ra=[RA] dec=[DEC] emin=100 emax=1000000 nenergies=41 thetamax=30 ntheta=300
+\end{lstlisting}
+Here \texttt{[RA]} and \texttt{[DEC]} are the coordinates of the burst, and \texttt{[IrfName]} is the Instrument Response Function, which depends on photon's event class (transient, source, clean, or ultraclean) and conversion type (back, or front).
+The list of IRF names can be obtained by using the \texttt{gtirfs}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtirfs.txt}} tool.
+We need to compute PSFs for all of them.
+
+Produced files contain the probability densities as a function of the angular distance to the source.
+But we need an integral value, that is given the distance $d$, we want to find a fraction of photons coming from this source, which are observed to be closer than $d$ to the source.
+To find that we sum over the probability densities:
+\begin{equation*}
+	p\left(n\right) = \sum_{i=0}^{n} \frac{\texttt{pdf[i]} + \texttt{pdf[i+1]}}{2} 2\pi \left(\cos\left(\texttt{angles[i]}\right) - \cos\left(\texttt{angles[i+1]}\right)\right)
+\end{equation*}
+Here \texttt{pdf[i]} is the observation probability density in the point separated with the source by the angle \texttt{angles[i]}.
+And $p\left(n\right)$ is the probability to observe the photon closer to the source than \texttt{angles[n]}.
+We find probabilities for intermediate angles by linear interpolation.
+Finally, we can compute the inverse value (that is distance as a function of probability) by binary search.
+
 \subsection{Exposure Maps}

### 2014-01-24 GRObservations/GRObservations/GRFermiLATDataServerQuery.cpp

@@ -392,10 +392,13 @@ int GRFermiLATDataServerQuery::gtmktime() {
     ostringstream cmd;
     cmd << fixed << "gtmktime" << " ";
     cmd << "scfile=" << hash << "/spacecraft.fits" << " ";
+    cmd << "sctable=" << "SC_DATA" << " ";
     cmd << "filter=" << "\"DATA_QUAL>0 && LAT_CONFIG==1\"" << " ";
     cmd << "roicut=" << "yes" << " ";
     cmd << "evfile=" << hash << "/filtered.fits" << " ";
+    cmd << "evtable=" << "EVENTS" << " ";
     cmd << "outfile=" << hash << "/timed.fits" << " ";
+    cmd << "apply_filter=" << "yes" << " ";
     cmd << "chatter=" << 0 << " ";
     return system(cmd.str().c_str());
 }
@@ -410,6 +413,32 @@ int GRFermiLATDataServerQuery::gtltcube() {
     cmd << "outfile=" << hash << "/ltcube.fits" << " ";
     cmd << "dcostheta=" << 0.025 << " ";
     cmd << "binsz=" << 1 << " ";
+    cmd << "phibins=" << 0 << " ";
+    cmd << "tmin=" << 0 << " ";
+    cmd << "tmax=" << 0 << " ";
+    cmd << "zmax=" << 180 << " ";
+    cmd << "zmin=" << 0 << " ";
+    cmd << "chatter=" << 0 << " ";
+    return system(cmd.str().c_str());
+}
+
+int GRFermiLATDataServerQuery::gtpsf(GRFermiEventClass eventClass, GRFermiConversionType conversionType) {
+    string irf = irfName(eventClass, conversionType);
+    string psfFilename = "psf_" + irf + ".fits";
+    
+    ostringstream cmd;
+    cmd << fixed << "gtpsf" << " ";
+    cmd << "expcube=" << hash << "/ltcube.fits" << " ";
+    cmd << "outfile=" << hash << "/" << psfFilename << " ";
+    cmd << "outtable=" << "PSF" << " ";
+    cmd << "irfs=" << irf << " ";
+    cmd << "ra=" << location.ra << " ";
+    cmd << "dec=" << location.dec << " ";
+    cmd << "emin=" << 100. << " ";
+    cmd << "emax=" << 1000000. << " ";
+    cmd << "nenergies=" << 41 << " ";
+    cmd << "thetamax=" << 30 << " ";
[…]

### 2014-01-24 Paper/gammaRays.tex
Exposure maps description drafted

@@
 
-\subsection{Point Spread Functions}
+\subsection{Point Spread Functions and Location Filtering}
 
@@
 
-Produced files contain the probability densities as a function of the angular distance to the source.
-But we need an integral value, that is given the distance $d$, we want to find a fraction of photons coming from this source, which are observed to be closer than $d$ to the source.
-To find that we sum over the probability densities:
+Imagine a stream of photons coming from some source locate at some coordinates $\theta, \phi$.
+Ideally, all the coordinates measured for the photons coming from the source should be the same.
+However, due to measurement errors, they would be spread across some region of the sky.
+If we find the region such that, say, $95\%$ of photons coming from the source are measured inside of it, we can filter out all the other photons without losing too much statistics, and significantly reducing background.
+\texttt{gtpsf} approximates these regions as circles, and the files generated in the previous step contain the observation probability densities as functions of distance from the source, and photon energies.
+We, however, need an integral quantity, that is given the radius, we want to know the fraction of photons observed inside the circle of this radius.
+We can compute this fraction by summing over the probability densities:
 \begin{equation*}
@@
 \end{equation*}
-Here \texttt{pdf[i]} is the observation probability density in the point separated with the source by the angle \texttt{angles[i]}.
-And $p\left(n\right)$ is the probability to observe the photon closer to the source than \texttt{angles[n]}.
+Here \texttt{pdf[i]} is the probability density in the point separated from the source by an angle \texttt{angles[i]}.
+And $p\left(n\right)$ is the probability to observe a photon closer to the source than \texttt{angles[n]}.
 We find probabilities for intermediate angles by linear interpolation.
-Finally, we can compute the inverse value (that is distance as a function of probability) by binary search.
+Finally, we can compute the inverse value $\theta_{\text{spread}}\left(p\right)$ (that is a distance as a function of probability) by binary search.
+Note, that $\theta_{\text{spread}}\left(p\right)$ depends on the photon energy, event class, and conversion type.
+We should use the appropriate function for each photon.
+The spread angles for intermediate energies are linearly interpolated.
 
-\subsection{Exposure Maps}
+Finally, to finish the filtering, we filter out all the photons, for which their location is separated from the source by more that $\theta_{\text{spread}}\left(0.95\right) + \texttt{location.err}$, where \texttt{location.err} is the uncertainty in the burst location (see table \ref{tab:bursts}).
+
+Now we have a decent set of photons, which we can use for analysis.
+However, some of the photons in this set are still originated from background, and not from the source.
+We cannot filter out these photons, however, we can substitute a linear component of background by estimating it from a long time period before the burst.
+In order to perform this computation, we'll need an exposure map, which will be discussed in the following section.
+
+\subsection{Exposure Maps and Background Estimation}
+
+The exposure map can be computed with the \texttt{gtexpcube2}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtexpcube2.txt}} tool. We use the following parameter set for it:
+\begin{lstlisting}
+gtexpcube2 infile=ltcube.fits cmap=none outfile=expcube_[IrfName].fits irfs=[IrfName]
+	nxpix=360 nypix=180 binsz=1 coordsys=CEL xref=0 yref=0 axisrot=0
+	proj=CAR ebinalg=log emin=100 emax=1000000 enumbins=40
+	ebinfile=NONE bincalc=EDGE ignorephi=no thmax=180 thmin=0 table=Exposure
+\end{lstlisting}
+The generated \texttt{expcube} files contain exposures as functions of energy and location on the sky\footnote{Note, that in the \texttt{expcube} FITS files R.A.
+coordinates are indexed in reverse order and starting from 180, for example, $\texttt{ra[0]}=180$, $\texttt{ra[1]}=179$, and so on.}.
+We use trilinear interpolation to compute exposures for all energies and locations.
+
+Knowing the exposures, we can now use the method introduced in the appendix of \cite{Rubtsov:2011qq} to estimate the linear component of background from the data observed before the burst.
+We compute two background estimates: for low and high energy bands.
+
 \subsection{KS-test}

### 2014-01-25 GRObservations/GRObservations/GRBurst.cpp
Refactored GRDistribution.cpp. Fixed bug with zenith angle filtering

@@ -16,13 +16,13 @@
 
 #include "GRBurst.h"
 
-#define TIME_EXTENTION_FACTOR 1.3
-#define PHOTONS_QUALITY       0.95
-#define LENGTHENING_MIN       0.1
-#define LENGTHENING_MAX       10.
-#define LENGTHENING_STEP      1.001
+#define TIME_EXTENTION_FACTOR   1.3
+#define PHOTONS_QUALITY         0.95
+#define STRETCHING_MIN          0.1
+#define STRETCHING_MAX          10.
+#define STRETCHING_STEP         1.001
 
-#define BACKGROUND_DURATION   86400
+#define BACKGROUND_DURATION     86400
 
 void GRBurst::init() {
     
@@ -99,11 +99,11 @@ void GRBurst::calculateBackground() {
         }
     }
         
-    mevDistribution.estimatedLinearComponent = mevExpectationValue;
+    mevDistribution.linearComponent = - mevExpectationValue + gevExpectationValue;
     mevDistribution.start = query.startTime - (time + startOffset);
     mevDistribution.end = query.endTime - (time + startOffset);
     
-    gevDistribution.estimatedLinearComponent = gevExpectationValue;
+    gevDistribution.linearComponent = 0.;
     gevDistribution.start = query.startTime - (time + startOffset);
     gevDistribution.end = query.endTime - (time + startOffset);
     
@@ -123,8 +123,6 @@ void GRBurst::read() {
         errorDescription = backgroundQuery.errorDescription;
         return;
     }
-    
-    cout << "event counts: " << query.events.size() << " " << backgroundQuery.events.size() << endl;
     
     for (int i = 0; i < query.events.size(); i++) {
         GRFermiLATPhoton photon = query.events[i];
@@ -148,6 +146,8 @@ void GRBurst::read() {
     sort(gevPhotons.begin(), gevPhotons.end());
     sort(gevBackgroundPhotons.begin(), gevBackgroundPhotons.end());
     
+    cout << "event counts: " << mevPhotons.size() + gevPhotons.size() << " " << mevBackgroundPhotons.size() + gevBackgroundPhotons.size() << endl;
+    
     calculateBackground();
     
     for (int i = 0; i < mevPhotons.size(); i++) {
@@ -168,39 +168,45 @@ void GRBurst::evaluate() {
         }
     }
     
+    ofstream mev((name + "/mev").c_str());
+    vector <struct GRDistributionCDFPoint> mevCdf = mevDistribution.cdf();
[…]

### 2014-01-25 GRObservations/GRObservations/GRBurst.h

@@ -45,8 +45,6 @@ public:
     double startOffset;
     double endOffset;
     
-    double redshift;
-        
 public:
     GRFermiLATDataServerQuery backgroundQuery;
     GRFermiLATDataServerQuery query;
@@ -62,21 +60,15 @@ public:
     GRDistribution gevDistribution;
     
 public:
-    double minLengthening[5];
-    double maxLengthening[5];
-    vector <float> lengtheningValues;
-    vector <float> lengtheningProbabilities;
+    double minStretching[5];
+    double maxStretching[5];
+    vector <float> stretchingValues;
+    vector <float> stretchingProbabilities;
     
     double trivialProbability;
     
 private:
     void calculateBackground();
-    double middleTime();
-    double duration();
-    double startTime();
-    double endTime();
-    double backgroundStartTime();
-    double backgroundEndTime();
             
 public:
     void init();

### 2014-01-25 GRObservations/GRObservations/GRDistribution.cpp

@@ -6,29 +6,12 @@
 //  Copyright (c) 2013 Maxim Piskunov. All rights reserved.
 //
 
-#include <math.h>
 #include <cmath>
+#include <algorithm>
 
 #include <fstream>
 
-#include <glpk.h>
-
 #include "GRDistribution.h"
-
-vector <GRDistributionCDFPoint> GRDistribution::cdf() {
-    vector <GRDistributionCDFPoint> result;
-    result.reserve(2*values.size());
-    for (int i = 0; i < values.size(); i++) {
-        GRDistributionCDFPoint point;
-        point.value = values[i];
-        point.probability = (double)i/values.size();
-        result.push_back(point);
-        point.value = values[i];
-        point.probability = (double)(i+1)/values.size();
-        result.push_back(point);
-    }
-    return result;
-}
 
 double GRDistribution::kolmogorovSmirnovProbability(double distance, int n1, int n2) {
     double n = (double)n1*n2/(n1+n2);
@@ -83,9 +66,92 @@ double GRDistribution::kolmogorovSmirnovDistance(double probability, int n1, int
     return z / (sqrt(n) + 0.12 + 0.11/sqrt(n));
 }
 
+pair<double, double> GRDistribution::cdfValueRange(double time) {
+    if (time < start) return make_pair(0, 0);
+    if (time > end) return make_pair(1, 1);
+    
+    vector<double>::iterator low = lower_bound(values.begin(), values.end(), time);
+    vector<double>::iterator high = upper_bound(values.begin(), values.end(), time);
+    
+    pair<double, double> result = make_pair(low-values.begin(), high-values.begin());
+    
+    result.first += linearComponent * (time - start) / (end - start);
+    result.second += linearComponent * (time - start) / (end - start);
+    
+    double total = values.size() + linearComponent;
+    
+    result.first /= total;
+    result.second /= total;
+    
+    return result;
+}
+
+vector <struct GRDistributionCDFPoint> GRDistribution::cdf() {
+    vector <double> times;
+    times.reserve(values.size() + 2);
+    times.push_back(start);
+    for (int i = 0; i != values.size(); ++i) {
+        times.push_back(values[i]);
+    }
+    times.push_back(end);
+    
+    vector <struct GRDistributionCDFPoint> result;
+    result.reserve(2*values.size() + 2);
+    
+    for (int i = 0; i != times.size(); ++i) {
+        pair<double, double> values = cdfValueRange(times[i]);
+        
[…]

### 2014-01-25 GRObservations/GRObservations/GRDistribution.h

@@ -23,7 +23,7 @@ struct GRDistributionCDFPoint {
 class GRDistribution {
 public:
     vector <double> values;
-    double estimatedLinearComponent;
+    double linearComponent;
     double start;
     double end;
     
@@ -31,9 +31,11 @@ private:
     double kolmogorovSmirnovProbability(double distance, int n1, int n2);
     double kolmogorovSmirnovDistance(double probability, int n1, int n2);
 public:
-    vector <GRDistributionCDFPoint> cdf();
-    float kolmogorovSmirnovTest(GRDistribution distribution, double lengthening = 1., bool plot = false, ostream &mev = cout, ostream &gev = cout);
-    pair<double, double> lengtheningLimits(GRDistribution distribution, float probability, bool *success);
+    pair<double, double> cdfValueRange(double time);
+    vector <struct GRDistributionCDFPoint> cdf();
+    
+    float kolmogorovSmirnovTest(GRDistribution distribution, double stretching = 1.);
+    
 };
 
 #endif /* defined(__Gamma_Rays__GRDistribution__) */

### 2014-01-25 GRObservations/GRObservations/GRExposureMap.cpp

@@ -47,6 +47,7 @@ float GRExposureMap::exposure(float energy, GRLocation location) {
 }
 
 float GRExposureMap::exposure(int energyIndex, GRLocation location) {
+#warning change to interpolation
     int raLow = location.ra;
     int raHigh = raLow+1;
     if (raHigh == 360) {

### 2014-01-25 GRObservations/GRObservations/GRFermiLATDataServerQuery.cpp

@@ -394,7 +394,8 @@ int GRFermiLATDataServerQuery::gtmktime() {
     cmd << "scfile=" << hash << "/spacecraft.fits" << " ";
     cmd << "sctable=" << "SC_DATA" << " ";
     cmd << "filter=" << "\"DATA_QUAL>0 && LAT_CONFIG==1\"" << " ";
-    cmd << "roicut=" << "yes" << " ";
+    //cmd << "roicut=" << "yes" << " ";
+    cmd << "roicut=" << "no" << " ";
     cmd << "evfile=" << hash << "/filtered.fits" << " ";
     cmd << "evtable=" << "EVENTS" << " ";
     cmd << "outfile=" << hash << "/timed.fits" << " ";

### 2014-01-25 GRObservations/GRObservations/main.cpp

@@ -93,11 +93,11 @@ int main(int argc, const char * argv[])
         }
         
         log << burstCatalog[i].name << ":" << endl;
-        log << "\tLow energy count  = " << burstCatalog[i].mevDistribution.values.size() - burstCatalog[i].mevDistribution.estimatedLinearComponent << endl;
-        log << "\tHigh energy count = " << burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.estimatedLinearComponent << endl;
+        log << "\tLow energy count  = " << burstCatalog[i].mevDistribution.values.size() - burstCatalog[i].mevDistribution.linearComponent << endl;
+        log << "\tHigh energy count = " << burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.linearComponent << endl;
         
         
-        if ((burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.estimatedLinearComponent) < 10) {
+        if ((burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.linearComponent) < 10) {
             log << "\tnot enought high energy photons" << endl << endl;
             continue;
         }
@@ -106,13 +106,13 @@ int main(int argc, const char * argv[])
         
         log << "\tprobability not to be stretched = " << sigma(burstCatalog[i].trivialProbability) << endl;
         for (int k = 0; k < 5; k++) {
-            log << "\t" << k+1 << "σ -> stretching factor in ranges (" << burstCatalog[i].minLengthening[k] << ", " << burstCatalog[i].maxLengthening[k] << ")" << endl;
+            log << "\t" << k+1 << "σ -> stretching factor in ranges (" << burstCatalog[i].minStretching[k] << ", " << burstCatalog[i].maxStretching[k] << ")" << endl;
         }
         log << endl;
         
         ofstream probs((burstCatalog[i].name + "/probs").c_str());
-        for (int j = 0; j < burstCatalog[i].lengtheningValues.size(); j++) {
-            probs << burstCatalog[i].lengtheningValues[j] << " " << burstCatalog[i].lengtheningProbabilities[j] << endl;
+        for (int j = 0; j < burstCatalog[i].stretchingValues.size(); j++) {
+            probs << burstCatalog[i].stretchingValues[j] << " " << burstCatalog[i].stretchingProbabilities[j] << endl;
         }
         probs.close();
     }

### 2014-03-03 GRObservations/GRObservations/GRBurst.cpp
change TIME_EXTENSION_FACTOR from 1.3 to 2.0

@@ -15,8 +15,7 @@
 #include <sstream>
 
 #include "GRBurst.h"
-
-#define TIME_EXTENTION_FACTOR   1.3
+#define TIME_EXTENSION_FACTOR   2.0
 #define PHOTONS_QUALITY         0.95
 #define STRETCHING_MIN          0.1
 #define STRETCHING_MAX          10.
@@ -27,12 +26,12 @@
 void GRBurst::init() {
     
     double middle = time + (startOffset + endOffset) / 2.;
-    double duration = (endOffset - startOffset) * TIME_EXTENTION_FACTOR;
+    double duration = (endOffset - startOffset) * TIME_EXTENSION_FACTOR;
     
     query.startTime = middle - duration/2.;
     query.endTime = middle + duration/2.;
     query.location = location;
-        
+    
     query.init();
     
     backgroundQuery.startTime = query.startTime - BACKGROUND_DURATION;
@@ -127,23 +126,25 @@ void GRBurst::read() {
     for (int i = 0; i < query.events.size(); i++) {
         GRFermiLATPhoton photon = query.events[i];
         photon.location.error = query.psfs[photon.eventClass][photon.conversionType].spread(photon.energy, PHOTONS_QUALITY);
+        if (photon.eventClass == GRFermiEventClassTransient) continue;
         if (location.isSeparated(photon.location)) continue;
         
         if (photon.energy < 1000.) mevPhotons.push_back(photon);
         else gevPhotons.push_back(photon);
     }
     sort(mevPhotons.begin(), mevPhotons.end());
-    sort(mevBackgroundPhotons.begin(), mevBackgroundPhotons.end());
+    sort(gevPhotons.begin(), gevPhotons.end());
     
     for (int i = 0; i < backgroundQuery.events.size(); i++) {
         GRFermiLATPhoton photon = backgroundQuery.events[i];
         photon.location.error = backgroundQuery.psfs[photon.eventClass][photon.conversionType].spread(photon.energy, PHOTONS_QUALITY);
+        if (photon.eventClass == GRFermiEventClassTransient) continue;
         if (location.isSeparated(photon.location)) continue;
         
         if (photon.energy < 1000.) mevBackgroundPhotons.push_back(photon);
         else gevBackgroundPhotons.push_back(photon);
     }
-    sort(gevPhotons.begin(), gevPhotons.end());
+    sort(mevBackgroundPhotons.begin(), mevBackgroundPhotons.end());
[…]

### 2014-03-03 GRObservations/GRObservations/GRBurst.h

@@ -64,6 +64,9 @@ public:
     double maxStretching[5];
     vector <float> stretchingValues;
     vector <float> stretchingProbabilities;
+    vector <double> maxDistanceTimes;
+    vector <double> maxDistanceGevValue;
+    vector <double> maxDistanceMevValue;
     
     double trivialProbability;
     

### 2014-03-03 GRObservations/GRObservations/GRDistribution.cpp

@@ -120,11 +120,13 @@ vector <struct GRDistributionCDFPoint> GRDistribution::cdf() {
 // this is supposed to be GeV distribution
 // distribution is MeV distribution
 
-float GRDistribution::kolmogorovSmirnovTest(GRDistribution distribution, double stretching) {
+float GRDistribution::kolmogorovSmirnovTest(GRDistribution distribution, double stretching, double *time, double *thisValue, double *distributionValue) {
     vector<GRDistributionCDFPoint> thisCDF = this->cdf();
     vector<GRDistributionCDFPoint> distributionCDF = distribution.cdf();
     
     double maxDistance = 0.;
+    double maxDistanceTime = 0.;
+    pair<double, double> maxDistanceRange = make_pair(0, 0);
     
     for (int i = 0; i != thisCDF.size(); ++i) {
         double time = thisCDF[i].value;
@@ -132,8 +134,17 @@ float GRDistribution::kolmogorovSmirnovTest(GRDistribution distribution, double 
         
         pair<double, double> distributionProbabilityRange = distribution.cdfValueRange(time / stretching);
         
-        maxDistance = max(maxDistance, abs(thisProbability - distributionProbabilityRange.first));
-        maxDistance = max(maxDistance, abs(thisProbability - distributionProbabilityRange.second));
+        if (abs(thisProbability - distributionProbabilityRange.first) > maxDistance) {
+            maxDistance = abs(thisProbability - distributionProbabilityRange.first);
+            maxDistanceTime = time;
+            maxDistanceRange = make_pair(thisProbability, distributionProbabilityRange.first);
+        }
+        
+        if (abs(thisProbability - distributionProbabilityRange.second) > maxDistance) {
+            maxDistance = abs(thisProbability - distributionProbabilityRange.second);
+            maxDistanceTime = time;
+            maxDistanceRange = make_pair(thisProbability, distributionProbabilityRange.second);
+        }
     }
     
     for (int i = 0; i != distributionCDF.size(); ++i) {
@@ -142,9 +153,22 @@ float GRDistribution::kolmogorovSmirnovTest(GRDistribution distribution, double 
         
         pair<double, double> thisProbabilityRange = this->cdfValueRange(time * stretching);
         
-        maxDistance = max(maxDistance, abs(distributionProbability - thisProbabilityRange.first));
-        maxDistance = max(maxDistance, abs(distributionProbability - thisProbabilityRange.second));
[…]

### 2014-03-03 GRObservations/GRObservations/GRDistribution.h

@@ -34,7 +34,7 @@ public:
     pair<double, double> cdfValueRange(double time);
     vector <struct GRDistributionCDFPoint> cdf();
     
-    float kolmogorovSmirnovTest(GRDistribution distribution, double stretching = 1.);
+    float kolmogorovSmirnovTest(GRDistribution distribution, double stretching = 1., double *time = NULL, double *thisValue = NULL, double *distributionValue = NULL);
     
 };
 

### 2014-03-03 GRObservations/GRObservations/main.cpp

@@ -55,11 +55,15 @@ int main(int argc, const char * argv[])
         float error;
         double startOffset;
         double endOffset;
+        char isReadable;
         
-        burstsFile >> name >> time >> ra >> dec >> error >> startOffset >> endOffset;
+        burstsFile >> isReadable >> name >> time >> ra >> dec >> error >> startOffset;
         
-        if (name == "") continue;
-        if (name[0] == '#') continue;
+        if (burstsFile.eof()) continue;
+
+        burstsFile >> endOffset;
+        
+        if (isReadable == '#') continue;
         
         if ((endOffset - startOffset) != 0) {
             GRBurst burst;
@@ -96,7 +100,6 @@ int main(int argc, const char * argv[])
         log << "\tLow energy count  = " << burstCatalog[i].mevDistribution.values.size() - burstCatalog[i].mevDistribution.linearComponent << endl;
         log << "\tHigh energy count = " << burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.linearComponent << endl;
         
-        
         if ((burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.linearComponent) < 10) {
             log << "\tnot enought high energy photons" << endl << endl;
             continue;
@@ -112,7 +115,7 @@ int main(int argc, const char * argv[])
         
         ofstream probs((burstCatalog[i].name + "/probs").c_str());
         for (int j = 0; j < burstCatalog[i].stretchingValues.size(); j++) {
-            probs << burstCatalog[i].stretchingValues[j] << " " << burstCatalog[i].stretchingProbabilities[j] << endl;
+            probs << burstCatalog[i].stretchingValues[j] << " " << burstCatalog[i].stretchingProbabilities[j] << " " << burstCatalog[i].maxDistanceTimes[j] << " " << burstCatalog[i].maxDistanceGevValue[j] << " " << burstCatalog[i].maxDistanceMevValue[j] << endl;
         }
         probs.close();
     }

### 2014-03-03 GRObservations/GRObservations/GRDistribution.cpp
Removed outdated KS-test code

@@ -172,117 +172,3 @@ float GRDistribution::kolmogorovSmirnovTest(GRDistribution distribution, double 
     
     return kolmogorovSmirnovProbability(maxDistance, (int)values.size(), (int)distribution.values.size());
 }
-
-//  self is GeV distribution
-//  distribution is MeV distribution
-/*float GRDistribution::kolmogorovSmirnovTest(GRDistribution distribution, double lengthening, bool plot, ostream &mev, ostream &gev) {
-    for (int i = 0; i < distribution.values.size(); i++) {
-        distribution.values[i] *= lengthening;
-    }
-    distribution.start *= lengthening;
-    distribution.end *= lengthening;
-        
-    double maxDistance = 0.;
-    
-    vector <int> selfCounts(values.size());
-    vector <int> distributionCounts(distribution.values.size());
-    int selfIndex = 0;
-    int distributionIndex = 0;
-    while (selfIndex != values.size() || distributionIndex != distribution.values.size()) {
-        bool selfIncreased;
-        if (selfIndex == values.size()) {
-            distributionIndex++;
-            selfIncreased = false;
-        } else if (distributionIndex == distribution.values.size()) {
-            selfIndex++;
-            selfIncreased = true;
-        } else if (values[selfIndex] > distribution.values[distributionIndex]) {
-            distributionIndex++;
-            selfIncreased = false;
-        } else {
-            selfIndex++;
-            selfIncreased = true;
-        }
-        
-        if (selfIncreased) {
-            selfCounts[selfIndex-1] = distributionIndex;
-        } else {
-            distributionCounts[distributionIndex-1] = selfIndex;
-        }
-    }
-        
-    double descretePart = (double)distribution.values.size() / (distribution.values.size() + linearComponent - distribution.linearComponent);
-    double selfPart = linearComponent / (distribution.values.size() + linearComponent - distribution.linearComponent);
-    double distributionPart = -distribution.linearComponent / (distribution.values.size() + linearComponent - distribution.linearComponent);
-    
-    vector <pair<double, double> > mevPoints;
-    vector <pair<double, double> > gevPoints;
-    
-    double bestSelfPoint, bestDistributionPoint, bestValue;
-    for (int i = 0; i < values.size(); i++) {
-        bool selfInProgress = (values[i] >= start) && (values[i] <= end);
-        bool distributionInProgress = (values[i] >= distribution.start) && (values[i] <= distribution.end);
-                
[…]

### 2014-03-07 Paper/gammaRays.tex
change TIME_EXTENSION_FACTOR from 1.3 to 2.0
KS-test drafted

@@
 
-\begin{figure}[modelOverview]
+\begin{figure}
         \centering
@@
 
-\begin{figure}[sampleLightCurve]
+\begin{figure}
 	\centering
@@
 
-\begin{table}[bursts]
+\begin{table}
 	\centering
@@
 We compute two background estimates: for low and high energy bands.
+Because the background of transient class photons may depend on the spacecraft position, we have to filter them as well.
 
 \subsection{KS-test}
+
+We take care of remaining background by using the following functions with the Kolmogorov-Smirnov test:
+\begin{equation}
+	\Phi_i\left(t\right) = \frac{p_i\left(T_1, t\right) - b_i \frac{t-T_1}{T_2-T_1}}{p_i\left(T_1, T_2\right) - b_i}
+\end{equation}
+Here $\left(T_1, T_2\right)$ is the time range of observations, $p_i\left(t_1, t_2\right)$ is the number of photons observed in the time range $t_1$ to $t_2$ in the $i$'s energy band ($i=\text{low},\text{high}$ are for the low and high energy bands respectfully), and $b_i$ is the estimated number of background photons for the time range $T_1$ to $T_2$ in the $i$'s energy band.
+
+The numbers of degrees of freedom (used as an input for the KS-test) are $p_i\left(T_1, T_2\right) - b_i$.
+
+While we did not prove it, we assume that the Kolmogorov-Smirnov statistics computed for these functions with provided number of degrees of freedom has a similar distribution to that of the KS-statistics for ordinary CDFs (that is CDFs with no background).
+
+Finally, we compute KS-probabilities for pairs $\Phi_\text{low}\left(t\right), \Phi_\text{high}\left(\kappa t\right)$ to obtain the allowed ranges for stretching factors $\kappa$.
+We evaluate stretching factors in the range $\texttt{STRETCHING\_MIN} = 0.1$ to $\texttt{STRETCHING\_MAX} = 10.$ with a logarithmic step of $\texttt{STRETCHING\_STEP} = 1.001$.
+
+As a result of such calculation we can constraint the range of allowed stretching factors to values for which the KS-probabilities are not too small, in other words, given the sigma-significance value, compute the range outside of which stretching factors are elliminated by observations.
+
+You can see the results of such computation in the following section.
+
 \subsection{Results}

### 2014-03-07 GRObservations/GRObservations/GRBurst.cpp

@@ -98,14 +98,13 @@ void GRBurst::calculateBackground() {
         }
     }
         
-    mevDistribution.linearComponent = - mevExpectationValue + gevExpectationValue;
+    mevDistribution.linearComponent = -mevExpectationValue;
     mevDistribution.start = query.startTime - (time + startOffset);
     mevDistribution.end = query.endTime - (time + startOffset);
     
-    gevDistribution.linearComponent = 0.;
+    gevDistribution.linearComponent = -gevExpectationValue;
     gevDistribution.start = query.startTime - (time + startOffset);
     gevDistribution.end = query.endTime - (time + startOffset);
-    
 }
 
 void GRBurst::read() {
@@ -147,7 +146,7 @@ void GRBurst::read() {
     sort(mevBackgroundPhotons.begin(), mevBackgroundPhotons.end());
     sort(gevBackgroundPhotons.begin(), gevBackgroundPhotons.end());
     
-    cout << "event counts: " << mevPhotons.size() + gevPhotons.size() << " " << mevBackgroundPhotons.size() + gevBackgroundPhotons.size() << endl;
+    cout << "event counts: signal {" << mevPhotons.size() << ", " << gevPhotons.size() << "} background {" << mevBackgroundPhotons.size() << ", " << gevBackgroundPhotons.size() << "}" << endl;
     
     calculateBackground();
     

### 2014-03-07 GRObservations/GRObservations/GRDistribution.cpp

@@ -13,7 +13,7 @@
 
 #include "GRDistribution.h"
 
-double GRDistribution::kolmogorovSmirnovProbability(double distance, int n1, int n2) {
+double GRDistribution::kolmogorovSmirnovProbability(double distance, double n1, double n2) {
     double n = (double)n1*n2/(n1+n2);
     double z = (sqrt(n) + 0.12 + 0.11/sqrt(n)) * distance;
     
@@ -27,7 +27,7 @@ double GRDistribution::kolmogorovSmirnovProbability(double distance, int n1, int
     }
 }
 
-double GRDistribution::kolmogorovSmirnovDistance(double probability, int n1, int n2) {
+double GRDistribution::kolmogorovSmirnovDistance(double probability, double n1, double n2) {
     double y,logy,yp,x,xp,f,ff,u,t;
     double z;
     if (probability == 1.) z = 0.;
@@ -170,5 +170,5 @@ float GRDistribution::kolmogorovSmirnovTest(GRDistribution distribution, double 
     if (thisValue != NULL) *thisValue = maxDistanceRange.first;
     if (distributionValue != NULL) *distributionValue = maxDistanceRange.second;
     
-    return kolmogorovSmirnovProbability(maxDistance, (int)values.size(), (int)distribution.values.size());
+    return kolmogorovSmirnovProbability(maxDistance, (double)values.size() + linearComponent, (double)distribution.values.size() + distribution.linearComponent);
 }

### 2014-03-07 GRObservations/GRObservations/GRDistribution.h

@@ -28,8 +28,8 @@ public:
     double end;
     
 private:
-    double kolmogorovSmirnovProbability(double distance, int n1, int n2);
-    double kolmogorovSmirnovDistance(double probability, int n1, int n2);
+    double kolmogorovSmirnovProbability(double distance, double n1, double n2);
+    double kolmogorovSmirnovDistance(double probability, double n1, double n2);
 public:
     pair<double, double> cdfValueRange(double time);
     vector <struct GRDistributionCDFPoint> cdf();

### 2014-03-07 GRObservations/GRObservations/main.cpp

@@ -105,9 +105,11 @@ int main(int argc, const char * argv[])
             continue;
         }
         
+        log << "\tBackground to signal ratio = " << - burstCatalog[i].mevDistribution.linearComponent / burstCatalog[i].mevPhotons.size() << endl;
+        
         burstCatalog[i].evaluate();
         
-        log << "\tprobability not to be stretched = " << sigma(burstCatalog[i].trivialProbability) << endl;
+        log << "\tProbability not to be stretched = " << sigma(burstCatalog[i].trivialProbability) << endl;
         for (int k = 0; k < 5; k++) {
             log << "\t" << k+1 << "σ -> stretching factor in ranges (" << burstCatalog[i].minStretching[k] << ", " << burstCatalog[i].maxStretching[k] << ")" << endl;
         }

### 2014-03-10 Paper/gammaRays.tex
Added / Redesigned some plots

@@
 In this paper we consider all the bursts observed by Fermi-LAT since 2008 August 4 to 2011 August 1, for which at least $10$ events with energies $\unit[1]{GeV}$ or higher were observed.
-There are $3$ of them: GRB080916C, GRB090902B, and GRB090926A.
+There are $4$ of them: 080916C, 090510, 090902B, and 090926A.
 We study their light curves in two bands, $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.
 The Kolmogorov-Smirnov test is used to check whether the light curves for these two bands are the same.
-No significant difference was found for GRB080916C and GRB090902B.
-However, we observed with statistical significance of $3.4 \sigma$, that the higher energy light curve of GRB090926A is stretched with respect to the lower-energy one.
+No significant difference was found for 080916C and 090510.
+However, we observed with statistical significance of $3.3 \sigma$, that the higher energy light curve of 090926A is stretched with respect to the lower-energy one, and with statistical significance of $2.2 \sigma$, that the lower energy light curve of 090902B is stretched with respect to the higher-energy one.
 
@@
 \subsection{Results}
+
+\begin{figure}
+        \centering
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{lightCurve080916C}
+                \label{fig:lightCurve080916C}
+        \end{subfigure}
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{probabilities080916C}
+                \label{fig:probabilities080916C}
+        \end{subfigure}
+        \caption{080916C}
+        \label{fig:grb080916C}
+\end{figure}
+
+\begin{figure}
+        \centering
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{lightCurve090510}
+                \label{fig:lightCurve090510}
+        \end{subfigure}
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{probabilities090510}
+                \label{fig:probabilities090510}
+        \end{subfigure}
+        \caption{090510}
+        \label{fig:grb090510}
+\end{figure}
+
+\begin{figure}
+        \centering
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{lightCurve090902B}
+                \label{fig:lightCurve090902B}
+        \end{subfigure}
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{probabilities090902B}
+                \label{fig:probabilities090902B}
+        \end{subfigure}
+        \caption{090902B}
+        \label{fig:grb090902B}
+\end{figure}
+
+\begin{figure}
+        \centering
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{lightCurve090926A}
+                \label{fig:lightCurve090926A}
+        \end{subfigure}
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{probabilities090926A}
+                \label{fig:probabilities090926A}
+        \end{subfigure}
+        \caption{090926A}
+        \label{fig:grb090926A}
+\end{figure}
 
@@
 
-\begin{figure}[kappaDistribution]
+\begin{figure}
         \centering
-        \begin{subfigure}[kappaDistributionHistogram]{0.49\textwidth}
+        \begin{subfigure}{0.49\textwidth}
                 \includegraphics[width=\textwidth]{kappaDistributionHistogram}
@@
         \end{subfigure}
-        \begin{subfigure}[kappaDistributionCDF]{0.49\textwidth}
+        \begin{subfigure}{0.49\textwidth}
                 \includegraphics[width=\textwidth]{kappaDistributionCDF}
@@
 The sample contains $600$ points.
-The red curve is an empirical CDF of the observed sample (3 points).}
+The gray curve is an empirical CDF of the observed sample (4 points).}
         \label{fig:kappaDistribution}

### 2014-03-15 Paper/gammaRays.tex
Drafted table with observation results

@@
 	\centering
-	\includegraphics[width=0.49\textwidth]{sampleLightCurve}
+	\begin{subfigure}{0.49\textwidth}
+		\includegraphics[width=\textwidth]{sampleLightCurveLogNegative}
+		\label{fig:sampleLightCurveLogNegative}
+	\end{subfigure}
+	\begin{subfigure}{0.49\textwidth}
+		\includegraphics[width=\textwidth]{sampleLightCurveLogPositive}
+		\label{fig:sampleLightCurveLogPosivie}
+	\end{subfigure}
 	\caption{
@@
 Redshift $z = 1$.
-Observer's off-axis angle $\chi = 0.01$.
+Observer's off-axis angles are $\chi = 0$ for the left figure, and $\chi = 0.01$ for the right one.
 	}
-	\label{fig:sampleLightCurve}
+	\label{fig:sampleLightCurves}
 \end{figure}
 
-Finally, our assumption is able to explain the non-trivial stretching factor of the GRB 090926A.
+Finally, our assumption is able to explain the non-trivial stretching factor of both GRBs 090902B and 090926A.
 Detailed computations will be described in the following sections, but you can see the main idea on the figure \ref{fig:modelOverview}.
 The computations confirm that this qualitative picture is indeed correct.
-You can see a sample light curve on the figure \ref{fig:sampleLightCurve}, and predicted distribution of stretching factors on the figure \ref{fig:kappaDistribution}.
+You can see a sample light curves on the figure \ref{fig:sampleLightCurves}, and predicted distribution of stretching factors on the figure \ref{fig:kappaDistribution}.
 
@@
 
+Out of 19 bursts studied, only 4 have at least 10 high energy events remaining after filtering, and thus eligible for the computation of stretching factors.
+The results of this computation are shown on figures \ref{fig:grb080916C}, \ref{fig:grb090510}, \ref{fig:grb090902B}, \ref{fig:grb090926A}.
+
 \begin{figure}
@@
 \end{figure}
+
+\begin{table}
+	\centering
+	\small
+	\begin{tabular}{ l | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] }
+		\multicolumn{1}{ c |}{GRB} & \multicolumn{2}{ c |}{$1\sigma$} & \multicolumn{2}{ c |}{$2\sigma$} & \multicolumn{2}{ c |}{$3\sigma$} & \multicolumn{2}{ c |}{$4\sigma$} & \multicolumn{2}{ c }{$5\sigma$} \\
+		\hline
+		080825C	&	1.04  & 2.24	&	0.67 & 3.32	&	0.42 & 5.83	&	0.19 & 14.9		&	0.087 & 35.7	\\
+		090501	&	0.58  & 1.1		&	0.43 & 1.61	&	0.32 & 2.29	&	0.22 &  3.03	&	0.17  &  5.11	\\
+		090902B	&	0.629 & 0.635	&	0.36 & 0.89	&	0.23 & 1.57	&	0.13 &  2.87	&	0.084 &  4.56	\\
+		090926A &	2.61  & 3.33	&	1.99 & 6.62	&	1.34 & 9.15	&	0.73 & 13.5		&	0.48  & 19.4
+	\end{tabular}
+	\caption{Caption}
+	\label{tab:observationResults}
+\end{table}
+
+%\begin{table}
+%	\centering
+%	\small
+%	\begin{tabular}{ l | S[table-format=1.3] | S[table-format=1.3] | S[table-format=1.3] | S[table-format=1.3] | S[table-format=1.3]}
+%		\multicolumn{1}{ c |}{GRB Name} & $1\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
+%		%\multicolumn{1}{ c |}{Name}& $\mathrm{MET,\ sec}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{deg}$ & $\mathrm{sec}$ & $\mathrm{sec}$ \\
+%		\hline
+%		080916C	&	1.042	&	1.200	&	1.300	&	1.400	&	1.500
+%	\end{tabular}
+%	\caption{Burst data used in our study.
+The first row contains the corresponding \texttt{GRBurst} class variable names.
+The data is taken from the tables 2 and 4 of \cite{Ackermann:2013zfa}.
+\url{https://github.com/maxitg/GammaRays/blob/master/GRObservations/bursts}}
+%	\label{tab:observationResults}
+%\end{table}
 

### 2014-03-15 GRObservations/GRObservations/GRBurst.cpp

@@ -17,9 +17,9 @@
 #include "GRBurst.h"
 #define TIME_EXTENSION_FACTOR   2.0
 #define PHOTONS_QUALITY         0.95
-#define STRETCHING_MIN          0.1
-#define STRETCHING_MAX          10.
-#define STRETCHING_STEP         1.001
+#define STRETCHING_MIN          0.01
+#define STRETCHING_MAX          100.
+#define STRETCHING_STEP         1.0001
 
 #define BACKGROUND_DURATION     86400
 

### 2014-03-23 Paper/gammaRays.tex
Drafted observation results

@@
 Out of 19 bursts studied, only 4 have at least 10 high energy events remaining after filtering, and thus eligible for the computation of stretching factors.
-The results of this computation are shown on figures \ref{fig:grb080916C}, \ref{fig:grb090510}, \ref{fig:grb090902B}, \ref{fig:grb090926A}.
+The results of this computation are shown on figures \ref{fig:grb080916C}, \ref{fig:grb090510}, \ref{fig:grb090902B}, \ref{fig:grb090926A} and table \ref{tab:observationResults}.
+
+Out of these 4 bursts, two (080916C and 090510) have stretching factors compatible with $\kappa = 1$ within $2\sigma$ range.
+
+GRB 090926A has, however, high energy radiation stretched with respect to low energy signal (that is $\kappa > 1$).
+In contrast to that, GRB 090902B has low energy radiation stretched with respect to high energy signal ($\kappa < 1$).
+
+So we obtained a preliminary (for significance is less than $5\sigma$) result that the stretching factors for observable bursts might be both larger and smaller than $1$.
 
@@
         \end{subfigure}
-        \caption{080916C}
+        \caption{GRB 080916C results. Stretching factor is compatible with $\kappa = 1$ within $2\sigma$.}
         \label{fig:grb080916C}
@@
         \end{subfigure}
-        \caption{090510}
+        \caption{GRB 090510 results. Stretching factor is compatible with $\kappa = 1$ within $1\sigma$.}
         \label{fig:grb090510}
@@
         \end{subfigure}
-        \caption{090902B}
+        \caption{GRB 090902B results. Low energy radiation is stretched ($\kappa < 1$) with significance of $2\sigma$.}
         \label{fig:grb090902B}
@@
         \end{subfigure}
-        \caption{090926A}
+        \caption{GRB 090926A results. High energy radiation is stretched ($\kappa > 1$) with significance of $3\sigma$.}
         \label{fig:grb090926A}
@@
 	\end{tabular}
-	\caption{Caption}
+	\caption{Burst observation results. Ranges of allowed stretching factors are shown for studied bursts for multiple levels of significance.}
 	\label{tab:observationResults}

### 2014-03-23 Paper/gammaRays.tex
Removed model results from model description

@@
 
-%\begin{table}
-%	\centering
-%	\small
-%	\begin{tabular}{ l | S[table-format=1.3] | S[table-format=1.3] | S[table-format=1.3] | S[table-format=1.3] | S[table-format=1.3]}
-%		\multicolumn{1}{ c |}{GRB Name} & $1\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
-%		%\multicolumn{1}{ c |}{Name}& $\mathrm{MET,\ sec}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{deg}$ & $\mathrm{sec}$ & $\mathrm{sec}$ \\
-%		\hline
-%		080916C	&	1.042	&	1.200	&	1.300	&	1.400	&	1.500
-%	\end{tabular}
-%	\caption{Burst data used in our study.
-The first row contains the corresponding \texttt{GRBurst} class variable names.
-The data is taken from the tables 2 and 4 of \cite{Ackermann:2013zfa}.
-\url{https://github.com/maxitg/GammaRays/blob/master/GRObservations/bursts}}
-%	\label{tab:observationResults}
-%\end{table}
-
 \section{Model}
@@
 \begin{itemize}
-\item{$v$, the velocity of the shell, $\gamma = \frac{1}{\sqrt{1-v^2}} \gg 1$.
-
-	If $\gamma$ is small enough, high photon densities would make the jet plasma optically thick to pair production.
-This in turn would suppress emitted flux, and this is not observed.
-From this argument the lower bound of $\gamma \ge 100$ was derived \cite{}.
-The limited total energy of the burst provides an upper bound: $\gamma \le \gamma_\text{max}$ (see section 3.6).
-Finally, the model dependent computations of $\gamma$ give the values ranging from $50$ for low-energy bursts to $1000$ for high-energy ones \cite{Ghirlanda:2011ux,Ghisellini:2012nc}.
-$\gamma$ may vary from burst to burst, and it should decrease during the burst due to the dissipation of energy.
-As a first approximation we select {\boldmath$\gamma = 100$} for this study.
-}
-\item{$\eta_0$, which defines the luminosity scale.
-
-	$\eta_0$ controls the number of photons we see from the bursts, and so can be roughly estimated from the number of luminous bursts we observe.
-We can use this and the measured burst redshifts to estimate both $\eta_0$ and the bursts density $\rho_0$, which will be discussed later.
-For this study we choose {\boldmath$\eta_0 = 1 \times 10^{45}$}.
-}
+\item{$v$, the velocity of the shell, $\gamma = \frac{1}{\sqrt{1-v^2}} \gg 1$.}
+\item{$\eta_0$, which defines the luminosity scale.}
 \item{$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
@@
 
+With this algorithm, we arrive to our second test / constraint for the burst parameters. The obtained stretching factors distribution should be compatible with observations.
+
+\subsection{High energy bursts fraction}
+
+Out final test is the comparison of answers to this question: given the bursts which were observed in a whole energy range $\left(\omega_1,\omega_3\right)$, which fraction of them can also be observed in a high energy range $\left(\omega_2,\omega_3\right)$?
+
+To calculate this value, we need to compute the number of bursts visible in a given energy range, which is the integral over space and jet directions:
+\begin{align}
+b\left(\omega_1,\omega_2\right) &= \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} 2\pi \sin\chi \dif\chi \nonumber\\
+&\approx 2\pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} \chi \dif\chi = \pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_2\right)
+\end{align}
+Here $z_\text{max}\left(\omega_1,\omega_2\right)$ and $\chi_\text{max}\left(z,\omega_1,\omega_2\right)$ are the same values we used in the previous section: the maximum redshift from which a burst can be observed in a given energy range, and the maximum off-axis angle with which the burst at redshift $z$ can be observed.
+
+The fraction we want to compute is the ratio of these integrals:
+\begin{equation}
+f\left(\omega_1,\omega_2,\omega_3\right) = \frac{b\left(\omega_2,\omega_3\right)}{b\left(\omega_1,\omega_3\right)} = \frac{\int_0^{z_\text{max}\left(\omega_2,\omega_3\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_2,\omega_3\right)}{\int_0^{z_\text{max}\left(\omega_1,\omega_3\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_3\right)}
+\end{equation}
+
+This is our 3rd test -- the ratio obtained should be compatible with the observed value.
+
+\subsection{Parameter values}
+\label{sec:parameters}
+
+Finally, we are in a position to use the tests to fit parameters of the model:
+
+\begin{itemize}
+	\item{
+		$v$, the velocity of the shell, $\gamma = \frac{1}{\sqrt{1-v^2}} \gg 1$.
+
+	}
+	\item{$\eta_0$, which defines the luminosity scale.}
+	\item{$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
+	\item{$n$, which determines the sharpness of the jet end, $n > 3$;}
+	\item{$\omega_0$, a characteristic radiation frequency;}
+	\item{$\theta_0$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;}
+	\item{$k$, which determines how much the opening angle changes with frequency, $k < 0$;}
+	\item{$\alpha$, the bare spectral index, $\alpha < -2k - 1$}
+\end{itemize}
+
+\subsection{Results}
+
 \begin{figure}
@@
 
-Our set of parameters led to a sample you can see on the figure \ref{fig:kappaDistribution}.
-However, the error margins of the observed stretching factors are too large to use the KS-test on them.
-Nevertheless, it appears that our model doesn't contradict existing observations.
-
-\subsection{Very high energy bursts fraction}
-
-Out final test is the comparison of answers to this question: given the bursts which were observed in a high energy range $\left(\omega_1,\omega_3\right)$, which fraction of them can also be observed in a very high energy range $\left(\omega_2,\omega_3\right)$?
-
-To calculate this value, we need to compute the number of bursts visible in a given energy range, which is the integral over space and jet directions:
-\begin{align}
-b\left(\omega_1,\omega_2\right) &= \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} 2\pi \sin\chi \dif\chi \nonumber\\
-&\approx 2\pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} \chi \dif\chi = \pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_2\right)
-\end{align}
-Here $z_\text{max}\left(\omega_1,\omega_2\right)$ and $\chi_\text{max}\left(z,\omega_1,\omega_2\right)$ are the same values we used in the previous section: the maximum redshift from which a burst can be observed in a given energy range, and the maximum off-axis angle with which the burst at redshift $z$ can be observed.
-
-The fraction we want to compute is the ratio of these integrals:
-\begin{equation}
-f\left(\omega_1,\omega_2,\omega_3\right) = \frac{b\left(\omega_2,\omega_3\right)}{b\left(\omega_1,\omega_3\right)} = \frac{\int_0^{z_\text{max}\left(\omega_2,\omega_3\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_2,\omega_3\right)}{\int_0^{z_\text{max}\left(\omega_1,\omega_3\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_3\right)}
-\end{equation}
-
-The ratio for our set of parameters is $f_m = 0.072$.
-The LAT burst catalog \cite{Ackermann:2013zfa} contains $n = 35$ bursts, $3$ of which were observed in the energy range $\left(\unit[1]{GeV}, \infty\right)$.
-So the ratio is $f_o = \frac{3}{35} \approx 0.086$.
-The binary distribution gives us the probability for it to happen:
-\begin{equation}
-\text{Pr}\left(f_o = \frac{k}{n}\right) = \binom{n}{k} f_m^k \left(1-f_m\right)^{n-k} \approx 0.22
-\end{equation}
-which means that the model doesn't contradict observations in this test.
-
-\subsection{Parameter values}
-\label{sec:parameters}
-
-\subsection{Results}
-
 \section{Discussion}

### 2014-03-25 Paper/gammaRays.tex
Fixed typo in table 2
Added reference to Castignani:2014gaa
Updated parameter values

@@
 The spectral lags were found between different low energy bands \cite{Yi:2005ht}.
-The very high energy radiation was discovered to be extended relative to x-ray emission \cite{Vianello:2013ela,Lange:2013uh}.
+The very high energy radiation was discovered to be extended relative to x-ray emission \cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}.
 Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
@@
 		\hline
-		080825C	&	1.04  & 2.24	&	0.67 & 3.32	&	0.42 & 5.83	&	0.19 & 14.9		&	0.087 & 35.7	\\
+		080916C	&	1.04  & 2.24	&	0.67 & 3.32	&	0.42 & 5.83	&	0.19 & 14.9		&	0.087 & 35.7	\\
 		090501	&	0.58  & 1.1		&	0.43 & 1.61	&	0.32 & 2.29	&	0.22 &  3.03	&	0.17  &  5.11	\\
@@
 \end{itemize}
+
+Numerical values we used for these parameters are summarized in the table \ref{tab:parameters}.
+
+\begin{table}
+	\centering
+	\small
+	\begin{tabular}{ r | l | l }
+		Parameter	&	Value			&	Comments \\
+		\hline
+		$\gamma$		&	$100$		&	\\
+		$\eta_0$		&	$10^{45}$	&
+	\end{tabular}
+	\caption{Parameter values.}
+	\label{tab:parameters}
+\end{table}
 
@@
 
-Finally, we are in a position to use the tests to fit parameters of the model:
+Finally, we can use the tests to fit parameters of the model:
 
@@
 
+		Observations suggest values for relativistic factors of plasma is a range from $\gamma = 100$ to $\gamma = 1000$.
+		We take $\gamma = 300$ as an order of magnitude estimate.
 	}
-	\item{$\eta_0$, which defines the luminosity scale.}
-	\item{$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
-	\item{$n$, which determines the sharpness of the jet end, $n > 3$;}
-	\item{$\omega_0$, a characteristic radiation frequency;}
-	\item{$\theta_0$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;}
-	\item{$k$, which determines how much the opening angle changes with frequency, $k < 0$;}
-	\item{$\alpha$, the bare spectral index, $\alpha < -2k - 1$}
+	\item{
+		$\eta_0$, which defines the luminosity scale.
+
+		From photon counts.
+	}
+	\item{
+		$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;
+
+		From durations.
+	}
+	\item{
+		$n$, which determines the sharpness of the jet end, $n > 3$;
+
+	}
+	\item{
+		$\omega_0$, a characteristic radiation frequency;
+
+		Non-physical
+	}
+	\item{
+		$\theta_0$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;
+
+	}
+	\item{
+		$k$, which determines how much the opening angle changes with frequency, $k < 0$;
+
+	}
+	\item{
+		$\alpha$, the bare spectral index, $\alpha < -2k - 1$
+
+	}
 \end{itemize}

### 2014-05-18 Paper/gammaRays.tex
Changed p to very slow but accurate version

@@
 \usepackage{listings}
+\usepackage{multicol}
 
@@
 		High and low energy light curves produced by the geometrical model.
-Parameter values are the same as discussed in the section \ref{sec:parameters}.
+Burst parameter values are the same as discussed in the section \ref{sec:parameters}.
 Redshift $z = 1$.
@@
 \begin{align}
+\label{eq:p}
 p\left( z, \chi; \tau_1, \tau_2; \omega_1, \omega_2 \right) &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)}
@@
 \subsection{Distribution of stretching factors}
+\label{sec:distribution}
 
@@
 	$z_\text{max}$ is defined by the farthest jet, which can be observed.
-Since the burst observability $p_\infty\left(z,\chi,\omega_2,\omega_3\right)$ decreases with both redshift and $\chi$, we can find the maximum redshift by solving the following equation:
-	\begin{equation}
-	p_\infty\left(z_\text{max},0,\omega_2,\omega_3\right) = p_\text{min}
+Since the burst observability $p_\infty\left(z,\chi;\omega_2,\omega_3\right)$ decreases with both redshift and $\chi$, we can find the maximum redshift by solving the following equation:
+	\begin{equation}
+	p_\infty\left(z_\text{max},0;\omega_2,\omega_3\right) = p_\text{min}
 	\end{equation}
@@
 	\begin{equation}
-	p_\infty\left(z_\text{min},\chi_\text{max},\omega_2,\omega_3\right) = p_\text{min}
+	p_\infty\left(z_\text{min},\chi_\text{max};\omega_2,\omega_3\right) = p_\text{min}
 	\end{equation}
@@
 }
-\item{Now we should check, if the burst in a selected position can be observed: $p_\infty\left(z,\chi,\omega_2,\omega_3\right) > p_\text{min}$.
+\item{Now we should check, if the burst in a selected position can be observed: $p_\infty\left(z,\chi;\omega_2,\omega_3\right) > p_\text{min}$.
 If it is, add $\kappa\left(z,\chi\right)$ to the sample.
@@
 
-\subsection{Parameter values}
+\subsection{Parameter fit}
 \label{sec:parameters}
 
-Finally, we can use the tests to fit parameters of the model:
-
+For now we discussed how to compute various observables of the model, and how to test the model against the data from observations.
+However, we never considered specific values for parameters of the burst.
+We will do that in this section.
+
+First of all, we need to determine, which parameters to fit.
+At most, the observables of a particular burst depend on 10 parameters: 8 of the burst $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k, \alpha\right)$, and 2 of the observer $\left(z, \chi\right)$.
+However, some of these parameters only change observables trivially, and, also, there are transformations of variables, which will not change observables at all (see equation \ref{eq:p}):
 \begin{itemize}
-	\item{
-		$v$, the velocity of the shell, $\gamma = \frac{1}{\sqrt{1-v^2}} \gg 1$.
-
-		Observations suggest values for relativistic factors of plasma is a range from $\gamma = 100$ to $\gamma = 1000$.
-		We take $\gamma = 300$ as an order of magnitude estimate.
-	}
-	\item{
-		$\eta_0$, which defines the luminosity scale.
-
-		From photon counts.
-	}
-	\item{
-		$r_0$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;
-
-		From durations.
-	}
-	\item{
-		$n$, which determines the sharpness of the jet end, $n > 3$;
-
-	}
-	\item{
-		$\omega_0$, a characteristic radiation frequency;
-
-		Non-physical
-	}
-	\item{
-		$\theta_0$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;
-
-	}
-	\item{
-		$k$, which determines how much the opening angle changes with frequency, $k < 0$;
-
-	}
-	\item{
-		$\alpha$, the bare spectral index, $\alpha < -2k - 1$
-
-	}
+	\item{The number of photons $p$ depends linearly on $\eta_0$, so $\eta_0$ doesn't affect other observables like duration, or stretching factor.
+We can easily fit $\eta_0$ to match the observed photon count.}
+	\item{If we make a simultaneous transformation of $r_0 \rightarrow \lambda r_0$ and $\eta_0 \rightarrow \frac{1}{\lambda^3}\eta_0$, where $\lambda$ is a parameter of transformation, then only the duration of the burst will change.
+Other observables, like the total number of photons, or stretching factor, will not be affected.}
+	\item{Finally, if we make a following transformation: $\omega_0 \rightarrow \lambda \omega_0$, $\theta_0 \rightarrow \lambda^k \theta_0$, $\eta_0 \rightarrow \lambda^\alpha \eta_0$, where $\lambda$ is another transformation parameter, no observables will change whatsoever.}
 \end{itemize}
+Given these transformations, we can reduce the number of parameters for fit to 7: $\left(\gamma, n, \theta_0, k, \alpha, z, \chi\right)$.
+
+Also, some of the parameters we know from observations:
+\begin{itemize}
+	\item{Redshifts $z$ of many bursts have been measured \cite{Ackermann:2013zfa}.}
+	\item{We know that Lorentz factors $\gamma$ are on the order of magnitude\footnote{Our minimization procedure can find similar minimums for other values of $\gamma$, at least from $100$ to $1000$} of $\gamma = 300$ \cite{Ghirlanda:2011ux}.}
+\end{itemize}
+This knowledge allows us to reduce the number of parameters to minimize against to just 5: $\left(n, \theta_0, k, \alpha, \chi\right)$.
+
+We want then our cost function $C\left(n, \theta_0, k, \alpha, \chi\right)$ to satisfy the following objectives:
+\begin{itemize}
+	\item {
+		Total energy of the burst is finite (that is $k+\frac{\alpha}{2}+1 < 0$) and smaller than $\unit[6 \times 10^{53}]{GeV} \approx \unit[10^{51}]{erg}$ \cite{Gehrels:2013xd}.
+	}
+	\item {
+		The stretching factor of the burst should be compatible with the value for the GRB 090926A.
+	}
+	\item {
+		We require all bursts to have the same burst parameters $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k, \alpha\right)$.
+And we know that bursts with small stretching factors like GRB 090902B should be possible to observe.
+So we require that bursts with the stretching factor of GRB 090902B or lower appear in random samples by $z$ and $\chi$ (sampling is done the same way as in section \ref{sec:distribution}).
+	}
+	\item {
+		The ratio of photons with high energies and low energies should be compatible with the value for the GRB 090926A, $\frac{p_{\infty} \left(z,\chi;\unit[1]{GeV},\infty\right)}{p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV}\right)} = 0.06$.
+	}
+	\item {
+		Finally, the burst should not be too faint compared to other bursts from the sample. That is the total number of observed photons should approximately equal to the median among the sample.
+	}
+\end{itemize}
+
+The cost function is then computed by the following procedure:
+\begin{enumerate}
+	\item{
+		Set $\gamma = 300$, $\omega_0 = 1$, $z = 2.1062$ (which is the redshift of GRB 090926A)
+	}
+	\item{
+		Set $\eta_0$ and $r_0$ such that the duration of the burst, and the total number of observed photons are compatible with GRB 090926A, $T_{0.99}\left(z,\chi;\unit[0.1]{GeV},\infty\right) = 219.5$ and $p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\infty\right) = 179.989$
+	}
+	\item{
+		If $k+\frac{\alpha}{2}+1 < 0$, or the energy of the burst $E\left(\unit[0.1]{GeV}\right) > \unit[6 \times 10^{53}]{GeV}$, the cost function equals to the penalization factor: $C\left(n, \theta_0, k, \alpha, \chi\right) = 400$.
+	}
+	\item{
+		\label{item:sample}
+		Compute the small sample of 10 bursts (we will need stretching factors and total photon counts) with the same fixed burst parameters $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k, \alpha\right)$, and with $z$ and $\chi$ representatively distributed (as discussed in section \ref{sec:distribution}).
+	}
+	\item{
+		Compute the cost due to the stretching factor:
+		\begin{equation}
+			C_\kappa = \frac{
+				\log{\kappa\left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV},\infty\right)} - \log{\kappa_{090926A}}
+			}{\Delta\log\kappa_{090926A}}
+		\end{equation}
+		Here $\log\kappa_{090926A} = \frac{1}{2}\left(\log\left(6.62\right) + \log\left(1.99\right)\right)$ and $\Delta\log\kappa_{090926A} = \frac{1}{2}\left(\log\left(6.62\right) - \log\left(1.99\right)\right)$.
+	}
+	\item{
+		Compute the cost due to the minimal stretching factor from the sample:
+		\begin{equation}
+			C_{\kappa\text{min}} = \max\left(0, 
+				\frac{
+					\log{\kappa_\text{min}} - \log{\kappa_{090902B}}
+				}{\Delta\log\kappa_{090902B}}\right)
+		\end{equation}
+		Here $\kappa_\text{min}$ is the minimal stretching factor from the sample computed in the step \ref{item:sample}, $\log\kappa_{090902B} = \frac{1}{2}\left(\log\left(0.89\right) + \log\left(0.36\right)\right)$ and $\Delta\log\kappa_{090902B} = \frac{1}{2}\left(\log\left(0.89\right) - \log\left(0.36\right)\right)$.
+	}
+	\item{
+		Compute the cost due to the fraction of high and low energy photon counts:
+		\begin{equation}
+			C_f = \frac{
+				\log{\frac{
+					p_{\infty} \left(z,\chi;\unit[1]{GeV},\infty\right)
+				}{
+					p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV}\right)}
+				} - \log f_{090926A}
+			}{\log(10)}
+		\end{equation}
+		Here $f_{090926A} = 0.0606777$.
+	}
+	\item{
+		Compute the cost due to the brightness of the burst compared to the median:
+		\begin{equation}
+			C_b = \frac{
+				\log{\frac{
+					p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\infty\right)
+				}{
+					p_\text{med}
+				}} - 0
+			}{\log(10)}
+		\end{equation}
+		Here $p_\text{med}$ is the median number of observed photons among the sample computed in the step \ref{item:sample}.
+	}
+	\item{
+		Finally, the value of the cost function is the sum of squares of the four:
+		\begin{equation}
+			C = C_\kappa^2 + C_{\kappa\text{min}}^2 + C_f^2 + C_b^2
+		\end{equation}
+	}
+\end{enumerate}
+
+\begin{table}
+	\centering
+	\small
+	\begin{tabular}{ S[table-format=2.4] | S[table-format=1.5] | S[table-format=-1.6] | S[table-format=-1.7] | S[table-format=1.8] }
+		$\mathrm{n}$ & $\mathrm{\theta_0}$ & $\mathrm{k}$ & $\mathrm{\alpha}$ & $\mathrm{\chi}$ \\
+		\hline
+		22.1525	&	$\mathrm{2.16424 \times 10^{-4}}$ 	&	-0.417021  	&	-1.3516300 	&	0.00704982 	\\
+		25.8640	&	$\mathrm{4.90222 \times 10^{-8}}$	&	-2.908750 	&	 3.5380100 	&	0.00028657 	\\
+		17.5380	&	$\mathrm{1.46284 \times 10^{-3}}$ 	&	-2.133600 	&	-2.1336000 	&	0.00152624 	\\
+		17.9736	&	$\mathrm{4.52631 \times 10^{-6}}$	&	-1.402620 	&	-0.0114132 	&	0.00032590 	\\
+		 5.0000	&	$\mathrm{1.12535 \times 10^{-7}}$	&	-0.200000	&	-2.0000000 	&	0.00473795 	\\
+		 7.0000	&	$\mathrm{2.00000 \times 10^{-12}}$	&	-3.000000	&	-3.0000000	&	0.00173795
+	\end{tabular}
+	\caption{}
+\end{table}
+
+We use the Nelder-Mead (downhill simplex) method for the minimization procedure.
+Initial points were chosen to include the areas in parameter space where each cost is close to $0$ (first $4$ points correspondingly), and to cover the large fraction of our parameter space:
+\begin{itemize}
+	\item{$n = 22.1525, \theta_0 = 0.000216424, k = -0.417021, \alpha = -1.35163, \chi = 0.00704982$}
+	\item{$n = 25.864, \theta_0 = 4.90222 \times 10^{-8}, k = -2.90875, \alpha = 3.53801, \chi = 0.00028657$}
+	\item{$n = 17.538, \theta_0 = 0.00146284, k = -0.20793, \alpha = -2.1336, \chi = 0.00152624$}
+	\item{$n = 17.9736, \theta_0 = 4.52631 \times 10^{-6}, k = -1.40262, \alpha = -0.0114132, \chi = 0.0003259$}
+	\item{$n = 5, \theta_0 = 1.12535 \times 10^{-7}, k = -0.2, \alpha = -2.0, \chi = 0.00473795$}
+	\item{$n = 7, \theta_0 = 2 \times 10^{-12}, k = -3, \alpha = -3, \chi = 0.00173795$}
+\end{itemize}
+
+The minimization then converged to the following parameter values:
+\begin{multicols}{2}
+\begin{itemize}
+		\item{p}
+		\item{p}
+		\item{p}
+		\item{p}
+		\item{p}
+		\item{p}
+		\item{p}
+		\item{p}
+		\item{p}
+		\item{p}
+\end{itemize}
+\end{multicols}
 

### 2014-05-21 Paper/gammaRays.tex
Plot \[Kappa]CDFPlot from sample16IA

@@
 \begin{figure}
-        \centering
-        \begin{subfigure}{0.49\textwidth}
+	\hspace*{\fill}
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width=\textwidth]{kappaDistributionHistogram}
 		\label{fig:kappaDistributionHistogram}
-        \end{subfigure}
-        \begin{subfigure}{0.49\textwidth}
+		\caption{Histogram (blue) obtained from the model. Bar lines (black) show observed stretching factors.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width=\textwidth]{kappaDistributionCDF}
 		\label{fig:kappaDistributionCDF}
-        \end{subfigure}
-        \caption{The histogram and the CDF of the stretching factors sample produced by our model (blue).
-The sample contains $600$ points.
-The gray curve is an empirical CDF of the observed sample (4 points).}
+		\caption{CDF (blue) obtained from the model. Gray curve and bar lines show observed stretching factors.}
+	\end{subfigure}
+	\hspace*{\fill}
+	\caption{Stretching factors histogram and CDF produced by our model. The sample contains $16$ inaccurate bursts.}
 	\label{fig:kappaDistribution}

### 2014-05-21 ITEPProceedings/piskunov.tex
Setup ITEP proceedings styling

\documentclass[
aps,%
12pt,%
final,%
notitlepage,%
oneside,%
onecolumn,%
nobibnotes,%
nofootinbib,% 
superscriptaddress,%
showpacs,%
centertags]%
{revtex4}

\newcommand{\exclude}[1]{}

\begin{document}
\selectlanguage{english}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

%\newcommand{\lll}{\lambda}

% Some useful journal names

\newcommand{\A}{\alpha}
\newcommand{\B}{\beta}
\newcommand{\T}{\theta}
\newcommand{\Ep}{\epsilon}
\newcommand{\beqn}{\begin{eqnarray}}
\newcommand{\eeqn}{\end{eqnarray}}

\title{ANOMALIES (SELECTED TOPICS)}

\author{\firstname{V.~I.}~\surname{Zakharov}}
\email{maxitg@icloud.com}
\affiliation{Institute for Theoretical and Experimental
Physics, Moscow, Russia
}

%\date{\today}

\begin {abstract}
The presentation is
intended  to be mainly pedagogical and covers the chiral anomaly
and
its most significant applications. Finally, we  
comment
on recent issues,
related to the lattice measurements.

\end{abstract}

\pacs{11.10.Ef,~11.15.Bt,~11.15.Ha}

\maketitle

\section{Introduction}
For the purposes of this review,
by anomalies one can understand violations
of the Ehrenfest theorem according to which
all matrix elements of the
classical equations of motion (considered as operators) vanish.
Thus, anomaly is established
 if we find, for example,
a nonvanishing
matrix element of the Dirac equation for a charged particle:
\begin{equation}
\langle \bar{\psi}(\hat{D}+im)\psi\rangle~\neq~0,
\end{equation}
where $\hat{D}=\gamma_{\mu}D_{\mu}$ and $D_{\mu}$ is the covariant derivative
and $m$
is the mass of the charged particle.

..........................

\bibliographystyle{maik}
\bibliography{piskunov}

\newpage

\selectlanguage{russian}

\begin{center}
\large \bfseries %\MakeTextUppercase{%
АНОМАЛИИ (ИЗБРАННЫЕ ТЕМЫ)
\end{center}
%
\begin{center}
{\bfseries
В.~И.~Захаров
}
\end{center}
%
\begin{center}
\begin{minipage}{\textwidth - 2cm}
\small

Данный обзор имеет образовательный характер и посвящен
киральным аномалиям и их наиболее существенным приложениям.
Также рассматриваются современные вопросы решеточных измерений.
\end{minipage}
\end{center}
\selectlanguage{english}
%
\newpage
{\bf Figure captions}
\vskip 1cm

{\bf Fig.~1.}~Anomalous triangle graph. Momenta $q,k_1,k_2$ are
carried by the axial current $a_\mu^5$ and
two photons, respectively. The triangle corresponds to a fermion of small mass $m$.

{\bf Fig.~2.}~Imaginary part of the anomalous triangle graph.
The intermediate fermions are on
mass shell. (The signs are the same as in Fig. 1)

{\bf Fig.~3.}~Three-body intermediate state in the imaginary part. (The signs are the same as in Fig. 1)

{\bf Fig.~4.}~Radiative correction to the two-body contribution to the
imaginary part. (The signs are the same as in Fig. 1)

{\bf Fig.~5.}~Graphic representation of the Banks--Casher evaluation of the quark condensate
$\langle \bar{q}q\rangle$. Dashed lines with crosses represent external (vacuum) gluonic
fields.

\newpage

%\begin{figure}
%\includegraphics{z1.eps}
%\caption{Zakharov, Fig. 1}
%\end{figure}

%\begin{figure}
%\includegraphics{z2.eps}
%\caption{Zakharov, Fig. 2}
%\end{figure}

%\begin{figure}
%\includegraphics{z3.eps}
%\caption{Zakharov, Fig. 3}
%\end{figure}

%\begin{figure}
%\includegraphics{z4.eps}
%\caption{Zakharov, Fig. 4}
%\end{figure}

%\begin{figure}
%\includegraphics{z5.eps}
%\caption{Zakharov, Fig. 5}
%\end{figure}

\end{document}
% \bibitem{kniel}
% Two-Loop Electroweak Corrections to the A0 gamma gamma and A0 g g Couplings of the CP-Odd Higgs Boson.
% Joachim Brod, Frank Fugel, Bernd A. Kniehl
%  Nucl.Phys.B807:188-209,2009.
% e-Print: arXiv:0807.1008 [hep-ph]

### 2014-05-21 ITEPProceedings/piskunov.tex
Added title, abstract and PAC numbers to ITEP

@@
 {revtex4}
+
+\usepackage{units}
 
@@
 
-\title{ANOMALIES (SELECTED TOPICS)}
+\title{TIME STRETCHING OF THE GEV EMISSION OF GRBS: FERMI LAT DATA VS GEOMETRICAL MODEL}
 
-\author{\firstname{V.~I.}~\surname{Zakharov}}
+\author{\firstname{M.~S.}~\surname{Piskunov}}
 \email{maxitg@icloud.com}
-\affiliation{Institute for Theoretical and Experimental
-Physics, Moscow, Russia
-}
+\affiliation{Institute for Nuclear Research, Moscow, Russia}
 
@@
 \begin {abstract}
-The presentation is
-intended  to be mainly pedagogical and covers the chiral anomaly
-and
-its most significant applications. Finally, we  
-comment
-on recent issues,
-related to the lattice measurements.
+	Numerous observations confirm that the high energy $(> \unit[100]{MeV})$ emission of gamma ray bursts is delayed with respect to the low energy emission.
+	However, the difference of light curves in various high energy bands has not been studied properly.
 
+	In this paper we consider all the bursts observed by Fermi-LAT since 2008 August 4 to 2011 August 1, for which at least $10$ events with energies $\unit[1]{GeV}$ or higher were observed.
+	There are $4$ of them: 080916C, 090510, 090902B, and 090926A.
+	We study their light curves in two bands, $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.
+	The Kolmogorov-Smirnov test is used to check whether the light curves for these two bands are the same.
+	No significant difference was found for 080916C and 090510.
+	However, we observed with statistical significance of $3.3 \sigma$, that the higher energy light curve of 090926A is stretched with respect to the lower-energy one, and with statistical significance of $2.2 \sigma$, that the lower energy light curve of 090902B is stretched with respect to the higher-energy one.
+
+	We suggest a simple geometrical model to explain this result.
+	The main assumption is the jet opening angle dependence on radiation energy -- the most energetic photons are emitted near the axis of the jet.
+	To test this model, we compute the total energy of the burst, and confirm that it is below the constraint.
+	We also compute the fraction of observable bursts in $(\unit[100]{MeV}, \unit[1]{GeV})$ band, which can also be observed in higher energies.
+	This fraction matches the observations.
+	Finally, we predict the distribution of observable stretching factors, which may be tested in the future when more observational data will be available.
 \end{abstract}
 
-\pacs{11.10.Ef,~11.15.Bt,~11.15.Ha}
+\pacs{95.85.Pw,~97.60.Bw}
 

### 2014-05-21 ITEPProceedings/piskunov.tex
Added sample light curves

@@
 	The main assumption is the jet opening angle dependence on radiation energy -- the most energetic photons are emitted near the axis of the jet.
+	We also assume that all bursts are the same in their reference frames (that is their light curves differ because of different redshifts and off-axis angles).
 	To test this model, we compute the total energy of the burst, and confirm that it is below the constraint.
@@
 \section{Introduction}
+	Sample citation: \cite{Ackermann:2013zfa}.
+
 For the purposes of this review,
@@
 \large \bfseries %\MakeTextUppercase{%
-АНОМАЛИИ (ИЗБРАННЫЕ ТЕМЫ)
+ВРЕМЕННОЕ РАСТЯЖЕНИЕ ГЭВ ИЗЛУЧЕНИЯ ГАММА-ВСПЛЕСКОВ: FERMI LAT И ГЕОМЕТРИЧЕСКАЯ МОДЕЛЬ
 \end{center}
@@
 {\bfseries
-В.~И.~Захаров
+М.~С.~Пискунов
 }
@@
 	\small
+	Наблюдения показывают, что излучение гамма-всплесков с энергией выше \unit[100]{МэВ} систематически наблюдается позже, чем низкоэнергичное излучение.
+	Различия же кривых блеска в различных диапазонах высокоэнергичного излучения (> \unit[100]{МэВ}) изучены хуже.
 
-Данный обзор имеет образовательный характер и посвящен
-киральным аномалиям и их наиболее существенным приложениям.
-Также рассматриваются современные вопросы решеточных измерений.
+	В данной работе мы изучаем различия кривых блеска в диапазонах (\unit[100]{МэВ}, \unit[1]{ГэВ}) и (\unit[1]{ГэВ}, \unit[300]{ГэВ}).
+	Для исследования выбраны все 4 гамма-всплеска из каталога Fermi, наблюдения которых содержат не менее 10 фотонов с энергией > \unit[1]{ГэВ}: 080916C, 090510, 090902B, 090926A.
+	Кривые блеска 080916C и 090510 в двух рассматриваемых диапазонах не различаются статистически значимо, однако для 090926A установлено со статистической значимостью $3.3\sigma$, что излучение в диапазоне (\unit[1]{ГэВ}, \unit[300]{ГэВ}) растянуто относительно менее энергичного, также для 090902B установлено со статистической значимостью $2.2\sigma$, что излучение в диапазоне (\unit[100]{МэВ}, \unit[1]{ГэВ}) растянуто относительно более энергичного.
+
+	Мы предлагаем простую геометрическую модель, объясняющую этот результат.
+	Основное предположение -- чем выше энергия излучения, тем ближе к оси джета оно излучается.
+	Мы также предполагаем, что все всплески одинаковы в собственной системе отсчета (то есть, различия кривых блеска объясняются исключительно разницой в красном смещении и углом наблюдения).
+	В рамках модели соблюдается ограничение на максимально допустимую энергию всплеска.
+	Также получена доля всплесков среди наблюдаемых в диапазоне (\unit[100]{МэВ}, \unit[1]{ГэВ}), которые также можно увидеть в диапазоне (\unit[1]{ГэВ}, \unit[300]{ГэВ}).
+	Полученная доля согласуется с наблюдениями.
+	Наконец, предсказано распределение коэффициентов растяжения, которое может быть проверено по мере накопления наблюдений.
 \end{minipage}
@@
 
-{\bf Fig.~1.}~Anomalous triangle graph. Momenta $q,k_1,k_2$ are
-carried by the axial current $a_\mu^5$ and
-two photons, respectively. The triangle corresponds to a fermion of small mass $m$.
+{\bf Fig.~1.}~Model Overview.
+Here the red and the blue cones represent the volumes through which low and high energy plasma propagates.
+In case depicted the observer's off-axis angle is smaller than the opening angle of the low energy jet, so, due to the relativistic beaming effect, the most of the observable low energy photons will travel along the straight line from the central engine.
+Also, the observer's off-axis angle is larger that the opening angle of the high energy jet, so the high energy radiation will still originate near the center of the jet (because it is the only place where there are high energy radiators).
+The observation time of a photon is a sum of two things: the time interval spent in plasma as a radiator (which approximately equals to the distance from the central engine to the point of emission); and the time interval from emission to detection (which is the distance from the point of emission to the observer's location).
+Given a position of the shock front, this sum is larger for high energy photons.
+Because of that, high energy emission will be observed later throughout the burst duration, therefore the high energy light curve will be stretched.
 
-{\bf Fig.~2.}~Imaginary part of the anomalous triangle graph.
-The intermediate fermions are on
-mass shell. (The signs are the same as in Fig. 1)
+%{\bf Fig.~2.}~Imaginary part of the anomalous triangle graph.
+%The intermediate fermions are on
+%mass shell. (The signs are the same as in Fig. 1)
 
-{\bf Fig.~3.}~Three-body intermediate state in the imaginary part. (The signs are the same as in Fig. 1)
+%{\bf Fig.~3.}~Three-body intermediate state in the imaginary part. (The signs are the same as in Fig. 1)
 
-{\bf Fig.~4.}~Radiative correction to the two-body contribution to the
-imaginary part. (The signs are the same as in Fig. 1)
+%{\bf Fig.~4.}~Radiative correction to the two-body contribution to the
+%imaginary part. (The signs are the same as in Fig. 1)
 
-{\bf Fig.~5.}~Graphic representation of the Banks--Casher evaluation of the quark condensate
-$\langle \bar{q}q\rangle$. Dashed lines with crosses represent external (vacuum) gluonic
-fields.
+%{\bf Fig.~5.}~Graphic representation of the Banks--Casher evaluation of the quark condensate
+%$\langle \bar{q}q\rangle$. Dashed lines with crosses represent external (vacuum) gluonic
+%fields.
 
@@
 
-%\begin{figure}
-%\includegraphics{z1.eps}
-%\caption{Zakharov, Fig. 1}
-%\end{figure}
+\begin{figure}
+\includegraphics{p1.pdf}
+\caption{Piskunov, Fig. 1}
+\end{figure}
 

### 2014-05-21 Paper/gammaRays.tex
Made distribution histogram and CDF accurate
Added correlation plots
Added line about off-axis prediction
Added sample light curves

@@
 	\centering
-	\begin{subfigure}{0.49\textwidth}
+	\hspace*{\fill}
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width=\textwidth]{sampleLightCurveLogNegative}
 		\label{fig:sampleLightCurveLogNegative}
+		\caption{Stretching factor $\kappa = 0.868$, redshift $z = 0.934$, off-axis angle $\chi = 4.65\times 10^{-4}$.}
 	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
+	\hfill
+	\begin{subfigure}{0.45\textwidth}
 		\includegraphics[width=\textwidth]{sampleLightCurveLogPositive}
 		\label{fig:sampleLightCurveLogPosivie}
+		\caption{Stretching factor $\kappa = 5.33$, redshift $z = 0.229$, off-axis angle $\chi = 8.97\times 10^{-3}$.}
 	\end{subfigure}
-	\caption{
-		High and low energy light curves produced by the geometrical model.
-Burst parameter values are the same as discussed in the section \ref{sec:parameters}.
-Redshift $z = 1$.
-Observer's off-axis angles are $\chi = 0$ for the left figure, and $\chi = 0.01$ for the right one.
+	\hspace*{\fill}
+	\caption{\the\linewidth
+		High and low energy light curves produced by the geometrical model. Burst parameter values are the same as discussed in section \ref{sec:parameters}.
 	}
@@
 	\hspace*{\fill}
-	\caption{Stretching factors histogram and CDF produced by our model. The sample contains $16$ inaccurate bursts.}
+	\caption{Stretching factors histogram and CDF produced by our model. The sample contains $16$ bursts.}
 	\label{fig:kappaDistribution}
@@
 
+\begin{figure}
+        \centering
+        \includegraphics[width=1.0\textwidth]{correlations}
+        \caption{Correlations between off-axis angles, stretching factors and high to low energy photon count ratios found in the sample produces by our model.
+The sample contains $16$ bursts.
+Black crosses show observations.
+This correlation allows to estimate off-axis angles of observed bursts.}
+        \label{fig:correlations}
+\end{figure}
+
 \section{Discussion}

### 2014-05-21 sample16.m
Updated the paper with 256 bursts sample

@@ -1,0 +1,32 @@
+sample16 = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
+      34.228215884194334, 222.73035829370284, 1.0280094289841748}, 
+     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
+      20.978799047458836, 123.38630837565341, 2.529581957694992}, 
+     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
+      10.070010357458859, 178.00778096254146, 1.784271416531604}, 
+     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
+      125.2194399101692, 107.40664116322054, 1.355228067650187}, 
+     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
+      12.708560050890373, 313.05858759493555, 1.1251344561349428}, 
+     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
+      18.983240726070672, 156.94967793679683, 1.6188869008349682}, 
+     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
+      434.73441714374775, 92.48671063445956, 1.085830862969823}, 
+     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
+      578.4999909315469, 94.50973498739872, 0.8908846736633591}, 
+     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
+      131.45761966748594, 125.400307231681, 1.040915010748597}, 
+     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
+      14.21984824761055, 176.44515329447125, 1.5781588059941152}, 
+     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
+      364.8701470277157, 100.1140576033533, 0.9741210674073871}, 
+     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
+      24.758770352965122, 223.3218204554101, 1.1155870882542687}, 
+     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
+      15.814107790469503, 258.1278608438802, 1.1647305773040832}, 
+     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
+      39.646577434687785, 150.48002884878116, 1.2902775201409398}, 
+     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-05-21 sample16IA.m

@@ -1,0 +1,32 @@
+sample16IA = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
+      34.228215884194334, 222.73035829370284, 1.0701575908785792}, 
+     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
+      20.978799047458836, 123.38630837565341, 3.049495182568959}, 
+     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
+      10.070010357458859, 178.00778096254146, 1.9224561249840593}, 
+     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
+      125.2194399101692, 107.40664116322054, 1.3821372215243484}, 
+     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
+      12.708560050890373, 313.05858759493555, 1.180678993898995}, 
+     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
+      18.983240726070672, 156.94967793679683, 1.70714273366365}, 
+     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
+      434.73441714374775, 92.48671063445956, 1.1077715331761284}, 
+     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
+      578.4999909315469, 94.50973498739872, 0.9364896504440356}, 
+     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
+      131.45761966748594, 125.400307231681, 1.0805165020789271}, 
+     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
+      14.21984824761055, 176.44515329447125, 1.6840634952055642}, 
+     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
+      364.8701470277157, 100.1140576033533, 1.0130197810371595}, 
+     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
+      24.758770352965122, 223.3218204554101, 1.1676582199615857}, 
+     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
+      15.814107790469503, 258.1278608438802, 1.225460126255487}, 
+     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
+      39.646577434687785, 150.48002884878116, 1.3439228024212373}, 
+     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-05-21 sample16pc40.m

@@ -1,0 +1,32 @@
+sample16pc40 = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
+      34.228215884194334, 222.73035829370284, 1.0280094289841748}, 
+     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
+      20.978799047458836, 123.38630837565341, 2.529581957694991}, 
+     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
+      10.070010357458859, 178.00778096254146, 1.7842714165316036}, 
+     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
+      125.2194399101692, 107.40664116322054, 1.355228067650187}, 
+     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
+      12.708560050890373, 313.05858759493555, 1.1251344561349426}, 
+     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
+      18.983240726070672, 156.94967793679683, 1.6188869008349682}, 
+     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
+      434.73441714374775, 92.48671063445956, 1.0858308629698226}, 
+     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
+      578.4999909315469, 94.50973498739872, 0.890884673663359}, 
+     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
+      131.45761966748594, 125.400307231681, 1.0409150107485967}, 
+     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
+      14.21984824761055, 176.44515329447125, 1.5781588059941158}, 
+     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
+      364.8701470277157, 100.1140576033533, 0.9741210674073869}, 
+     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
+      24.758770352965122, 223.3218204554101, 1.1155870882542687}, 
+     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
+      15.814107790469503, 258.1278608438802, 1.1647305773040835}, 
+     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
+      39.646577434687785, 150.48002884878116, 1.2902775201409395}, 
+     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-05-21 sample256.m

@@ -1,0 +1,512 @@
+sample256 = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
+      34.228215884194334, 222.73035829370284, 1.0280094289841748}, 
+     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
+      20.978799047458836, 123.38630837565341, 2.529581957694992}, 
+     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
+      10.070010357458859, 178.00778096254146, 1.784271416531604}, 
+     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
+      125.2194399101692, 107.40664116322054, 1.355228067650187}, 
+     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
+      12.708560050890373, 313.05858759493555, 1.1251344561349428}, 
+     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
+      18.983240726070672, 156.94967793679683, 1.6188869008349682}, 
+     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
+      434.73441714374775, 92.48671063445956, 1.085830862969823}, 
+     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
+      578.4999909315469, 94.50973498739872, 0.8908846736633591}, 
+     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
+      131.45761966748594, 125.400307231681, 1.040915010748597}, 
+     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
+      14.21984824761055, 176.44515329447125, 1.5781588059941152}, 
+     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
+      364.8701470277157, 100.1140576033533, 0.9741210674073871}, 
+     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
+      24.758770352965122, 223.3218204554101, 1.1155870882542687}, 
+     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
+      15.814107790469503, 258.1278608438802, 1.1647305773040832}, 
+     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
+      39.646577434687785, 150.48002884878116, 1.2902775201409398}, 
+     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-05-21 sample256IA.m

@@ -1,0 +1,512 @@
+sample256IA = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
+      34.228215884194334, 222.73035829370284, 1.0701575908785792}, 
+     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
+      20.978799047458836, 123.38630837565341, 3.049495182568959}, 
+     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
+      10.070010357458859, 178.00778096254146, 1.9224561249840593}, 
+     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
+      125.2194399101692, 107.40664116322054, 1.3821372215243484}, 
+     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
+      12.708560050890373, 313.05858759493555, 1.180678993898995}, 
+     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
+      18.983240726070672, 156.94967793679683, 1.70714273366365}, 
+     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
+      434.73441714374775, 92.48671063445956, 1.1077715331761284}, 
+     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
+      578.4999909315469, 94.50973498739872, 0.9364896504440356}, 
+     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
+      131.45761966748594, 125.400307231681, 1.0805165020789271}, 
+     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
+      14.21984824761055, 176.44515329447125, 1.6840634952055642}, 
+     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
+      364.8701470277157, 100.1140576033533, 1.0130197810371595}, 
+     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
+      24.758770352965122, 223.3218204554101, 1.1676582199615857}, 
+     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
+      15.814107790469503, 258.1278608438802, 1.225460126255487}, 
+     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
+      39.646577434687785, 150.48002884878116, 1.3439228024212373}, 
+     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-05-25 ITEPProceedings/piskunov.tex
Drafted ITEP Proceedings

@@
 
-\newcommand{\beq}{\begin{equation}}
-\newcommand{\eeq}{\end{equation}}
+%\newcommand{\beq}{\begin{equation}}
+%\newcommand{\eeq}{\end{equation}}
 
@@
 
-\newcommand{\A}{\alpha}
-\newcommand{\B}{\beta}
-\newcommand{\T}{\theta}
-\newcommand{\Ep}{\epsilon}
-\newcommand{\beqn}{\begin{eqnarray}}
-\newcommand{\eeqn}{\end{eqnarray}}
+%\newcommand{\A}{\alpha}
+%\newcommand{\B}{\beta}
+%\newcommand{\T}{\theta}
+%\newcommand{\Ep}{\epsilon}
+%\newcommand{\beqn}{\begin{eqnarray}}
+%\newcommand{\eeqn}{\end{eqnarray}}
 
@@
 
-\begin {abstract}
+\begin{abstract}
 	Numerous observations confirm that the high energy $(> \unit[100]{MeV})$ emission of gamma ray bursts is delayed with respect to the low energy emission.
@@
 \section{Introduction}
-	Sample citation: \cite{Ackermann:2013zfa}.
+	Gamma Ray Bursts (GRBs) are among the most energetic events in the Universe, therefore they might provide new knowledge for particle physics.
+	Modern observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} made possible to study these explosions extensively \cite{Vianello:2013ela,Gehrels:2013xd}.
+	These and previous studies led to several interesting results.
+	For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}.
+	This suggests that most of the bursts must have similar energetics in their rest frame.
+	Temporal variations of spectra were also studied.
+	The spectral lags were found between different low energy bands \cite{Yi:2005ht}.
+	The very high energy radiation was discovered to be extended relative to x-ray emission \cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}.
+	In this study we continue this series of results and explore the spectral variations between high energy bands $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
+	We use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
+	Results obtained with this approach are discussed in the section \ref{sec:data}.
+	Then we propose a simple geometrical model to explain these results in section \ref{sec:model}.
+	Finally, our study is summarized in the section \ref{sec:discussion}.
 
-For the purposes of this review,
-by anomalies one can understand violations
-of the Ehrenfest theorem according to which
-all matrix elements of the
-classical equations of motion (considered as operators) vanish.
-Thus, anomaly is established
- if we find, for example,
-a nonvanishing
-matrix element of the Dirac equation for a charged particle:
-\begin{equation}
-\langle \bar{\psi}(\hat{D}+im)\psi\rangle~\neq~0,
-\end{equation}
-where $\hat{D}=\gamma_{\mu}D_{\mu}$ and $D_{\mu}$ is the covariant derivative
-and $m$
-is the mass of the charged particle.
+\section{Fermi LAT Data}
+	\label{sec:data}
 
-..........................
+	We take the GRB list from the Fermi-LAT catalog \cite{Ackermann:2013zfa}, and use the technique introduced in \cite{Rubtsov:2011qq} to estimate background radiation in both energy bands.
+	Most of the bursts in the catalog, however, do not have enough events with over $\unit[1]{GeV}$ energies to do desired computations, so we choose only those bursts, from which at least $10$ photons were detected in the high energy band.
+	This leaves $4$ of them: 080916C \cite{Tajima:2009az}, 090510 \cite{Ackermann:2010us}, 090902B \cite{Abdo:2009pg} and 090926A \cite{Bregeon:2011bu}.
+
+	For these 4 bursts we compute both high and low energy distributions of photon times.
+	We subtract the low energy background estimate from the low energy CDF, and add the high energy background estimate to it.
+	This makes the low energy CDF non-monotonous and, rigorously speaking, we cannot use the 2-sample Kolmogorov-Smirnov test on it.
+	However, since there is a much greater number of photons on lower energies, we can think of the low energy CDF as continuous.
+	Therefore, this non-monotonicity will be negligible, and will not harm the KS-test results much.
+
+	Finally, we stretch the high energy CDF by different factors, and compare it to the low energy CDF using the KS-test.
+	If we require $2\sigma$-significant probability to exclude a particular stretching factor, then the stretching factor of $1$ (which means no stretching) is allowed for GRBs 080916C and 090510.
+	For the other two bursts the stretching factor of $1$ is, however, prohibited.
+	For GRB 090902B, it is smaller than $1$, so the low energy light curve is stretched with respect to high energy one (see fig. \ref{fig:grb090902B}).
+	And for GRB 090926A, even more dramatic deviation for the stretching factor of $1$ is observed in the opposite direction.
+	The high energy light curve of GRB 090926A is stretched with respect to low energy one with a factor of at least $1.99$ (see fig. \ref{fig:grb090926A}).
+	You can see allowed ranges for stretching factors for all the bursts studied in table \ref{tab:bursts}.
+
+	So we obtained a preliminary (for significance is less than $5\sigma$) result that the stretching factors for observable bursts might be both larger and smaller than $1$.
+
+\section{Model}
+	\label{sec:model}
+
+	We propose that time stretching of GRB light curves can be explained by the curvature effects (that is the effects of jet geometry) even if all the bursts are the same in their rest frames.
+	These effects were explored by multiple authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.
+	However, these studies were only concerned with x-ray radiation, and, even more importantly, they assumed that the distribution of radiation sources is homogeneous throughout the jet.
+	We propose a contrary idea: that the highest energy radiators are concentrated near the axis of the burst, meaning that the burst opening angle depends on energy.
+	We cannot prove this assumption rigorously, but have arguments supporting it.
+
+	First of all, there are around 750 GRBs detected by GBM, half of which were in the LAT field of view at the moment of observation \cite{Vianello:2013ela}.
+	However, only about 30 of them were detected by the LAT, and only 4 of them were bright in the high energy band.
+	If we extrapolate the uniform jet model to very high energies, this observation would mean that these groups of bursts are internally different: some of them produce VHE radiation, and other do not.
+	These differences are hard to explain given that burst energetics are similar \cite{Bloom:2003wy}.
+	Nevertheless, these differences in burst counts can easily be explained by our model.
+	In our model, the opening angle of a jet is inversely proportional to the energy of photons it radiates.
+	Therefore the most common scenario is that the off-axis angle of the observer is smaller than the low energy jet opening angle, but larger than the opening angle of a high energy jet.
+	Because of that, most of the bursts can only be seen at low energies.
+	The 4 bursts we study in this paper were seen, according to our model, from the lowest off-axis angles.
+
+	Second, consider the plasma right after its ejection from the central engine.
+	Two processes happen there simultaneously:
+	\begin{enumerate}
+		\item{Particles near the jet boundary change their movement directions, therefore increasing the opening angle of the jet.}
+		\item{Particles lose energy, therefore decreasing radiation frequency emitted by the jet.}
+	\end{enumerate}
+	We argue that these two processes are correlated, for they happen due to the same particle interaction processes.
+	And since they are correlated, jets with larger opening angles should have lower energies, which is the assumption we are trying to justify.
+
+	You can see why this assumption allows for non-trivial stretching factors in the figure \ref{fig:modelOverview}.
+	More rigorously, the model is defined by the following assumptions:
+	\begin{enumerate}
+		\item{
+			Time $t = 0$, a spherical shell of plasma is emitted.
+			The center is called the central engine.
+		}
+		\item{The shell points propagate with a constant velocity $v = \frac{\sqrt{\gamma^2-1}}{\gamma} \sim 1$, so at the time $t$ the radius of the shall is $v t$.}
+		\item{Each point of the shell is an isotropic radiator in its rest frame.}
+		\item{
+			The radiation intensity is a function of the radiator position and the radiation frequency:
+			\begin{equation*}
+				\eta\left(r,\theta,\omega\right) = 
+				\frac{\eta_0}{1 + \left(\frac{r}{r_0}\right)^n}
+				\exp\left(
+					-\left(\frac{\theta}{\theta_0}\right)^2
+					\left(\frac{\omega}{\omega_0}\right)^{-2k}
+				\right)
+				\left(\frac{\omega}{\omega_0}\right)^\alpha
+			\end{equation*}
+			$\eta$ is a number of particles emitted per volume per solid angle per frequency.
+			It is a function of the distance $r$ from the central engine, of the off-axis angle $\theta$, and of the radiation frequency $\omega$.
+		}
+	\end{enumerate}
+
+	The burst is fully specified by the following set of parameters (we assume these parameters to be the same for all bursts, the values are obtained by fitting to make possible observed bursts from section \ref{sec:data}):
+	\begin{itemize}
+		\item{$\gamma = 300$, the relativistic factor of the shell, $\gamma \gg 1$.}
+		\item{$\eta_0 = \unit[8.15 \times 10^{34}]{sec^{-3} GeV^{-1}}$, which defines the luminosity scale.}
+		\item{$r_0 = \unit[3.36 \times 10^6]{sec^3}$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
+		\item{$n = 22.5$, which determines the sharpness of the jet end, $n > 3$;}
+		\item{$\omega_0 = \unit[1]{GeV}$, a characteristic radiation frequency;}
+		\item{$\theta_0 = 5.02 \times 10^{-5}$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;}
+		\item{$k = -0.700$, which determines how much the opening angle changes with frequency, $k < 0$;}
+		\item{$\alpha = -0.627$, the bare spectral index, $\alpha < -2k - 1$}
+	\end{itemize}
+
+	Numerical simulations based on these assumptions are able to reproduce both low (like of GRB 090902B) and high (like of GRB 090926A) stretching factors (see fig. \ref{fig:sampleLightCurves}).
+
+	To test the model agreement to observations, we computed a few more things in its framework:
+	\begin{itemize}
+		\item{
+			Total energy emitted in $\unit[100]{MeV}$ and more energetic gamma rays $E < \unit[5.89 \times 10^{53}]{GeV}$, which is in agreement with \cite{Gehrels:2013xd}.
+		}
+		\item{
+			Distribution of stretching factors of observable bursts.
+			You can see the PDF and CDF on fig. \ref{fig:kappaDistribution}.
+			This distribution doesn't contradict with the values of stretching factors computed in section \ref{sec:data}.
+		}
+		\item{
+			The fraction of bursts observable in low energy band, which are also observed in high energy band $f_m = 0.116$.
+			The Fermi LAT catalog contains $35$ bursts out of which $4$ can be observed in high energy band.
+			Therefore the observed value $f_o = 0.11 \pm 0.05$ (error obtained from binomial distribution) agrees with the model.
+		}
+		\item{
+			There is a correlation between the fractions of high energy photons (that is high energy photon count divided by total photon count), stretching factors and off-axis angles of observers (see fig.
+\ref{fig:correlations}).
+It allows one to predict off-axis angles.
+		}
+	\end{itemize}
+
+	So our model, though with some discrepancies, is able to explain the time stretching result from section \ref{sec:data}.
+
+\section{Discussion}
+	\label{sec:discussion}
+
+	The main results of our study may be summarized as following:
+	\begin{enumerate}
+		\item{
+			The time stretching of GRB light curves between different high energy bands is confirmed with statistical significance of $3.3\sigma$.
+		}
+		\item{
+			This time stretching can be explained by geometry of the jet, without introducing any new spectral components.
+		}
+		\item{
+			All GRBs are assumed the same in their rest frames.
+			There is no need to vary internal burst parameters.
+		}
+	\end{enumerate}
+
+	There, however, directions for improvement:
+	\begin{enumerate}
+		\item{
+			The model prediction and observations on fig. \ref{fig:correlations} are, though close, not exactly agree.
+		}
+		\item{
+			Shapes of light curves produced by model differ considerably from observed ones.
+		}
+		\item{
+			Assumption of energy dependence on off-axis angle should be better justified.
+		}
+		\item{
+			Much more results can be obtained from the model (like energy spectrum, and its variation with time).
+		}
+	\end{enumerate}
+
+	We believe, however, that these problems are not unresolvable, and may be addressed by refining the model without changing its main assumptions.
+	And in the end, we think the main ideas of this paper will be close to truth.
 
@@
 
-{\bf Fig.~1.}~Model Overview.
+{\bf Fig.~1.}~GRB 090902B results. Low energy radiation is stretched ($\kappa < 1$) with significance of $2.2\sigma$.
+
+{\bf Fig.~2.}~GRB 090926A results. High energy radiation is stretched ($\kappa > 1$) with significance of $3.3\sigma$.
+
+{\bf Tab.~1.}~Stretching factor computation results for the bursts studied.
+Ranges of allowed stretching factors are shown for studied bursts for multiple levels of significance.
+The values for trigger time and location are taken from the tables 2 and 4 of \cite{Ackermann:2013zfa}.
+
+{\bf Fig.~3.}~Model Overview.
 Here the red and the blue cones represent the volumes through which low and high energy plasma propagates.
@@
 
-%{\bf Fig.~2.}~Imaginary part of the anomalous triangle graph.
-%The intermediate fermions are on
-%mass shell. (The signs are the same as in Fig. 1)
+{\bf Fig.~4.}~High and low energy light curves produced by the geometrical model.
+Burst parameter values are the same as discussed in section \ref{sec:model}.
+On the left panel: the stretching factor $\kappa = 0.821$, redshift $z = 1.82$, off-axis angle $\chi = 0$.
+On the right: stretching factor $\kappa = 2.26$, redshift $z = 2.106$, off-axis angle $\chi = 5.41 \times 10^{-3}$.
 
-%{\bf Fig.~3.}~Three-body intermediate state in the imaginary part. (The signs are the same as in Fig. 1)
+{\bf Fig.~5.}~Stretching factors histogram and CDF produced by our model. Bar lines (black) show observed stretching factors. The sample contains $4096$ bursts.
 
-%{\bf Fig.~4.}~Radiative correction to the two-body contribution to the
-%imaginary part. (The signs are the same as in Fig. 1)
-
-%{\bf Fig.~5.}~Graphic representation of the Banks--Casher evaluation of the quark condensate
-%$\langle \bar{q}q\rangle$. Dashed lines with crosses represent external (vacuum) gluonic
-%fields.
+{\bf Fig.~6.}~Correlations between high energy photon ratios, off-axis angles and stretching factors found in the sample produces by our model.
+The sample contains $4096$ bursts.
+Black crosses show observations.
+This correlation allows one to estimate off-axis angles of observed bursts.
 
@@
 \begin{figure}
-\includegraphics{p1.pdf}
+\includegraphics[width=\textwidth]{p1.pdf}
 \caption{Piskunov, Fig. 1}
+\label{fig:grb090902B}
 \end{figure}
 
-%\begin{figure}
-%\includegraphics{z2.eps}
-%\caption{Zakharov, Fig. 2}
-%\end{figure}
+\begin{figure}
+\includegraphics[width=\textwidth]{p2.pdf}
+\caption{Piskunov, Fig. 2}
+\label{fig:grb090926A}
+\end{figure}
 
-%\begin{figure}
-%\includegraphics{z3.eps}
-%\caption{Zakharov, Fig. 3}
-%\end{figure}
+\begin{table}
+	\centering
+	\small
+	\begin{tabular}{ l | c | c | c | c | c | c | c}
+		GRB 	& 	Redshift	&	GBM Trigger Time	& R.A.			& Dec.			&	Stretching factor	&	Stretching factor	\\
+		Name 	& 				&	MET sec				& J2000, deg	& J2000, deg	&	$2\sigma$ range		&	$3\sigma$ range		\\
+		\hline
+		080916C	&	$4.35$		&	243216766.614		&	119.85 		&	-56.64 		&	0.67 -- 3.32		&	0.42 -- 5.83		\\
+		090510 	&	$0.903$		&	263607781.971		&	333.55 		&	-26.583		&	0.43 -- 1.61		&	0.32 -- 2.29		\\
+		090902B	&	$1.822$		&	273582310.313		&	264.94 		&	 27.324		&	0.36 -- 0.89		&	0.23 -- 1.57		\\
+		090926A	&	$2.1062$	&	275631628.990		&	353.4  		&	-66.32 		&	1.99 -- 6.62		&	1.34 -- 9.15		\\
+	\end{tabular}
+	\caption{Piskunov, Tab. 1}
+	\label{tab:bursts}
+\end{table}
 
-%\begin{figure}
-%\includegraphics{z4.eps}
-%\caption{Zakharov, Fig. 4}
-%\end{figure}
+\begin{figure}
+\includegraphics[width=.7\textwidth]{p3.pdf}
+\caption{Piskunov, Fig. 3}
+\label{fig:modelOverview}
+\end{figure}
 
-%\begin{figure}
-%\includegraphics{z5.eps}
-%\caption{Zakharov, Fig. 5}
-%\end{figure}
+\begin{figure}
+\includegraphics[width=\textwidth]{p4.pdf}
+\caption{Piskunov, Fig. 4}
+\label{fig:sampleLightCurves}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{p5.pdf}
+\caption{Piskunov, Fig. 5}
+\label{fig:kappaDistribution}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{p6.pdf}
+\caption{Piskunov, Fig. 6}
+\label{fig:correlations}
+\end{figure}
 
 \end{document}
-% \bibitem{kniel}
-% Two-Loop Electroweak Corrections to the A0 gamma gamma and A0 g g Couplings of the CP-Odd Higgs Boson.
-% Joachim Brod, Frank Fugel, Bernd A. Kniehl
-%  Nucl.Phys.B807:188-209,2009.
-% e-Print: arXiv:0807.1008 [hep-ph]

### 2014-05-25 GRObservations/GRObservations/GRBurst.cpp

@@ -107,7 +107,7 @@ void GRBurst::calculateBackground() {
     gevDistribution.end = query.endTime - (time + startOffset);
 }
 
-double GRBurst::photonRatio() {
+double GRBurst::highEnergyPhotonFraction() {
     double mevCount;
     double gevCount;
     
@@ -123,7 +123,7 @@ double GRBurst::photonRatio() {
         }
     }
     
-    return gevCount / mevCount;
+    return gevCount / (mevCount + gevCount);
 }
 
 void GRBurst::read() {

### 2014-05-25 GRObservations/GRObservations/GRBurst.h

@@ -55,7 +55,7 @@ public:
     vector <GRFermiLATPhoton> gevPhotons;
     vector <GRFermiLATPhoton> gevBackgroundPhotons;
     
-    double photonRatio();
+    double highEnergyPhotonFraction();
     
 public:
     GRDistribution mevDistribution;

### 2014-05-25 GRObservations/GRObservations/main.cpp

@@ -99,7 +99,7 @@ int main(int argc, const char * argv[])
         log << burstCatalog[i].name << ":" << endl;
         log << "\tLow energy count  = " << burstCatalog[i].mevDistribution.values.size() - burstCatalog[i].mevDistribution.linearComponent << endl;
         log << "\tHigh energy count = " << burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.linearComponent << endl;
-        log << "\tPhoton Ratio       = " << burstCatalog[i].photonRatio() << endl;
+        log << "\tPhoton Ratio       = " << burstCatalog[i].highEnergyPhotonFraction() << endl;
         
         if ((burstCatalog[i].gevDistribution.values.size() - burstCatalog[i].gevDistribution.linearComponent) < 10) {
             log << "\tnot enought high energy photons" << endl << endl;

### 2014-05-25 sample256pc40.m

@@ -1,0 +1,512 @@
+sample256pc40 = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
+      34.228215884194334, 222.73035829370284, 1.0280094289841748}, 
+     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
+      20.978799047458836, 123.38630837565341, 2.529581957694991}, 
+     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
+      10.070010357458859, 178.00778096254146, 1.7842714165316036}, 
+     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
+      125.2194399101692, 107.40664116322054, 1.355228067650187}, 
+     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
+      12.708560050890373, 313.05858759493555, 1.1251344561349426}, 
+     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
+      18.983240726070672, 156.94967793679683, 1.6188869008349682}, 
+     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
+      434.73441714374775, 92.48671063445956, 1.0858308629698226}, 
+     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
+      578.4999909315469, 94.50973498739872, 0.890884673663359}, 
+     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
+      131.45761966748594, 125.400307231681, 1.0409150107485967}, 
+     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
+      14.21984824761055, 176.44515329447125, 1.5781588059941158}, 
+     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
+      364.8701470277157, 100.1140576033533, 0.9741210674073869}, 
+     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
+      24.758770352965122, 223.3218204554101, 1.1155870882542687}, 
+     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
+      15.814107790469503, 258.1278608438802, 1.1647305773040835}, 
+     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
+      39.646577434687785, 150.48002884878116, 1.2902775201409395}, 
+     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-05-25 ITEPProceedings/piskunov.tex
Revised Introduction and Fermi LAT Data

@@
 	The very high energy radiation was discovered to be extended relative to x-ray emission \cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}.
-	In this study we continue this series of results and explore the spectral variations between high energy bands $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
+	In this study we continue this research and explore the spectral variations between energy bands $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
 	We use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
-	Results obtained with this approach are discussed in the section \ref{sec:data}.
+	Results obtained with this approach are discussed in section \ref{sec:data}.
 	Then we propose a simple geometrical model to explain these results in section \ref{sec:model}.
-	Finally, our study is summarized in the section \ref{sec:discussion}.
+	Finally, our study is summarized in section \ref{sec:discussion}.
 
@@
 
-	We take the GRB list from the Fermi-LAT catalog \cite{Ackermann:2013zfa}, and use the technique introduced in \cite{Rubtsov:2011qq} to estimate background radiation in both energy bands.
+	We take the GRB list from the Fermi-LAT catalog \cite{Ackermann:2013zfa}, and use the technique introduced in \cite{Rubtsov:2011qq} to estimate background radiation in both energy bands.\footnote{Detailed description of computations in this and the next section (including the source code) can be found on GitHub: \url{https://github.com/maxitg/GammaRays}}
 	Most of the bursts in the catalog, however, do not have enough events with over $\unit[1]{GeV}$ energies to do desired computations, so we choose only those bursts, from which at least $10$ photons were detected in the high energy band.

### 2014-05-25 ITEPProceedings/piskunov.tex
Revised Model section of ITEP Proceedings

@@
 	Second, consider the plasma right after its ejection from the central engine.
-	Two processes happen there simultaneously:
+	Two processes happen here simultaneously:
 	\begin{enumerate}
@@
 			\end{equation*}
-			$\eta$ is a number of particles emitted per volume per solid angle per frequency.
+			$\eta$ is the number of particles emitted per volume per solid angle per frequency.
 			It is a function of the distance $r$ from the central engine, of the off-axis angle $\theta$, and of the radiation frequency $\omega$.
@@
 
-	The burst is fully specified by the following set of parameters (we assume these parameters to be the same for all bursts, the values are obtained by fitting to make possible observed bursts from section \ref{sec:data}):
+	The burst is fully specified by the following set of parameters (we assume these parameters to be the same for all bursts, the specified values are obtained by fitting to make possible observed bursts from section \ref{sec:data}):
 	\begin{itemize}
@@
 
-	Numerical simulations based on these assumptions are able to reproduce both low (like of GRB 090902B) and high (like of GRB 090926A) stretching factors (see fig. \ref{fig:sampleLightCurves}).
+	Numerical simulations based on these assumptions are able to reproduce both stretching factors smaller than $1$ (like of GRB 090902B) and larger than $1$ (like of GRB 090926A), see fig.
+\ref{fig:sampleLightCurves}.
 
-	To test the model agreement to observations, we computed a few more things in its framework:
+	To test the model against observations, we computed a few more things:
 	\begin{itemize}
@@
 			Distribution of stretching factors of observable bursts.
-			You can see the PDF and CDF on fig. \ref{fig:kappaDistribution}.
-			This distribution doesn't contradict with the values of stretching factors computed in section \ref{sec:data}.
+			You can see the PDF and the CDF of it on fig. \ref{fig:kappaDistribution}.
+			This distribution doesn't contradict with the values of stretching factors obtained in section \ref{sec:data}.
 		}
@@
 		\item{
-			There is a correlation between the fractions of high energy photons (that is high energy photon count divided by total photon count), stretching factors and off-axis angles of observers (see fig.
+			There is a correlation between the fraction of high energy photons (that is high energy photon count divided by the total photon count), the stretching factor and the observer's off-axis angle (see fig.
 \ref{fig:correlations}).
-It allows one to predict off-axis angles.
+It allows one to predict observer's off-axis angles.
 		}
@@
 	There, however, directions for improvement:
-	\begin{enumerate}
+	\begin{itemize}
 		\item{
@@
 		}
-	\end{enumerate}
+	\end{itemize}
 

### 2014-05-25 ITEPProceedings/piskunov.tex
Revised Model

@@
 			$\eta$ is the number of particles emitted per volume per solid angle per frequency.
-			It is a function of the distance $r$ from the central engine, of the off-axis angle $\theta$, and of the radiation frequency $\omega$.
+			It is a function of the distance $r$ from the central engine, the off-axis angle $\theta$, and the radiation frequency $\omega$.
 		}
@@
 		\item{$\eta_0 = \unit[8.15 \times 10^{34}]{sec^{-3} GeV^{-1}}$, which defines the luminosity scale.}
-		\item{$r_0 = \unit[3.36 \times 10^6]{sec^3}$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
+		\item{$r_0 = \unit[3.36 \times 10^6]{sec}$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
 		\item{$n = 22.5$, which determines the sharpness of the jet end, $n > 3$;}
@@
 
-	Numerical simulations based on these assumptions are able to reproduce both stretching factors smaller than $1$ (like of GRB 090902B) and larger than $1$ (like of GRB 090926A), see fig.
-\ref{fig:sampleLightCurves}.
+	Numerical simulations based on these assumptions are able to reproduce both stretching factors smaller than $1$ (like of GRB 090902B) and larger than $1$ (like of GRB 090926A) depending on redshift and observer's off-axis angle (see fig.
+\ref{fig:sampleLightCurves}).
 

### 2014-05-25 ITEPProceedings/piskunov.tex
Refined ITEP Proceedings

@@
 	The main assumption is the jet opening angle dependence on radiation energy -- the most energetic photons are emitted near the axis of the jet.
-	We also assume that all bursts are the same in their reference frames (that is their light curves differ because of different redshifts and off-axis angles).
+	We also assume that all bursts are the same in their rest frames (that is their light curves differ because of different redshifts and observer's off-axis angles).
 	To test this model, we compute the total energy of the burst, and confirm that it is below the constraint.
@@
 		\item{
-			The time stretching of GRB light curves between different high energy bands is confirmed with statistical significance of $3.3\sigma$.
+			The time stretching of GRB light curves between different VHE bands is confirmed with statistical significance of $3.3\sigma$.
 		}
@@
 		\item{
-			All GRBs are assumed the same in their rest frames.
+			All GRBs are assumed to be the same in their rest frames.
 			There is no need to vary internal burst parameters.
@@
 
-	There, however, directions for improvement:
+	There are, however, directions for improvement:
 	\begin{itemize}
@@
 		\item{
-			Shapes of light curves produced by model differ considerably from observed ones.
+			Shapes of light curves produced by the model differ considerably from observed ones.
 		}
@@
 		\item{
-			Much more results can be obtained from the model (like energy spectrum, and its variation with time).
+			While we believe that all bursts are almost the same in their rest frames, they are certainly not exactly the same.
+			The minor differences between GRBs might be responsible for observable deviations from the model predictions.
 		}
@@
 
-{\bf Fig.~1.}~GRB 090902B results. Low energy radiation is stretched ($\kappa < 1$) with significance of $2.2\sigma$.
+{\bf Fig.~1.}~GRB 090902B results. Low energy radiation is stretched ($\kappa < 1$) with statistical significance of $2.2\sigma$.
 
-{\bf Fig.~2.}~GRB 090926A results. High energy radiation is stretched ($\kappa > 1$) with significance of $3.3\sigma$.
+{\bf Fig.~2.}~GRB 090926A results. High energy radiation is stretched ($\kappa > 1$) with statistical significance of $3.3\sigma$.
 
@@
 
-{\bf Fig.~6.}~Correlations between high energy photon ratios, off-axis angles and stretching factors found in the sample produces by our model.
-The sample contains $4096$ bursts.
+{\bf Fig.~6.}~Correlation between the high energy photon fraction, the observer's off-axis angle, and the stretching factor found in the sample produces by our model.
 Black crosses show observations.
-This correlation allows one to estimate off-axis angles of observed bursts.
+This correlation allows one to estimate observer's off-axis angles of real bursts.
 

### 2014-06-01 Paper/gammaRays.tex
Updated the paper with 256 bursts sample
Updated paper with 4096 points sample

@@
 		\label{fig:sampleLightCurveLogNegative}
-		\caption{Stretching factor $\kappa = 0.868$, redshift $z = 0.934$, off-axis angle $\chi = 4.65\times 10^{-4}$.}
+		\caption{Stretching factor $\kappa = 0.821$, redshift $z = 1.82$, off-axis angle $\chi = 0$.}
 	\end{subfigure}
@@
 		\label{fig:sampleLightCurveLogPosivie}
-		\caption{Stretching factor $\kappa = 5.33$, redshift $z = 0.229$, off-axis angle $\chi = 8.97\times 10^{-3}$.}
+		\caption{Stretching factor $\kappa = 2.26$, redshift $z = 2.106$, off-axis angle $\chi = 5.41 \times 10^{-3}$.}
 	\end{subfigure}
 	\hspace*{\fill}
-	\caption{\the\linewidth
+	\caption{
 		High and low energy light curves produced by the geometrical model. Burst parameter values are the same as discussed in section \ref{sec:parameters}.
@@
 	\hspace*{\fill}
-	\caption{Stretching factors histogram and CDF produced by our model. The sample contains $16$ bursts.}
+	\caption{Stretching factors histogram and CDF produced by our model. The sample contains $4096$ bursts.}
 	\label{fig:kappaDistribution}
@@
         \caption{Correlations between off-axis angles, stretching factors and high to low energy photon count ratios found in the sample produces by our model.
-The sample contains $16$ bursts.
+The sample contains $4096$ bursts.
 Black crosses show observations.

### 2014-06-11 Paper/gammaRays.tex
Drafted Fit section

@@
 \usepackage{multicol}
+\usepackage{bm}
 
@@
 	\item{
-		Set $\gamma = 300$, $\omega_0 = 1$, $z = 2.1062$ (which is the redshift of GRB 090926A)
-	}
-	\item{
-		Set $\eta_0$ and $r_0$ such that the duration of the burst, and the total number of observed photons are compatible with GRB 090926A, $T_{0.99}\left(z,\chi;\unit[0.1]{GeV},\infty\right) = 219.5$ and $p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\infty\right) = 179.989$
+		Set $\gamma = 300$, $\omega_0 = \unit[1]{GeV}$, $z = 2.1062$ (which is the redshift of GRB 090926A)
+	}
+	\item{
+		Set $\eta_0$ and $r_0$ such that the duration of the burst, and the total number of observed photons are compatible with GRB 090926A, $T_{0.99}\left(z,\chi;\unit[0.1]{GeV},\infty\right) = \unit[219.5]{sec}$ and $p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\infty\right) = 179.989$
 	}
@@
 		\hline
-		22.1525	&	$\mathrm{2.16424 \times 10^{-4}}$ 	&	-0.417021  	&	-1.3516300 	&	0.00704982 	\\
-		25.8640	&	$\mathrm{4.90222 \times 10^{-8}}$	&	-2.908750 	&	 3.5380100 	&	0.00028657 	\\
-		17.5380	&	$\mathrm{1.46284 \times 10^{-3}}$ 	&	-2.133600 	&	-2.1336000 	&	0.00152624 	\\
-		17.9736	&	$\mathrm{4.52631 \times 10^{-6}}$	&	-1.402620 	&	-0.0114132 	&	0.00032590 	\\
-		 5.0000	&	$\mathrm{1.12535 \times 10^{-7}}$	&	-0.200000	&	-2.0000000 	&	0.00473795 	\\
-		 7.0000	&	$\mathrm{2.00000 \times 10^{-12}}$	&	-3.000000	&	-3.0000000	&	0.00173795
+		22.1525	&	$\mathrm{2.16424 \times 10^{-4}}$ 	&	-0.417021  	&	-1.3516300 	&	$\mathrm{7.04982 \times 10^{-3}}$	\\
+		25.8640	&	$\mathrm{4.90222 \times 10^{-8}}$	&	-2.908750 	&	 3.5380100 	&	$\mathrm{2.86570 \times 10^{-4}}$ 	\\
+		17.5380	&	$\mathrm{1.46284 \times 10^{-3}}$ 	&	-2.133600 	&	-2.1336000 	&	$\mathrm{1.52624 \times 10^{-3}}$ 	\\
+		17.9736	&	$\mathrm{4.52631 \times 10^{-6}}$	&	-1.402620 	&	-0.0114132 	&	$\mathrm{3.25900 \times 10^{-4}}$ 	\\
+		 5.0000	&	$\mathrm{1.12535 \times 10^{-7}}$	&	-0.200000	&	-2.0000000 	&	$\mathrm{4.73795 \times 10^{-3}}$ 	\\
+		 7.0000	&	$\mathrm{2.00000 \times 10^{-12}}$	&	-3.000000	&	-3.0000000	&	$\mathrm{1.73795 \times 10^{-3}}$
 	\end{tabular}
-	\caption{}
+	\caption{Initial points used for minimization procedure from section \ref{sec:parameters}.}
+	\label{tab:fitInitialPoints}
 \end{table}
@@
 We use the Nelder-Mead (downhill simplex) method for the minimization procedure.
-Initial points were chosen to include the areas in parameter space where each cost is close to $0$ (first $4$ points correspondingly), and to cover the large fraction of our parameter space:
-\begin{itemize}
-	\item{$n = 22.1525, \theta_0 = 0.000216424, k = -0.417021, \alpha = -1.35163, \chi = 0.00704982$}
-	\item{$n = 25.864, \theta_0 = 4.90222 \times 10^{-8}, k = -2.90875, \alpha = 3.53801, \chi = 0.00028657$}
-	\item{$n = 17.538, \theta_0 = 0.00146284, k = -0.20793, \alpha = -2.1336, \chi = 0.00152624$}
-	\item{$n = 17.9736, \theta_0 = 4.52631 \times 10^{-6}, k = -1.40262, \alpha = -0.0114132, \chi = 0.0003259$}
-	\item{$n = 5, \theta_0 = 1.12535 \times 10^{-7}, k = -0.2, \alpha = -2.0, \chi = 0.00473795$}
-	\item{$n = 7, \theta_0 = 2 \times 10^{-12}, k = -3, \alpha = -3, \chi = 0.00173795$}
-\end{itemize}
+Initial points were chosen to include the areas in parameter space where each cost is close to $0$ (first $4$ points correspondingly), and to cover the large fraction of the parameter space.
+You can see the values in table \ref{tab:fitInitialPoints}.
 
@@
 \begin{itemize}
-		\item{p}
-		\item{p}
-		\item{p}
-		\item{p}
-		\item{p}
-		\item{p}
-		\item{p}
-		\item{p}
-		\item{p}
-		\item{p}
+		\item{$\gamma = 300$}
+		\item{$\eta_0 = \unit[8.14796255480381 \times 10^{34}]{sec^{-3} GeV^{-1}}$}
+		\item{$r_0 = \unit[3.357930930824906 \times 10^6]{sec}$}
+		\item{$\bm{n = 22.5269}$}
+		\item{$\omega_0 = \unit[1]{GeV}$}
+		\item{$\bm{\theta_0 = 5.02035 \times 10^{-5}}$}
+		\item{$\bm{k = -0.699813}$}
+		\item{$\bm{\alpha = -0.626944}$}
+		\item{$z = 2.1062$}
+		\item{$\bm{\chi = 0.00541037}$}
 \end{itemize}

### 2014-06-21 Paper/gammaRays.tex
Revised abstract

@@
 	We study their light curves in two bands, $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.
+
 	The Kolmogorov-Smirnov test is used to check whether the light curves for these two bands are the same.
@@
 	The main assumption is the jet opening angle dependence on radiation energy -- the most energetic photons are emitted near the axis of the jet.
+	We also assume that all bursts are the same in their rest frames (that is their light curves differ only because of different redshifts and different off-axis angles of observers).
 	To test this model, we compute the total energy of the burst, and confirm that it is below the constraint.
@@
 	This fraction matches the observations.
-Finally, we predict the distribution of observable stretching factors, which may be tested in the future when more observational data will be available.
+	We predict the distribution of observable stretching factors, which may be tested in the future when more observational data become available.
+	Finally, we propose a way to estimate observer's off-axis angles based on stretching factors and fractions of high energy photons.
 \end{abstract}

### 2014-06-21 Paper/gammaRays.tex
Added line brakes to Introduction
Revised Introduction

@@
 	For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}.
-This suggests that most of the bursts must have similar energetics in their rest frame.
-Temporal variations of spectra were also studied.
-The spectral lags were found between different low energy bands \cite{Yi:2005ht}.
+	This suggests that most of the bursts must have similar energetics in their rest frames.
+	Temporal variations of spectra were also studied. The spectral lags were found between different low energy bands \cite{Yi:2005ht}.
 	The very high energy radiation was discovered to be extended relative to x-ray emission \cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}.
-Moreover, the Large Area Telescope (LAT) of Fermi allows us to explore the spectral variations even between high energy bands like $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
-In this study we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
+	We decided to elaborate on that result, and, using the data from Large Area Telescope (LAT) of Fermi, explore the spectral variations between different high energy bands, specifically $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
+	In particular, we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
 
@@
 	We also download events during a day before the burst, and use the technique introduced in \cite{Rubtsov:2011qq} to estimate background radiation in both energy bands.
+
 	Most of the bursts in the catalog, however, do not have enough events with over $\unit[1]{GeV}$ energies to do desired computations, so we choose only those bursts, from which at least $10$ photons were detected in the high energy band.
@@
 
-Second, consider the plasma right after its ejection from the central engine. Two processes happen there simultaneously:
+	Second, consider the plasma right after its ejection from the central engine.
+	Two processes happen there simultaneously:
 	\begin{enumerate}
@@
 		\caption{
-		High and low energy light curves produced by the geometrical model. Burst parameter values are the same as discussed in section \ref{sec:parameters}.
+			High and low energy light curves produced by the geometrical model.
+			Burst parameter values are the same as discussed in section \ref{sec:parameters}.
 		}

### 2014-06-21 Paper/gammaRays.tex
Changed 3 to 4 observations in Introduction

@@
 	Most of the bursts in the catalog, however, do not have enough events with over $\unit[1]{GeV}$ energies to do desired computations, so we choose only those bursts, from which at least $10$ photons were detected in the high energy band.
-	This leaves $3$ of them: 080916C \cite{Tajima:2009az}, 090902B \cite{Abdo:2009pg} and 090926A \cite{Bregeon:2011bu}.
-
-	For these 3 bursts we compute both high and low energy distributions of photon times.
-	We subtract the low energy background estimate from the low energy CDF, and add the high energy background estimate to it.
-	This makes the low energy CDF non-monotonous and, rigorously speaking, we cannot use the 2-sample Kolmogorov-Smirnov test on it.
-	However, since there is a much greater number of photons on lower energies, we can think of the low energy CDF as continuous.
-	Therefore, this non-monotonicity will be negligible, and will not harm the KS-test results much.
+	It leaves $4$ of them: 080916C \cite{Tajima:2009az}, 090510 \cite{Ackermann:2010us}, 090902B \cite{Abdo:2009pg} and 090926A \cite{Bregeon:2011bu}.
+
+	For these 4 bursts we compute both high and low energy distributions of photon times, and subtract the background estimates from them.
+	This makes CDFs non-monotonous and, rigorously speaking, we cannot use the 2-sample Kolmogorov-Smirnov test on it.
+	However, since the number of photons is much higher than estimated background, this non-monotonicity will be negligible, and will not harm the KS-test results much.
 

### 2014-06-21 Paper/gammaRays.tex
Revised prohibited stretching factors

@@
 
-	For these 4 bursts we compute both high and low energy distributions of photon times, and subtract the background estimates from them.
+	For these 4 bursts we compute both high and low energy distributions of photon times, and subtract background estimates from them.
 	This makes CDFs non-monotonous and, rigorously speaking, we cannot use the 2-sample Kolmogorov-Smirnov test on it.
-	However, since the number of photons is much higher than estimated background, this non-monotonicity will be negligible, and will not harm the KS-test results much.
+	However, since the number of photons is much higher than estimated background, this non-monotonicity is negligible, and will not harm KS-test results much.
 
 	Finally, we stretch the high energy CDF by different factors, and compare it to the low energy CDF using the KS-test.
-	If we require $2\sigma$-significant probability to exclude a particular stretching factor, then the stretching factor of $1$ (which means no stretching) is allowed for GRBs 080916C and 090902B.
-	However, for 090926A, the stretching factor of $1$ is prohibited, so the high energy light curve should be stretched with respect to the low energy one.
+	If we require $2\sigma$-significant probability to exclude a particular stretching factor, then the stretching factor of $1$ (which means no stretching) is allowed for GRBs 080916C and 090510.
+	For the other two bursts the stretching factor of 1 is, however, prohibited.
+	For the GRB 090902B it is smaller than 1, so the low energy light curve is stretched with respect to the high energy one (see fig. \ref{fig:grb090902B}).
+	And for the GRB 090926A even more dramatic deviation from the stretching factor of 1 is observed in the opposite direction.
+	The high energy light curve of GRB 090926A is stretched with respect to the low energy one by at least a factor of 1.99 (see fig. \ref{fig:grb090926A}).
+	All observed stretching factors are summarized in table \ref{tab:observationResults}.
 

### 2014-06-23 Paper/gammaRays.tex
Added incomplete model results and discussion

@@
 	For the other two bursts the stretching factor of 1 is, however, prohibited.
-	For the GRB 090902B it is smaller than 1, so the low energy light curve is stretched with respect to the high energy one (see fig. \ref{fig:grb090902B}).
-	And for the GRB 090926A even more dramatic deviation from the stretching factor of 1 is observed in the opposite direction.
+	For GRB 090902B it is smaller than 1, so the low energy light curve is stretched with respect to the high energy one (see fig. \ref{fig:grb090902B}).
+	And for GRB 090926A even more dramatic deviation from the stretching factor of 1 is observed in the opposite direction.
 	The high energy light curve of GRB 090926A is stretched with respect to the low energy one by at least a factor of 1.99 (see fig. \ref{fig:grb090926A}).
@@
 	However, only about 30 of them were detected by the LAT, and only 3 of them were bright in the high energy band.
-	If we extrapolate the uniform jet model to very high energies, this observation would mean that these groups of bursts are internally different: some of them produce VHE radiation, and other do not.
+	If we extrapolate the uniform jet model to very high energies, this observation would mean that these groups of bursts are internally different: some of them produce VHE radiation, while others do not.
 	These differences are hard to explain given that burst energetics are similar \cite{Bloom:2003wy}.
@@
 	In our model, the opening angle of a jet is inversely proportional to the energy of photons it radiates.
-	Therefore the most common scenario is that the off-axis angle of the observer is smaller than the low energy jet opening angle, but larger than the opening angle of a high energy jet.
+	Therefore the most common scenario is that the off-axis angle of an observer is smaller than the low energy jet opening angle, but much larger than the opening angle of a high energy jet.
 	Because of that, most of the bursts can only be seen at low energies.
-	The 3 bursts we study in this paper, were seen, according to our model, from the lowest off-axis angles.
+	The 4 bursts we study in this paper were seen, according to our model, from the lowest off-axis angles.
 
@@
         	Model Overview.
-        	Here the red and the blue cones represent the volumes through which low and high energy plasma propagates.
+        	Here red and blue cones represent the regions through which low and high energy plasma propagates.
         	In case depicted the observer's off-axis angle is smaller than the opening angle of the low energy jet, so, due to the relativistic beaming effect, the most of the observable low energy photons will travel along the straight line from the central engine.
@@
 	Finally, our assumption is able to explain the non-trivial stretching factor of both GRBs 090902B and 090926A.
-	Detailed computations will be described in the following sections, but you can see the main idea on the figure \ref{fig:modelOverview}.
+	Detailed computations will be described in the following sections, but you can see the main idea on figure \ref{fig:modelOverview}.
 	The computations confirm that this qualitative picture is indeed correct.
-	You can see a sample light curves on the figure \ref{fig:sampleLightCurves}, and predicted distribution of stretching factors on the figure \ref{fig:kappaDistribution}.
+	You can see a sample light curves on figure \ref{fig:sampleLightCurves}, and predicted distribution of stretching factors on figure \ref{fig:kappaDistribution}.
 
 \section{Observations}
+\label{sec:observations}
 
@@
 	\begin{tabular}{ l | S[table-format=9.3] | S[table-format=3.3] | S[table-format=-2.3] | S[table-format=1.5] | S[table-format=2.1] | S[table-format=3.1] }
-		\multicolumn{1}{ c |}{\texttt{name}} & $\mathrm{\texttt{time}}$ & \texttt{location.ra} & \texttt{location.dec} & \texttt{location.err} & \texttt{startOffset} & \texttt{endOffset} \\
 		\multicolumn{1}{ c |}{GRB} & $\mathrm{GBM\ Trigger\ Time}$ & $\mathrm{R.A.}$ & $\mathrm{Dec.}$ & $\mathrm{Location\ Error}$ & $T_{05}$ & $T_{95}$ \\
 		\multicolumn{1}{ c |}{Name}& $\mathrm{MET,\ sec}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{deg}$ & $\mathrm{sec}$ & $\mathrm{sec}$ \\
+		\multicolumn{1}{ c |}{\texttt{name}} & $\mathrm{\texttt{time}}$ & \texttt{location.ra} & \texttt{location.dec} & \texttt{location.err} & \texttt{startOffset} & \texttt{endOffset} \\
 		\hline
@@
 	\caption{Burst data used in our study.
-The first row contains the corresponding \texttt{GRBurst} class variable names.
+The third row contains the corresponding \texttt{GRBurst} class variable names.
 The data is taken from the tables 2 and 4 of \cite{Ackermann:2013zfa}.
@@
 \begin{itemize}
-	\item{GBM trigger time (table 2). It is used as a reference point for the GRB time.}
-	\item{Location of the GRB (table 2). We need burst locations to filter out photons coming from other sources.}
-	\item{Location error (table 2). Location errors are used to improve accuracy of filtering, specifically to avoid losing statistics by filtering too much.}
-	\item{$T_{05}$ -- $T_{95}$ interval of the LAT-detected emission (table 4).
+	\item{
+		GBM trigger time (table 2).
+		It is used as a reference point for the GRB time.
+	}
+	\item{
+		Location of the GRB (table 2).
+		We need burst locations to filter out photons coming from other sources.
+	}
+	\item{
+		Location error (table 2).
+		Location errors are used to improve accuracy of filtering, specifically to avoid losing statistics by filtering too much.
+	}
+	\item{
+		$T_{05}$ -- $T_{95}$ interval of the LAT-detected emission (table 4).
 		We need precise intervals for two things.
 		First, to understand where is the fixed point of the stretching ($T_{05}$ is used as a proxy for that).
-Second, the burst interval is used to download events data from the data server.}
+		Second, the burst interval is used to download events data from the data server.
+	}
 \end{itemize}
-The data obtained is summarized in the table \ref{tab:bursts}.
+The data obtained is summarized in table \ref{tab:bursts}.
 
@@
 \begin{itemize}
-	\item{{\bf Object name or coordinates}. Here the coordinates are filled from the table \ref{tab:bursts}.}
-	\item{{\bf Coordinate system}. J2000.}
-	\item{{\bf Search radius (degrees)}. $60$. Events will be filtered by location later on by using the point spread functions (PSFs) of the LAT.}
+	\item{
+		{\bf Object name or coordinates}.
+		Here the coordinates are filled from the table \ref{tab:bursts}.
+	}
+	\item{
+		{\bf Coordinate system}.
+		J2000.
+	}
+	\item{
+		{\bf Search radius (degrees)}.
+		$60$.
+		Events will be filtered by location later by using the point spread functions (PSFs) of the LAT.
+	}
 	\item{
 		{\bf Observation dates}.
-To have a decent safety margin, we extend the duration of the burst by $15\%$ to both past and future relative to the table \ref{tab:bursts} time ranges.
+		To have a decent safety margin, we extend the duration of the burst by $50\%$ to both past and future relative to the table \ref{tab:bursts} time ranges.
 		So, we fill in the following values:
 		\begin{align*}
-			\texttt{time} + \texttt{startOffset} &- 0.15\left(\texttt{endOffset}-\texttt{startOffset}\right),\\
-			\texttt{time} + \texttt{endOffset} &+ 0.15\left(\texttt{endOffset}-\texttt{startOffset}\right)
+			\texttt{time} + \texttt{startOffset} &- 0.5\left(\texttt{endOffset}-\texttt{startOffset}\right),\\
+			\texttt{time} + \texttt{endOffset} &+ 0.5\left(\texttt{endOffset}-\texttt{startOffset}\right)
 		\end{align*}
 	}
-	\item{{\bf Time system}. MET.}
-	\item{{\bf Energy range (MeV)}. $30,\,300000$. The whole energy range is required for our study.}
-	\item{{\bf LAT data type}. Extended, because we need transient photons for better statistics.}
-	\item{{\bf Spacecraft data}. Checked. It is required for both events filtering, and calculation of the PSFs and exposure maps.}
+	\item{
+		{\bf Time system}.
+		MET.
+	}
+	\item{
+		{\bf Energy range (MeV)}.
+		$30,\,300000$.
+		The whole energy range is required for our study.
+	}
+	\item{
+		{\bf LAT data type}.
+		Extended.
+	}
+	\item{
+		{\bf Spacecraft data}.
+		Checked.
+		It is required for both events filtering, and calculation of the PSFs and exposure maps.
+	}
 \end{itemize}
 
-For background estimation, we also download observational data during a day before the burst. All the values are put the same into the form, except the time range:
+For background estimation, we also download observational data during a day before the burst.
+All the values are put the same into the form, except the time range:
 \begin{align*}
-	\texttt{time} + \texttt{startOffset} &- 0.15\left(\texttt{endOffset} - \texttt{startOffset}\right) - \unit[86400]{sec},\\
-	\texttt{time} + \texttt{startOffset} &- 0.15\left(\texttt{endOffset} - \texttt{startOffset}\right)
+	\texttt{time} + \texttt{startOffset} &- 0.5\left(\texttt{endOffset} - \texttt{startOffset}\right) - 86400,\\
+	\texttt{time} + \texttt{startOffset} &- 0.5\left(\texttt{endOffset} - \texttt{startOffset}\right)
 \end{align*}
-All other values filled in the form are the same.
 
@@
 This file, however, contains observations of a large area of the sky.
-Before using the data obtained from it, we should perform a more elaborate filtering by location using Fermi LAT's point spread function, which will be described in the following section.
+Before using the data obtained from it, we should perform a more elaborate filtering by location using Fermi LAT's point spread function.
+The procedure is described in the following section.
 
@@
 \end{lstlisting}
-Here \texttt{[RA]} and \texttt{[DEC]} are the coordinates of the burst, and \texttt{[IrfName]} is the Instrument Response Function, which depends on photon's event class (transient, source, clean, or ultraclean) and conversion type (back, or front).
-The list of IRF names can be obtained by using the \texttt{gtirfs}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtirfs.txt}} tool.
+Here \texttt{[RA]} and \texttt{[DEC]} are the coordinates of the burst, and \texttt{[IrfName]} is the Instrument Response Function, which depends on photon's event class (transient, source, clean, or ultraclean) and conversion type (back or front).
+The list of IRF names can be obtained with the \texttt{gtirfs}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtirfs.txt}} tool.
 We need to compute PSFs for all of them.
 
-Imagine a stream of photons coming from some source locate at some coordinates $\theta, \phi$.
+To understand how to use obtained PSFs, imagine a stream of photons coming from some source located at some coordinates $\theta, \phi$.
 Ideally, all the coordinates measured for the photons coming from the source should be the same.
@@
 If we find the region such that, say, $95\%$ of photons coming from the source are measured inside of it, we can filter out all the other photons without losing too much statistics, and significantly reducing background.
-\texttt{gtpsf} approximates these regions as circles, and the files generated in the previous step contain the observation probability densities as functions of distance from the source, and photon energies.
+\texttt{gtpsf} approximates these regions as circles, and the files generated in the previous step contain the observation probability densities as functions of distance from the source and photon energies.
 We, however, need an integral quantity, that is given the radius, we want to know the fraction of photons observed inside the circle of this radius.
 We can compute this fraction by summing over the probability densities:
-\begin{equation*}
-	p\left(n\right) = \sum_{i=0}^{n} \frac{\texttt{pdf[i]} + \texttt{pdf[i+1]}}{2} 2\pi \left(\cos\left(\texttt{angles[i]}\right) - \cos\left(\texttt{angles[i+1]}\right)\right)
-\end{equation*}
+\begin{align*}
+	p\left(n\right) &= \int_0^n \texttt{pdf[i]} 2\pi \sin\left(\texttt{angles[i]}\right) \dif i \\
+	&\approx \sum_{i=0}^{n} \frac{\texttt{pdf[i]} + \texttt{pdf[i+1]}}{2} 2\pi \left(\cos\left(\texttt{angles[i]}\right) - \cos\left(\texttt{angles[i+1]}\right)\right)
+\end{align*}
 Here \texttt{pdf[i]} is the probability density in the point separated from the source by an angle \texttt{angles[i]}.
@@
 However, some of the photons in this set are still originated from background, and not from the source.
-We cannot filter out these photons, however, we can substitute a linear component of background by estimating it from a long time period before the burst.
+We cannot filter out these photons, but we can substitute a linear component of background by estimating it from a long time period before the burst.
 In order to perform this computation, we'll need an exposure map, which will be discussed in the following section.
@@
 \end{lstlisting}
-The generated \texttt{expcube} files contain exposures as functions of energy and location on the sky\footnote{Note, that in the \texttt{expcube} FITS files R.A.
-coordinates are indexed in reverse order and starting from 180, for example, $\texttt{ra[0]}=180$, $\texttt{ra[1]}=179$, and so on.}.
+The generated \texttt{expcube} files contain exposures as functions of energy and location on the sky\footnote{Note, that R.A.
+coordinates in the \texttt{expcube} FITS files are indexed in reverse order and starting from 180.
+For example, $\texttt{ra[0]}=180$, $\texttt{ra[1]}=179$, and so on.}.
 We use trilinear interpolation to compute exposures for all energies and locations.
@@
 We compute two background estimates: for low and high energy bands.
-Because the background of transient class photons may depend on the spacecraft position, we have to filter them as well.
+Because the background of transient class photons may depend on the spacecraft position, we have to filter them completely.
 
@@
 
-As a result of such calculation we can constraint the range of allowed stretching factors to values for which the KS-probabilities are not too small, in other words, given the sigma-significance value, compute the range outside of which stretching factors are elliminated by observations.
+As a result of such calculation we can constrain the range of allowed stretching factors to values for which the KS-probabilities are not too small, in other words, given the sigma-significance value, compute the range outside of which stretching factors are elliminated by observations.
 
@@
 
-So we obtained a preliminary (for significance is less than $5\sigma$) result that the stretching factors for observable bursts might be both larger and smaller than $1$.
+So we obtained a preliminary (for significance is only $2\sigma$) result that the stretching factors for observable bursts might be both larger and smaller than $1$.
 
@@
         \end{subfigure}
-        \caption{GRB 090902B results. Low energy radiation is stretched ($\kappa < 1$) with significance of $2\sigma$.}
+        \caption{GRB 090902B results. Low energy radiation is stretched ($\kappa < 1$) with significance of $2.2\sigma$.}
         \label{fig:grb090902B}
@@
         \end{subfigure}
-        \caption{GRB 090926A results. High energy radiation is stretched ($\kappa > 1$) with significance of $3\sigma$.}
+        \caption{GRB 090926A results. High energy radiation is stretched ($\kappa > 1$) with significance of $3.3\sigma$.}
         \label{fig:grb090926A}
@@
 \end{itemize}
-
-Numerical values we used for these parameters are summarized in the table \ref{tab:parameters}.
-
-\begin{table}
-	\centering
-	\small
-	\begin{tabular}{ r | l | l }
-		Parameter	&	Value			&	Comments \\
-		\hline
-		$\gamma$		&	$100$		&	\\
-		$\eta_0$		&	$10^{45}$	&
-	\end{tabular}
-	\caption{Parameter values.}
-	\label{tab:parameters}
-\end{table}
 
@@
 
+	TODO: Connect this part.
+
+	Numerical simulations based on these assumptions are able to reproduce both stretching factors smaller than $1$ (like of GRB 090902B) and larger than $1$ (like of GRB 090926A) depending on redshift and observer's off-axis angle (see fig.
+\ref{fig:sampleLightCurves}).
+
+	To test the model against observations, we computed a few more things:
+	\begin{itemize}
+		\item{
+			Total energy emitted in $\unit[100]{MeV}$ and more energetic gamma rays $E < \unit[5.89 \times 10^{53}]{GeV}$, which is in agreement with \cite{Gehrels:2013xd}.
+		}
+		\item{
+			Distribution of stretching factors of observable bursts.
+			You can see the PDF and the CDF of it on fig. \ref{fig:kappaDistribution}.
+			This distribution doesn't contradict with the values of stretching factors obtained in section \ref{sec:observations}.
+		}
+		\item{
+			The fraction of bursts observable in low energy band, which are also observed in high energy band $f_m = 0.116$.
+			The Fermi LAT catalog contains $35$ bursts out of which $4$ can be observed in high energy band.
+			Therefore the observed value $f_o = 0.11 \pm 0.05$ (error obtained from binomial distribution) agrees with the model.
+		}
+		\item{
+			There is a correlation between the fraction of high energy photons (that is high energy photon count divided by the total photon count), the stretching factor and the observer's off-axis angle (see fig.
+\ref{fig:correlations}).
+It allows one to predict observer's off-axis angles.
+		}
+	\end{itemize}
+
+	So our model, though with some discrepancies, is able to explain the time stretching result from section \ref{sec:observations}.
+
 \begin{figure}
@@
 
+\label{sec:discussion}
+
+	The main results of our study may be summarized as following:
+	\begin{enumerate}
+		\item{
+			The time stretching of GRB light curves between different VHE bands is confirmed with statistical significance of $3.3\sigma$.
+		}
+		\item{
+			This time stretching can be explained by geometry of the jet, without introducing any new spectral components.
+		}
+		\item{
+			All GRBs are assumed to be the same in their rest frames.
+			There is no need to vary internal burst parameters.
+		}
+	\end{enumerate}
+
+	There are, however, directions for improvement:
+	\begin{itemize}
+		\item{
+			The model prediction and observations on fig. \ref{fig:correlations} are, though close, not exactly agree.
+		}
+		\item{
+			Shapes of light curves produced by the model differ considerably from observed ones.
+		}
+		\item{
+			Assumption of energy dependence on off-axis angle should be better justified.
+		}
+		\item{
+			While we believe that all bursts are almost the same in their rest frames, they are certainly not exactly the same.
+			The minor differences between GRBs might be responsible for observable deviations from the model predictions.
+		}
+	\end{itemize}
+
+	We believe, however, that these problems are not unresolvable, and may be addressed by refining the model without changing its main assumptions.
+	And in the end, we think the main ideas of this paper will be close to truth.
+
+	{\small {\it Acknowledgments.} The work was supported by Russian Science Foundation grant 14-12-01340.}
+
 \bibliographystyle{hep}

### 2014-07-21 Poster/main.tex
Added poster

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Jacobs Landscape Poster
% LaTeX Template
% Version 1.1 (14/06/14)
%
% Created by:
% Computational Physics and Biophysics Group, Jacobs University
% https://teamwork.jacobs-university.de:8443/confluence/display/CoPandBiG/LaTeX+Poster
% 
% Further modified by:
% Nathaniel Johnston (nathaniel@njohnston.ca)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[final]{beamer}

\usepackage[size=a0,orientation=portrait,scale=1.24]{beamerposter} % Use the beamerposter package for laying out the poster

\usetheme{confposter} % Use the confposter theme supplied with this template

\setbeamercolor{block title}{fg=ngreen,bg=white} % Colors of the block titles
\setbeamercolor{block body}{fg=black,bg=white} % Colors of the body of blocks
\setbeamercolor{block alerted title}{fg=white,bg=dblue!70} % Colors of the highlighted block titles
\setbeamercolor{block alerted body}{fg=black,bg=dblue!10} % Colors of the body of highlighted blocks
% Many more colors are available for use in beamerthemeconfposter.sty

%-----------------------------------------------------------
% Define the column widths and overall poster size
% To set effective sepwid, onecolwid and twocolwid values, first choose how many columns you want and how much separation you want between columns
% In this template, the separation width chosen is 0.024 of the paper width and a 4-column layout
% onecolwid should therefore be (1-(# of columns+1)*sepwid)/# of columns e.g. (1-(4+1)*0.024)/4 = 0.22
% Set twocolwid to be (2*onecolwid)+sepwid = 0.464
% Set threecolwid to be (3*onecolwid)+2*sepwid = 0.708

\newlength{\sepwid}
\newlength{\onecolwid}
\newlength{\twocolwid}
\newlength{\threecolwid}
\setlength{\paperwidth}{33.1in} % A0 width: 46.8in
\setlength{\paperheight}{46.8in} % A0 height: 33.1in
\setlength{\sepwid}{0.066\paperwidth} % Separation width (white space) between columns
\setlength{\onecolwid}{0.4\paperwidth} % Width of one column
\setlength{\twocolwid}{0.4\paperwidth} % Width of two columns
\setlength{\threecolwid}{0.8\paperwidth} % Width of three columns
\setlength{\topmargin}{-0.5in} % Reduce the top margin size
%-----------------------------------------------------------

\usepackage{graphicx}  % Required for including images

\usepackage{booktabs} % Top and bottom rules for tables

\usepackage{units}
\usepackage{subfig}

%----------------------------------------------------------------------------------------
%	TITLE SECTION 
%----------------------------------------------------------------------------------------

\title{Time Stretching of GeV Emission from GRBs:\\[0.2ex]Fermi LAT vs Geometrical Model} % Poster title

\author{Maxim Piskunov and Grigory Rubtsov \\[.3ex] \normalsize \href{mailto:maxit@ms2.inr.ac.ru}{maxit@ms2.inr.ac.ru}} % Author(s)

\institute{Institute for Nuclear Research RAS \\[2.0ex] \normalsize \url{https://github.com/maxitg/GammaRays}} % Institution(s)

%----------------------------------------------------------------------------------------

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{2ex}} % White space under blocks
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}} % White space under highlighted (alert) blocks

\setlength{\belowcaptionskip}{2ex} % White space under figures
\setlength\belowdisplayshortskip{2ex} % White space under equations

\begin{frame}[t] % The whole poster is enclosed in one beamer frame

\begin{columns}[t] % The whole poster consists of three major columns, the second of which is split into two columns twice - the [t] option aligns each column's content to the top

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\onecolwid} % The first column

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\begin{block}{Introduction}

Observations confirm that the high energy $(> \unit[100]{MeV})$ emission of gamma ray bursts is delayed with respect to the low energy emission.
However, the difference of light curves in various high energy bands has not been studied properly.
We study GRBs in two energy bands: \unit[100]{MeV} to \unit[1]{GeV}, and \unit[1]{GeV} to \unit[300]{GeV}.

\end{block}

%------------------------------------------------

\begin{block}{Observations}

\begin{figure}
{\large Kolmogorov-Smirnov test results for GRB 090926A} \\[0.5ex]
\subfloat{\includegraphics[width=\textwidth]{lightCurveArrow090926A}} \\
\subfloat{\includegraphics[width=\textwidth]{lightCurveStretchedArrow090926A.pdf}} \\
\end{figure}

\begin{itemize}
\item{4 GRBs were studied: 080916C, 090510, 090902B and 090926A.}
\item{080916C and 090902B have stretching factors compatible with 1 within $2\sigma$.}
\item{High energy light curve of 090926A is stretched with respect to low energy one (with $3.3\sigma$ significance)}
\item{Low energy light curve of 090902B is stretched with respect to high energy one (with $2.2\sigma$ significance)}
\end{itemize}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of the first column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\begin{column}{\twocolwid} % Begin a column which is two columns wide (column 2)

\begin{columns}[t,totalwidth=\twocolwid] % Split up the two columns wide column

\begin{column}{\onecolwid}\vspace{-.6in} % The first column within column 2 (column 2.1)

%----------------------------------------------------------------------------------------
%	MATERIALS
%----------------------------------------------------------------------------------------

\begin{block}{Model}

\begin{figure}
\includegraphics[width=\textwidth]{modelOverview.pdf}
\end{figure}

We suggest a simple geometrical model to explain this result.
The main hypothesis is the jet opening angle dependence on radiation energy -- the most energetic photons are emitted near the axis of the jet.

\begin{enumerate}
	\item{Time $t = 0$, the central engine emits a spherical shell of plasma.}
	\item{The shell propagates with relativistic velocity.}
	\item{Each jet point is an isotropic radiator in its inertial reference frame.}
	\item{
		Intensity depends on coordinates in space and frequencies:
		\begin{equation*}
			\eta\left(r,\theta,\omega\right) = 
			\frac{\eta_0}{1 + \left(r/r_0\right)^n}
			e^{
				-\left(\theta/\theta_0\right)^2
				\left(\omega/\omega_0\right)^{-2k}
			}
			\left(\omega/\omega_0\right)^\alpha
		\end{equation*}
	}
\end{enumerate}

\end{block}

\begin{block}{Results}

The simple geometrical model can explain observed stretching:

\begin{figure}
	\includegraphics[width=\textwidth]{kappaDistributionHistogram}
\end{figure}

We also computed (for some parameter values):

\begin{itemize}
	\item{Total energy of the burst above \unit[1]{GeV}. The result is reasonable.}
	\item{Fraction of bursts observable above \unit[100]{MeV}, which are also observed above \unit[1]{GeV} is $f_m = 0.072$. Observed value is $f_o = 0.086$.}
\end{itemize}

\end{block}

%----------------------------------------------------------------------------------------

\end{column} % End of column 2.1

\end{columns} % End of the split of column 2 - any content after this will now take up 2 columns width

%----------------------------------------------------------------------------------------

\end{column} % End of the second column

\begin{column}{\sepwid}\end{column} % Empty spacer column

\end{columns} % End of all the columns in the poster

\end{frame} % End of the enclosing frame

\end{document}

### 2014-08-11 Paper/gammaRays.tex
Small refinements throughout the paper

@@
 \item{Time $t = 0$, a spherical shell of plasma is emitted. The center is called the central engine.}
-\item{The shell points propagate with a constant velocity $v \sim 1$, so at the time $t$ the radius of the shall is $v t$.}
+\item{The shell points propagate with a constant velocity $v = \frac{\sqrt{\gamma^2 - 1}}{\gamma} \sim 1$, so at the time $t$ the radius of the shall is $v t$.}
 \item{Each point of the shell is an isotropic radiator in its rest frame.}
@@
 \begin{itemize}
-\item{$v$, the velocity of the shell, $\gamma = \frac{1}{\sqrt{1-v^2}} \gg 1$.}
+\item{$\gamma$, the relativistic factor of the shell, $\gamma \gg 1$.}
 \item{$\eta_0$, which defines the luminosity scale.}
@@
 This time is a function of the radiator location $\left(r, \theta, \phi\right)$, as well as the observer location $\left(d, \chi, 0\right)$.
-(Here we choose coordinates so that the rotation angle of the observer is $0$.) Lets assume for now that the observer is far enough to resolve the jet geometry, yet close enough so that the expansion of space is negligible.
+(Here we choose coordinates so that the rotation angle of the observer is $0$.) Lets assume for now that the observer is too far from the jet to resolve its geometry, yet close enough so that the expansion of space is negligible.
 This assumption allows us to ignore the effects of cosmology for this calculation.
@@
 Note that $\tau$ doesn't depend on $d$ anymore.
-But this is only true if the scale factor did not change since emission to observation.
+But this is only true if the scale factor did not change from emission to observation.
 For a distant observer we should account for the change in the scale factor, which will stretch the distances between photons:
@@
 \end{align*}
-Here $H_\text{obs} = H\left(t_\text{obs}\right) = \dot{a}\left( t_\text{obs} \right)$ is the Hubble parameter at the observation time.
+Here $H_\text{obs} = H\left(t_\text{obs}\right) = \frac{\dot{a}\left( t_\text{obs} \right)}{a\left( t_\text{obs} \right)} = \dot{a}\left( t_\text{obs} \right)$ is the Hubble parameter at the observation time.
 
@@
 For that we need to integrate the radiation intensity $\eta$ over four things.
-Two burst-related things: radiator positions and frequencies at emission.
-Two observer-related things: observation time and frequencies at observation.
+Two burst-related things: radiator positions $\left(r, \theta, \phi\right)$ and frequencies at emission $\omega'$.
+Two observer-related things: observation time $\tau$ and frequencies at observation $\omega$.
 We relate the first two and the second two with the delta functions:
@@
 
-Furthermore, we can do the integrals over $\omega$ and $\tau$, if we write $\eta$ explicitly:
+Furthermore, if we write $\eta$ explicitly we can do the integrals over $\omega$ and $\tau$:
 \begin{align}
@@
 \end{align}
-Here $I$ and $J$ are the indefinite integrals over $\omega$ and $\tau$:
+Here $I$ and $J$ are indefinite integrals over $\omega$ and $\tau$:
 \begin{align}
@@
 	\item{Redshifts $z$ of many bursts have been measured \cite{Ackermann:2013zfa}.}
-	\item{We know that Lorentz factors $\gamma$ are on the order of magnitude\footnote{Our minimization procedure can find similar minimums for other values of $\gamma$, at least from $100$ to $1000$} of $\gamma = 300$ \cite{Ghirlanda:2011ux}.}
+	\item{We know that relativistic factors $\gamma$ are on the order of magnitude\footnote{Our minimization procedure can find similar minimums for other values of $\gamma$, at least from $100$ to $1000$} of $\gamma = 300$ \cite{Ghirlanda:2011ux}.}
 \end{itemize}
@@
 
-We want then our cost function $C\left(n, \theta_0, k, \alpha, \chi\right)$ to satisfy the following objectives:
+We want our cost function $C\left(n, \theta_0, k, \alpha, \chi\right)$ to satisfy the following objectives:
 \begin{itemize}
@@
 	\item {
-		Finally, the burst should not be too faint compared to other bursts from the sample. That is the total number of observed photons should approximately equal to the median among the sample.
+		Finally, the burst should not be too faint compared to other bursts from the sample. In other words, the total number of observed photons should approximately equal to the median among the sample.
 	}
@@
 
+Note, that these parameter values (except for $z$ and $\chi$) are universal, and they can describe the whole population of observed bursts, as will be shown in the next section.
+
 \subsection{Results}
 
-	TODO: Connect this part.
-
-	Numerical simulations based on these assumptions are able to reproduce both stretching factors smaller than $1$ (like of GRB 090902B) and larger than $1$ (like of GRB 090926A) depending on redshift and observer's off-axis angle (see fig.
+	As was discussed in the previous section our model can reproduce both stretching factors smaller than $1$ (like of GRB 090902B) and larger than $1$ (like of GRB 090926A) depending on redshift and observer's off-axis angle (see fig.
 \ref{fig:sampleLightCurves}).
@@
         \includegraphics[width=1.0\textwidth]{correlations}
-        \caption{Correlations between off-axis angles, stretching factors and high to low energy photon count ratios found in the sample produces by our model.
+        \caption{Correlations between off-axis angles, stretching factors and high to low energy photon count ratios found in the sample produced by our model.
 The sample contains $4096$ bursts.
-Black crosses show observations.
-This correlation allows to estimate off-axis angles of observed bursts.}
+Black crosses show observations with $2\sigma$ error bars.
+This correlation allows one to estimate off-axis angles of observed bursts.}
         \label{fig:correlations}
@@
 		\item{
-			All GRBs are assumed to be the same in their rest frames.
+			All GRBs can be assumed to be the same in their rest frames.
 			There is no need to vary internal burst parameters.
@@
 		\item{
-			Assumption of energy dependence on off-axis angle should be better justified.
+			Assumption of energy dependence on off-axis angle should be justified better.
 		}
@@
 	We believe, however, that these problems are not unresolvable, and may be addressed by refining the model without changing its main assumptions.
-	And in the end, we think the main ideas of this paper will be close to truth.
 

### 2014-08-11 GRObservations/GRObservations/GRExposureMap.cpp

@@ -47,7 +47,6 @@ float GRExposureMap::exposure(float energy, GRLocation location) {
 }
 
 float GRExposureMap::exposure(int energyIndex, GRLocation location) {
-#warning change to interpolation
     int raLow = location.ra;
     int raHigh = raLow+1;
     if (raHigh == 360) {

### 2014-08-11 GRObservations/GRObservations/GRFermiLATDataServerQuery.cpp

@@ -28,7 +28,7 @@ using namespace std;
 
 #include <curl/curl.h>
 
-#include <fitsio.h>
+#include "fitsio.h"
 
 #include "GRFermiLATDataServerQuery.h"
 
@@ -174,13 +174,13 @@ void GRFermiLATDataServerQuery::download() {
     curl_formadd(&formpost,
                     &lastptr,
                     CURLFORM_COPYNAME, "energyfield",
-                    CURLFORM_COPYCONTENTS, "30, 300000",
+                    CURLFORM_COPYCONTENTS, "100, 300000",
                     CURLFORM_END);
     
     curl_formadd(&formpost,
                     &lastptr,
                     CURLFORM_COPYNAME, "photonOrExtendedOrNone",
-                    CURLFORM_COPYCONTENTS, "Extended",
+                    CURLFORM_COPYCONTENTS, "Photon",
                     CURLFORM_END);
     
     curl_formadd(&formpost,

### 2014-08-11 GRObservations/GRObservations/GRPsf.cpp

@@ -11,7 +11,7 @@
 #include <memory>
 #include <sstream>
 
-#include <fitsio.h>
+#include "fitsio.h"
 
 #include "GRPsf.h"
 

### 2014-08-12 GRObservations/GRObservations/GRFermiLATDataServerQuery.cpp
Changed query energy range in paper

@@ -180,7 +180,7 @@ void GRFermiLATDataServerQuery::download() {
     curl_formadd(&formpost,
                     &lastptr,
                     CURLFORM_COPYNAME, "photonOrExtendedOrNone",
-                    CURLFORM_COPYCONTENTS, "Photon",
+                    CURLFORM_COPYCONTENTS, "Extended",
                     CURLFORM_END);
     
     curl_formadd(&formpost,

### 2014-08-13 Paper/gammaRays.tex
Changed query energy range in paper
Off-axis angles -> view directions in abstract

@@
 	The main assumption is the jet opening angle dependence on radiation energy -- the most energetic photons are emitted near the axis of the jet.
-	We also assume that all bursts are the same in their rest frames (that is their light curves differ only because of different redshifts and different off-axis angles of observers).
+	We also assume that all bursts are the same in their rest frames (that is their light curves differ only because of different redshifts and different view directions).
 	To test this model, we compute the total energy of the burst, and confirm that it is below the constraint.
@@
 	We predict the distribution of observable stretching factors, which may be tested in the future when more observational data become available.
-	Finally, we propose a way to estimate observer's off-axis angles based on stretching factors and fractions of high energy photons.
+	Finally, we propose a way to estimate observer's off-axis angles (view directions) based on stretching factors and fractions of high energy photons.
 \end{abstract}
@@
 		{\bf Energy range (MeV)}.
-		$30,\,300000$.
-		The whole energy range is required for our study.
+		$100,\,300000$.
+		This includes our both energy ranges.
 	}

### 2014-08-13 Paper/gammaRays.tex
Corrections to Introduction

@@
 
-	For these 4 bursts we compute both high and low energy distributions of photon times, and subtract background estimates from them.
+	For these 4 bursts we compute both high and low energy distributions of photon arrival times, and subtract background estimates from them.
 	This makes CDFs non-monotonous and, rigorously speaking, we cannot use the 2-sample Kolmogorov-Smirnov test on it.
@@
 	If we require $2\sigma$-significant probability to exclude a particular stretching factor, then the stretching factor of $1$ (which means no stretching) is allowed for GRBs 080916C and 090510.
-	For the other two bursts the stretching factor of 1 is, however, prohibited.
-	For GRB 090902B it is smaller than 1, so the low energy light curve is stretched with respect to the high energy one (see fig. \ref{fig:grb090902B}).
-	And for GRB 090926A even more dramatic deviation from the stretching factor of 1 is observed in the opposite direction.
-	The high energy light curve of GRB 090926A is stretched with respect to the low energy one by at least a factor of 1.99 (see fig. \ref{fig:grb090926A}).
+	For the other two bursts the stretching factor of 1 is, however, excluded.
+	For GRB 090902B it should be smaller than 1, so the low energy light curve is stretched with respect to the high energy one (see fig. \ref{fig:grb090902B}).
+	And for GRB 090926A even more significant deviation from the stretching factor of 1 is observed in the opposite direction.
+	The high energy light curve of GRB 090926A is stretched with respect to the low energy one by a factor of at least 1.99 (see fig. \ref{fig:grb090926A}).
 	All observed stretching factors are summarized in table \ref{tab:observationResults}.
@@
 	However, these studies were only concerned with x-ray radiation, and, even more importantly, they assumed that the distribution of radiation sources is homogeneous throughout the jet.
-	We propose a contrary idea: that the highest energy radiators are concentrated near the axis of the burst, meaning that the burst opening angle depends on energy.
-	We cannot prove this assumption rigorously, but have arguments supporting it.
+	We propose a contrary idea: that the highest energy radiators are concentrated near the axis of the jet, so that the jet opening angle depends on energy.
+	We cannot prove this assumption rigorously, but we have arguments supporting it.
 
@@
 	\begin{enumerate}
-		\item{Particles near the jet boundary change their movement directions, therefore increasing the opening angle of the jet.}
+		\item{Particles near the jet boundary collide and change their movement directions, therefore increasing the opening angle of the jet.}
 		\item{Particles lose energy, therefore decreasing radiation frequency emitted by the jet.}

### 2014-08-13 Paper/gammaRays.tex
Correction to Burst Data Download

@@
 For background estimation, we also download observational data during a day before the burst.
-All the values are put the same into the form, except the time range:
+All the values are put the same into the form, except for the time range:
 \begin{align*}
@@
 \end{lstlisting}
-where \texttt{[spacecraft]} is the FITS containing the spacecraft data.
+where \texttt{[spacecraft]} is the FITS file containing the spacecraft data.
 
@@
 This file, however, contains observations of a large area of the sky.
-Before using the data obtained from it, we should perform a more elaborate filtering by location using Fermi LAT's point spread function.
+Before using the data obtained from it, we should perform a more elaborate filtering by location using the point spread function of Fermi LAT.
 The procedure is described in the following section.

### 2014-08-13 Paper/gammaRays.tex
Correction to PSFs and Location Filtering

@@
 However, some of the photons in this set are still originated from background, and not from the source.
-We cannot filter out these photons, but we can substitute a linear component of background by estimating it from a long time period before the burst.
-In order to perform this computation, we'll need an exposure map, which will be discussed in the following section.
+We cannot filter out these photons, but we can substitute a linear component of background estimating it from a long time period before the burst.
+In order to perform this computation, we need an exposure map, which will be discussed in the following section.
 

### 2014-08-13 Paper/gammaRays.tex
Correction to Exposure Maps
Corrections to KS-test

@@
 We compute two background estimates: for low and high energy bands.
-Because the background of transient class photons may depend on the spacecraft position, we have to filter them completely.
+Also, we have to filter out transient class photons completely, because their background appers to depend on the spacecraft position.
 
@@
 \end{equation}
-Here $\left(T_1, T_2\right)$ is the time range of observations, $p_i\left(t_1, t_2\right)$ is the number of photons observed in the time range $t_1$ to $t_2$ in the $i$'s energy band ($i=\text{low},\text{high}$ are for the low and high energy bands respectfully), and $b_i$ is the estimated number of background photons for the time range $T_1$ to $T_2$ in the $i$'s energy band.
+Here $\left(T_1, T_2\right)$ is the time range of observations, $p_i\left(t_1, t_2\right)$ is the number of photons observed in the time range $t_1$ to $t_2$ in the $i$'s energy band ($i$ is either low or high for the low and high energy bands respectfully), and $b_i$ is the estimated number of background photons for the time range $T_1$ to $T_2$ in the $i$'s energy band.
 
@@
 
-As a result of such calculation we can constrain the range of allowed stretching factors to values for which the KS-probabilities are not too small, in other words, given the sigma-significance value, compute the range outside of which stretching factors are elliminated by observations.
+As a result of such calculation we can constrain the range of allowed stretching factors to values for which the KS-probabilities are not too small, in other words, given the sigma-significance value, compute the range outside of which stretching factors are excluded by observations.
 

### 2014-08-13 Paper/gammaRays.tex
Corrections to Photon observation time

@@
 
-Out of 19 bursts studied, only 4 have at least 10 high energy events remaining after filtering, and thus eligible for the computation of stretching factors.
+Out of 19 bursts studied, only 4 have at least 10 high energy events remaining after filtering, and thus eligible for computating the stretching factors.
 The results of this computation are shown on figures \ref{fig:grb080916C}, \ref{fig:grb090510}, \ref{fig:grb090902B}, \ref{fig:grb090926A} and table \ref{tab:observationResults}.
@@
 
-GRB 090926A has, however, high energy radiation stretched with respect to low energy signal (that is $\kappa > 1$).
-In contrast to that, GRB 090902B has low energy radiation stretched with respect to high energy signal ($\kappa < 1$).
+GRB 090926A has, however, high energy radiation stretched with respect to low energy radiation (that is $\kappa > 1$).
+In contrast to that, GRB 090902B has low energy radiation stretched with respect to high energy radiation ($\kappa < 1$).
 
@@
 Note that $\tau$ doesn't depend on $d$ anymore.
-But this is only true if the scale factor did not change from emission to observation.
-For a distant observer we should account for the change in the scale factor, which will stretch the distances between photons:
+But this is only true if we did not account for cosmology, or if the scale factor did not change from emission to observation.
+For a distant observer it will change, however, which will stretch the distances between photons:
 \begin{equation}

### 2014-08-13 Paper/gammaRays.tex
Removed r = r proof from Cosmology
Also other corrections to Cosmology

@@
 
-We also need to know the areas of two spheres: the photon sphere, a sphere over which photons emitted in a particular burst are spread; and the bursts sphere, a sphere over which bursts at a particular redshift are distributed.
+We also need to know the areas of two spheres: the photon sphere, a sphere over which photons emitted from a particular burst are spread; and the bursts sphere, a sphere over which bursts at a particular redshift are distributed.
 
@@
 Here $_2F_1\left(a, b; c; z\right) = \frac{\Gamma\left(c\right)}{\Gamma\left(b\right)\Gamma\left(c-b\right)}\int_0^1 \frac{t^{b-1}\left(1-t\right)^{c-b-1}}{\left(1-t z\right)^a} \dif t$ is a hypergeometric function.
-Here $a\left(0\right) = \frac{1}{1+z}$, so, finally,
+We can substitute $a\left(0\right) = \frac{1}{1+z}$, so, finally,
 \begin{equation}
@@
 
-The second sphere is the bursts sphere.
-It has its origin at the observer's position.
-We define a new primed set of coordinates, so that $r' = 0$ and $t' = 0$ at the observation event.
-We can relate the primed coordinates of the burst to the observer's unprimed coordinates:
-\begin{align*}
-t'_\text{burst} &= -t_\text{obs} \\
-r'\left(t'_\text{burst}\right) &= r'\left(-t_\text{obs}\right) = \int_{-t_\text{obs}}^0\frac{\dif t'}{a'\left(t'\right)} = \int_0^{t_\text{obs}}\frac{\dif t}{a\left(t\right)} = r\left(t_\text{obs}\right)
-\end{align*}
-We see that the radius of the bursts sphere is the same as the radius of the photon sphere. Finally, the area is:
-\begin{equation}
-A_\text{b}\left(z\right) = 4 \pi a'^2\left( -t_\text{obs} \right) r^2\left(z\right) = 4 \pi a^2\left(0\right) r^2\left(z\right) = \frac{4 \pi r^2\left(z\right)}{\left(1+z\right)^2}
-\end{equation}
-
-We will need to know one more thing about the bursts sphere -- the volume of the infinitesimal shell surrounding it. For that we can again use the metric:
+The second sphere is the bursts sphere. It has its origin at the observer's position, and the radius the same as the photon sphere. Therefore, its area is:
+\begin{equation}
+A_\text{b}\left(z\right) = 4 \pi a^2\left(0\right) r^2\left(z\right) = \frac{4 \pi r^2\left(z\right)}{\left(1+z\right)^2}
+\end{equation}
+
+We need to know one more thing about the bursts sphere -- the volume of the infinitesimal shell surrounding it. For that we can again use the metric:
 \begin{align}

### 2014-08-13 Paper/gammaRays.tex
Corrections to Light curve

@@
 Here $A_\text{det}$ is the effective area of the detector.
-We have taken account for the four relativistic effects, which affect intensities and frequencies of the radiators: relativistic aberration; time dilation of the radiators relative to the observer; relativistic blue/redshift; and cosmological redshift.
+We have taken account for four relativistic effects, which affect intensities and frequencies of the radiators: relativistic aberration; time dilation of the radiators relative to the observer; relativistic blue/redshift; and cosmological redshift.
 
@@
 \end{align*}
-After the transformations the expression for $p$ takes the following form:
+After the transformations expression for $p$ takes the following form:
 \begin{align}
@@
 \end{align}
-where $E_n$ function $E_n\left(x\right) = \int_1^\infty \frac{e^{-x t} \dif t}{t^n}$.
+where exponential integral function $E_n\left(x\right) = \int_1^\infty \frac{e^{-x t} \dif t}{t^n}$.
 
@@
 \end{align*}
-Using that, and the observation that this integral is an even function of $\theta$, we arrive to the optimized expressions for $p$, $I$ and $J$:
+Using this assumption, and the evenness of the integral as a function of $\theta$, we arrive to the optimized expressions for $p$, $I$ and $J$:
 \begin{align}
@@
 
-Now, when we know $p$, we can compute many different things with it. Examples include:
+Now, when we know $p$, we can compute many different things with it. For example:
 \begin{itemize}

### 2014-08-13 Paper/gammaRays.tex
Correction to Stretching Factor

@@
 \end{equation}
-And this value of $\kappa$ also makes the $\max\left({D_+, -D_-}\right)$ minimal, since $D_+$ monotonously decrease, and $-D_-$ monotonously increase with $\kappa$.
+And this value of $\kappa$ also makes the $\max\left({D_+, -D_-}\right)$ minimal, since $D_+$ monotonously decrease, and $\left(-D_-\right)$ monotonously increase with $\kappa$.
 

### 2014-08-13 Paper/gammaRays.tex
Corrections to Total Energy
Corrections to Distribution of stretching factors

@@
 \end{align*}
-The integrals over $r$ and $\theta$ can be done symbolically:
+The integrals over $r$ and $\theta$ can be computed symbolically:
 \begin{equation*}
@@
 \end{equation*}
-Now we only have one integral left:
+Now we have only one integral left:
 \begin{equation}
@@
 
-The second test is to calculate the distribution of the stretching factors of the observable bursts, and to compare it with observations.
-
-The computation of the exact and precise distribution is technically hard (because stretching factors neither change monotonously with redshift, nor with off-axis angle), and computationally intensive.
+The second test is to calculate the distribution of stretching factors of observable bursts, and to compare it to observations.
+
+The computation of the exact and precise distribution is technically hard (because stretching factors change monotonously neither with redshifts, nor with off-axis angles), and computationally intensive.
 So we will not compute the precise distribution.
@@
 We still have one ingredient missing though -- the evolution of the bursts density.
-We assume that the density is roughly proportional to the stars density, and the stars density is roughly proportional to the matter density, which changes with $z$ as $\left(1+z\right)^3$.
+We assume that the bursts density is roughly proportional to the stars density, and the stars density is roughly proportional to the matter density, which changes with $z$ as $\left(1+z\right)^3$.
 It is clear, however, that since no stars existed at very small redshifts, the burst density should decline there, so we add an exponential cutoff to it:
@@
 	\end{align}
-	Here $y$ is an another random variable uniformly distributed in the range $0$ to $1$.
+	Here $y$ is an another independent random variable uniformly distributed in the range $0$ to $1$.
 }

### 2014-08-13 Paper/gammaRays.tex
Correction to High energy bursts fraction

@@
 
-Out final test is the comparison of answers to this question: given the bursts which were observed in a whole energy range $\left(\omega_1,\omega_3\right)$, which fraction of them can also be observed in a high energy range $\left(\omega_2,\omega_3\right)$?
+Our final test is comparison of answers to this question: given the bursts which were observed in a whole energy range $\left(\omega_1,\omega_3\right)$, which fraction of them can also be observed in a high energy range $\left(\omega_2,\omega_3\right)$?
 

### 2014-08-13 Paper/gammaRays.tex
Corrections to parameter fit

@@
 We can easily fit $\eta_0$ to match the observed photon count.}
-	\item{If we make a simultaneous transformation of $r_0 \rightarrow \lambda r_0$ and $\eta_0 \rightarrow \frac{1}{\lambda^3}\eta_0$, where $\lambda$ is a parameter of transformation, then only the duration of the burst will change.
+	\item{If we make a simultaneous transformation of $r_0 \rightarrow \lambda r_0$ and $\eta_0 \rightarrow \frac{1}{\lambda^3}\eta_0$, where $\lambda$ is a parameter of transformation, only duration of the burst will change.
 Other observables, like the total number of photons, or stretching factor, will not be affected.}
@@
 
-Also, some of the parameters we know from observations:
+Also, some of these parameters we know from observations:
 \begin{itemize}
@@
 		\label{fig:kappaDistributionHistogram}
-		\caption{Histogram (blue) obtained from the model. Bar lines (black) show observed stretching factors.}
+		\caption{Histogram (blue) obtained from the model. Black error bars show observed stretching factors.}
 	\end{subfigure}
@@
 		\label{fig:kappaDistributionCDF}
-		\caption{CDF (blue) obtained from the model. Gray curve and bar lines show observed stretching factors.}
+		\caption{CDF (blue) obtained from the model. Gray curve and black error bars show observed stretching factors.}
 	\end{subfigure}

### 2014-08-13 Paper/gammaRays.tex
Corrections to Results

@@
 
-	To test the model against observations, we computed a few more things:
+	Here are results of our tests:
 	\begin{itemize}
@@
 		\item{
-			Distribution of stretching factors of observable bursts.
-			You can see the PDF and the CDF of it on fig. \ref{fig:kappaDistribution}.
-			This distribution doesn't contradict with the values of stretching factors obtained in section \ref{sec:observations}.
+			The distribution of observable bursts' stretching factors is shown on fig. \ref{fig:kappaDistribution}.
+			This distribution does not contradict to the values of stretching factors obtained in section \ref{sec:observations}.
 		}
@@
 		}
-		\item{
-			There is a correlation between the fraction of high energy photons (that is high energy photon count divided by the total photon count), the stretching factor and the observer's off-axis angle (see fig.
+	\end{itemize}
+
+	Also, there is a correlation between the fraction of high energy photons (that is high energy photon count divided by the total photon count), the stretching factor and the observer's off-axis angle (see fig.
 \ref{fig:correlations}).
 	It allows one to predict observer's off-axis angles.
-		}
-	\end{itemize}
-
-	So our model, though with some discrepancies, is able to explain the time stretching result from section \ref{sec:observations}.
+
+	So our model is able to explain the time stretching result from section \ref{sec:observations}.
 

### 2014-08-21 Paper/gammaRays.tex
Renamed Discussion to Summary
Also corrections to summary
Fixed zmax=180 -> zmax=100 in paper
Corrections to code upcoming

@@
 	outfile=ltcube.fits
-	dcostheta=0.025 binsz=1 phibins=0 tmin=0 tmax=0 zmax=180 zmin=0
+	dcostheta=0.025 binsz=1 phibins=0 tmin=0 tmax=0 zmax=100 zmin=0
 \end{lstlisting}
@@
 
-\section{Discussion}
-
-\label{sec:discussion}
+\section{Summary}
+
+\label{sec:summary}
 
@@
 		\item{
-			The time stretching of GRB light curves between different VHE bands is confirmed with statistical significance of $3.3\sigma$.
+			Time stretching of GRB light curves between different VHE bands is discovered with statistical significance of $3.3\sigma$.
 		}

### 2014-08-21 GRObservations/GRObservations/GRFermiLATDataServerQuery.cpp
Change zmax 180 -> 100 in code
Results corrections upcoming

@@ -417,7 +417,7 @@ int GRFermiLATDataServerQuery::gtltcube() {
     cmd << "phibins=" << 0 << " ";
     cmd << "tmin=" << 0 << " ";
     cmd << "tmax=" << 0 << " ";
-    cmd << "zmax=" << 180 << " ";
+    cmd << "zmax=" << 100 << " ";
     cmd << "zmin=" << 0 << " ";
     cmd << "chatter=" << 0 << " ";
     return system(cmd.str().c_str());

### 2014-08-22 Paper/gammaRays.tex
Merged sections 2.1 and 2.2
Also partially written Appendix A (about Fermi tools usage)

@@
 \usepackage{bm}
+\usepackage{appendix}
 
@@
 
-\subsection{Burst Data Download}
+\subsection{Fermi LAT Photon Selection}
+\label{sec:photonSelection}
 
@@
 This file, however, contains observations of a large area of the sky.
-Before using the data obtained from it, we should perform a more elaborate filtering by location using the point spread function of Fermi LAT.
-The procedure is described in the following section.
-
-\subsection{Point Spread Functions and Location Filtering}
-
-\texttt{gtpsf}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtpsf.txt}} tool requires the livetime map to generate the PSFs.
-This map is computed by \texttt{gtltcube}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtltcube.txt}} with the following parameters:
+Before using the data obtained from it, we should perform a more elaborate filtering by location using the point spread function of Fermi LAT, which can be computed with \texttt{gtpsf}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtpsf.txt}} tool.
+
+Before computing it, however, we should calculate the livetime map with \texttt{gtltcube}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtltcube.txt}} with the following parameters:
 \begin{lstlisting}
@@
 
+\begin{appendices}
+
+\section{Details of using Fermi Science tools}
+This appendix contains two things: exact values we are passing to a web form on the Fermi LAT data server, and the code we use to invoke the Fermi tools (including full list of options).
+
+Parameters that are passed to the web form of the Fermi LAT data server\footnote{\url{http://fermi.gsfc.nasa.gov/cgi-bin/ssc/LAT/LATDataQuery.cgi}} are the following:
+\begin{itemize}
+	\item{
+		{\bf Object name or coordinates}.
+		Coordinates are filled from the table \ref{tab:bursts}.
+	}
+	\item{
+		{\bf Coordinate system}.
+		J2000.
+	}
+	\item{
+		{\bf Search radius (degrees)}.
+		$60$.
+		Events are filtered by location separately by using the point spread functions (PSFs) of the LAT (as described in section \ref{sec:photonSelection}).
+	}
+	\item{
+		{\bf Observation dates}.
+		To have a decent safety margin, we extend the duration of the burst by $50\%$ to both past and future relative to the table \ref{tab:bursts} time ranges.
+		So, we fill in the following values:
+		\begin{align*}
+			\texttt{time} + \texttt{startOffset} &- 0.5\left(\texttt{endOffset}-\texttt{startOffset}\right),\\
+			\texttt{time} + \texttt{endOffset} &+ 0.5\left(\texttt{endOffset}-\texttt{startOffset}\right)
+		\end{align*}
+	}
+	\item{
+		{\bf Time system}.
+		MET.
+	}
+	\item{
+		{\bf Energy range (MeV)}.
+		$100,\,300000$.
+		This includes our both energy ranges.
+	}
+	\item{
+		{\bf LAT data type}.
+		Extended.
+	}
+	\item{
+		{\bf Spacecraft data}.
+		Checked.
+		It is required for both events filtering, and calculation of the PSFs and exposure maps.
+	}
+\end{itemize}
+
+The commands we use to invoke Fermi Science tools are the following:
+\begin{lstlisting}
+gtselect infile=[eventFile] outfile=filtered.fits
+	ra=INDEF dec=INDEF rad=180 tmin=INDEF tmax=INDEF emin=100 emax=300000
+	zmax=100 evclass=0 convtype=-1 evtable=EVENTS
+\end{lstlisting}
+
+\begin{lstlisting}
+gtmktime scfile=[spacecraft] sctable=SC_DATA
+	filter="DATA_QUAL>0 && LAT_CONFIG==1" roicut=yes
+	evfile=filtered.fits evtable=EVENTS outfile=timed.fits
+	apply_fiter=yes
+\end{lstlisting}
+
+\end{appendices}
+
 \bibliographystyle{hep}

### 2014-08-22 Paper/gammaRays.tex
Drafted fermiCode appendix

@@
 \section{Details of using Fermi Science tools}
+\label{sec:fermiCode}
 This appendix contains two things: exact values we are passing to a web form on the Fermi LAT data server, and the code we use to invoke the Fermi tools (including full list of options).
@@
 \end{lstlisting}
+Here \texttt{[eventFile]} is the file downloaded from the Fermi LAT data server.
 
@@
 \end{lstlisting}
+where \texttt{[spacecraft]} is the FITS file containing the spacecraft data (also downloaded from the Fermi LAT data server).
+
+\begin{lstlisting}
+gtltcube evfile=timed.fits evtable=EVENTS scfile=[spacecraft] sctable=SC_DATA
+	outfile=ltcube.fits
+	dcostheta=0.025 binsz=1 phibins=0 tmin=0 tmax=0 zmax=100 zmin=0
+\end{lstlisting}
+
+\begin{lstlisting}
+gtpsf expcube=ltcube.fits outfile=psf_[IrfName].fits outtable=PSF irfs=[IrfName]
+	ra=[RA] dec=[DEC] emin=100 emax=1000000 nenergies=41 thetamax=30 ntheta=300
+\end{lstlisting}
+\texttt{[RA]} and \texttt{[DEC]} here are the coordinates of the burst, and \texttt{[IrfName]} is the Instrument Response Function, which depends on photon's event class (transient, source, clean, or ultraclean) and conversion type (back or front).
+The list of IRF names can be obtained with the \texttt{gtirfs}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtirfs.txt}} tool.
+We need to compute PSFs for all of them.
+
+\begin{lstlisting}
+gtexpcube2 infile=ltcube.fits cmap=none outfile=expcube_[IrfName].fits irfs=[IrfName]
+	nxpix=360 nypix=180 binsz=1 coordsys=CEL xref=0 yref=0 axisrot=0
+	proj=CAR ebinalg=log emin=100 emax=1000000 enumbins=40
+	ebinfile=NONE bincalc=EDGE ignorephi=no thmax=180 thmin=0 table=Exposure
+\end{lstlisting}
+Note, that R.A. coordinates in the \texttt{expcube} FITS files are indexed in reverse order and starting from 180. For example, $\texttt{ra[0]}=180$, $\texttt{ra[1]}=179$, and so on.
 

### 2014-08-22 Paper/gammaRays.tex
Removed code from photonSelection section

@@
 \usepackage{appendix}
+\usepackage{textcomp}
+\usepackage{gensymb}
 
@@
 
-Finally, to download the data about observed photons, we fill the form at the LAT data server front page with the following values:
-\begin{itemize}
-	\item{
-		{\bf Object name or coordinates}.
-		Here the coordinates are filled from the table \ref{tab:bursts}.
-	}
-	\item{
-		{\bf Coordinate system}.
-		J2000.
-	}
-	\item{
-		{\bf Search radius (degrees)}.
-		$60$.
-		Events will be filtered by location later by using the point spread functions (PSFs) of the LAT.
-	}
-	\item{
-		{\bf Observation dates}.
-		To have a decent safety margin, we extend the duration of the burst by $50\%$ to both past and future relative to the table \ref{tab:bursts} time ranges.
-		So, we fill in the following values:
-		\begin{align*}
-			\texttt{time} + \texttt{startOffset} &- 0.5\left(\texttt{endOffset}-\texttt{startOffset}\right),\\
-			\texttt{time} + \texttt{endOffset} &+ 0.5\left(\texttt{endOffset}-\texttt{startOffset}\right)
-		\end{align*}
-	}
-	\item{
-		{\bf Time system}.
-		MET.
-	}
-	\item{
-		{\bf Energy range (MeV)}.
-		$100,\,300000$.
-		This includes our both energy ranges.
-	}
-	\item{
-		{\bf LAT data type}.
-		Extended.
-	}
-	\item{
-		{\bf Spacecraft data}.
-		Checked.
-		It is required for both events filtering, and calculation of the PSFs and exposure maps.
-	}
-\end{itemize}
-
+To have a decent safety margin, we download the data about observed photons extended in time by $50\%$ to both past and future relative to the table \ref{tab:bursts}.
 For background estimation, we also download observational data during a day before the burst.
-All the values are put the same into the form, except for the time range:
-\begin{align*}
-	\texttt{time} + \texttt{startOffset} &- 0.5\left(\texttt{endOffset} - \texttt{startOffset}\right) - 86400,\\
-	\texttt{time} + \texttt{startOffset} &- 0.5\left(\texttt{endOffset} - \texttt{startOffset}\right)
-\end{align*}
-
-Before reading the data, we need to apply the basic filtering as described in Fermi guidelines.
+Exact values we put into the web form are listed in appendix \ref{sec:fermiCode}.
+
+Before reading the data, we need to apply the basic filtering as described in Fermi guidelines (the actual code, and all parameters values we use to invoke Fermi Science tools are listed in appendix \ref{sec:fermiCode}).
 Two tools from the Fermi Science Tools package are used for that task (the same filtering is applied to both burst and background data).
 
-The first one is \texttt{gtselect}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtselect.txt}}.
-It filters the Earth limb emission.
-We run the tool with the following parameters\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/explore_latdata_burst.html}}:
-\begin{lstlisting}
-gtselect infile=[eventFile] outfile=filtered.fits
-	ra=INDEF dec=INDEF rad=180 tmin=INDEF tmax=INDEF emin=100 emax=300000
-	zmax=100 evclass=0 convtype=-1 evtable=EVENTS
-\end{lstlisting}
-Here \texttt{[eventFile]} is the file downloaded in the previous step.
-
-The second one is \texttt{gtmktime}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtmktime.txt}}.
+First, we use \texttt{gtselect}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtselect.txt}}.
+We filter the Earth limb emission with it below the zenith angle of $100\degree$.
+
+The second one we use is \texttt{gtmktime}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtmktime.txt}}.
 It uses the spacecraft file as well as events file, and filters out time periods when Fermi LAT was not operational.
 Running \texttt{gtmktime} is also required for exposure maps calculation, which we'll need later on.
-Parameters are the following\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/data_preparation.html}}:
-\begin{lstlisting}
-gtmktime scfile=[spacecraft] sctable=SC_DATA
-	filter="DATA_QUAL>0 && LAT_CONFIG==1" roicut=yes
-	evfile=filtered.fits evtable=EVENTS outfile=timed.fits
-	apply_fiter=yes
-\end{lstlisting}
-where \texttt{[spacecraft]} is the FITS file containing the spacecraft data.
-
-Now, with basic filtering done, we can read events data from \texttt{timed.fits}.
-This file, however, contains observations of a large area of the sky.
+
+Now, with basic filtering done, we can read events data from \texttt{timed.fits}. This file, however, contains observations of a large area of the sky.
 Before using the data obtained from it, we should perform a more elaborate filtering by location using the point spread function of Fermi LAT, which can be computed with \texttt{gtpsf}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtpsf.txt}} tool.
-
-Before computing it, however, we should calculate the livetime map with \texttt{gtltcube}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtltcube.txt}} with the following parameters:
-\begin{lstlisting}
-gtltcube evfile=timed.fits evtable=EVENTS scfile=[spacecraft] sctable=SC_DATA
-	outfile=ltcube.fits
-	dcostheta=0.025 binsz=1 phibins=0 tmin=0 tmax=0 zmax=100 zmin=0
-\end{lstlisting}
-
-Having the maps, we can compute the PSFs:
-\begin{lstlisting}
-gtpsf expcube=ltcube.fits outfile=psf_[IrfName].fits outtable=PSF irfs=[IrfName]
-	ra=[RA] dec=[DEC] emin=100 emax=1000000 nenergies=41 thetamax=30 ntheta=300
-\end{lstlisting}
-Here \texttt{[RA]} and \texttt{[DEC]} are the coordinates of the burst, and \texttt{[IrfName]} is the Instrument Response Function, which depends on photon's event class (transient, source, clean, or ultraclean) and conversion type (back or front).
-The list of IRF names can be obtained with the \texttt{gtirfs}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtirfs.txt}} tool.
-We need to compute PSFs for all of them.
+We use 41 energy bins from \unit[100]{MeV} to \unit[1]{TeV}, and 300 angular bins with maximal angle being $30\degree$.
+We use \texttt{gtltcube}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtltcube.txt}} tool to compute livetime maps, which are \texttt{gtpsf} need.
+Note, that we need PSFs for all Instrument Response Functions of Fermi, that is for all event classes (transient, source, clean and ultraclean) and conversion types (back and front).
 
@@
 
-The exposure map can be computed with the \texttt{gtexpcube2}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtexpcube2.txt}} tool. We use the following parameter set for it:
-\begin{lstlisting}
-gtexpcube2 infile=ltcube.fits cmap=none outfile=expcube_[IrfName].fits irfs=[IrfName]
-	nxpix=360 nypix=180 binsz=1 coordsys=CEL xref=0 yref=0 axisrot=0
-	proj=CAR ebinalg=log emin=100 emax=1000000 enumbins=40
-	ebinfile=NONE bincalc=EDGE ignorephi=no thmax=180 thmin=0 table=Exposure
-\end{lstlisting}
-The generated \texttt{expcube} files contain exposures as functions of energy and location on the sky\footnote{Note, that R.A.
-coordinates in the \texttt{expcube} FITS files are indexed in reverse order and starting from 180.
-For example, $\texttt{ra[0]}=180$, $\texttt{ra[1]}=179$, and so on.}.
-We use trilinear interpolation to compute exposures for all energies and locations.
+The exposure map can be computed with the \texttt{gtexpcube2}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtexpcube2.txt}} tool.
+We compute it with 40 energy bins from \unit[100]{MeV} to \unit[1]{TeV}.
+
+The generated \texttt{expcube} files contain exposures as functions of energy and location on the sky. We use trilinear interpolation to compute exposures for all energies and locations.
 
@@
 
-Parameters that are passed to the web form of the Fermi LAT data server\footnote{\url{http://fermi.gsfc.nasa.gov/cgi-bin/ssc/LAT/LATDataQuery.cgi}} are the following:
+Parameters that are passed to the web form of the Fermi LAT data server to download the bursts data are the following:
 \begin{itemize}
@@
 
+For background, all parameters are the same, except for the time ranges, which are the following:
+\begin{align*}
+	\texttt{time} + \texttt{startOffset} &- 0.5\left(\texttt{endOffset} - \texttt{startOffset}\right) - 86400,\\
+	\texttt{time} + \texttt{startOffset} &- 0.5\left(\texttt{endOffset} - \texttt{startOffset}\right)
+\end{align*}
+
 The commands we use to invoke Fermi Science tools are the following:

### 2014-08-22 Paper/gammaRays.tex
Fixed maximal energy from 1000000 to 300000
1000000 is not allowed in Fermi tools documentation

@@
 Before using the data obtained from it, we should perform a more elaborate filtering by location using the point spread function of Fermi LAT, which can be computed with \texttt{gtpsf}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtpsf.txt}} tool.
-We use 41 energy bins from \unit[100]{MeV} to \unit[1]{TeV}, and 300 angular bins with maximal angle being $30\degree$.
+We use 41 energy bins from \unit[100]{MeV} to \unit[300]{GeV}, and 300 angular bins with maximal angle being $30\degree$.
 We use \texttt{gtltcube}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtltcube.txt}} tool to compute livetime maps, which are \texttt{gtpsf} need.
@@
 The exposure map can be computed with the \texttt{gtexpcube2}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtexpcube2.txt}} tool.
-We compute it with 40 energy bins from \unit[100]{MeV} to \unit[1]{TeV}.
+We compute it with 41 energy bins from \unit[100]{MeV} to \unit[300]{GeV}.
 
@@
 gtpsf expcube=ltcube.fits outfile=psf_[IrfName].fits outtable=PSF irfs=[IrfName]
-	ra=[RA] dec=[DEC] emin=100 emax=1000000 nenergies=41 thetamax=30 ntheta=300
+	ra=[RA] dec=[DEC] emin=100 emax=300000 nenergies=41 thetamax=30 ntheta=300
 \end{lstlisting}
@@
 	nxpix=360 nypix=180 binsz=1 coordsys=CEL xref=0 yref=0 axisrot=0
-	proj=CAR ebinalg=log emin=100 emax=1000000 enumbins=40
+	proj=CAR ebinalg=log emin=100 emax=300000 enumbins=40
 	ebinfile=NONE bincalc=EDGE ignorephi=no thmax=180 thmin=0 table=Exposure

### 2014-08-22 GRObservations/GRObservations/GRFermiLATDataServerQuery.cpp

@@ -436,7 +436,7 @@ int GRFermiLATDataServerQuery::gtpsf(GRFermiEventClass eventClass, GRFermiConver
     cmd << "ra=" << location.ra << " ";
     cmd << "dec=" << location.dec << " ";
     cmd << "emin=" << 100. << " ";
-    cmd << "emax=" << 1000000. << " ";
+    cmd << "emax=" << 300000. << " ";
     cmd << "nenergies=" << 41 << " ";
     cmd << "thetamax=" << 30 << " ";
     cmd << "ntheta=" << 300 << " ";
@@ -464,7 +464,7 @@ int GRFermiLATDataServerQuery::gtexpcube(GRFermiEventClass eventClass, GRFermiCo
     cmd << "proj=" << "CAR" << " ";
     cmd << "ebinalg=" << "LOG" << " ";
     cmd << "emin=" << 100. << " ";
-    cmd << "emax=" << 1000000. << " ";
+    cmd << "emax=" << 300000. << " ";
     cmd << "enumbins=" << 40 << " ";
     cmd << "ebinfile=" << "NONE" << " ";
     cmd << "bincalc=" << "EDGE" << " ";

### 2014-08-22 GRObservations/GRObservations/GRFermiLATDataServerQuery.cpp
Changed section names
Also increased the number of energy bins

@@ -437,7 +437,7 @@ int GRFermiLATDataServerQuery::gtpsf(GRFermiEventClass eventClass, GRFermiConver
     cmd << "dec=" << location.dec << " ";
     cmd << "emin=" << 100. << " ";
     cmd << "emax=" << 300000. << " ";
-    cmd << "nenergies=" << 41 << " ";
+    cmd << "nenergies=" << 300 << " ";
     cmd << "thetamax=" << 30 << " ";
     cmd << "ntheta=" << 300 << " ";
     cmd << "chatter=" << 0 << " ";
@@ -465,7 +465,7 @@ int GRFermiLATDataServerQuery::gtexpcube(GRFermiEventClass eventClass, GRFermiCo
     cmd << "ebinalg=" << "LOG" << " ";
     cmd << "emin=" << 100. << " ";
     cmd << "emax=" << 300000. << " ";
-    cmd << "enumbins=" << 40 << " ";
+    cmd << "enumbins=" << 300 << " ";
     cmd << "ebinfile=" << "NONE" << " ";
     cmd << "bincalc=" << "EDGE" << " ";
     cmd << "ignorephi=" << "no" << " ";

### 2014-08-22 Paper/gammaRays.tex
Changed section names
Also increased the number of energy bins
Added "GRB " to GRB names

@@
 	In this paper we consider all the bursts observed by Fermi-LAT since 2008 August 4 to 2011 August 1, for which at least $10$ events with energies $\unit[1]{GeV}$ or higher were observed.
-	There are $4$ of them: 080916C, 090510, 090902B, and 090926A.
+	There are $4$ of them: GRB 080916C, GRB 090510, GRB 090902B, and GRB 090926A.
 	We study their light curves in two bands, $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.
@@
 	The Kolmogorov-Smirnov test is used to check whether the light curves for these two bands are the same.
-	No significant difference was found for 080916C and 090510.
-	However, we observed with statistical significance of $3.3 \sigma$, that the higher energy light curve of 090926A is stretched with respect to the lower-energy one, and with statistical significance of $2.2 \sigma$, that the lower energy light curve of 090902B is stretched with respect to the higher-energy one.
+	No significant difference was found for GRB 080916C and GRB 090510.
+	However, we observed with statistical significance of $3.3 \sigma$, that the higher energy light curve of GRB 090926A is stretched with respect to the lower-energy one, and with statistical significance of $2.2 \sigma$, that the lower energy light curve of GRB 090902B is stretched with respect to the higher-energy one.
 
@@
 	Most of the bursts in the catalog, however, do not have enough events with over $\unit[1]{GeV}$ energies to do desired computations, so we choose only those bursts, from which at least $10$ photons were detected in the high energy band.
-	It leaves $4$ of them: 080916C \cite{Tajima:2009az}, 090510 \cite{Ackermann:2010us}, 090902B \cite{Abdo:2009pg} and 090926A \cite{Bregeon:2011bu}.
+	It leaves $4$ of them: GRB 080916C \cite{Tajima:2009az}, GRB 090510 \cite{Ackermann:2010us}, GRB 090902B \cite{Abdo:2009pg} and GRB 090926A \cite{Bregeon:2011bu}.
 
@@
 	Finally, we stretch the high energy CDF by different factors, and compare it to the low energy CDF using the KS-test.
-	If we require $2\sigma$-significant probability to exclude a particular stretching factor, then the stretching factor of $1$ (which means no stretching) is allowed for GRBs 080916C and 090510.
+	If we require $2\sigma$-significant probability to exclude a particular stretching factor, then the stretching factor of $1$ (which means no stretching) is allowed for GRB 080916C and GRB 090510.
 	For the other two bursts the stretching factor of 1 is, however, excluded.
@@
 
-	Finally, our assumption is able to explain the non-trivial stretching factor of both GRBs 090902B and 090926A.
+	Finally, our assumption is able to explain the non-trivial stretching factor of both GRB 090902B and GRB 090926A.
 	Detailed computations will be described in the following sections, but you can see the main idea on figure \ref{fig:modelOverview}.
@@
 
-\subsection{Results}
+\subsection{Stretching factors}
 
@@
 
-Out of these 4 bursts, two (080916C and 090510) have stretching factors compatible with $\kappa = 1$ within $2\sigma$ range.
+Out of these 4 bursts, two (GRB 080916C and GRB 090510) have stretching factors compatible with $\kappa = 1$ within $2\sigma$ range.
 
@@
 
+\section{Results}
+
 \subsection{Parameter fit}
@@
 			C_\kappa = \frac{
-				\log{\kappa\left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV},\infty\right)} - \log{\kappa_{090926A}}
-			}{\Delta\log\kappa_{090926A}}
-		\end{equation}
-		Here $\log\kappa_{090926A} = \frac{1}{2}\left(\log\left(6.62\right) + \log\left(1.99\right)\right)$ and $\Delta\log\kappa_{090926A} = \frac{1}{2}\left(\log\left(6.62\right) - \log\left(1.99\right)\right)$.
+				\log{\kappa\left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV},\infty\right)} - \log{\kappa_{\text{GRB090926A}}}
+			}{\Delta\log\kappa_{\text{GRB090926A}}}
+		\end{equation}
+		Here $\log\kappa_{\text{GRB090926A}} = \frac{1}{2}\left(\log\left(6.62\right) + \log\left(1.99\right)\right)$ and $\Delta\log\kappa_{\text{GRB090926A}} = \frac{1}{2}\left(\log\left(6.62\right) - \log\left(1.99\right)\right)$.
 	}
@@
 				\frac{
-					\log{\kappa_\text{min}} - \log{\kappa_{090902B}}
-				}{\Delta\log\kappa_{090902B}}\right)
-		\end{equation}
-		Here $\kappa_\text{min}$ is the minimal stretching factor from the sample computed in the step \ref{item:sample}, $\log\kappa_{090902B} = \frac{1}{2}\left(\log\left(0.89\right) + \log\left(0.36\right)\right)$ and $\Delta\log\kappa_{090902B} = \frac{1}{2}\left(\log\left(0.89\right) - \log\left(0.36\right)\right)$.
+					\log{\kappa_\text{min}} - \log{\kappa_{\text{GRB090902B}}}
+				}{\Delta\log\kappa_{\text{GRB090902B}}}\right)
+		\end{equation}
+		Here $\kappa_\text{min}$ is the minimal stretching factor from the sample computed in the step \ref{item:sample}, $\log\kappa_{\text{GRB090902B}} = \frac{1}{2}\left(\log\left(0.89\right) + \log\left(0.36\right)\right)$ and $\Delta\log\kappa_{\text{GRB090902B}} = \frac{1}{2}\left(\log\left(0.89\right) - \log\left(0.36\right)\right)$.
 	}
@@
 					p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV}\right)}
-				} - \log f_{090926A}
+				} - \log f_{\text{GRB090926A}}
 			}{\log(10)}
 		\end{equation}
-		Here $f_{090926A} = 0.0606777$.
+		Here $f_{\text{GRB090926A}} = 0.0606777$.
 	}
@@
 
-\subsection{Results}
+\subsection{Results of the tests}
 
@@
 
-	So our model is able to explain the time stretching result from section \ref{sec:observations}.
+	So our model is able to explain the time stretching effect from section \ref{sec:observations}.
 

### 2014-08-22 Paper/gammaRays.tex
Changed transient filtering out wording

@@
 We compute two background estimates: for low and high energy bands.
-Also, we have to filter out transient class photons completely, because their background appers to depend on the spacecraft position.
+Also, our method doesn't work for transient class photons, because their background appears to depend on the spacecraft position.
+So we had to filter them out completely.
 

### 2014-08-24 Paper/gammaRays.tex
Moved cosmology to Appendix
Added elaboration on spheres to Cosmology

@@
 Here $z$ is the redshift of the burst from the point of view of the observer.
-
-\subsection{Cosmology}
-Before calculating light curves and stretching factors, we will need some equations from cosmology.
-
-First of all, the metric of the expanding Universe:
-\begin{equation}
-\dif s^2 = -\dif t^2 + a^2\left(t\right) \dif r^2 + a^2\left(t\right) r^2 \dif \Omega_2
-\end{equation}
-We define $a\left(t_\text{obs}\right) = 1$, where $t_\text{obs}$ is the observation time.
-
-We need to understand how the scale factor changes with time.
-For that we assume that the energy content of the Universe consists of matter $\Omega_m$ and vacuum energy $\Omega_\Lambda$ only, so that the Friedmann equation takes the following form:
-\begin{align*}
-\left( \frac{\dot{a}\left(t\right)}{a\left(t\right)} \right)^2 &= \Omega_m H_\text{obs}^2 \frac{1}{a^3\left(t\right)} + H_\text{obs}^2 \Omega_\Lambda \\
-\dot{a}\left(t\right) &= a\left(t\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda} \\
-\dif t &= \frac{\dif a}{a\left(t\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}}
-\end{align*}
-Here $H_\text{obs} = H\left(t_\text{obs}\right) = \frac{\dot{a}\left( t_\text{obs} \right)}{a\left( t_\text{obs} \right)} = \dot{a}\left( t_\text{obs} \right)$ is the Hubble parameter at the observation time.
-
-We also need to know the areas of two spheres: the photon sphere, a sphere over which photons emitted from a particular burst are spread; and the bursts sphere, a sphere over which bursts at a particular redshift are distributed.
-
-Lets begin with the photon sphere.
-This sphere has the origin at $r = 0$ and $t = 0$, at the central engine of a particular burst.
-Taking $\dif s = 0$ and $\dif \Omega_2 = 0$ in the equation for metric, we get:
-\begin{equation*}
-\dif r = \frac{\dif t}{a\left(t\right)}
-\end{equation*}
-And now we integrate over time to find the observer's position:
-\begin{align*}
-r\left(t_\text{obs}\right) &= \int_0^{t_\text{obs}}\frac{\dif t}{a\left(t\right)} \\
-&= \int_{a\left(0\right)}^1\frac{\dif a}{a^2\left(t\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}} \\
-&= \frac{\, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{a^3\left(0\right) \Omega_\Lambda }\right)-a\left(0\right) \, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{\Omega_\Lambda }\right)}{a\left(0\right) H_\text{obs} \sqrt{\Omega_\Lambda }}
-\end{align*}
-Here $_2F_1\left(a, b; c; z\right) = \frac{\Gamma\left(c\right)}{\Gamma\left(b\right)\Gamma\left(c-b\right)}\int_0^1 \frac{t^{b-1}\left(1-t\right)^{c-b-1}}{\left(1-t z\right)^a} \dif t$ is a hypergeometric function.
-We can substitute $a\left(0\right) = \frac{1}{1+z}$, so, finally,
-\begin{equation}
-r\left(z\right) = \frac{\left(1+z\right)\, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{\Omega_\Lambda}\left(1+z\right)^3\right) - \, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{\Omega_\Lambda }\right)}{H_\text{obs} \sqrt{\Omega_\Lambda }}
-\end{equation}
-The area of the photon sphere is then:
-\begin{equation}
-A_\text{ph}\left(z\right) = 4 \pi a^2\left(t_\text{obs}\right) r^2\left(z\right) = 4 \pi r^2\left(z\right)
-\end{equation}
-
-The second sphere is the bursts sphere. It has its origin at the observer's position, and the radius the same as the photon sphere. Therefore, its area is:
-\begin{equation}
-A_\text{b}\left(z\right) = 4 \pi a^2\left(0\right) r^2\left(z\right) = \frac{4 \pi r^2\left(z\right)}{\left(1+z\right)^2}
-\end{equation}
-
-We need to know one more thing about the bursts sphere -- the volume of the infinitesimal shell surrounding it. For that we can again use the metric:
-\begin{align}
-\dif V\left(z\right) &= - A_\text{b}\left(z\right) a\left(0\right) \dif r \nonumber\\
-&= - A_\text{b}\left(z\right) a\left(0\right) \frac{\dif a}{a^2\left(0\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}} \nonumber\\
-&= A_\text{b}\left(z\right) \frac{\frac{1}{\left(1+z\right)^2}\dif z \left(1+z\right)}{H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}} \nonumber\\
-&= A_\text{b}\left(z\right) \frac{1}{\left(1+z\right)} \frac{\dif z}{H_\text{obs} \sqrt{\Omega_m\left(1+z\right)^3 + \Omega_\Lambda}} \nonumber\\
-&= \frac{4\pi r^2\left(z\right)}{\left(1+z\right)^3} \frac{\dif z}{H_\text{obs} \sqrt{\Omega_m\left(1+z\right)^3 + \Omega_\Lambda}}
-\end{align}
 
@@
 
+\section{Cosmology}
+Before calculating light curves and stretching factors, we will need some equations from cosmology.
+
+First of all, the metric of the expanding Universe:
+\begin{equation}
+\dif s^2 = -\dif t^2 + a^2\left(t\right) \dif r^2 + a^2\left(t\right) r^2 \dif \Omega_2
+\end{equation}
+We define $a\left(t_\text{obs}\right) = 1$, where $t_\text{obs}$ is the observation time.
+
+We need to understand how the scale factor changes with time.
+For that we assume that the energy content of the Universe consists of matter $\Omega_m$ and vacuum energy $\Omega_\Lambda$ only, so that the Friedmann equation takes the following form:
+\begin{align*}
+\left( \frac{\dot{a}\left(t\right)}{a\left(t\right)} \right)^2 &= \Omega_m H_\text{obs}^2 \frac{1}{a^3\left(t\right)} + H_\text{obs}^2 \Omega_\Lambda \\
+\dot{a}\left(t\right) &= a\left(t\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda} \\
+\dif t &= \frac{\dif a}{a\left(t\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}}
+\end{align*}
+Here $H_\text{obs} = H\left(t_\text{obs}\right) = \frac{\dot{a}\left( t_\text{obs} \right)}{a\left( t_\text{obs} \right)} = \dot{a}\left( t_\text{obs} \right)$ is the Hubble parameter at the observation time.
+
+We also need to know the areas of two spheres: the photon sphere, a sphere over which photons emitted in a particular burst are spread; and the bursts sphere, a sphere over which bursts at a particular redshift are distributed.
+These two spheres have the same radii -- a distance from the observer to the burst, but different centers: the photon sphere is centered on the burst, while the bursts sphere is centered on the observer.
+Also, they have different areas, because the scale factor differs at locations of burst and observer.
+
+Lets begin with the photon sphere.
+This sphere has the origin at $r = 0$ and $t = 0$, at the central engine of a particular burst.
+Taking $\dif s = 0$ and $\dif \Omega_2 = 0$ in the equation for metric, we get:
+\begin{equation*}
+\dif r = \frac{\dif t}{a\left(t\right)}
+\end{equation*}
+And now we integrate over time to find the observer's position:
+\begin{align*}
+r\left(t_\text{obs}\right) &= \int_0^{t_\text{obs}}\frac{\dif t}{a\left(t\right)} \\
+&= \int_{a\left(0\right)}^1\frac{\dif a}{a^2\left(t\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}} \\
+&= \frac{\, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{a^3\left(0\right) \Omega_\Lambda }\right)-a\left(0\right) \, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{\Omega_\Lambda }\right)}{a\left(0\right) H_\text{obs} \sqrt{\Omega_\Lambda }}
+\end{align*}
+Here $_2F_1\left(a, b; c; z\right) = \frac{\Gamma\left(c\right)}{\Gamma\left(b\right)\Gamma\left(c-b\right)}\int_0^1 \frac{t^{b-1}\left(1-t\right)^{c-b-1}}{\left(1-t z\right)^a} \dif t$ is a hypergeometric function.
+We can substitute $a\left(0\right) = \frac{1}{1+z}$, so, finally,
+\begin{equation}
+r\left(z\right) = \frac{\left(1+z\right)\, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{\Omega_\Lambda}\left(1+z\right)^3\right) - \, _2F_1\left(\frac{1}{3},\frac{1}{2};\frac{4}{3};-\frac{\Omega_m}{\Omega_\Lambda }\right)}{H_\text{obs} \sqrt{\Omega_\Lambda }}
+\end{equation}
+The area of the photon sphere is then:
+\begin{equation}
+A_\text{ph}\left(z\right) = 4 \pi a^2\left(t_\text{obs}\right) r^2\left(z\right) = 4 \pi r^2\left(z\right)
+\end{equation}
+
+The second sphere is the bursts sphere. It has its origin at the observer's position, and the radius the same as the photon sphere. Therefore, its area is:
+\begin{equation}
+A_\text{b}\left(z\right) = 4 \pi a^2\left(0\right) r^2\left(z\right) = \frac{4 \pi r^2\left(z\right)}{\left(1+z\right)^2}
+\end{equation}
+
+We need to know one more thing about the bursts sphere -- the volume of the infinitesimal shell surrounding it. For that we can again use the metric:
+\begin{align}
+\dif V\left(z\right) &= - A_\text{b}\left(z\right) a\left(0\right) \dif r \nonumber\\
+&= - A_\text{b}\left(z\right) a\left(0\right) \frac{\dif a}{a^2\left(0\right) H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}} \nonumber\\
+&= A_\text{b}\left(z\right) \frac{\frac{1}{\left(1+z\right)^2}\dif z \left(1+z\right)}{H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}} \nonumber\\
+&= A_\text{b}\left(z\right) \frac{1}{\left(1+z\right)} \frac{\dif z}{H_\text{obs} \sqrt{\Omega_m\left(1+z\right)^3 + \Omega_\Lambda}} \nonumber\\
+&= \frac{4\pi r^2\left(z\right)}{\left(1+z\right)^3} \frac{\dif z}{H_\text{obs} \sqrt{\Omega_m\left(1+z\right)^3 + \Omega_\Lambda}}
+\end{align}
+
 \end{appendices}

### 2014-08-24 Paper/gammaRays.tex
Added Intro to Cosmology appendix

@@
 \section{Cosmology}
-Before calculating light curves and stretching factors, we will need some equations from cosmology.
+This appendix contains a few formulas we need from cosmology, in particular the area of a photon sphere $A_{ph}\left(z\right)$, a sphere over which photons emitted in a particular burst at redshift $z$ are spread; and the volume of the infinitesimal shell surrounding a burst sphere $\dif V\left(z\right)$, a sphere over which bursts at redshift $z$ are distributed.
 

### 2014-08-24 Paper/gammaRays.tex
Add notes about cosmology where it first appear

--- 2014-08-24 Paper/gammaRays.tex
@@
 \end{align}
-Here $A_\text{det}$ is the effective area of the detector.
+Here $A_\text{det}$ is the effective area of the detector, and $A_\text{ph}$ is an area of the sphere over which photons emitted in the burst are spread (see appendix \ref{sec:cosmology} for derivation).
 We have taken account for four relativistic effects, which affect intensities and frequencies of the radiators: relativistic aberration; time dilation of the radiators relative to the observer; relativistic blue/redshift; and cosmological redshift.
@@
 	\end{equation}
+	where $\dif V\left(z'\right)$ is the volume of the infinitesimal shell surrounding a sphere over which bursts at redshift $z'$ are distributed (see appendix \ref{sec:cosmology} for derivation).
 
@@
 \section{Cosmology}
-Before calculating light curves and stretching factors, we will need some equations from cosmology.
+\label{sec:cosmology}
+This appendix contains a few formulas we need from cosmology, in particular the area of a photon sphere $A_\text{ph}\left(z\right)$, a sphere over which photons emitted in a particular burst at redshift $z$ are spread; and the volume of the infinitesimal shell surrounding a burst sphere $\dif V\left(z\right)$, a sphere over which bursts at redshift $z$ are distributed.
 

### 2014-08-24 Paper/gammaRays.tex
Fix 3 to 4 observed GRBs in Intro justification
Also partly written Model Description

@@
 	First of all, there are around 750 GRBs detected by GBM, half of which were in the LAT field of view at the moment of observation \cite{Vianello:2013ela}.
-	However, only about 30 of them were detected by the LAT, and only 3 of them were bright in the high energy band.
+	However, only about 30 of them were detected by the LAT, and only 4 of them were bright in the high energy band.
 	If we extrapolate the uniform jet model to very high energies, this observation would mean that these groups of bursts are internally different: some of them produce VHE radiation, while others do not.
@@
 \section{Model}
+
+\subsection{Description}
+
+The main idea behind our model is to assume that the burst opening angle depends on energy of emitted photons, or, equivalently, that the most energetic plasma particles are concentrated near the axis of the jet, while low energy particle are on the periphery.
+As you can see on figure \ref{fig:modelOverview} this leads to a time stretching effect in some cases due to relativistic beaming.
+
+At the moment, we do not have any rigorous proof of our assumption.
+There are arguments, however, supporting it.
+
+First of all, the number of observed GRBs rapidly decreases with energy.
+There are around 750 bursts detected by GBM in hard X-rays and low energy gamma rays, and only 4
 

### 2014-08-24 Paper/gammaRays.tex
Drafted Model Description section

@@
 
-The main idea behind our model is to assume that the burst opening angle depends on energy of emitted photons, or, equivalently, that the most energetic plasma particles are concentrated near the axis of the jet, while low energy particle are on the periphery.
+The main idea behind our model is to assume that the burst opening angle depends on energy of emitted photons, or, equivalently, that the most energetic plasma particles are concentrated near the axis of the jet, while low energy particles are on the periphery.
 As you can see on figure \ref{fig:modelOverview} this leads to a time stretching effect in some cases due to relativistic beaming.
@@
 First of all, the number of observed GRBs rapidly decreases with energy.
-There are around 750 bursts detected by GBM in hard X-rays and low energy gamma rays, and only 4
+There are around 750 bursts detected by GBM in hard X-rays and low energy gamma rays \cite{Vianello:2013ela} (half of them were in a LAT field of view), only 30 or so were detected by LAT above \unit[100]{MeV}, and only about 4 of them were detected above GeV.
+This observation is hard to explain in a framework of uniform jet model given that burst energetics are similar \cite{Bloom:2003wy}.
+It is, however, easy to explain in our model, given that small observation angles are rare, and the only way to produce high energy radiation is to have the observation angle small enough.
+
+Another argument comes from consideration of processes happening in jets, specifically the scattering of particles near the jet boundary.
+While plasma particles scatter, they simultaneously lose energy, and change directions, sometimes propagating beyond the jet boundary, and therefore increasing the jet opening angle.
+So, the processes of energy loss and increase of jet size are correlated, therefore the low energy particles should be closer to the jet boundary.
+
+In the following sections we will rigorously derive the procedure for computing time stretching factors from our model, and compare them to observations obtained in the previous section.
 

### 2014-08-25 Paper/gammaRays.tex
Added reference to Observations from Intro
Added links to model, results and summary to Intro

@@
 	The high energy light curve of GRB 090926A is stretched with respect to the low energy one by a factor of at least 1.99 (see fig. \ref{fig:grb090926A}).
-	All observed stretching factors are summarized in table \ref{tab:observationResults}.
+	All observed stretching factors are summarized in table \ref{tab:observationResults}, and the detailed procedure for computing them is described in section \ref{sec:observations} of the paper.
 
@@
 	Finally, our assumption is able to explain the non-trivial stretching factor of both GRB 090902B and GRB 090926A.
-	Detailed computations will be described in the following sections, but you can see the main idea on figure \ref{fig:modelOverview}.
-	The computations confirm that this qualitative picture is indeed correct.
+	Detailed computations are described in section \ref{sec:model}, but you can see the main idea on figure \ref{fig:modelOverview}.
+	Computations confirm that this qualitative picture is indeed correct.
 	You can see a sample light curves on figure \ref{fig:sampleLightCurves}, and predicted distribution of stretching factors on figure \ref{fig:kappaDistribution}.
+
+	Results of our study, including a few predictions of our model and a new method for estimating jet observation angles, are summarized in section \ref{sec:results}, and the caveats are discussed in section \ref{sec:summary}.
 
@@
 \section{Model}
+\label{sec:model}
 
@@
 \section{Results}
+\label{sec:results}
 

### 2014-08-26 Paper/gammaRays.tex
Added open source notes to title and abstract

@@
 
-\title{Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model}
+\title{
+	Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model \\
+	{\large \url{https://github.com/maxitg/GammaRays}}
+}
+
 \author{
@@
 \begin{document}
+
 \maketitle
@@
 	Finally, we propose a way to estimate observer's off-axis angles (view directions) based on stretching factors and fractions of high energy photons.
+
+	All our code is open source, and is available freely on GitHub for you to download and play with.
 \end{abstract}

### 2014-08-27 Paper/gammaRays.tex
Fixed the number of energy bins in text to 300

@@
 Before using the data obtained from it, we should perform a more elaborate filtering by location using the point spread function of Fermi LAT, which can be computed with \texttt{gtpsf}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtpsf.txt}} tool.
-We use 41 energy bins from \unit[100]{MeV} to \unit[300]{GeV}, and 300 angular bins with maximal angle being $30\degree$.
+We use 300 energy bins from \unit[100]{MeV} to \unit[300]{GeV}, and 300 angular bins with maximal angle being $30\degree$.
 We use \texttt{gtltcube}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtltcube.txt}} tool to compute livetime maps, which are \texttt{gtpsf} need.
@@
 The exposure map can be computed with the \texttt{gtexpcube2}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtexpcube2.txt}} tool.
-We compute it with 41 energy bins from \unit[100]{MeV} to \unit[300]{GeV}.
+We compute it with 300 energy bins from \unit[100]{MeV} to \unit[300]{GeV}.
 

### 2014-08-27 Paper/gammaRays.tex
Changed evclass to 2 in paper and in code

@@
 We filter the Earth limb emission with it below the zenith angle of $100\degree$.
+We also filter out photons with transient event class for a reason described in the next section.
 
@@
 We compute two background estimates: for low and high energy bands.
-Also, our method doesn't work for transient class photons, because their background appears to depend on the spacecraft position.
-So we had to filter them out completely.
+Finally, our method doesn't work for transient class photons, because their background appears to depend on the spacecraft position.
+That is why we had to filter them out completely.
 
@@
 	ra=INDEF dec=INDEF rad=180 tmin=INDEF tmax=INDEF emin=100 emax=300000
-	zmax=100 evclass=0 convtype=-1 evtable=EVENTS
+	zmax=100 evclass=2 convtype=-1 evtable=EVENTS
 \end{lstlisting}

### 2014-08-27 GRObservations/GRObservations/GRBurst.cpp

@@ -144,7 +144,10 @@ void GRBurst::read() {
     for (int i = 0; i < query.events.size(); i++) {
         GRFermiLATPhoton photon = query.events[i];
         photon.location.error = query.psfs[photon.eventClass][photon.conversionType].spread(photon.energy, PHOTONS_QUALITY);
-        if (photon.eventClass == GRFermiEventClassTransient) continue;
+        if (photon.eventClass == GRFermiEventClassTransient) {
+            cout << "Warning. Transients appeared even though they should have been filtered out by gtselect." << endl;
+            continue;
+        }
         if (location.isSeparated(photon.location)) continue;
         
         if (photon.energy < 1000.) mevPhotons.push_back(photon);
@@ -156,7 +159,10 @@ void GRBurst::read() {
     for (int i = 0; i < backgroundQuery.events.size(); i++) {
         GRFermiLATPhoton photon = backgroundQuery.events[i];
         photon.location.error = backgroundQuery.psfs[photon.eventClass][photon.conversionType].spread(photon.energy, PHOTONS_QUALITY);
-        if (photon.eventClass == GRFermiEventClassTransient) continue;
+        if (photon.eventClass == GRFermiEventClassTransient) {
+            cout << "Warning. Transients appeared even though they should have been filtered out by gtselect." << endl;
+            continue;
+        }
         if (location.isSeparated(photon.location)) continue;
         
         if (photon.energy < 1000.) mevBackgroundPhotons.push_back(photon);

### 2014-08-27 GRObservations/GRObservations/GRFermiLATDataServerQuery.cpp

@@ -381,7 +381,7 @@ int GRFermiLATDataServerQuery::gtselect() {
     cmd << "emin=" << 100. << " ";
     cmd << "emax=" << 300000. << " ";
     cmd << "zmax=" << 100. << " ";
-    cmd << "evclass=" << 0 << " ";
+    cmd << "evclass=" << 2 << " ";
     cmd << "convtype=" << -1 << " ";
     cmd << "evtable=" << "EVENTS" << " ";
     cmd << "chatter=" << 0 << " ";

### 2014-08-28 Paper/gammaRays.tex
Replaced many links to Tools with two links

@@
 
-Before reading the data, we need to apply the basic filtering as described in Fermi guidelines (the actual code, and all parameters values we use to invoke Fermi Science tools are listed in appendix \ref{sec:fermiCode}).
-Two tools from the Fermi Science Tools package are used for that task (the same filtering is applied to both burst and background data).
-
-First, we use \texttt{gtselect}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtselect.txt}}.
+Before reading the data, we need to apply the basic filtering as described in Fermi guidelines\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/data_preparation.html}} (the actual code, and all parameters values we use to invoke Fermi Science tools are listed in appendix \ref{sec:fermiCode}).
+Two tools from the Fermi Science Tools package\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/overview.html}} are used for that task (the same filtering is applied to both burst and background data).
+
+First, we use \texttt{gtselect}.
 We filter the Earth limb emission with it below the zenith angle of $100\degree$.
@@
 
-The second one we use is \texttt{gtmktime}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtmktime.txt}}.
+The second one we use is \texttt{gtmktime}.
 It uses the spacecraft file as well as events file, and filters out time periods when Fermi LAT was not operational.
@@
 Now, with basic filtering done, we can read events data from \texttt{timed.fits}. This file, however, contains observations of a large area of the sky.
-Before using the data obtained from it, we should perform a more elaborate filtering by location using the point spread function of Fermi LAT, which can be computed with \texttt{gtpsf}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtpsf.txt}} tool.
+Before using the data obtained from it, we should perform a more elaborate filtering by location using the point spread function of Fermi LAT, which can be computed with \texttt{gtpsf} tool.
 We use 300 energy bins from \unit[100]{MeV} to \unit[300]{GeV}, and 300 angular bins with maximal angle being $30\degree$.
-We use \texttt{gtltcube}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtltcube.txt}} tool to compute livetime maps, which are \texttt{gtpsf} need.
+We use \texttt{gtltcube} tool to compute livetime maps, which are \texttt{gtpsf} need.
 Note, that we need PSFs for all Instrument Response Functions of Fermi, that is for all event classes (transient, source, clean and ultraclean) and conversion types (back and front).
@@
 
-The exposure map can be computed with the \texttt{gtexpcube2}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtexpcube2.txt}} tool.
-We compute it with 300 energy bins from \unit[100]{MeV} to \unit[300]{GeV}.
+The exposure map can be computed with the \texttt{gtexpcube2} tool. We compute it with 300 energy bins from \unit[100]{MeV} to \unit[300]{GeV}.
 
@@
 \texttt{[RA]} and \texttt{[DEC]} here are the coordinates of the burst, and \texttt{[IrfName]} is the Instrument Response Function, which depends on photon's event class (transient, source, clean, or ultraclean) and conversion type (back or front).
-The list of IRF names can be obtained with the \texttt{gtirfs}\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/help/gtirfs.txt}} tool.
+The list of IRF names can be obtained with the \texttt{gtirfs} tool.
 We need to compute PSFs for all of them.

### 2014-08-29 sample16.m
Updated samples (except sample4096pc40)

@@ -1,32 +1,32 @@
-sample16 = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
-      34.228215884194334, 222.73035829370284, 1.0280094289841748}, 
-     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
-      20.978799047458836, 123.38630837565341, 2.529581957694992}, 
-     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
-      10.070010357458859, 178.00778096254146, 1.784271416531604}, 
-     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
-      125.2194399101692, 107.40664116322054, 1.355228067650187}, 
-     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
-      12.708560050890373, 313.05858759493555, 1.1251344561349428}, 
-     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
-      18.983240726070672, 156.94967793679683, 1.6188869008349682}, 
-     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
-      434.73441714374775, 92.48671063445956, 1.085830862969823}, 
-     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
-      578.4999909315469, 94.50973498739872, 0.8908846736633591}, 
-     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
-      131.45761966748594, 125.400307231681, 1.040915010748597}, 
-     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
-      14.21984824761055, 176.44515329447125, 1.5781588059941152}, 
-     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
-      364.8701470277157, 100.1140576033533, 0.9741210674073871}, 
-     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
-      24.758770352965122, 223.3218204554101, 1.1155870882542687}, 
-     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
-      15.814107790469503, 258.1278608438802, 1.1647305773040832}, 
-     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
-      39.646577434687785, 150.48002884878116, 1.2902775201409398}, 
-     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-08-29 sample16IA.m

@@ -1,32 +1,32 @@
-sample16IA = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
-      34.228215884194334, 222.73035829370284, 1.0701575908785792}, 
-     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
-      20.978799047458836, 123.38630837565341, 3.049495182568959}, 
-     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
-      10.070010357458859, 178.00778096254146, 1.9224561249840593}, 
-     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
-      125.2194399101692, 107.40664116322054, 1.3821372215243484}, 
-     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
-      12.708560050890373, 313.05858759493555, 1.180678993898995}, 
-     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
-      18.983240726070672, 156.94967793679683, 1.70714273366365}, 
-     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
-      434.73441714374775, 92.48671063445956, 1.1077715331761284}, 
-     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
-      578.4999909315469, 94.50973498739872, 0.9364896504440356}, 
-     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
-      131.45761966748594, 125.400307231681, 1.0805165020789271}, 
-     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
-      14.21984824761055, 176.44515329447125, 1.6840634952055642}, 
-     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
-      364.8701470277157, 100.1140576033533, 1.0130197810371595}, 
-     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
-      24.758770352965122, 223.3218204554101, 1.1676582199615857}, 
-     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
-      15.814107790469503, 258.1278608438802, 1.225460126255487}, 
-     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
-      39.646577434687785, 150.48002884878116, 1.3439228024212373}, 
-     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-08-29 sample16pc40.m

@@ -1,32 +1,32 @@
-sample16pc40 = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
-      34.228215884194334, 222.73035829370284, 1.0280094289841748}, 
-     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
-      20.978799047458836, 123.38630837565341, 2.529581957694991}, 
-     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
-      10.070010357458859, 178.00778096254146, 1.7842714165316036}, 
-     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
-      125.2194399101692, 107.40664116322054, 1.355228067650187}, 
-     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
-      12.708560050890373, 313.05858759493555, 1.1251344561349426}, 
-     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
-      18.983240726070672, 156.94967793679683, 1.6188869008349682}, 
-     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
-      434.73441714374775, 92.48671063445956, 1.0858308629698226}, 
-     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
-      578.4999909315469, 94.50973498739872, 0.890884673663359}, 
-     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
-      131.45761966748594, 125.400307231681, 1.0409150107485967}, 
-     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
-      14.21984824761055, 176.44515329447125, 1.5781588059941158}, 
-     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
-      364.8701470277157, 100.1140576033533, 0.9741210674073869}, 
-     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
-      24.758770352965122, 223.3218204554101, 1.1155870882542687}, 
-     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
-      15.814107790469503, 258.1278608438802, 1.1647305773040835}, 
-     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
-      39.646577434687785, 150.48002884878116, 1.2902775201409395}, 
-     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-08-29 sample256.m

@@ -1,512 +1,512 @@
-sample256 = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
-      34.228215884194334, 222.73035829370284, 1.0280094289841748}, 
-     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
-      20.978799047458836, 123.38630837565341, 2.529581957694992}, 
-     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
-      10.070010357458859, 178.00778096254146, 1.784271416531604}, 
-     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
-      125.2194399101692, 107.40664116322054, 1.355228067650187}, 
-     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
-      12.708560050890373, 313.05858759493555, 1.1251344561349428}, 
-     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
-      18.983240726070672, 156.94967793679683, 1.6188869008349682}, 
-     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
-      434.73441714374775, 92.48671063445956, 1.085830862969823}, 
-     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
-      578.4999909315469, 94.50973498739872, 0.8908846736633591}, 
-     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
-      131.45761966748594, 125.400307231681, 1.040915010748597}, 
-     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
-      14.21984824761055, 176.44515329447125, 1.5781588059941152}, 
-     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
-      364.8701470277157, 100.1140576033533, 0.9741210674073871}, 
-     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
-      24.758770352965122, 223.3218204554101, 1.1155870882542687}, 
-     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
-      15.814107790469503, 258.1278608438802, 1.1647305773040832}, 
-     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
-      39.646577434687785, 150.48002884878116, 1.2902775201409398}, 
-     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-08-29 sample256IA.m

@@ -1,512 +1,512 @@
-sample256IA = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
-      34.228215884194334, 222.73035829370284, 1.0701575908785792}, 
-     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
-      20.978799047458836, 123.38630837565341, 3.049495182568959}, 
-     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
-      10.070010357458859, 178.00778096254146, 1.9224561249840593}, 
-     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
-      125.2194399101692, 107.40664116322054, 1.3821372215243484}, 
-     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
-      12.708560050890373, 313.05858759493555, 1.180678993898995}, 
-     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
-      18.983240726070672, 156.94967793679683, 1.70714273366365}, 
-     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
-      434.73441714374775, 92.48671063445956, 1.1077715331761284}, 
-     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
-      578.4999909315469, 94.50973498739872, 0.9364896504440356}, 
-     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
-      131.45761966748594, 125.400307231681, 1.0805165020789271}, 
-     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
-      14.21984824761055, 176.44515329447125, 1.6840634952055642}, 
-     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
-      364.8701470277157, 100.1140576033533, 1.0130197810371595}, 
-     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
-      24.758770352965122, 223.3218204554101, 1.1676582199615857}, 
-     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
-      15.814107790469503, 258.1278608438802, 1.225460126255487}, 
-     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
-      39.646577434687785, 150.48002884878116, 1.3439228024212373}, 
-     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-08-29 sample256pc40.m

@@ -1,512 +1,513 @@
-sample256pc40 = {{4.577676114922525, 0.002146172294505203, 198.4196620066235, 
-      34.228215884194334, 222.73035829370284, 1.0280094289841748}, 
-     {0.38242879878225633, 0.007303018195367234, 2510.4548818606413, 
-      20.978799047458836, 123.38630837565341, 2.529581957694991}, 
-     {1.790678523643586, 0.004771211764882174, 291.8119917380803, 
-      10.070010357458859, 178.00778096254146, 1.7842714165316036}, 
-     {0.6679554361909841, 0.004633224927380444, 1634.8277140318862, 
-      125.2194399101692, 107.40664116322054, 1.355228067650187}, 
-     {6.7438565048704895, 0.0025153880238022266, 114.41219610878554, 
-      12.708560050890373, 313.05858759493555, 1.1251344561349426}, 
-     {1.5334501903844784, 0.004552168302441222, 402.761616478708, 
-      18.983240726070672, 156.94967793679683, 1.6188869008349682}, 
-     {0.6730575558627483, 0.0033384774004186746, 2098.3853992699323, 
-      434.73441714374775, 92.48671063445956, 1.0858308629698226}, 
-     {1.0880990456375839, 0.001605553355172647, 1254.8786426044062, 
-      578.4999909315469, 94.50973498739872, 0.890884673663359}, 
-     {1.61092789076918, 0.0026727127210554205, 612.3130849666738, 
-      131.45761966748594, 125.400307231681, 1.0409150107485967}, 
-     {1.9742563479671753, 0.004326147803631458, 290.25136432853395, 
-      14.21984824761055, 176.44515329447125, 1.5781588059941158}, 
-     {1.0632763208205827, 0.0023948982757103594, 1172.1006546668993, 
-      364.8701470277157, 100.1140576033533, 0.9741210674073869}, 
-     {4.167417372575645, 0.0026216289538748623, 189.43015298651406, 
-      24.758770352965122, 223.3218204554101, 1.1155870882542687}, 
-     {4.933123699945455, 0.0027710318657595307, 147.57157347627398, 
-      15.814107790469503, 258.1278608438802, 1.1647305773040835}, 
-     {1.7903411064295107, 0.003648417441486108, 414.66398652831725, 
-      39.646577434687785, 150.48002884878116, 1.2902775201409395}, 
-     {1.992142656294731, 0.003909920015542307, 327.6715685897391, 
[…]

### 2014-08-29 Paper/gammaRays.tex
Added GRB names to plots (as numbers)

@@
 	\hspace*{\fill}
-	\caption{Stretching factors histogram and CDF produced by our model. The sample contains $4096$ bursts.}
+	\caption{
+		Stretching factors histogram and CDF produced by our model.
+		The sample contains $4096$ bursts.
+		Numbered error bars correspond to GRB 080916C, GRB 090510, GRB 090902B and GRB 090926A (in this order).
+	}
 	\label{fig:kappaDistribution}
@@
         \includegraphics[width=1.0\textwidth]{correlations}
-        \caption{Correlations between off-axis angles, stretching factors and high to low energy photon count ratios found in the sample produced by our model.
+        \caption{
+        	Correlations between off-axis angles, stretching factors and high to low energy photon count ratios found in the sample produced by our model.
         	The sample contains $4096$ bursts.
-Black crosses show observations with $2\sigma$ error bars.
-This correlation allows one to estimate off-axis angles of observed bursts.}
+        	Numbered black crosses show data for GRB 080916C, GRB 090510, GRB 090902B and GRB 090926A (in this order) with $2\sigma$ error bars.
+        	This correlation allows one to estimate off-axis angles of observed bursts.
+        }
         \label{fig:correlations}

### 2014-08-31 Paper/gammaRays.tex
Updated paper with numbers from fixed zmax results

@@
 	No significant difference was found for GRB 080916C and GRB 090510.
-	However, we observed with statistical significance of $3.3 \sigma$, that the higher energy light curve of GRB 090926A is stretched with respect to the lower-energy one, and with statistical significance of $2.2 \sigma$, that the lower energy light curve of GRB 090902B is stretched with respect to the higher-energy one.
+	However, we observed with statistical significance of $3.3 \sigma$, that the higher energy light curve of GRB 090926A is stretched with respect to the lower-energy one, and with statistical significance of $2.3 \sigma$, that the lower energy light curve of GRB 090902B is stretched with respect to the higher-energy one.
 
@@
 			\label{fig:sampleLightCurveLogNegative}
-			\caption{Stretching factor $\kappa = 0.821$, redshift $z = 1.82$, off-axis angle $\chi = 0$.}
+			\caption{Stretching factor $\kappa = 0.819$, redshift $z = 1.82$, off-axis angle $\chi = 0$.}
 		\end{subfigure}
@@
 			\label{fig:sampleLightCurveLogPosivie}
-			\caption{Stretching factor $\kappa = 2.26$, redshift $z = 2.106$, off-axis angle $\chi = 5.41 \times 10^{-3}$.}
+			\caption{Stretching factor $\kappa = 2.43$, redshift $z = 2.106$, off-axis angle $\chi = 5.90 \times 10^{-3}$.}
 		\end{subfigure}
@@
         \end{subfigure}
-        \caption{GRB 090902B results. Low energy radiation is stretched ($\kappa < 1$) with significance of $2.2\sigma$.}
+        \caption{GRB 090902B results. Low energy radiation is stretched ($\kappa < 1$) with significance of $2.3\sigma$.}
         \label{fig:grb090902B}
@@
 		090501	&	0.58  & 1.1		&	0.43 & 1.61	&	0.32 & 2.29	&	0.22 &  3.03	&	0.17  &  5.11	\\
-		090902B	&	0.629 & 0.635	&	0.36 & 0.89	&	0.23 & 1.57	&	0.13 &  2.87	&	0.084 &  4.56	\\
+		090902B	&	0.604 & 0.613	&	0.35 & 0.89	&	0.22 & 1.53	&	0.13 &  2.86	&	0.078 &  4.49	\\
 		090926A &	2.61  & 3.33	&	1.99 & 6.62	&	1.34 & 9.15	&	0.73 & 13.5		&	0.48  & 19.4
@@
 	\item {
-		The ratio of photons with high energies and low energies should be compatible with the value for the GRB 090926A, $\frac{p_{\infty} \left(z,\chi;\unit[1]{GeV},\infty\right)}{p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV}\right)} = 0.06$.
+		The ratio of photons with high energies and low energies should be compatible with the value for the GRB 090926A, $\frac{p_{\infty} \left(z,\chi;\unit[1]{GeV},\infty\right)}{p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV}\right)} = 0.057$.
 	}
@@
 	\item{
-		Set $\eta_0$ and $r_0$ such that the duration of the burst, and the total number of observed photons are compatible with GRB 090926A, $T_{0.99}\left(z,\chi;\unit[0.1]{GeV},\infty\right) = \unit[219.5]{sec}$ and $p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\infty\right) = 179.989$
+		Set $\eta_0$ and $r_0$ such that the duration of the burst, and the total number of observed photons are compatible with GRB 090926A, $T_{0.99}\left(z,\chi;\unit[0.1]{GeV},\infty\right) = \unit[219.5]{sec}$ and $p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\infty\right) = 179.996$
 	}
@@
 		\end{equation}
-		Here $\kappa_\text{min}$ is the minimal stretching factor from the sample computed in the step \ref{item:sample}, $\log\kappa_{\text{GRB090902B}} = \frac{1}{2}\left(\log\left(0.89\right) + \log\left(0.36\right)\right)$ and $\Delta\log\kappa_{\text{GRB090902B}} = \frac{1}{2}\left(\log\left(0.89\right) - \log\left(0.36\right)\right)$.
+		Here $\kappa_\text{min}$ is the minimal stretching factor from the sample computed in the step \ref{item:sample}, $\log\kappa_{\text{GRB090902B}} = \frac{1}{2}\left(\log\left(0.89\right) + \log\left(0.35\right)\right)$ and $\Delta\log\kappa_{\text{GRB090902B}} = \frac{1}{2}\left(\log\left(0.89\right) - \log\left(0.35\right)\right)$.
 	}
@@
 		\end{equation}
-		Here $f_{\text{GRB090926A}} = 0.0606777$.
+		Here $f_{\text{GRB090926A}} = 0.0570003$.
 	}
@@
 		\item{$\gamma = 300$}
-		\item{$\eta_0 = \unit[8.14796255480381 \times 10^{34}]{sec^{-3} GeV^{-1}}$}
-		\item{$r_0 = \unit[3.357930930824906 \times 10^6]{sec}$}
-		\item{$\bm{n = 22.5269}$}
+		\item{$\eta_0 = \unit[4.629640817921788 \times 10^{34}]{sec^{-3} GeV^{-1}}$}
+		\item{$r_0 = \unit[3.142715108207946 \times 10^6]{sec}$}
+		\item{$\bm{n = 22.9977}$}
 		\item{$\omega_0 = \unit[1]{GeV}$}
-		\item{$\bm{\theta_0 = 5.02035 \times 10^{-5}}$}
-		\item{$\bm{k = -0.699813}$}
-		\item{$\bm{\alpha = -0.626944}$}
+		\item{$\bm{\theta_0 = 7.93393 \times 10^{-5}}$}
+		\item{$\bm{k = -0.65356}$}
+		\item{$\bm{\alpha = -0.724394}$}
 		\item{$z = 2.1062$}
-		\item{$\bm{\chi = 0.00541037}$}
+		\item{$\bm{\chi = 0.00590163}$}
 \end{itemize}
@@
 		\item{
-			The fraction of bursts observable in low energy band, which are also observed in high energy band $f_m = 0.116$.
+			The fraction of bursts observable in low energy band, which are also observed in high energy band $f_m = 0.110$.
 			The Fermi LAT catalog contains $35$ bursts out of which $4$ can be observed in high energy band.

### 2014-09-01 Paper/gammaRays.tex
Changed summary from bullet list to text

@@
 
-	The main results of our study may be summarized as following:
-	\begin{enumerate}
-		\item{
-			Time stretching of GRB light curves between different VHE bands is discovered with statistical significance of $3.3\sigma$.
-		}
-		\item{
-			This time stretching can be explained by geometry of the jet, without introducing any new spectral components.
-		}
-		\item{
-			All GRBs can be assumed to be the same in their rest frames.
-			There is no need to vary internal burst parameters.
-		}
-	\end{enumerate}
-
-	There are, however, directions for improvement:
-	\begin{itemize}
-		\item{
-			The model prediction and observations on fig. \ref{fig:correlations} are, though close, not exactly agree.
-		}
-		\item{
-			Shapes of light curves produced by the model differ considerably from observed ones.
-		}
-		\item{
-			Assumption of energy dependence on off-axis angle should be justified better.
-		}
-		\item{
-			While we believe that all bursts are almost the same in their rest frames, they are certainly not exactly the same.
-			The minor differences between GRBs might be responsible for observable deviations from the model predictions.
-		}
-	\end{itemize}
-
-	We believe, however, that these problems are not unresolvable, and may be addressed by refining the model without changing its main assumptions.
+	Our study can be summarized with 3 main conclusions.
+
+	First, the time stretching of GRB light curves between different VHE (in particular $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$) bands is discovered with statistic significance of $3.3\sigma$.
+	This time stretching might be positive or negative (that is higher or lower energy light curve is stretched) depending on the burst.
+
+	Second, this time stretching effect can be explained with curvature effects, that is effects of jet geometry.
+	There is no need to introduce any new spectral components.
+
+	Finally, we can assume all GRBs to be the same in their rest frames.
+	The internal burst parameters (such as $\gamma$, $\eta_0$, $r_0$, $n$, $\omega_0$, $\theta_0$, $k$ and $\alpha$) might stay the same for all bursts, and this assumption is more or less consistent with existing observations.
+
+	However, there are some caveats, and directions for improvement.
+
+	First of all, the model prediction and observations are not completely consistent.
+	If you look at fig. \ref{fig:correlations}, you will see that the stretching factor of GRB 090926A is too high compared to the model prediction.
+	It appears that parameters of the model cannot be fitted to account for it (without ruining agreement with other observations).
+	More data should be analyzed to see whether the problem will be exacerbated, and some refinement of the model might also be required.
+
+	Second, the shapes of light curves produced by the model (see fig. \ref{fig:sampleLightCurves}) differ considerably from observed ones (e.g. fig. \ref{fig:grb090902B}).
+	Most importantly, the slope of light curves produced by the model is zero at $t = 0$, even though the slope of observed light curves at $t = 0$ is not only non-zero, but appears to have a maximum.
+
+	Third, the assumption of plasma energy dependence on off-axis angle should be better justified.
+	This might probably be done using hydrodynamic simulation of the jet.
+
+	Finally, even though it appears that bursts are similar in their rest frames, they cannot be exactly the same.
+	The differences in masses, metallicities, angular momenta of progenitor stars, and the differences in local conditions around them might account for differences in light curves of GRBs (or, even, for especially large stretching factor of GRB 090926A).
+	It means that minor variations of model parameters might be necessary on burst per burst basis.
+
+	It seems, however, that all these problems are not unresolvable, and may be addressed by refining the model without changing its main assumptions.
 

### 2014-10-30 QuarksProceedings/piskunov.tex
Added Quarks proceedings

--- 2014-05-25 ITEPProceedings/piskunov.tex
@@
-\documentclass[
-aps,%
-12pt,%
-final,%
-notitlepage,%
-oneside,%
-onecolumn,%
-nobibnotes,%
-nofootinbib,% 
-superscriptaddress,%
-showpacs,%
-centertags]%
-{revtex4}
+\documentclass[11pt,a4paper]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{epsfig,graphicx,units,hyperref,subcaption}
 
-\usepackage{units}
-
-\newcommand{\exclude}[1]{}
+\topmargin -0.4cm
+\headsep=0.0cm
+\headheight=0.0cm
+\textheight 24.6cm
+\oddsidemargin -0.3cm
+\evensidemargin -0.3cm
+\textwidth 15.9cm
 
 \begin{document}
-\selectlanguage{english}
 
-%\newcommand{\beq}{\begin{equation}}
-%\newcommand{\eeq}{\end{equation}}
-
-%\newcommand{\lll}{\lambda}
-
-% Some useful journal names
-
-%\newcommand{\A}{\alpha}
-%\newcommand{\B}{\beta}
-%\newcommand{\T}{\theta}
-%\newcommand{\Ep}{\epsilon}
-%\newcommand{\beqn}{\begin{eqnarray}}
-%\newcommand{\eeqn}{\end{eqnarray}}
-
-\title{TIME STRETCHING OF THE GEV EMISSION OF GRBS: FERMI LAT DATA VS GEOMETRICAL MODEL}
-
-\author{\firstname{M.~S.}~\surname{Piskunov}}
-\email{maxitg@icloud.com}
-\affiliation{Institute for Nuclear Research, Moscow, Russia}
-
-%\date{\today}
+\title{\bf Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model}
+\author{M.~S.~Piskunov$^a$\footnote{{\bf e-mail}: maxitg@icloud.com},
+G.~I.~Rubtsov$^{a}$\footnote{{\bf e-mail}: grisha@ms2.inr.ac.ru}
+\\
+$^a$ \small{\em Institute for Nuclear Research RAS} \\
+\small{\em Prospekt 60-letiya Oktyabrya 7a, Moscow 117312}
+}
+\date{}
+\maketitle
 
@@
 In this paper we consider all the bursts observed by Fermi-LAT since 2008 August 4 to 2011 August 1, for which at least $10$ events with energies $\unit[1]{GeV}$ or higher were observed.
-	There are $4$ of them: 080916C, 090510, 090902B, and 090926A.
+There are $4$ of them: GRB 080916C, GRB 090510, GRB 090902B, and GRB 090926A.
 We study their light curves in two bands, $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.
+
 The Kolmogorov-Smirnov test is used to check whether the light curves for these two bands are the same.
-	No significant difference was found for 080916C and 090510.
-	However, we observed with statistical significance of $3.3 \sigma$, that the higher energy light curve of 090926A is stretched with respect to the lower-energy one, and with statistical significance of $2.2 \sigma$, that the lower energy light curve of 090902B is stretched with respect to the higher-energy one.
+No significant difference was found for GRB 080916C and GRB 090510.
+However, we observed with statistical significance of $3.3 \sigma$, that the higher energy light curve of GRB 090926A is stretched with respect to the lower-energy one, and with statistical significance of $2.3 \sigma$, that the lower energy light curve of GRB 090902B is stretched with respect to the higher-energy one.
 
@@
 The main assumption is the jet opening angle dependence on radiation energy -- the most energetic photons are emitted near the axis of the jet.
-	We also assume that all bursts are the same in their reference frames (that is their light curves differ because of different redshifts and off-axis angles).
+We also assume that all bursts are the same in their rest frames (that is their light curves differ only because of different redshifts and different view directions).
 To test this model, we compute the total energy of the burst, and confirm that it is below the constraint.
@@
 This fraction matches the observations.
-	Finally, we predict the distribution of observable stretching factors, which may be tested in the future when more observational data will be available.
+We predict the distribution of observable stretching factors, which may be tested in the future when more observational data become available.
+Finally, we propose a way to estimate observer's off-axis angles (view directions) based on stretching factors and fractions of high energy photons.
+
+All our code is open source, and is available freely on GitHub\footnote{\url{https://github.com/maxitg/GammaRays}}.
 \end{abstract}
-
-\pacs{95.85.Pw,~97.60.Bw}
-
-\maketitle
 
@@
 Gamma Ray Bursts (GRBs) are among the most energetic events in the Universe, therefore they might provide new knowledge for particle physics.
-	Modern observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} made possible to study these explosions extensively \cite{Vianello:2013ela,Gehrels:2013xd}.
+Modern observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift, \cite{Gehrels:2004gu} made it possible to study these explosions extensively \cite{Vianello:2013ela,Gehrels:2013xd}.
 These and previous studies led to several interesting results.
 For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}.
-	This suggests that most of the bursts must have similar energetics in their rest frame.
+This suggests that the majority of the bursts must have similar energetics in their rest frame.
 Temporal variations of spectra were also studied.
 The spectral lags were found between different low energy bands \cite{Yi:2005ht}.
-	The very high energy radiation was discovered to be extended relative to x-ray emission \cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}.
+The very high energy radiation was discovered to be extended relative to the x-ray emission \cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}.
 In this study we continue this research and explore the spectral variations between energy bands $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
@@
 Then we propose a simple geometrical model to explain these results in section \ref{sec:model}.
-	Finally, our study is summarized in section \ref{sec:discussion}.
+Finally, our study is summarized in section \ref{sec:summary}.
 
@@
 
-	We take the GRB list from the Fermi-LAT catalog \cite{Ackermann:2013zfa}, and use the technique introduced in \cite{Rubtsov:2011qq} to estimate background radiation in both energy bands.\footnote{Detailed description of computations in this and the next section (including the source code) can be found on GitHub: \url{https://github.com/maxitg/GammaRays}}
+We take the GRB list from the Fermi-LAT catalog \cite{Ackermann:2013zfa}, and use the technique introduced in \cite{Rubtsov:2011qq} to estimate background radiation in both energy bands.\footnote{Detailed description of computations in this and the next section can be found in the extended version of the paper on GitHub.}
 Most of the bursts in the catalog, however, do not have enough events with over $\unit[1]{GeV}$ energies to do desired computations, so we choose only those bursts, from which at least $10$ photons were detected in the high energy band.
@@
 
-	So we obtained a preliminary (for significance is less than $5\sigma$) result that the stretching factors for observable bursts might be both larger and smaller than $1$.
+\begin{figure}
+        \centering
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{lightCurve090902B}
+                \label{fig:lightCurve090902B}
+        \end{subfigure}
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{probabilities090902B}
+                \label{fig:probabilities090902B}
+        \end{subfigure}
+        \caption{GRB 090902B results. Low energy radiation is stretched ($\kappa < 1$) with significance of $2.3\sigma$.}
+        \label{fig:grb090902B}
+\end{figure}
+
+\begin{figure}
+        \centering
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{lightCurve090926A}
+                \label{fig:lightCurve090926A}
+        \end{subfigure}
+        \begin{subfigure}{0.49\textwidth}
+                \includegraphics[width=\textwidth]{probabilities090926A}
+                \label{fig:probabilities090926A}
+        \end{subfigure}
+        \caption{GRB 090926A results. High energy radiation is stretched ($\kappa > 1$) with significance of $3.3\sigma$.}
+        \label{fig:grb090926A}
+\end{figure}
+
+\begin{table}
+	\centering
+	\small
+	\begin{tabular}{ l | c | c | c | c | c | c | c}
+		GRB 	& 	Redshift	&	GBM Trigger Time	& R.A., Dec.			&	Stretching factor	&	Stretching factor	\\
+		Name 	& 				&	MET sec				& J2000, deg			&	$2\sigma$ range		&	$3\sigma$ range		\\
+		\hline
+		080916C	&	$4.35$		&	243216766.614		&	119.85,	-56.64 		&	0.67 -- 3.32		&	0.42 -- 5.83		\\
+		090510 	&	$0.903$		&	263607781.971		&	333.55, -26.58		&	0.43 -- 1.61		&	0.32 -- 2.29		\\
+		090902B	&	$1.822$		&	273582310.313		&	264.94,  27.32		&	0.35 -- 0.89		&	0.22 -- 1.53		\\
+		090926A	&	$2.1062$	&	275631628.990		&	353.40, -66.32 		&	1.99 -- 6.62		&	1.34 -- 9.15		\\
+	\end{tabular}
+	\caption{
+		Stretching factor computation results for the bursts studied. Ranges of allowed stretching factors are shown for studied bursts for multiple levels of significance.
+		The values for trigger time and location are taken from the tables 2 and 4 of \cite{Ackermann:2013zfa}.
+	}
+	\label{tab:bursts}
+\end{table}
+
+So we obtained a preliminary (note that its significance is less than $5\sigma$) result that the stretching factors for observable bursts might be both larger and smaller than $1$.
 
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.7\textwidth]{modelOverview}
+	\caption{
+        Model Overview.
+        Here lighter and darker cones represent the regions through which low and high energy plasma propagates.
+        In case depicted the observer's off-axis angle is smaller than the opening angle of the low energy jet, so, due to the relativistic beaming effect, the most of the observable low energy photons will travel along the straight line from the central engine.
+        Also, the observer's off-axis angle is larger that the opening angle of the high energy jet, so the high energy radiation will still originate near the center of the jet (because it is the only place where there are high energy radiators).
+        The observation time of a photon is a sum of two things: the time interval spent in plasma as a radiator (which approximately equals to the distance from the central engine to the point of emission); and the time interval from emission to detection (which is the distance from the point of emission to the observer's location).
+        Given a position of the shock front, this sum is larger for high energy photons.
+        Because of that, high energy emission will be observed later throughout the burst duration, therefore the high energy light curve will be stretched.
+        }
+    \label{fig:modelOverview}
+\end{figure}
+
 The burst is fully specified by the following set of parameters (we assume these parameters to be the same for all bursts, the specified values are obtained by fitting to make possible observed bursts from section \ref{sec:data}):
@@
 	\item{$\gamma = 300$, the relativistic factor of the shell, $\gamma \gg 1$.}
-		\item{$\eta_0 = \unit[8.15 \times 10^{34}]{sec^{-3} GeV^{-1}}$, which defines the luminosity scale.}
-		\item{$r_0 = \unit[3.36 \times 10^6]{sec}$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
-		\item{$n = 22.5$, which determines the sharpness of the jet end, $n > 3$;}
+	\item{$\eta_0 = \unit[4.63 \times 10^{34}]{sec^{-3} GeV^{-1}}$, which defines the luminosity scale.}
+	\item{$r_0 = \unit[3.14 \times 10^6]{sec}$, the characteristic jet length; $r_0 \ll \frac{1}{H\left(0\right)}$, $H\left(t\right)$ is the Hubble parameter;}
+	\item{$n = 23.0$, which determines the sharpness of the jet end, $n > 3$;}
 	\item{$\omega_0 = \unit[1]{GeV}$, a characteristic radiation frequency;}
-		\item{$\theta_0 = 5.02 \times 10^{-5}$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;}
-		\item{$k = -0.700$, which determines how much the opening angle changes with frequency, $k < 0$;}
-		\item{$\alpha = -0.627$, the bare spectral index, $\alpha < -2k - 1$}
+	\item{$\theta_0 = 7.93 \times 10^{-5}$, the opening angle of the jet for radiation with frequency $\omega_0$, $\theta_0 \ll 1$;}
+	\item{$k = -0.654$, which determines how much the opening angle changes with frequency, $k < 0$;}
+	\item{$\alpha = -0.724$, the bare spectral index, $\alpha < -2k - 1$}
 \end{itemize}
@@
 \ref{fig:sampleLightCurves}).
+
+\begin{figure}
+	\centering
+	\hspace*{\fill}
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[width=\textwidth]{sampleLightCurveLogNegative}
+		\label{fig:sampleLightCurveLogNegative}
+		\caption{Stretching factor $\kappa = 0.819$, redshift $z = 1.82$, off-axis angle $\chi = 0$.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[width=\textwidth]{sampleLightCurveLogPositive}
+		\label{fig:sampleLightCurveLogPosivie}
+		\caption{Stretching factor $\kappa = 2.43$, redshift $z = 2.106$, off-axis angle $\chi = 5.90 \times 10^{-3}$.}
+	\end{subfigure}
+	\hspace*{\fill}
+	\caption{
+		High and low energy light curves produced by the geometrical model.
+		Burst parameter values are listed in section \ref{sec:model}.
+	}
+	\label{fig:sampleLightCurves}
+\end{figure}
 
@@
 	\item{
-			The fraction of bursts observable in low energy band, which are also observed in high energy band $f_m = 0.116$.
+		The fraction of bursts observable in low energy band, which are also observed in high energy band $f_m = 0.110$.
 		The Fermi LAT catalog contains $35$ bursts out of which $4$ can be observed in high energy band.
@@
 
-	So our model, though with some discrepancies, is able to explain the time stretching result from section \ref{sec:data}.
-
-\section{Discussion}
-	\label{sec:discussion}
-
-	The main results of our study may be summarized as following:
-	\begin{enumerate}
-		\item{
-			The time stretching of GRB light curves between different high energy bands is confirmed with statistical significance of $3.3\sigma$.
+\begin{figure}
+	\hspace*{\fill}
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[width=\textwidth]{kappaDistributionHistogram}
+		\label{fig:kappaDistributionHistogram}
+		\caption{Histogram obtained from the model. Error bars show observed stretching factors.}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}{0.45\textwidth}
+		\includegraphics[width=\textwidth]{kappaDistributionCDF}
+		\label{fig:kappaDistributionCDF}
+		\caption{CDF obtained from the model. Steps and error bars show observed stretching factors.}
+	\end{subfigure}
+	\hspace*{\fill}
+	\caption{
+		Stretching factors histogram and CDF produced by our model.
+		The sample contains $4096$ bursts.
+		Numbered error bars correspond respectively to GRB 080916C, GRB 090510, GRB 090902B and GRB 090926A.
 	}
-		\item{
-			This time stretching can be explained by geometry of the jet, without introducing any new spectral components.
-		}
-		\item{
-			All GRBs are assumed the same in their rest frames.
-			There is no need to vary internal burst parameters.
-		}
-	\end{enumerate}
-
-	There, however, directions for improvement:
-	\begin{itemize}
-		\item{
-			The model prediction and observations on fig. \ref{fig:correlations} are, though close, not exactly agree.
-		}
-		\item{
-			Shapes of light curves produced by model differ considerably from observed ones.
-		}
-		\item{
-			Assumption of energy dependence on off-axis angle should be better justified.
-		}
-		\item{
-			Much more results can be obtained from the model (like energy spectrum, and its variation with time).
-		}
-	\end{itemize}
-
-	We believe, however, that these problems are not unresolvable, and may be addressed by refining the model without changing its main assumptions.
-	And in the end, we think the main ideas of this paper will be close to truth.
-
-\bibliographystyle{maik}
-\bibliography{piskunov}
-
-\newpage
-
-\selectlanguage{russian}
-
-\begin{center}
-\large \bfseries %\MakeTextUppercase{%
-ВРЕМЕННОЕ РАСТЯЖЕНИЕ ГЭВ ИЗЛУЧЕНИЯ ГАММА-ВСПЛЕСКОВ: FERMI LAT И ГЕОМЕТРИЧЕСКАЯ МОДЕЛЬ
-\end{center}
-%
-\begin{center}
-{\bfseries
-М.~С.~Пискунов
-}
-\end{center}
-%
-\begin{center}
-\begin{minipage}{\textwidth - 2cm}
-	\small
-	Наблюдения показывают, что излучение гамма-всплесков с энергией выше \unit[100]{МэВ} систематически наблюдается позже, чем низкоэнергичное излучение.
-	Различия же кривых блеска в различных диапазонах высокоэнергичного излучения (> \unit[100]{МэВ}) изучены хуже.
-
-	В данной работе мы изучаем различия кривых блеска в диапазонах (\unit[100]{МэВ}, \unit[1]{ГэВ}) и (\unit[1]{ГэВ}, \unit[300]{ГэВ}).
-	Для исследования выбраны все 4 гамма-всплеска из каталога Fermi, наблюдения которых содержат не менее 10 фотонов с энергией > \unit[1]{ГэВ}: 080916C, 090510, 090902B, 090926A.
-	Кривые блеска 080916C и 090510 в двух рассматриваемых диапазонах не различаются статистически значимо, однако для 090926A установлено со статистической значимостью $3.3\sigma$, что излучение в диапазоне (\unit[1]{ГэВ}, \unit[300]{ГэВ}) растянуто относительно менее энергичного, также для 090902B установлено со статистической значимостью $2.2\sigma$, что излучение в диапазоне (\unit[100]{МэВ}, \unit[1]{ГэВ}) растянуто относительно более энергичного.
-
-	Мы предлагаем простую геометрическую модель, объясняющую этот результат.
-	Основное предположение -- чем выше энергия излучения, тем ближе к оси джета оно излучается.
-	Мы также предполагаем, что все всплески одинаковы в собственной системе отсчета (то есть, различия кривых блеска объясняются исключительно разницой в красном смещении и углом наблюдения).
-	В рамках модели соблюдается ограничение на максимально допустимую энергию всплеска.
-	Также получена доля всплесков среди наблюдаемых в диапазоне (\unit[100]{МэВ}, \unit[1]{ГэВ}), которые также можно увидеть в диапазоне (\unit[1]{ГэВ}, \unit[300]{ГэВ}).
-	Полученная доля согласуется с наблюдениями.
-	Наконец, предсказано распределение коэффициентов растяжения, которое может быть проверено по мере накопления наблюдений.
-\end{minipage}
-\end{center}
-\selectlanguage{english}
-%
-\newpage
-{\bf Figure captions}
-\vskip 1cm
-
-{\bf Fig.~1.}~GRB 090902B results. Low energy radiation is stretched ($\kappa < 1$) with significance of $2.2\sigma$.
-
-{\bf Fig.~2.}~GRB 090926A results. High energy radiation is stretched ($\kappa > 1$) with significance of $3.3\sigma$.
-
-{\bf Tab.~1.}~Stretching factor computation results for the bursts studied.
-Ranges of allowed stretching factors are shown for studied bursts for multiple levels of significance.
-The values for trigger time and location are taken from the tables 2 and 4 of \cite{Ackermann:2013zfa}.
-
-{\bf Fig.~3.}~Model Overview.
-Here the red and the blue cones represent the volumes through which low and high energy plasma propagates.
-In case depicted the observer's off-axis angle is smaller than the opening angle of the low energy jet, so, due to the relativistic beaming effect, the most of the observable low energy photons will travel along the straight line from the central engine.
-Also, the observer's off-axis angle is larger that the opening angle of the high energy jet, so the high energy radiation will still originate near the center of the jet (because it is the only place where there are high energy radiators).
-The observation time of a photon is a sum of two things: the time interval spent in plasma as a radiator (which approximately equals to the distance from the central engine to the point of emission); and the time interval from emission to detection (which is the distance from the point of emission to the observer's location).
-Given a position of the shock front, this sum is larger for high energy photons.
-Because of that, high energy emission will be observed later throughout the burst duration, therefore the high energy light curve will be stretched.
-
-{\bf Fig.~4.}~High and low energy light curves produced by the geometrical model.
-Burst parameter values are the same as discussed in section \ref{sec:model}.
-On the left panel: the stretching factor $\kappa = 0.821$, redshift $z = 1.82$, off-axis angle $\chi = 0$.
-On the right: stretching factor $\kappa = 2.26$, redshift $z = 2.106$, off-axis angle $\chi = 5.41 \times 10^{-3}$.
-
-{\bf Fig.~5.}~Stretching factors histogram and CDF produced by our model. Bar lines (black) show observed stretching factors. The sample contains $4096$ bursts.
-
-{\bf Fig.~6.}~Correlations between high energy photon ratios, off-axis angles and stretching factors found in the sample produces by our model.
-The sample contains $4096$ bursts.
-Black crosses show observations.
-This correlation allows one to estimate off-axis angles of observed bursts.
-
-\newpage
-
-\begin{figure}
-\includegraphics[width=\textwidth]{p1.pdf}
-\caption{Piskunov, Fig. 1}
-\label{fig:grb090902B}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{p2.pdf}
-\caption{Piskunov, Fig. 2}
-\label{fig:grb090926A}
-\end{figure}
-
-\begin{table}
-	\centering
-	\small
-	\begin{tabular}{ l | c | c | c | c | c | c | c}
-		GRB 	& 	Redshift	&	GBM Trigger Time	& R.A.			& Dec.			&	Stretching factor	&	Stretching factor	\\
-		Name 	& 				&	MET sec				& J2000, deg	& J2000, deg	&	$2\sigma$ range		&	$3\sigma$ range		\\
-		\hline
-		080916C	&	$4.35$		&	243216766.614		&	119.85 		&	-56.64 		&	0.67 -- 3.32		&	0.42 -- 5.83		\\
-		090510 	&	$0.903$		&	263607781.971		&	333.55 		&	-26.583		&	0.43 -- 1.61		&	0.32 -- 2.29		\\
-		090902B	&	$1.822$		&	273582310.313		&	264.94 		&	 27.324		&	0.36 -- 0.89		&	0.23 -- 1.57		\\
-		090926A	&	$2.1062$	&	275631628.990		&	353.4  		&	-66.32 		&	1.99 -- 6.62		&	1.34 -- 9.15		\\
-	\end{tabular}
-	\caption{Piskunov, Tab. 1}
-	\label{tab:bursts}
-\end{table}
-
-\begin{figure}
-\includegraphics[width=.7\textwidth]{p3.pdf}
-\caption{Piskunov, Fig. 3}
-\label{fig:modelOverview}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{p4.pdf}
-\caption{Piskunov, Fig. 4}
-\label{fig:sampleLightCurves}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{p5.pdf}
-\caption{Piskunov, Fig. 5}
 	\label{fig:kappaDistribution}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{p6.pdf}
-\caption{Piskunov, Fig. 6}
+        \centering
+        \includegraphics[width=1.0\textwidth]{correlations}
+        \caption{
+        	Correlations between off-axis angles, stretching factors and high to low energy photon count ratios found in the sample produced by our model.
+        	The sample contains $4096$ bursts.
+        	Numbered black crosses show data respectively for GRB 080916C, GRB 090510, GRB 090902B and GRB 090926A with $2\sigma$ error bars.
+        	This correlation allows one to estimate off-axis angles of observed bursts.
+        }
         \label{fig:correlations}
@@
 
+So our model, though with some discrepancies, is able to explain the time stretching result from section \ref{sec:data}.
+
+\section{Summary}
+
+\label{sec:summary}
+
+Our study can be summarized in 3 main points.
+
+First, the time stretching of GRB light curves between different VHE bands, in particular $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$, is discovered with statistical significance of $3.3\sigma$.
+This time stretching might be positive or negative (that is higher or lower energy light curve is stretched) depending on the burst.
+
+Second, this time stretching effect can be explained with curvature effects, that is effects of jet geometry.
+There is no need to introduce any new spectral components.
+
+Finally, we can assume all GRBs to be the same in their rest frames.
+The internal burst parameters (such as $\gamma$, $\eta_0$, $r_0$, $n$, $\omega_0$, $\theta_0$, $k$ and $\alpha$) might stay the same for all bursts, and this assumption is more or less consistent with existing observations.
+
+However, there are some caveats, and directions for improvement.
+
+First of all, the model prediction and observations are not completely consistent.
+If you look at fig. \ref{fig:correlations}, you will see that the stretching factor of GRB 090926A is too high compared to the model prediction.
+It appears that parameters of the model cannot be fitted to account for it (without ruining agreement with other observations).
+More data should be analyzed to see whether the problem will be exacerbated, and some refinement of the model might also be required.
+
+Second, the shapes of light curves produced by the model (see fig. \ref{fig:sampleLightCurves}) differ considerably from observed ones (e.g. fig. \ref{fig:grb090902B}).
+Most importantly, the slope of light curves produced by the model is zero at $t = 0$, even though the slope of observed light curves at $t = 0$ is not only non-zero, but appears to have a maximum.
+
+Third, the assumption of plasma energy dependence on off-axis angle should be better justified.
+This might probably be done using hydrodynamic simulation of the jet.
+
+Finally, even though it appears that bursts are similar in their rest frames, they cannot be exactly the same.
+The differences in masses, metallicities, angular momenta of progenitor stars, and the differences in local conditions around them might account for differences in light curves of GRBs (or, even, for especially large stretching factor of GRB 090926A).
+It means that minor variations of model parameters might be necessary on burst per burst basis.
+
+It seems, however, that all these problems are not unresolvable, and may be addressed by refining the model without changing its main assumptions.
+
+{\small {\it Acknowledgments.} The work was supported by Russian Science Foundation grant 14-12-01340.}
+
+\bibliographystyle{plain}
+\bibliography{piskunov}
+
 \end{document}

### 2014-11-03 Paper/gammaRays.tex
Abstract revised

@@
 \usepackage{gensymb}
+\usepackage{booktabs}
 
@@
 \title{
-	Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model \\
-	{\large \url{https://github.com/maxitg/GammaRays}}
+	Time stretching of the GeV emission of GRBs: Fermi LAT data vs geometrical model
 }
@@
 \author{
-	Maxim S. Piskunov$^{1}$\thanks{\href{mailto:maxit@ms2.inr.ac.ru}{maxit@ms2.inr.ac.ru}} and
-	Grigory I. Rubtsov$^{1}$\thanks{\href{mailto:grisha@ms2.inr.ac.ru}{grisha@ms2.inr.ac.ru}} \\
-	$^{1}$Institute for Nuclear Research RAS, 117312, Moscow, Russia\\
+	Maxim S. Piskunov$^{1,2}$\thanks{\href{mailto:maxit@ms2.inr.ac.ru}{maxit@ms2.inr.ac.ru}} ~and
+	Grigory I. Rubtsov$^{2,3}$\thanks{\href{mailto:grisha@ms2.inr.ac.ru}{grisha@ms2.inr.ac.ru}} \\
+	{\small $^{1}$Faculty of Physics, Moscow State University, Moscow, Russia}\\
+	{\small $^{2}$Institute for Nuclear Research of the RAS, Moscow, Russia}\\
+	{\small $^{3}$Novosibirsk State University, Novosibirsk, Russia}
 }
@@
 \begin{abstract}
-	Numerous observations confirm that the high energy $(> \unit[100]{MeV})$ emission of gamma ray bursts is delayed with respect to the low energy emission.
-	However, the difference of light curves in various high energy bands has not been studied properly.
-
-	In this paper we consider all the bursts observed by Fermi-LAT since 2008 August 4 to 2011 August 1, for which at least $10$ events with energies $\unit[1]{GeV}$ or higher were observed.
-	There are $4$ of them: GRB 080916C, GRB 090510, GRB 090902B, and GRB 090926A.
-	We study their light curves in two bands, $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$.
-
-	The Kolmogorov-Smirnov test is used to check whether the light curves for these two bands are the same.
-	No significant difference was found for GRB 080916C and GRB 090510.
-	However, we observed with statistical significance of $3.3 \sigma$, that the higher energy light curve of GRB 090926A is stretched with respect to the lower-energy one, and with statistical significance of $2.3 \sigma$, that the lower energy light curve of GRB 090902B is stretched with respect to the higher-energy one.
-
-	We suggest a simple geometrical model to explain this result.
-	The main assumption is the jet opening angle dependence on radiation energy -- the most energetic photons are emitted near the axis of the jet.
-	We also assume that all bursts are the same in their rest frames (that is their light curves differ only because of different redshifts and different view directions).
-	To test this model, we compute the total energy of the burst, and confirm that it is below the constraint.
-	We also compute the fraction of observable bursts in $(\unit[100]{MeV}, \unit[1]{GeV})$ band, which can also be observed in higher energies.
-	This fraction matches the observations.
-	We predict the distribution of observable stretching factors, which may be tested in the future when more observational data become available.
-	Finally, we propose a way to estimate observer's off-axis angles (view directions) based on stretching factors and fractions of high energy photons.
-
-	All our code is open source, and is available freely on GitHub for you to download and play with.
+	It is known that the high energy $(> \unit[100]{MeV})$
+        emission of gamma-ray bursts is delayed with respect to the
+        low energy emission.  However, the dependence of light curves
+        on energy has not been studied for the high energy bands.  In
+        this paper we consider the bursts observed by Fermi-LAT from
+        2008 August 4 to 2011 August 1, for which at least $10$
+        photons were observed with energy greater than
+        $\unit[1]{GeV}$. These include $4$ bursts: GRB 080916C, GRB
+        090510, GRB 090902B, and GRB 090926A.  We use the
+        Kolmogorov-Smirnov test to compare the light curves in the two
+        bands, $\unit[100]{MeV} < E < \unit[1]{GeV}$ and
+        $\unit[1]{GeV} < E < \unit[300]{GeV}$. For GRB 080916C and GRB
+        090510 the light curves in two bands are statistically
+        compatible. However, for GRB 090926A, the higher-energy light
+        curve is stretched compared to the lower-energy one with a
+        statistical significance of $3.3 \sigma$ and, for GRB 090902B,
+        on the contrary, the lower-energy curve is stretched with $2.3
+        \sigma$ significance. We argue that the observed diversity of
+        stretching factors may be explained in a simple geometrical
+        model. The model assumes that the jet opening angle depends on
+        the emission energy in a way that the most energetic photons
+        are radiated near the axis of the jet. All the bursts are
+        considered equivalent in their rest frames and the observed
+        light curves differ only due to different redshifts and view
+        directions. The model conforms to the total burst energy
+        constraint and matches the Fermi LAT observations of the
+        fraction of GRBs visible in $\unit[100]{MeV} < E <
+        \unit[1]{GeV}$ band, which may be observed at higher
+        energies. The model predicts the distribution of observable
+        stretching factors, which may be tested in the future
+        data. Finally, we propose a way to estimate observer's
+        off-axis angle based on stretching factor and fraction of the
+        high-energy photons. The code for modeling is open source and
+        is publicly available on GitHub
+        (\url{https://github.com/maxitg/GammaRays}).
 \end{abstract}
 
+%TODO: add grants
+%? ref to Lemoine GRB1
+
 \section{Introduction}
 
-	Gamma Ray Bursts (GRBs) are among the most energetic events in the Universe, therefore they might provide new knowledge for particle physics.
-	Modern observatories, specifically Fermi \cite{Ackermann:2012kna} and Swift \cite{Gehrels:2004gu} made possible to study these explosions extensively \cite{Vianello:2013ela,Gehrels:2013xd}.
-	These and previous studies led to several interesting results.
-	For example, the total energy emitted in gamma-rays during a burst was found to be similar among different bursts within an order of magnitude \cite{Bloom:2003wy}.
-	This suggests that most of the bursts must have similar energetics in their rest frames.
-	Temporal variations of spectra were also studied. The spectral lags were found between different low energy bands \cite{Yi:2005ht}.
-	The very high energy radiation was discovered to be extended relative to x-ray emission \cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}.
-	We decided to elaborate on that result, and, using the data from Large Area Telescope (LAT) of Fermi, explore the spectral variations between different high energy bands, specifically $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
+Gamma-ray bursts (GRBs) are among the most energetic events in the
+Universe. Therefore, the studies of the emission mechanism and
+phenomenology of GRBs may provide the new knowledge in particle
+physics. Moreover, the GRBs are observed from cosmological distances
+and therefore bear an imprint of the late-time evolution of the
+Universe. An extensive studies of these explosions led to a number of
+interesting results, see~\cite{Vianello:2013ela,Gehrels:2013xd} for a
+review. In particular, the total energy emitted in gamma-rays during a
+burst was found to be similar for the different GRBs within an order
+of magnitude~\cite{Bloom:2003wy}. This indicates that the most of the
+bursts have similar energetics in their rest frames. Several
+observations are related to the temporal variations of spectra. This
+way, the spectral lags were found between different low energy bands
+\cite{Yi:2005ht} and the very-high-energy radiation was discovered to
+be extended relative to the x-ray emission
+\cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}.
+
+%% change the next statement
+	We decided to elaborate on that result, and, using the data
+        from Large Area Telescope (LAT) of Fermi satellite~\cite{Ackermann:2012kna}, explore the spectral variations between different high energy bands, specifically $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
 	In particular, we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
@@
 
-\bibliographystyle{hep}
+\bibliographystyle{ieeetr}
 \bibliography{gammaRays.bib}

### 2014-12-11 Paper/gammaRays.tex
updated introduction

@@
         2008 August 4 to 2011 August 1, for which at least $10$
-        photons were observed with energy greater than
+        photons were observed with the energy greater than
         $\unit[1]{GeV}$. These include $4$ bursts: GRB 080916C, GRB
@@
         bands, $\unit[100]{MeV} < E < \unit[1]{GeV}$ and
-        $\unit[1]{GeV} < E < \unit[300]{GeV}$. For GRB 080916C and GRB
-        090510 the light curves in two bands are statistically
+        $E > \unit[1]{GeV}$. For GRB 080916C and GRB
+        090510 the light curves in the two bands are statistically
         compatible. However, for GRB 090926A, the higher-energy light
@@
         stretching factors, which may be tested in the future
-        data. Finally, we propose a way to estimate observer's
-        off-axis angle based on stretching factor and fraction of the
+        data. Finally, we propose a method to estimate observer's
+        off-axis angle based on the stretching factor and the fraction of the
         high-energy photons. The code for modeling is open source and
@@
 
-%TODO: add grants
-%? ref to Lemoine GRB1
-
-\section{Introduction}
+\section{Introduction and summary}
 
@@
 of magnitude~\cite{Bloom:2003wy}. This indicates that the most of the
-bursts have similar energetics in their rest frames. Several
+bursts have similar energetics in their rest frames. Moreover, the
+high energy radiation of GRBs is shown to follow the predictions of
+the synchrotron radiation scenario~\cite{Wang:2013ptaw}.  Several
 observations are related to the temporal variations of spectra. This
@@
 be extended relative to the x-ray emission
-\cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}.
-
-%% change the next statement
-	We decided to elaborate on that result, and, using the data
-        from Large Area Telescope (LAT) of Fermi satellite~\cite{Ackermann:2012kna}, explore the spectral variations between different high energy bands, specifically $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
-	In particular, we use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
-
-	We take the GRB list from the Fermi-LAT catalog \cite{Ackermann:2013zfa}.
-	Tables 2 and 4 in \cite{Ackermann:2013zfa} provide the time, durations and locations of the bursts, which we use to download observational data from the LAT Data Server\footnote{\url{http://fermi.gsfc.nasa.gov/cgi-bin/ssc/LAT/LATDataQuery.cgi}}.
-	We also download events during a day before the burst, and use the technique introduced in \cite{Rubtsov:2011qq} to estimate background radiation in both energy bands.
-
-	Most of the bursts in the catalog, however, do not have enough events with over $\unit[1]{GeV}$ energies to do desired computations, so we choose only those bursts, from which at least $10$ photons were detected in the high energy band.
-	It leaves $4$ of them: GRB 080916C \cite{Tajima:2009az}, GRB 090510 \cite{Ackermann:2010us}, GRB 090902B \cite{Abdo:2009pg} and GRB 090926A \cite{Bregeon:2011bu}.
-
-	For these 4 bursts we compute both high and low energy distributions of photon arrival times, and subtract background estimates from them.
-	This makes CDFs non-monotonous and, rigorously speaking, we cannot use the 2-sample Kolmogorov-Smirnov test on it.
-	However, since the number of photons is much higher than estimated background, this non-monotonicity is negligible, and will not harm KS-test results much.
-
-	Finally, we stretch the high energy CDF by different factors, and compare it to the low energy CDF using the KS-test.
-	If we require $2\sigma$-significant probability to exclude a particular stretching factor, then the stretching factor of $1$ (which means no stretching) is allowed for GRB 080916C and GRB 090510.
-	For the other two bursts the stretching factor of 1 is, however, excluded.
-	For GRB 090902B it should be smaller than 1, so the low energy light curve is stretched with respect to the high energy one (see fig. \ref{fig:grb090902B}).
-	And for GRB 090926A even more significant deviation from the stretching factor of 1 is observed in the opposite direction.
-	The high energy light curve of GRB 090926A is stretched with respect to the low energy one by a factor of at least 1.99 (see fig. \ref{fig:grb090926A}).
-	All observed stretching factors are summarized in table \ref{tab:observationResults}, and the detailed procedure for computing them is described in section \ref{sec:observations} of the paper.
-
-	We propose that this result can be explained by the curvature effects (that is the effects of the jet geometry).
-	These effects were explored by multiple authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.
-	However, these studies were only concerned with x-ray radiation, and, even more importantly, they assumed that the distribution of radiation sources is homogeneous throughout the jet.
-	We propose a contrary idea: that the highest energy radiators are concentrated near the axis of the jet, so that the jet opening angle depends on energy.
-	We cannot prove this assumption rigorously, but we have arguments supporting it.
-
-	First of all, there are around 750 GRBs detected by GBM, half of which were in the LAT field of view at the moment of observation \cite{Vianello:2013ela}.
-	However, only about 30 of them were detected by the LAT, and only 4 of them were bright in the high energy band.
-	If we extrapolate the uniform jet model to very high energies, this observation would mean that these groups of bursts are internally different: some of them produce VHE radiation, while others do not.
-	These differences are hard to explain given that burst energetics are similar \cite{Bloom:2003wy}.
-	Nevertheless, these differences in burst counts can easily be explained by our model.
-	In our model, the opening angle of a jet is inversely proportional to the energy of photons it radiates.
-	Therefore the most common scenario is that the off-axis angle of an observer is smaller than the low energy jet opening angle, but much larger than the opening angle of a high energy jet.
-	Because of that, most of the bursts can only be seen at low energies.
-	The 4 bursts we study in this paper were seen, according to our model, from the lowest off-axis angles.
-
-	Second, consider the plasma right after its ejection from the central engine.
-	Two processes happen there simultaneously:
+\cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}. In this paper
+we go further and compare the temporal extension of GRB gamma-ray
+radiation in the two energy bands $\unit[100]{MeV} < E <
+\unit[1]{GeV}$ and $E > \unit[1]{GeV}$ (hereafter, low and high-energy
+bands) using the data from the Large Area Telescope (LAT) of the Fermi
+satellite~\cite{Ackermann:2012kna}.
+
+We start with the location, time and duration of bursts from the
+Fermi-LAT Gamma-Ray Burst Catalog \cite{Ackermann:2013zfa}. The
+photons during the burst and the photons for 24 hours before the burst
+are downloaded from the LAT Data
+Server~\footnote{\url{http://fermi.gsfc.nasa.gov/cgi-bin/ssc/LAT/LATDataQuery.cgi}}.
+The latter are used for the background estimation in both energy bands
+with the technique introduced in \cite{Rubtsov:2011qq}. In order to
+have enough statistics we require that at least 10 photons are
+detected with the energy greater than $\unit[1]{GeV}$. This leaves are
+with the four bursts for the time span of the Fermi-LAT Gamma-Ray
+Burst Catalog. Namely, GRB~080916C \cite{Tajima:2009az}, GRB~090510
+\cite{Ackermann:2010us}, GRB~090902B \cite{Abdo:2009pg} and
+GRB~090926A \cite{Bregeon:2011bu}. For these bursts we stretch the
+high energy light curve by the arbitrary factor and compare with the
+low energy light curve using the Kolmogorov-Smirnov (KS) test. Within
+the $95\%$ confidence level the stretching factor is compatible to $1$
+(no stretching) for the GRB 080916C and GRB 090510. For the other two
+bursts the stretching factor of 1 is, however, excluded.  For GRB
+090902B it should be smaller than 1, so the low energy light curve is
+stretched with respect to the high energy one, see
+fig. \ref{fig:grb090902B}.  Even more significant deviation from the
+stretching factor of 1 in the opposite direction is observed for the
+GRB 090926A. The high energy light curve of GRB 090926A is stretched
+with respect to the low energy one by a factor of at least 1.99 (see
+fig. \ref{fig:grb090926A}).  The stretching of the GRB 090926A high
+energy light is found with $3.3\sigma$ significance.  All the observed
+stretching factors are summarized in table
+\ref{tab:observationResults}, and the detailed procedure for computing
+them is described in section \ref{sec:observations}.
+
+We propose that the observed energy dependence of the time profiles
+may be explained without introduction of the new effects. We argue
+that the stretching may originate from the effects of the jet geometry
+called curvature effects. These effects were explored by multiple
+authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.
+%% TODO here 
+ However,
+these studies concerned x-ray radiation, and, even more
+importantly, they assumed that the distribution of radiation sources
+is homogeneous throughout the jet.  We propose a contrary idea: that
+the highest energy radiators are concentrated near the axis of the
+jet, so that the jet opening angle depends on energy.  We cannot prove
+this assumption rigorously, but we have arguments supporting it.
+
+First of all, there are around 750 GRBs detected by GBM, half of which
+were in the LAT field of view at the moment of observation
+\cite{Vianello:2013ela}.  However, only about 30 of them were detected
+by the LAT, and only 4 of them were bright in the high energy band.
+If we extrapolate the uniform jet model to very high energies, this
+observation would mean that these groups of bursts are internally
+different: some of them produce VHE radiation, while others do not.
+These differences are hard to explain given that burst energetics are
+similar \cite{Bloom:2003wy}.  Nevertheless, these differences in burst
+counts can easily be explained by our model.  In our model, the
+opening angle of a jet is inversely proportional to the energy of
+photons it radiates.  Therefore the most common scenario is that the
+off-axis angle of an observer is smaller than the low energy jet
+opening angle, but much larger than the opening angle of a high energy
+jet.  Because of that, most of the bursts can only be seen at low
+energies.  The 4 bursts we study in this paper were seen, according to
+our model, from the lowest off-axis angles.
+
+Second, consider the plasma right after its ejection from the central
+engine.  Two processes happen there simultaneously:
 \begin{enumerate}
@@
 \end{enumerate}
-	We argue that these two processes are correlated, for they happen due to the same particle interaction processes.
-	And since they are correlated, jets with larger opening angles should have lower energies, which is the assumption we are trying to justify.
+We argue that these two processes are correlated, for they happen due
+to the same particle interaction processes.  And since they are
+correlated, jets with larger opening angles should have lower
+energies, which is the assumption we are trying to justify.
 
@@
 
-	Finally, our assumption is able to explain the non-trivial stretching factor of both GRB 090902B and GRB 090926A.
-	Detailed computations are described in section \ref{sec:model}, but you can see the main idea on figure \ref{fig:modelOverview}.
-	Computations confirm that this qualitative picture is indeed correct.
-	You can see a sample light curves on figure \ref{fig:sampleLightCurves}, and predicted distribution of stretching factors on figure \ref{fig:kappaDistribution}.
-
-	Results of our study, including a few predictions of our model and a new method for estimating jet observation angles, are summarized in section \ref{sec:results}, and the caveats are discussed in section \ref{sec:summary}.
+Finally, our assumption is able to explain the non-trivial stretching
+factor of both GRB 090902B and GRB 090926A.  Detailed computations are
+described in section \ref{sec:model}, but you can see the main idea on
+figure \ref{fig:modelOverview}.  Computations confirm that this
+qualitative picture is indeed correct.  You can see a sample light
+curves on figure \ref{fig:sampleLightCurves}, and predicted
+distribution of stretching factors on figure
+\ref{fig:kappaDistribution}.
+
+Results of our study, including a few predictions of our model and a
+new method for estimating jet observation angles, are summarized in
+section \ref{sec:results}, and the caveats are discussed in section
+\ref{sec:conclusions}.
 
@@
 
-Before we start to compute the Kolmogorov-Smirnov probabilities, we need to download the actual observational data from the Fermi LAT data server.
-We take the data about the bursts from the catalog \cite{Ackermann:2013zfa}, which contains all the bright bursts seen by the LAT since 2008 August 4 to 2011 August 1.
-Specifically, we take 4 pieces of information from the tables 2 and 4 there:
+%% TODO: change, add Pass 7REP (V15), ``SOURCE'' class
+Before we start to compute the Kolmogorov-Smirnov probabilities, we
+need to download the actual observational data from the Fermi LAT data
+server.  We take the data about the bursts from the catalog
+\cite{Ackermann:2013zfa}, which contains all the bright bursts seen by
+the LAT since 2008 August 4 to 2011 August 1.  Specifically, we take 4
+pieces of information from the tables 2 and 4 there:
 \begin{itemize}
@@
 That is why we had to filter them out completely.
+%% TODO: rephrase positively, Method works for SOURCE-class photons,
+%% but ..
 
@@
 
-\section{Summary}
-
-\label{sec:summary}
+\section{Conclusions}
+
+\label{sec:conclusions}
+
+%% TODO: add more descriptive words
 
@@
 
-	It seems, however, that all these problems are not unresolvable, and may be addressed by refining the model without changing its main assumptions.
-
-	{\small {\it Acknowledgments.} The work was supported by Russian Science Foundation grant 14-12-01340.}
+It seems, however, that all these problems are not unresolvable, and
+may be addressed by refining the model without changing its main
+assumptions.
+%% TODO: remove unresolvable, rephrase
+
+{\bf Acknowledgments.} We thank M.~Pshirkov and S.~Troitsky for
+comments and inspiring discussions.  The work was supported by Russian
+Science Foundation grant 14-12-01340. The analysis is based on data
+and software provided by the Fermi Science Support Center (FSSC). The
+numerical part of the work was done at the cluster of the Theoretical
+Division of INR RAS. G.R. acknowledges the fellowship of the Dynasty
+foundation.
 
@@
 \section{Cosmology}
+%% TODO: change title
 \label{sec:cosmology}
@@
 
-We also need to know the areas of two spheres: the photon sphere, a sphere over which photons emitted in a particular burst are spread; and the bursts sphere, a sphere over which bursts at a particular redshift are distributed.
-These two spheres have the same radii -- a distance from the observer to the burst, but different centers: the photon sphere is centered on the burst, while the bursts sphere is centered on the observer.
-Also, they have different areas, because the scale factor differs at locations of burst and observer.
-
-Lets begin with the photon sphere.
+We also need to know the areas of two spheres: the photon sphere, a
+sphere over which photons emitted in a particular burst are spread;
+and the bursts sphere, a sphere over which bursts at a particular
+redshift are distributed. These two spheres have the same radii -- a
+distance from the observer to the burst, but different centers: the
+photon sphere is centered on the burst, while the bursts sphere is
+centered on the observer. Also, they have different areas, because the
+scale factor differs at locations of burst and observer.
+%% TODO rephrase, mention back in time
+
+Let us begin with the photon sphere.
 This sphere has the origin at $r = 0$ and $t = 0$, at the central engine of a particular burst.
@@
 
-The second sphere is the bursts sphere. It has its origin at the observer's position, and the radius the same as the photon sphere. Therefore, its area is:
+The second sphere is the bursts sphere. It has its origin at the
+observer's position, and the radius the same as the photon
+sphere. Therefore, its area is:
+%% TODO: mention that this is the sphere in the past
 \begin{equation}

### 2014-12-12 Paper/gammaRays.tex
revised section 2

@@
 	Maxim S. Piskunov$^{1,2}$\thanks{\href{mailto:maxit@ms2.inr.ac.ru}{maxit@ms2.inr.ac.ru}} ~and
-	Grigory I. Rubtsov$^{2,3}$\thanks{\href{mailto:grisha@ms2.inr.ac.ru}{grisha@ms2.inr.ac.ru}} \\
+	Grigory I. Rubtsov$^{2}$\thanks{\href{mailto:grisha@ms2.inr.ac.ru}{grisha@ms2.inr.ac.ru}} \\
 	{\small $^{1}$Faculty of Physics, Moscow State University, Moscow, Russia}\\
-	{\small $^{2}$Institute for Nuclear Research of the RAS, Moscow, Russia}\\
-	{\small $^{3}$Novosibirsk State University, Novosibirsk, Russia}
+	{\small $^{2}$Institute for Nuclear Research of the RAS, Moscow, Russia}
 }
@@
 \begin{document}
-
-\maketitle
+\begin{flushright}  
+INR-TH/2014-034
+\end{flushright}
+\vskip -0.9cm
+{\let\newpage\relax \maketitle}
 
@@
 bands) using the data from the Large Area Telescope (LAT) of the Fermi
-satellite~\cite{Ackermann:2012kna}.
+satellite~\cite{2009ApJ...697.1071A,Ackermann:2012kna}.
 
@@
 energy light is found with $3.3\sigma$ significance.  All the observed
-stretching factors are summarized in table
-\ref{tab:observationResults}, and the detailed procedure for computing
-them is described in section \ref{sec:observations}.
+stretching factors are summarized in Table
+\ref{tab:observationResults}, and the detailed procedure for their
+calculation is described in Section \ref{sec:observations}.
 
@@
 called curvature effects. These effects were explored by multiple
-authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.
-%% TODO here 
- However,
-these studies concerned x-ray radiation, and, even more
-importantly, they assumed that the distribution of radiation sources
-is homogeneous throughout the jet.  We propose a contrary idea: that
-the highest energy radiators are concentrated near the axis of the
-jet, so that the jet opening angle depends on energy.  We cannot prove
-this assumption rigorously, but we have arguments supporting it.
-
-First of all, there are around 750 GRBs detected by GBM, half of which
-were in the LAT field of view at the moment of observation
-\cite{Vianello:2013ela}.  However, only about 30 of them were detected
-by the LAT, and only 4 of them were bright in the high energy band.
-If we extrapolate the uniform jet model to very high energies, this
-observation would mean that these groups of bursts are internally
-different: some of them produce VHE radiation, while others do not.
-These differences are hard to explain given that burst energetics are
-similar \cite{Bloom:2003wy}.  Nevertheless, these differences in burst
-counts can easily be explained by our model.  In our model, the
-opening angle of a jet is inversely proportional to the energy of
-photons it radiates.  Therefore the most common scenario is that the
-off-axis angle of an observer is smaller than the low energy jet
-opening angle, but much larger than the opening angle of a high energy
-jet.  Because of that, most of the bursts can only be seen at low
-energies.  The 4 bursts we study in this paper were seen, according to
-our model, from the lowest off-axis angles.
-
-Second, consider the plasma right after its ejection from the central
-engine.  Two processes happen there simultaneously:
-\begin{enumerate}
-	\item{Particles near the jet boundary collide and change their movement directions, therefore increasing the opening angle of the jet.}
-	\item{Particles lose energy, therefore decreasing radiation frequency emitted by the jet.}
-\end{enumerate}
-We argue that these two processes are correlated, for they happen due
-to the same particle interaction processes.  And since they are
-correlated, jets with larger opening angles should have lower
-energies, which is the assumption we are trying to justify.
+authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.  These
+studies concerned x-ray radiation, and considered the homogeneous
+distribution of radiation sources throughout the jet. Our model is
+based on the reverse assumption that the highest energy radiators are
+concentrated near the axis of the jet and consequently the jet opening
+angle depends on energy. The latter is motivated by observations if
+all bursts are considered equivalent. In this case the energy
+dependence of the jet opening angle resolves the contradiction between
+the small fraction of GRBs visible above GeV and the hard
+spectrum.
 
@@
         \includegraphics[width=0.7\textwidth]{modelOverview}
-        \caption{
-        	Model Overview.
-        	Here red and blue cones represent the regions through which low and high energy plasma propagates.
-        	In case depicted the observer's off-axis angle is smaller than the opening angle of the low energy jet, so, due to the relativistic beaming effect, the most of the observable low energy photons will travel along the straight line from the central engine.
-        	Also, the observer's off-axis angle is larger that the opening angle of the high energy jet, so the high energy radiation will still originate near the center of the jet (because it is the only place where there are high energy radiators).
-        	The observation time of a photon is a sum of two things: the time interval spent in plasma as a radiator (which approximately equals to the distance from the central engine to the point of emission); and the time interval from emission to detection (which is the distance from the point of emission to the observer's location).
-        	Given a position of the shock front, this sum is larger for high energy photons.
-        	Because of that, high energy emission will be observed later throughout the burst duration, therefore the high energy light curve will be stretched.
-        }
+        \caption{ Model Overview.  The red and blue cones represent
+          the regions through which low and high energy charged
+          particles propagate.  In the case shown the observer's
+          off-axis angle is smaller than the opening angle of the low
+          energy jet, so, due to the relativistic beaming effect, the
+          most of the observable low energy photons will travel along
+          the straight line from the central engine.  Also, the
+          observer's off-axis angle is larger that the opening angle
+          of the high energy jet, but the high energy radiation will
+          still originate near the center of the jet as this is the
+          only place where the high energy radiators are located.  The
+          observation time of a photon is a sum of two terms. First, the
+          time interval spent in plasma as a radiator which
+          approximately equals to the distance from the central engine
+          to the point of emission. Second, the time interval from
+          emission to detection which equals to the distance from the point
+          of emission to the observer's location.  For the given position
+          of the observer the sum is larger for high energy
+          photons.  Therefore, the high energy emission will be
+          observed later throughout the burst duration and the
+          high energy light curve will be stretched.  }
         \label{fig:modelOverview}
@@
 
-Finally, our assumption is able to explain the non-trivial stretching
-factor of both GRB 090902B and GRB 090926A.  Detailed computations are
-described in section \ref{sec:model}, but you can see the main idea on
-figure \ref{fig:modelOverview}.  Computations confirm that this
-qualitative picture is indeed correct.  You can see a sample light
-curves on figure \ref{fig:sampleLightCurves}, and predicted
-distribution of stretching factors on figure
-\ref{fig:kappaDistribution}.
-
-Results of our study, including a few predictions of our model and a
-new method for estimating jet observation angles, are summarized in
-section \ref{sec:results}, and the caveats are discussed in section
-\ref{sec:conclusions}.
-
-\section{Observations}
+The basic idea of the model is illustrated in Figure
+\ref{fig:modelOverview}. The model explains both the stretching
+factors lower and higher than 1 with the difference of the position of
+the observer with respect to the burst. The sample light curves are
+shown in Figure \ref{fig:sampleLightCurves}. The detailed calculations
+show that the stretching factors of GRB 090902B and GRB 090926A appear
+naturally in the model. The predicted distribution of stretching
+factors is shown in Figure \ref{fig:kappaDistribution}.
+
+The rest of the paper is organized as follows. We describe the Fermi
+LAT data and the analysis procedures in Section
+\ref{sec:observations}. The model and the details on calculations are
+explained in Section \ref{sec:model}. The Section \label{sec:tests}
+introduces three phenomenological tests for the model. The parameters
+of the model and the new method for estimating the jet observation
+angles are shown in Section \ref{sec:results}. The code for modeling
+is open source and is publicly available on GitHub
+(\url{https://github.com/maxitg/GammaRays}).
+
+\section{Data analysis}
 \label{sec:observations}
@@
 	\end{tabular}
-	\caption{Burst data used in our study.
-The third row contains the corresponding \texttt{GRBurst} class variable names.
-The data is taken from the tables 2 and 4 of \cite{Ackermann:2013zfa}.
-\url{https://github.com/maxitg/GammaRays/blob/master/GRObservations/bursts}}
+	\caption{Burst data used in our study taken from~\cite{Ackermann:2013zfa}. The third row contains
+          the corresponding \texttt{GRBurst} class variable names of
+          the software package \url{https://github.com/maxitg/GammaRays/blob/master/GRObservations/bursts}}
 	\label{tab:bursts}
@@
 
-%% TODO: change, add Pass 7REP (V15), ``SOURCE'' class
-Before we start to compute the Kolmogorov-Smirnov probabilities, we
-need to download the actual observational data from the Fermi LAT data
-server.  We take the data about the bursts from the catalog
-\cite{Ackermann:2013zfa}, which contains all the bright bursts seen by
-the LAT since 2008 August 4 to 2011 August 1.  Specifically, we take 4
-pieces of information from the tables 2 and 4 there:
+First, we download the Pass 7REP (V15) ``SOURCE'' class photons from
+the Fermi-LAT data
+server~\cite{2009ApJ...697.1071A,Ackermann:2012kna}. In order to
+define the regions of interest we use the catalog
+\cite{Ackermann:2013zfa}, which contains all the bright gamma-ray
+bursts seen by the Fermi LAT since 2008 August 4 to 2011 August 1.
+Specifically, we take 4 pieces of information from the tables 2 and 4
+of Ref. \cite{Ackermann:2013zfa}:
 \begin{itemize}
 	\item{
-		GBM trigger time (table 2).
+		GBM trigger time.
 		It is used as a reference point for the GRB time.
@@
 	\item{
-		Location of the GRB (table 2).
-		We need burst locations to filter out photons coming from other sources.
-	}
-	\item{
-		Location error (table 2).
-		Location errors are used to improve accuracy of filtering, specifically to avoid losing statistics by filtering too much.
-	}
-	\item{
-		$T_{05}$ -- $T_{95}$ interval of the LAT-detected emission (table 4).
-		We need precise intervals for two things.
-		First, to understand where is the fixed point of the stretching ($T_{05}$ is used as a proxy for that).
-		Second, the burst interval is used to download events data from the data server.
+		Location of the GRB.
+		The burst locations are used to filter out photons
+                coming from the other sources.
+	}
+	\item{
+		Location error.
+		Location errors are used to improve the accuracy of filtering, specifically to avoid losing statistics by filtering too much.
+	}
+	\item{
+		$T_{05}$ -- $T_{95}$ interval of the LAT-detected
+          emission. The beginning of the interval is used as a fixed
+          point of stretching. The Fermi LAT photons are downloaded
+          for the period extended in time by $50\%$ towards both past
+          and future compared to this interval.
 	}
 \end{itemize}
-The data obtained is summarized in table \ref{tab:bursts}.
-
-To have a decent safety margin, we download the data about observed photons extended in time by $50\%$ to both past and future relative to the table \ref{tab:bursts}.
-For background estimation, we also download observational data during a day before the burst.
-Exact values we put into the web form are listed in appendix \ref{sec:fermiCode}.
-
-Before reading the data, we need to apply the basic filtering as described in Fermi guidelines\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/data_preparation.html}} (the actual code, and all parameters values we use to invoke Fermi Science tools are listed in appendix \ref{sec:fermiCode}).
-Two tools from the Fermi Science Tools package\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/overview.html}} are used for that task (the same filtering is applied to both burst and background data).
-
-First, we use \texttt{gtselect}.
-We filter the Earth limb emission with it below the zenith angle of $100\degree$.
-We also filter out photons with transient event class for a reason described in the next section.
-
-The second one we use is \texttt{gtmktime}.
-It uses the spacecraft file as well as events file, and filters out time periods when Fermi LAT was not operational.
-Running \texttt{gtmktime} is also required for exposure maps calculation, which we'll need later on.
-
-Now, with basic filtering done, we can read events data from \texttt{timed.fits}. This file, however, contains observations of a large area of the sky.
-Before using the data obtained from it, we should perform a more elaborate filtering by location using the point spread function of Fermi LAT, which can be computed with \texttt{gtpsf} tool.
-We use 300 energy bins from \unit[100]{MeV} to \unit[300]{GeV}, and 300 angular bins with maximal angle being $30\degree$.
-We use \texttt{gtltcube} tool to compute livetime maps, which are \texttt{gtpsf} need.
-Note, that we need PSFs for all Instrument Response Functions of Fermi, that is for all event classes (transient, source, clean and ultraclean) and conversion types (back and front).
-
-To understand how to use obtained PSFs, imagine a stream of photons coming from some source located at some coordinates $\theta, \phi$.
-Ideally, all the coordinates measured for the photons coming from the source should be the same.
-However, due to measurement errors, they would be spread across some region of the sky.
-If we find the region such that, say, $95\%$ of photons coming from the source are measured inside of it, we can filter out all the other photons without losing too much statistics, and significantly reducing background.
-\texttt{gtpsf} approximates these regions as circles, and the files generated in the previous step contain the observation probability densities as functions of distance from the source and photon energies.
-We, however, need an integral quantity, that is given the radius, we want to know the fraction of photons observed inside the circle of this radius.
-We can compute this fraction by summing over the probability densities:
-\begin{align*}
-	p\left(n\right) &= \int_0^n \texttt{pdf[i]} 2\pi \sin\left(\texttt{angles[i]}\right) \dif i \\
-	&\approx \sum_{i=0}^{n} \frac{\texttt{pdf[i]} + \texttt{pdf[i+1]}}{2} 2\pi \left(\cos\left(\texttt{angles[i]}\right) - \cos\left(\texttt{angles[i+1]}\right)\right)
-\end{align*}
-Here \texttt{pdf[i]} is the probability density in the point separated from the source by an angle \texttt{angles[i]}.
+The data are summarized in table \ref{tab:bursts}. For the background
+estimation, we also download observational data for 24 hours before
+the burst.  The analysis is performed with the Fermi Science Tools
+v9r32p5
+package~\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/overview.html}}
+following the
+guidelines~\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/data_preparation.html}}.
+We require Earth zenith angle to be below $100^\circ$ to remove the
+Earth limb emission. After the basic filtering by \texttt{gtselect}
+and \texttt{gtmktime} tools is done we perform more elaborate
+selection using the point spread function of Fermi LAT, calculated
+with \texttt{gtpsf} tool. We use 300 energy bins from \unit[100]{MeV}
+to \unit[300]{GeV}, and 300 angular bins with maximal angle being
+$30\degree$.  The \texttt{gtltcube} tool is used to compute required
+livetime maps. Note, that we calculate PSFs separately for both
+conversion types: front and back. We define $\mathrm{PSF}_{95}$ as an
+angular distance from the source containing 95\% of the emitted
+photons. We require that the photon is closer than $\mathrm{PSF}_{95}
++ \texttt{location.err}$ to the location of the GRB (see Table
+\ref{tab:bursts}). The details of the data request and
+analysis are given in appendix \ref{sec:fermiCode}.
+
+%To understand how to use obtained PSFs, imagine a stream of photons
+%coming from some source located at some coordinates $\theta,
+%\phi$. Ideally, all the coordinates measured for the photons coming
+%from the source should be the same. However, due to measurement
+%errors, they would be spread across some region of the sky. If we find
+%the region such that, say, $95\%$ of photons coming from the source
+%are measured inside of it, we can filter out all the other photons
+%without losing too much statistics, and significantly reducing
+%background. \texttt{gtpsf} approximates these regions as circles, and
+%the files generated in the previous step contain the observation
+%probability densities as functions of distance from the source and
+%photon energies. We, however, need an integral quantity, that is given
+%the radius, we want to know the fraction of photons observed inside
+%the circle of this radius. We can compute this fraction by summing
+%over the probability densities:
+%\begin{align*}
+%	p\left(n\right) &= \int_0^n \texttt{pdf[i]} 2\pi \sin\left(\texttt{angles[i]}\right) \dif i \\
+%	&\approx \sum_{i=0}^{n} \frac{\texttt{pdf[i]} + \texttt{pdf[i+1]}}{2} 2\pi \left(\cos\left(\texttt{angles[i]}\right) - \cos\left(\texttt{angles[i+1]}\right)\right)
+%\end{align*}
+%Here \texttt{pdf[i]} is the probability density in the point separated from the source by an angle \texttt{angles[i]}.
 And $p\left(n\right)$ is the probability to observe a photon closer to the source than \texttt{angles[n]}.
@@
 
-Finally, to finish the filtering, we filter out all the photons, for which their location is separated from the source by more that $\theta_{\text{spread}}\left(0.95\right) + \texttt{location.err}$, where \texttt{location.err} is the uncertainty in the burst location (see table \ref{tab:bursts}).
-
-Now we have a decent set of photons, which we can use for analysis.
-However, some of the photons in this set are still originated from background, and not from the source.
-We cannot filter out these photons, but we can substitute a linear component of background estimating it from a long time period before the burst.
-In order to perform this computation, we need an exposure map, which will be discussed in the following section.
+The set will be used for the analysis with the reservation that it
+unavoidably contains some background photons. We consider the
+background flux constant in time and estimate it using the data for 24
+hours before the burst. In order to calculate the number of background
+photons we need an exposure map, which will be discussed in the
+following section.
 
@@
 
-The exposure map can be computed with the \texttt{gtexpcube2} tool. We compute it with 300 energy bins from \unit[100]{MeV} to \unit[300]{GeV}.
-
-The generated \texttt{expcube} files contain exposures as functions of energy and location on the sky. We use trilinear interpolation to compute exposures for all energies and locations.
-
-Knowing the exposures, we can now use the method introduced in the appendix of \cite{Rubtsov:2011qq} to estimate the linear component of background from the data observed before the burst.
-We compute two background estimates: for low and high energy bands.
-Finally, our method doesn't work for transient class photons, because their background appears to depend on the spacecraft position.
-That is why we had to filter them out completely.
-%% TODO: rephrase positively, Method works for SOURCE-class photons,
-%% but ..
+The exposure map is computed with the \texttt{gtexpcube2} tool. We
+compute it with 300 energy bins from \unit[100]{MeV} to
+\unit[300]{GeV}. The produced \texttt{expcube} files contain exposures
+as functions of energy and location on the sky. We use trilinear
+interpolation to compute exposures for all energies and
+locations. Knowing the exposures, we use the method introduced in the
+appendix of \cite{Rubtsov:2011qq} to estimate the background for low
+and high-energy bands. Note, that the ``TRANSIENT'' class photons are
+not included in the analysis due to the strong dependence of the
+background on the spacecraft position.
 
@@
 
-We take care of remaining background by using the following functions with the Kolmogorov-Smirnov test:
+Before applying the KS-test we subtract the background, which is
+considered linear for the duration of the burst :
+
 \begin{equation}
@@
 \end{equation}
-Here $\left(T_1, T_2\right)$ is the time range of observations, $p_i\left(t_1, t_2\right)$ is the number of photons observed in the time range $t_1$ to $t_2$ in the $i$'s energy band ($i$ is either low or high for the low and high energy bands respectfully), and $b_i$ is the estimated number of background photons for the time range $T_1$ to $T_2$ in the $i$'s energy band.
-
-The numbers of degrees of freedom (used as an input for the KS-test) are $p_i\left(T_1, T_2\right) - b_i$.
-
-While we did not prove it, we assume that the Kolmogorov-Smirnov statistics computed for these functions with provided number of degrees of freedom has a similar distribution to that of the KS-statistics for ordinary CDFs (that is CDFs with no background).
-
-Finally, we compute KS-probabilities for pairs $\Phi_\text{low}\left(t\right), \Phi_\text{high}\left(\kappa t\right)$ to obtain the allowed ranges for stretching factors $\kappa$.
-We evaluate stretching factors in the range $\texttt{STRETCHING\_MIN} = 0.1$ to $\texttt{STRETCHING\_MAX} = 10.$ with a logarithmic step of $\texttt{STRETCHING\_STEP} = 1.001$.
-
-As a result of such calculation we can constrain the range of allowed stretching factors to values for which the KS-probabilities are not too small, in other words, given the sigma-significance value, compute the range outside of which stretching factors are excluded by observations.
-
-You can see the results of such computation in the following section.
+Here $\left(T_1, T_2\right)$ is the time range of observations,
+$p_i\left(t_1, t_2\right)$ is the number of photons observed in the
+time range $t_1$ to $t_2$ in the $i$'s energy band ($i$ is either low
+or high for the low and high-energy bands respectively), and $b_i$ is
+the estimated number of background photons for the time range $T_1$ to
+$T_2$ in the $i$'s energy band.
+
+The number of degrees of freedom used as an input for the KS-test is
+$p_i\left(T_1, T_2\right) - b_i$. This is lower than the number of
+independent photons and therefore we are conservative when estimating
+the probability and significance of the effect. The KS-probabilities
+are computed for pairs $\Phi_\text{low}\left(t\right),
+\Phi_\text{high}\left(\kappa t\right)$ to obtain the allowed ranges
+for stretching factors $\kappa$ for a given confidence level.
+% We evaluate stretching factors in the
+%range $\texttt{STRETCHING\_MIN} = 0.1$ to $\texttt{STRETCHING\_MAX} =
+%10.$ with a logarithmic step of $\texttt{STRETCHING\_STEP} = 1.001$.
 
@@
 
-Out of 19 bursts studied, only 4 have at least 10 high energy events remaining after filtering, and thus eligible for computating the stretching factors.
-The results of this computation are shown on figures \ref{fig:grb080916C}, \ref{fig:grb090510}, \ref{fig:grb090902B}, \ref{fig:grb090926A} and table \ref{tab:observationResults}.
-
-Out of these 4 bursts, two (GRB 080916C and GRB 090510) have stretching factors compatible with $\kappa = 1$ within $2\sigma$ range.
-
-GRB 090926A has, however, high energy radiation stretched with respect to low energy radiation (that is $\kappa > 1$).
-In contrast to that, GRB 090902B has low energy radiation stretched with respect to high energy radiation ($\kappa < 1$).
-
-So we obtained a preliminary (for significance is only $2\sigma$) result that the stretching factors for observable bursts might be both larger and smaller than $1$.
+Out of 19 bursts studied, only 4 have at least 10 high energy events
+remaining after the filtering, and thus eligible for computating the
+stretching factors. The results of this computation are shown in
+Figures \ref{fig:grb080916C}, \ref{fig:grb090510},
+\ref{fig:grb090902B}, \ref{fig:grb090926A} and table
+\ref{tab:observationResults}.
+
+Out of these 4 bursts, two (GRB 080916C and GRB 090510) have
+stretching factors compatible with $\kappa = 1$ within $2\sigma$
+range.
+
+GRB 090926A has, however, high energy radiation stretched with respect
+to low energy radiation (that is $\kappa > 1$). In contrast to that,
+GRB 090902B has low energy radiation stretched with respect to high
+energy radiation ($\kappa < 1$).
+
+Therefore we obtained an indication that the stretching factors for
+observable bursts may take values both larger and smaller than $1$.
 
@@
         \end{subfigure}
-        \caption{GRB 080916C results. Stretching factor is compatible with $\kappa = 1$ within $2\sigma$.}
+        \caption{GRB 080916C light curves. Stretching factor is compatible with $\kappa = 1$ within $2\sigma$.}
         \label{fig:grb080916C}
@@
         \end{subfigure}
-        \caption{GRB 090510 results. Stretching factor is compatible with $\kappa = 1$ within $1\sigma$.}
+        \caption{GRB 090510 light curves. Stretching factor is compatible with $\kappa = 1$ within $1\sigma$.}
         \label{fig:grb090510}
@@
         \end{subfigure}
-        \caption{GRB 090902B results. Low energy radiation is stretched ($\kappa < 1$) with significance of $2.3\sigma$.}
+        \caption{GRB 090902B light curves. Low energy radiation is stretched ($\kappa < 1$) with significance of $2.3\sigma$.}
         \label{fig:grb090902B}
@@
         \end{subfigure}
-        \caption{GRB 090926A results. High energy radiation is stretched ($\kappa > 1$) with significance of $3.3\sigma$.}
+        \caption{GRB 090926A light curves. High energy radiation is stretched ($\kappa > 1$) with significance of $3.3\sigma$.}
         \label{fig:grb090926A}
@@
 	\end{tabular}
-	\caption{Burst observation results. Ranges of allowed stretching factors are shown for studied bursts for multiple levels of significance.}
+	\caption{The ranges of allowed stretching factors for multiple
+          levels of significance for the GRBs studied.}
 	\label{tab:observationResults}
@@
 
-\subsection{Description}
-
 The main idea behind our model is to assume that the burst opening angle depends on energy of emitted photons, or, equivalently, that the most energetic plasma particles are concentrated near the axis of the jet, while low energy particles are on the periphery.
-As you can see on figure \ref{fig:modelOverview} this leads to a time stretching effect in some cases due to relativistic beaming.
+As you can see in Figure \ref{fig:modelOverview} this leads to a time stretching effect in some cases due to relativistic beaming.
 
@@
 First of all, the number of observed GRBs rapidly decreases with energy.
-There are around 750 bursts detected by GBM in hard X-rays and low energy gamma rays \cite{Vianello:2013ela} (half of them were in a LAT field of view), only 30 or so were detected by LAT above \unit[100]{MeV}, and only about 4 of them were detected above GeV.
+There are around 750 bursts detected by GBM in hard X-rays and low
+energy gamma rays \cite{Vianello:2013ela} (half of them were in a LAT
+field of view), only 30 or so were detected by LAT above
+\unit[100]{MeV}, and only about 4 of them were detected above GeV.
+
+% TODO merge 2
+First of all, the number of observed GRBs rapidly decreases with energy.
+There are around 750 bursts detected by GBM in hard X-rays and low
+energy gamma rays \cite{Vianello:2013ela} (half of them were in a LAT
+field of view), only 30 or so were detected by LAT above
+\unit[100]{MeV}, and only about 4 of them were detected above GeV.
+
+First of all, there are around 750 GRBs detected by GBM, half of which
+were in the LAT field of view at the moment of observation
+\cite{Vianello:2013ela}.  However, only about 30 of them were detected
+by the LAT, and only 4 of them were bright in the high energy band.
+If we extrapolate the uniform jet model to very high energies, this
+observation would mean that these groups of bursts are internally
+different: some of them produce VHE radiation, while others do not.
+These differences are hard to explain given that burst energetics are
+similar \cite{Bloom:2003wy}.  Nevertheless, these differences in burst
+counts can easily be explained by our model.  In our model, the
+opening angle of a jet is inversely proportional to the energy of
+photons it radiates.  Therefore the most common scenario is that the
+off-axis angle of an observer is smaller than the low energy jet
+opening angle, but much larger than the opening angle of a high energy
+jet.  Because of that, most of the bursts can only be seen at low
+energies.  The 4 bursts we study in this paper were seen, according to
+our model, from the lowest off-axis angles.
+
 This observation is hard to explain in a framework of uniform jet model given that burst energetics are similar \cite{Bloom:2003wy}.
@@
 
-Another argument comes from consideration of processes happening in jets, specifically the scattering of particles near the jet boundary.
-While plasma particles scatter, they simultaneously lose energy, and change directions, sometimes propagating beyond the jet boundary, and therefore increasing the jet opening angle.
-So, the processes of energy loss and increase of jet size are correlated, therefore the low energy particles should be closer to the jet boundary.
+% merge 
+Another argument comes from consideration of processes happening in
+jets, specifically the scattering of particles near the jet boundary.
+While plasma particles scatter, they simultaneously lose energy, and
+change directions, sometimes propagating beyond the jet boundary, and
+therefore increasing the jet opening angle.  So, the processes of
+energy loss and increase of jet size are correlated, therefore the low
+energy particles should be closer to the jet boundary.
+
+Second, consider the plasma right after its ejection from the central
+engine.  Two processes happen there simultaneously:
+\begin{enumerate}
+	\item{Particles near the jet boundary collide and change their
+          movement directions, therefore increasing the opening angle
+          of the jet.}
+	\item{Particles lose energy, therefore decreasing radiation
+          frequency emitted by the jet.}
+\end{enumerate}
+We argue that these two processes are correlated, for they happen due
+to the same particle interaction processes.  And since they are
+correlated, jets with larger opening angles should have lower
+energies, which is the assumption we are trying to justify.
 
 In the following sections we will rigorously derive the procedure for computing time stretching factors from our model, and compare them to observations obtained in the previous section.
-
-\subsection{Assumptions}
 
@@
 
+\section{Observational tests}
+\label{sec:tests}
 \subsection{Total energy}
@@
 
-\section{Conclusions}
-
-\label{sec:conclusions}
+%\section{Conclusions}
+
+%\label{sec:conclusions}
+%% TODO: remove section
 

### 2014-12-12 Paper/gammaRays.tex
revised sections 3,4

@@
 
-%To understand how to use obtained PSFs, imagine a stream of photons
-%coming from some source located at some coordinates $\theta,
-%\phi$. Ideally, all the coordinates measured for the photons coming
-%from the source should be the same. However, due to measurement
-%errors, they would be spread across some region of the sky. If we find
-%the region such that, say, $95\%$ of photons coming from the source
-%are measured inside of it, we can filter out all the other photons
-%without losing too much statistics, and significantly reducing
-%background. \texttt{gtpsf} approximates these regions as circles, and
-%the files generated in the previous step contain the observation
-%probability densities as functions of distance from the source and
-%photon energies. We, however, need an integral quantity, that is given
-%the radius, we want to know the fraction of photons observed inside
-%the circle of this radius. We can compute this fraction by summing
-%over the probability densities:
-%\begin{align*}
-%	p\left(n\right) &= \int_0^n \texttt{pdf[i]} 2\pi \sin\left(\texttt{angles[i]}\right) \dif i \\
-%	&\approx \sum_{i=0}^{n} \frac{\texttt{pdf[i]} + \texttt{pdf[i+1]}}{2} 2\pi \left(\cos\left(\texttt{angles[i]}\right) - \cos\left(\texttt{angles[i+1]}\right)\right)
-%\end{align*}
-%Here \texttt{pdf[i]} is the probability density in the point separated from the source by an angle \texttt{angles[i]}.
-And $p\left(n\right)$ is the probability to observe a photon closer to the source than \texttt{angles[n]}.
-We find probabilities for intermediate angles by linear interpolation.
-Finally, we can compute the inverse value $\theta_{\text{spread}}\left(p\right)$ (that is a distance as a function of probability) by binary search.
-Note, that $\theta_{\text{spread}}\left(p\right)$ depends on the photon energy, event class, and conversion type.
-We should use the appropriate function for each photon.
-The spread angles for intermediate energies are linearly interpolated.
-
 The set will be used for the analysis with the reservation that it
@@
 for stretching factors $\kappa$ for a given confidence level.
-% We evaluate stretching factors in the
-%range $\texttt{STRETCHING\_MIN} = 0.1$ to $\texttt{STRETCHING\_MAX} =
-%10.$ with a logarithmic step of $\texttt{STRETCHING\_STEP} = 1.001$.
 
@@
 
-The main idea behind our model is to assume that the burst opening angle depends on energy of emitted photons, or, equivalently, that the most energetic plasma particles are concentrated near the axis of the jet, while low energy particles are on the periphery.
-As you can see in Figure \ref{fig:modelOverview} this leads to a time stretching effect in some cases due to relativistic beaming.
-
-At the moment, we do not have any rigorous proof of our assumption.
-There are arguments, however, supporting it.
-
-First of all, the number of observed GRBs rapidly decreases with energy.
-There are around 750 bursts detected by GBM in hard X-rays and low
-energy gamma rays \cite{Vianello:2013ela} (half of them were in a LAT
-field of view), only 30 or so were detected by LAT above
-\unit[100]{MeV}, and only about 4 of them were detected above GeV.
-
-% TODO merge 2
-First of all, the number of observed GRBs rapidly decreases with energy.
-There are around 750 bursts detected by GBM in hard X-rays and low
-energy gamma rays \cite{Vianello:2013ela} (half of them were in a LAT
-field of view), only 30 or so were detected by LAT above
-\unit[100]{MeV}, and only about 4 of them were detected above GeV.
+The main idea behind our model is to assume that the burst opening
+angle depends on the energy of emitted photons, or, equivalently, that the
+most energetic plasma particles are concentrated near the axis of the
+jet, while low energy particles are on the periphery.  As it's seen 
+in Figure \ref{fig:modelOverview} this leads to a time stretching
+effect due to relativistic beaming.
+
+Although the details of the exact mechanism of the GRB are not known
+the are arguments in support of our model.
 
 First of all, there are around 750 GRBs detected by GBM, half of which
-were in the LAT field of view at the moment of observation
+are in the LAT field of view at the moment of observation
 \cite{Vianello:2013ela}.  However, only about 30 of them were detected
 by the LAT, and only 4 of them were bright in the high energy band.
-If we extrapolate the uniform jet model to very high energies, this
-observation would mean that these groups of bursts are internally
-different: some of them produce VHE radiation, while others do not.
-These differences are hard to explain given that burst energetics are
-similar \cite{Bloom:2003wy}.  Nevertheless, these differences in burst
-counts can easily be explained by our model.  In our model, the
-opening angle of a jet is inversely proportional to the energy of
+If one extrapolates the uniform jet model to the very high energies,
+this observation would mean that these groups of bursts are internally
+different: some of them produce VHE radiation, while the others do
+not.  These differences are hard to explain given that the burst
+energetics are similar \cite{Bloom:2003wy}.  These differences in
+burst counts find natural explanation in our model, in which the
+opening angle of a jet is inversely proportional to the energy of the
 photons it radiates.  Therefore the most common scenario is that the
@@
 opening angle, but much larger than the opening angle of a high energy
-jet.  Because of that, most of the bursts can only be seen at low
+jet.  Due to that, most of the bursts can only be seen at low
 energies.  The 4 bursts we study in this paper were seen, according to
@@
 
-This observation is hard to explain in a framework of uniform jet model given that burst energetics are similar \cite{Bloom:2003wy}.
-It is, however, easy to explain in our model, given that small observation angles are rare, and the only way to produce high energy radiation is to have the observation angle small enough.
-
-% merge 
 Another argument comes from consideration of processes happening in
@@
 
-Second, consider the plasma right after its ejection from the central
-engine.  Two processes happen there simultaneously:
+First, let us list the assumptions of the model:
+
 \begin{enumerate}
-	\item{Particles near the jet boundary collide and change their
-          movement directions, therefore increasing the opening angle
-          of the jet.}
-	\item{Particles lose energy, therefore decreasing radiation
-          frequency emitted by the jet.}
-\end{enumerate}
-We argue that these two processes are correlated, for they happen due
-to the same particle interaction processes.  And since they are
-correlated, jets with larger opening angles should have lower
-energies, which is the assumption we are trying to justify.
-
-In the following sections we will rigorously derive the procedure for computing time stretching factors from our model, and compare them to observations obtained in the previous section.
-
-To understand the observed spectral lag, lets explore geometry of the jet. First of all, lets make some assumptions:
-\begin{enumerate}
-\item{Time $t = 0$, a spherical shell of plasma is emitted. The center is called the central engine.}
-\item{The shell points propagate with a constant velocity $v = \frac{\sqrt{\gamma^2 - 1}}{\gamma} \sim 1$, so at the time $t$ the radius of the shall is $v t$.}
-\item{Each point of the shell is an isotropic radiator in its rest frame.}
-\item{
-	The radiation intensity is a function of the radiator position and the radiation frequency:
-	\begin{equation}
+\item{Time $t = 0$, a spherical shell of plasma is emitted. The center
+  is called the central engine.}
+\item{The shell points propagate with a constant velocity $v =
+  \frac{\sqrt{\gamma^2 - 1}}{\gamma} \sim 1$, so at the time $t$ the
+  radius of the shall is $v t$.}
+\item{Each point of the shell is an isotropic radiator in its rest
+  frame.}
+\item{ The radiation intensity is a function of the radiator position
+  and the radiation frequency:
+	\begin{equation}\label{eq:eta}
 		\eta\left(r,\theta,\omega\right) = 
@@
 	\end{equation}
-	$\eta$ is a number of particles emitted per volume per solid angle per frequency.
-It is a function of the distance $r$ from the central engine, of the off-axis angle $\theta$, and of the radiation frequency $\omega$.
-}
+	$\eta$ is a number of particles emitted per volume per solid
+        angle per frequency. It is a function of the distance $r$ from
+        the central engine, of the off-axis angle $\theta$, and of the
+        radiation frequency $\omega$.  }
 \end{enumerate}
 
-The burst is fully specified by the following set of parameters:
+The burst is fully defined by the following set of parameters:
 \begin{itemize}
@@
 
-Now we have all we need to calculate the observed light curves, and then the stretching factors.
+In the next step we calculate the observed light curves, and the
+stretching factors.
 
@@
 
-Lets begin by computing a time at which some particular photon is observed.
-This time is a function of the radiator location $\left(r, \theta, \phi\right)$, as well as the observer location $\left(d, \chi, 0\right)$.
-(Here we choose coordinates so that the rotation angle of the observer is $0$.) Lets assume for now that the observer is too far from the jet to resolve its geometry, yet close enough so that the expansion of space is negligible.
-This assumption allows us to ignore the effects of cosmology for this calculation.
-
-The observation time is a sum of two terms: the time interval from $t = 0$ to the photon emission (the plasma time), and the time interval between the emission and the observation (the photon time):
+We begin with computing a time at which some particular photon is
+observed. This time is a function of the radiator location $\left(r,
+\theta, \phi\right)$, as well as the observer location $\left(d, \chi,
+0\right)$ with respect to the burst center (we choose the coordinates
+so that the rotation angle of the observer is $0$.) We assume for now
+that the observer is too far from the jet to resolve its geometry, yet
+close enough so that the expansion of space is negligible. The latter
+assumption is made for simplicity of presentation. The cosmological
+expansion effect will be restored at the end of this section.
+
+The observation time is a sum of two terms: the time interval from $t
+= 0$ to the photon emission (the plasma time), and the time interval
+between the emission and the observation (the photon time):
 \begin{equation*}
@@
 \end{equation*}
-Finally, we can use the spherical law of cosines to write this in terms of the great-circle distance $\sigma\left( \theta, \phi, \chi \right)$ between the points $\left( \theta, \phi \right)$ and $\left( \chi, 0 \right)$:
+Finally, one can use the spherical law of cosines to write this in terms of the great-circle distance $\sigma\left( \theta, \phi, \chi \right)$ between the points $\left( \theta, \phi \right)$ and $\left( \chi, 0 \right)$:
 \begin{equation}
@@
 \subsection{Light curve}
-We now know enough to compute the observable quantity - the number of observed photons $p$.
-For that we need to integrate the radiation intensity $\eta$ over four things.
-Two burst-related things: radiator positions $\left(r, \theta, \phi\right)$ and frequencies at emission $\omega'$.
-Two observer-related things: observation time $\tau$ and frequencies at observation $\omega$.
-We relate the first two and the second two with the delta functions:
+We now know enough to compute the observable quantity - the number of
+observed photons $p$. For that we need to integrate the radiation
+intensity $\eta$ over the four groups of variables. Two groups are
+burst-related: radiator position $\left(r, \theta, \phi\right)$ and
+frequency at emission $\omega'$. The other two variables are related
+to observer: observation time $\tau$ and frequency at observation
+$\omega$. We relate the first two and the second two with the delta
+functions:
 \begin{align}
@@
 
-We can use the delta functions to do the integrals over $\omega'$ and $r$. For that lets transform the first one:
+We integrate over $\omega'$ and $r$ using the delta functions. For
+that we transform them as follows:
 \begin{align*}
 \delta\left( \frac{\omega'}{\gamma \left( 1 - v\cos\sigma \right)\left(1+z\right)} - \omega \right) &= \delta\left( \frac{1}{\gamma\left( 1-v\cos\sigma \right)\left(1+z\right)} \left( \omega' - \gamma\left( 1 - v\cos\sigma \right) \left(1+z\right) \omega \right) \right) \\
-&= \gamma\left( 1 - v\cos\sigma \right) \left(1+z\right) \delta\left( \omega' - \gamma\left( 1 - v\cos\sigma \right) \left(1+z\right) \omega \right)
-\end{align*}
-And the second one:
-\begin{align*}
+&= \gamma\left( 1 - v\cos\sigma \right) \left(1+z\right) \delta\left( \omega' - \gamma\left( 1 - v\cos\sigma \right) \left(1+z\right) \omega \right)\\
 \delta\left(\tau - r\left( \frac{1}{v} - \cos\sigma \right)\left(1+z\right)\right) &= \delta\left( \left( \frac{1}{v} - \cos\sigma \right) \left( 1+z \right) \left( r - \frac{\tau}{\left(\frac{1}{v} - \cos\sigma\right)\left(1+z\right)} \right) \right) \\
@@
 
-Furthermore, if we write $\eta$ explicitly we can do the integrals over $\omega$ and $\tau$:
+Furthermore, one may integrate over $\omega$ and $\tau$ after using an
+explicit form for $\eta$ Eq. \ref{eq:eta}:
 \begin{align}
@@
 
-Finally, we will need to compute limits of $p$ for $\omega_2 \rightarrow \infty$ and $\tau_2 \rightarrow \infty$.
-It requires us to know the limit of $I$ for $\omega \rightarrow \infty$ (we assumed that $k < 0$):
+Finally, we need to compute the limits of $p$ for $\omega_2
+\rightarrow \infty$ and $\tau_2 \rightarrow \infty$. It requires us to
+know the limit of $I$ for $\omega \rightarrow \infty$ (here we assume that
+$k < 0$):
 \begin{equation}
@@
 
-Now, when we know $p$, we can compute many different things with it. For example:
+The following quantities of phenomenological importance may be
+computed with the use of $p$:
 \begin{itemize}
-\item{the total number of particles observed in a given energy range, $p_\infty\left(z,\chi; \omega_1,\omega_2\right) = p\left( z,\chi; 0,\infty; \omega_1,\omega_2 \right)$;}
-\item{the fraction of photons observed during a given time interval, $\Phi\left(z,\chi; \tau_1,\tau_2; \omega_1,\omega_2\right) = \frac{p\left( z,\chi; \tau_1,\tau_2; \omega_1,\omega_2\right)}{p_\infty\left( z,\chi; \omega_1,\omega_2 \right)}$;}
-\item{the duration of the burst $T_f\left( z,\chi;\omega_1,\omega_2 \right)$, that is the time by which the fraction $f$ of photons is observed.
-We can compute it by solving the following equation for $T_f$: $p\left( z,\chi; 0,T_f; \omega_1,\omega_2\right) = f p_\infty\left( z,\chi; \omega_1,\omega_2 \right)$;}
-\item{the stretching factor, which we will discuss in the next section.}
+\item{the total number of particles observed in a given energy range,
+  $p_\infty\left(z,\chi; \omega_1,\omega_2\right) = p\left( z,\chi;
+  0,\infty; \omega_1,\omega_2 \right)$;}
+\item{the fraction of photons observed during a given time interval,
+  $\Phi\left(z,\chi; \tau_1,\tau_2; \omega_1,\omega_2\right) =
+  \frac{p\left( z,\chi; \tau_1,\tau_2;
+    \omega_1,\omega_2\right)}{p_\infty\left( z,\chi; \omega_1,\omega_2
+    \right)}$;}
+\item{the duration of the burst $T_f\left( z,\chi;\omega_1,\omega_2
+  \right)$, that is the time by which the fraction $f$ of photons is
+  observed. One may compute it by solving the following equation for
+  $T_f$: $p\left( z,\chi; 0,T_f; \omega_1,\omega_2\right) = f
+  p_\infty\left( z,\chi; \omega_1,\omega_2 \right)$;}
+\item{the stretching factor, which will be discussed in the next
+  section.}
 \end{itemize}
@@
 \subsection{Stretching factor}
-The stretching factor for a continuous light curve is defined exactly like the stretching factor for a discrete one. It is the value of $\kappa$ which makes the KS-distance minimal:
+The stretching factor for a continuous light curve is defined exactly
+like the stretching factor for a discrete one. It is the value of
+$\kappa$ which makes the KS-distance minimal:
 \begin{equation}
@@
 
-The maximum of an absolute value cannot be differentiated, so the computation of $\kappa$ by the given definition is complicated.
-Lets instead rewrite this expression in terms of a positive and a negative KS-distances:
+The maximum of an absolute value is not differentiable, so the
+computation of $\kappa$ by the given definition is complicated. One
+may instead rewrite this expression in terms of a positive and a
+negative KS-distances:
 \begin{align*}
@@
 
-So we have now a simpler way to compute $\kappa$ by solving an equation instead of computing the minimum.
-And $\Phi$ can be differentiated, which makes it easier to compute maximums and minimums of their differences.
-
-Being able to compute the stretching factor, we could now compare our model predictions with observations given the position of an observer.
-We don't know the observer's off-axis angle $\chi$, however, so we cannot check the stretching factor prediction directly.
-Instead, we will focus on a series of tests, which will ensure that our model doesn't contradict existing observations.
-These tests will be discussed in the following sections.
+So we arrive at a simpler way to compute $\kappa$ by solving an
+equation instead of computing the minimum.
+
+Being able to compute the stretching factor, we could now compare our
+model predictions with observations given the position of an
+observer. We don't know the observer's off-axis angle $\chi$, however,
+so we cannot check the stretching factor prediction on the
+burst-to-burst basis. Instead, we focus on a series of tests, which
+will ensure that our model doesn't contradict to existing
+observations. Namely the test of the total energy and comparison with
+Fermi-LAT data of the distribution of the stretching factors and the
+fraction of the high-energy bursts. These tests will be discussed in
+the following section.
 
@@
 
-The first test is to compute the total energy radiated from the burst, and to ensure that it doesn't exceed the mass of the star from which the burst originated.
-
-To compute the total energy, we need to multiply the radiation intensity by frequency, and integrate it over frequencies $\omega$, volume $\left(r,\theta,\phi\right)$, and observer positions $\left(\sigma, \xi\right)$.
-We assumed $\theta \ll 1$, and have taken into account the same relativistic effects as we did in the observed particle count computation:
+In the first test we compute the total energy radiated from the burst,
+and to ensure that it doesn't exceed the mass of the star from which
+the burst originated.
+
+To compute the total energy, we need to multiply the radiation
+intensity by frequency, and integrate it over frequencies $\omega$,
+volume $\left(r,\theta,\phi\right)$, and observer positions
+$\left(\sigma, \xi\right)$. We assume $\theta \ll 1$, and have taken
+into account the same relativistic effects as we did in the observed
+particle count computation:
 \begin{align}
@@
 
-Note, however, that since $\alpha < -2k-1$ this integral diverges for $\omega \rightarrow 0$.
-But our model is not expected to describe low-energy radiation of the burst, so we should not integrate over low-frequency radiators.
-Nevertheless, we should ensure that the total energy of the high-energy radiation is not too large.
-To compute this energy we will only integrate over those radiators which emission we can observe, that is over the radiators with frequencies greater than $\frac{\omega_1}{\gamma}$, where $\omega_1$ is the smallest observable photon frequency.
-Now we can compute the last integral and get the final expression for $E$:
+Note, however, that since $\alpha < -2k-1$ this integral diverges for
+$\omega \rightarrow 0$. This is not a problem as long as our model is
+not expected to describe low-energy radiation of the burst. One should
+ensure that the total energy of the high-energy radiation is not too
+large. For this test, we integrate only over those radiators which
+emission is observed by Fermi LAT, that is over the frequencies
+greater than $\frac{\omega_1}{\gamma}$, where $\omega_1$ is the
+smallest observable photon frequency. Now we compute the last integral
+and get the final expression for $E$:
 \begin{align}
@@
 
-This energy should not be larger than a typical mass $M_s$ of a massive star. So, finally, we arrive at the first constraint for the burst parameters:
+This energy should not exceed a typical mass $M_s$ of a massive
+star. So, finally, we arrive at the first constraint for the burst
+parameters:
 \begin{equation}
@@
 
-The second test is to calculate the distribution of stretching factors of observable bursts, and to compare it to observations.
-
-The computation of the exact and precise distribution is technically hard (because stretching factors change monotonously neither with redshifts, nor with off-axis angles), and computationally intensive.
-So we will not compute the precise distribution.
-
-Instead, we will use the Monte-Carlo method to produce a large representative sample of stretching factors.
-Then, we will calculate an empirical CDF of this sample, which can be compared with observations using the KS-test.
-Since we can compute much larger sample than that of observations, we will not lose much precision due to this simplification.
-
-We still have one ingredient missing though -- the evolution of the bursts density.
-We assume that the bursts density is roughly proportional to the stars density, and the stars density is roughly proportional to the matter density, which changes with $z$ as $\left(1+z\right)^3$.
-It is clear, however, that since no stars existed at very small redshifts, the burst density should decline there, so we add an exponential cutoff to it:
-\begin{equation}
-\rho = \rho_0 \left(1+z\right)^3 \exp\left(-\frac{z}{z_c}\right)
-\end{equation}
-Here $\rho_0$ is a normalization factor, and $z_c$ is a redshift scale, after which the density is cut off.
-
-Now when we have all the ingredients, we can compute the sample with the following steps:
+The second test is to calculate the distribution of stretching factors
+of observable bursts, and to compare it to observations.
+
+The computation of the exact and precise distribution is technically
+hard and computationally intensive because stretching factor is not a
+monotonic function of neither redshift nor off-axis angles.  We do not
+compute the distribution directly. Instead, we use the Monte-Carlo
+method to produce a large representative sample of stretching
+factors. Then, we calculate an empirical CDF of this sample, which can
+be compared with observations using the KS-test. Since one may compute
+much larger sample than that of observations, there is no precision
+loss due to this simplification.
+
+We still have one ingredient missing though -- the evolution of the
+bursts density. We assume that the bursts density is roughly
+proportional to the stars density, and the latter is roughly
+proportional to the matter density, which changes with $z$ as
+$\left(1+z\right)^3$. It is clear, however, that since no stars
+existed at very small redshifts, the burst density should decline
+there, so we add an exponential cutoff to it:
+\begin{equation}\label{eq:redshift}
+\rho = \rho_0 \left(1+z\right)^3 \exp\left(-\frac{z}{z_c}\right)\,
+\end{equation}
+where $\rho_0$ is a normalization factor, and $z_c$ is a redshift
+scale, after which the density is cut off.
+
+For the modeling we first need to define the range of redshifts and
+off-axis angles. The range should include all the observable jet
+positions, but one should keep the range as small as possible in order
+to avoid production of too many points corresponding to an invisible
+bursts. For simplicity we consider the rectangular region.
+
+We start with selecting a range for redshifts. Ideally, one would
+expect that this range starts with $z=0$. Then, however, there will be
+visible jets for all possible off-axis angles, which will make our
+angles range too large. Note also, that since the bursts count
+increases with redshift as $z^3$, the probability to observe a
+low-redshift burst is very small. So instead of selecting the smallest
+redshift to be $0$, we select it to be a small number $z_\text{min}$.
+
+$z_\text{max}$ is defined by the farthest jet, which can be
+observed. Since the burst observability
+$p_\infty\left(z,\chi;\omega_2,\omega_3\right)$ decreases with both
+redshift and $\chi$, we can find the maximum redshift by solving the
+following equation:
+\begin{equation}
+p_\infty\left(z_\text{max},0;\omega_2,\omega_3\right) = p_\text{min} \,,
+\end{equation}
+where $p_\text{min}$ is the minimum number of particles required to claim an observation.
+
+The observability declines with $z$ and $\chi$, so the observable
+burst with the largest $\chi$ should be located at the redshift
+$z_\text{min}$. We can find this angle by solving the similar
+equation, as we did for the redshifts:
+\begin{equation}
+  	p_\infty\left(z_\text{min},\chi_\text{max};\omega_2,\omega_3\right) = p_\text{min}
+\end{equation}
+
+The simulation is performed by repetition of the following 3 steps
+before the required number of bursts is attained:
 \begin{enumerate}
 \item{
-	First of all we need to define the range from which to select redshifts and off-axis angles.
-We should include all observable jet positions in this range.
-However, to avoid dropping too many points corresponding to invisible bursts, we should keep the range as small as possible.
-We also want to make the region rectangular to be able to select redshifts and angles separately.
-
-	Lets start by selecting a range for redshifts.
-Ideally, we want this range to start with $z=0$.
-Then, however, there will be visible jets for all possible off-axis angles, which will make our angles range too large.
-Note also, that since the bursts count increases with redshift as $z^3$, the probability to observe a low-redshift burst is very small.
-So instead of selecting the smallest redshift to be $0$, we select it to be a small number $z_\text{min}$.
-
-	$z_\text{max}$ is defined by the farthest jet, which can be observed.
-Since the burst observability $p_\infty\left(z,\chi;\omega_2,\omega_3\right)$ decreases with both redshift and $\chi$, we can find the maximum redshift by solving the following equation:
-	\begin{equation}
-	p_\infty\left(z_\text{max},0;\omega_2,\omega_3\right) = p_\text{min}
-	\end{equation}
-	where $p_\text{min}$ is the minimum number of particles required to claim an observation.
-}
-\item{
-	The range for off-axis angles is easier to compute since nothing prohibits us from selecting the smallest angle to be $0$.
-
-	The observability declines with $z$ and $\chi$, so the observable burst with the largest $\chi$ should be located at the redshift $z_\text{min}$.
-We can find this angle by solving the similar equation, as we did for the redshifts:
-	\begin{equation}
-	p_\infty\left(z_\text{min},\chi_\text{max};\omega_2,\omega_3\right) = p_\text{min}
-	\end{equation}
-}
-\item{
-	Now we are in a position to select a particular properly distributed random redshift. The CDF of the distribution of redshifts is a ratio of redshift counts in different space volumes:
+First, the random redshift is generated following the distribution
+\ref{eq:redshift}. This may be done with the Inverse transform sampling
+method. The CDF of the distribution of redshifts is a ratio of
+ redshift counts in different space volumes:
 	\begin{equation}
@@
 	\end{equation}
-	where $\dif V\left(z'\right)$ is the volume of the infinitesimal shell surrounding a sphere over which bursts at redshift $z'$ are distributed (see appendix \ref{sec:cosmology} for derivation).
-
-	To generate a redshift, we should uniformly select a value of $\Phi_z$, and solve the corresponding equation for $z$:
+	where $\dif V\left(z'\right)$ is the volume of the
+        infinitesimal shell surrounding a sphere over which bursts at
+        redshift $z'$ are distributed (see appendix
+        \ref{sec:cosmology} for derivation).
+
+To generate a redshift, one should uniformly select a value of
+$\Phi_z$, and solve the corresponding equation for $z$:
 	\begin{equation}
@@
 }
-\item{
-	An off-axis angle can be selected in a similar way. The CDF of the angles distribution is a ratio of spherical areas:
-	\begin{equation}
-	\Phi_\chi\left(\chi\right) = \frac{\int_0^\chi \sin\chi' \dif\chi'}{\int_0^{\chi_\text{max}} \sin\chi' \dif\chi'} \approx \frac{\int_0^\chi \chi' \dif\chi'}{\int_0^{\chi_\text{max}} \chi' \dif\chi'} = \left(\frac{\chi}{\chi_\text{max}}\right)^2
+\item{ The random off-axis angle may be then generated with a similar
+  method. The CDF of the angles distribution is a ratio of spherical
+  areas:
+	\begin{equation}
+	\Phi_\chi\left(\chi\right) = \frac{\int_0^\chi \sin\chi' \dif\chi'}{\int_0^{\chi_\text{max}} \sin\chi' \dif\chi'} \approx \frac{\int_0^\chi \chi' \dif\chi'}{\int_0^{\chi_\text{max}} \chi' \dif\chi'} = \left(\frac{\chi}{\chi_\text{max}}\right)^2\,.
 	\end{equation}
@@
 	\begin{align}
-	\Phi_\chi\left(\chi\right) = \left(\frac{\chi}{\chi_\text{max}}\right)^2 = y \\
-	\chi = \chi_\text{max}\sqrt{y}
+	\Phi_\chi\left(\chi\right) = \left(\frac{\chi}{\chi_\text{max}}\right)^2 = y\,, \\
+	\chi = \chi_\text{max}\sqrt{y}\,,
 	\end{align}
-	Here $y$ is an another independent random variable uniformly distributed in the range $0$ to $1$.
-}
-\item{Now we should check, if the burst in a selected position can be observed: $p_\infty\left(z,\chi;\omega_2,\omega_3\right) > p_\text{min}$.
-If it is, add $\kappa\left(z,\chi\right)$ to the sample.
-If not, repeat from the step $3$.}
-\item{If the sample is not as big as we want yet, repeat from the step $3$.}
+	where $y$ is an another independent random variable uniformly distributed in the range $0$ to $1$.
+}
+\item{Now we check, whether the burst in a selected position may be
+  observed: $p_\infty\left(z,\chi;\omega_2,\omega_3\right) >
+  p_\text{min}$. If it is, add $\kappa\left(z,\chi\right)$ to the
+  sample.}
 \end{enumerate}
 
-With this algorithm, we arrive to our second test / constraint for the burst parameters. The obtained stretching factors distribution should be compatible with observations.
+With this algorithm, we arrive to the stretching factor distribution
+which may be compared with the observed one.
 
@@
 
-Our final test is comparison of answers to this question: given the bursts which were observed in a whole energy range $\left(\omega_1,\omega_3\right)$, which fraction of them can also be observed in a high energy range $\left(\omega_2,\omega_3\right)$?
-
-To calculate this value, we need to compute the number of bursts visible in a given energy range, which is the integral over space and jet directions:
+Our final test is based on the answer to the following question: given
+the bursts observed in the whole energy range
+$\left(\omega_1,\omega_3\right)$, which fraction of them may also be
+observed in the high energy range $\left(\omega_2,\omega_3\right)$?
+
+To calculate this ratio, one needs to compute the number of bursts
+visible in a given energy range, which is the integral over space and
+jet directions:
 \begin{align}
 b\left(\omega_1,\omega_2\right) &= \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} 2\pi \sin\chi \dif\chi \nonumber\\
-&\approx 2\pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} \chi \dif\chi = \pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_2\right)
+&\approx 2\pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} \chi \dif\chi = \pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_2\right)\,,
 \end{align}
-Here $z_\text{max}\left(\omega_1,\omega_2\right)$ and $\chi_\text{max}\left(z,\omega_1,\omega_2\right)$ are the same values we used in the previous section: the maximum redshift from which a burst can be observed in a given energy range, and the maximum off-axis angle with which the burst at redshift $z$ can be observed.
-
-The fraction we want to compute is the ratio of these integrals:
-\begin{equation}
-f\left(\omega_1,\omega_2,\omega_3\right) = \frac{b\left(\omega_2,\omega_3\right)}{b\left(\omega_1,\omega_3\right)} = \frac{\int_0^{z_\text{max}\left(\omega_2,\omega_3\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_2,\omega_3\right)}{\int_0^{z_\text{max}\left(\omega_1,\omega_3\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_3\right)}
-\end{equation}
-
-This is our 3rd test -- the ratio obtained should be compatible with the observed value.
+where $z_\text{max}\left(\omega_1,\omega_2\right)$ and
+$\chi_\text{max}\left(z,\omega_1,\omega_2\right)$ are the same values
+we used in the previous section: the maximum redshift from which a
+burst can be observed in a given energy range, and the maximum
+off-axis angle with which the burst at redshift $z$ can be observed.
+
+The required fraction is the ratio of these integrals:
+\begin{equation}
+f\left(\omega_1,\omega_2,\omega_3\right) = \frac{b\left(\omega_2,\omega_3\right)}{b\left(\omega_1,\omega_3\right)} = \frac{\int_0^{z_\text{max}\left(\omega_2,\omega_3\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_2,\omega_3\right)}{\int_0^{z_\text{max}\left(\omega_1,\omega_3\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_3\right)}\,.
+\end{equation}
+
+Our 3rd test is the comparison of the obtained ratio with the observed value.
 
@@
 
-For now we discussed how to compute various observables of the model, and how to test the model against the data from observations.
-However, we never considered specific values for parameters of the burst.
-We will do that in this section.
-
-First of all, we need to determine, which parameters to fit.
-At most, the observables of a particular burst depend on 10 parameters: 8 of the burst $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k, \alpha\right)$, and 2 of the observer $\left(z, \chi\right)$.
-However, some of these parameters only change observables trivially, and, also, there are transformations of variables, which will not change observables at all (see equation \ref{eq:p}):
+For now we have discussed how to compute various observables of the model,
+and how to test the model against the data from observations. Now we
+come to the discussion of specific values for the parameters of the
+burst.
+
+First of all, we need to determine, which parameters to fit. At most,
+the observables of a particular burst depend on 10 parameters: 8 of
+the burst $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k,
+\alpha\right)$, and 2 of the observers position related to the burst
+$\left(z, \chi\right)$. However, some of these parameters only change
+observables trivially, and, also, there are transformations of
+variables, which do not change observables at all (see equation
+\ref{eq:p}):
 \begin{itemize}
-	\item{The number of photons $p$ depends linearly on $\eta_0$, so $\eta_0$ doesn't affect other observables like duration, or stretching factor.
-We can easily fit $\eta_0$ to match the observed photon count.}
-	\item{If we make a simultaneous transformation of $r_0 \rightarrow \lambda r_0$ and $\eta_0 \rightarrow \frac{1}{\lambda^3}\eta_0$, where $\lambda$ is a parameter of transformation, only duration of the burst will change.
-Other observables, like the total number of photons, or stretching factor, will not be affected.}
-	\item{Finally, if we make a following transformation: $\omega_0 \rightarrow \lambda \omega_0$, $\theta_0 \rightarrow \lambda^k \theta_0$, $\eta_0 \rightarrow \lambda^\alpha \eta_0$, where $\lambda$ is another transformation parameter, no observables will change whatsoever.}
+	\item{The number of photons $p$ depends linearly on $\eta_0$,
+          so $\eta_0$ doesn't affect other observables like duration,
+          or stretching factor. One can easily fit $\eta_0$ to match
+          the observed photon count.}
+	\item{If we make a simultaneous transformation of $r_0
+          \rightarrow \lambda r_0$ and $\eta_0 \rightarrow
+          \frac{1}{\lambda^3}\eta_0$, where $\lambda$ is a parameter
+          of transformation, only duration of the burst will
+          change. Other observables, like the total number of photons,
+          or stretching factor, will not be affected.}
+	\item{Finally, if we make a following transformation:
+          $\omega_0 \rightarrow \lambda \omega_0$, $\theta_0
+          \rightarrow \lambda^k \theta_0$, $\eta_0 \rightarrow
+          \lambda^\alpha \eta_0$, where $\lambda$ is another
+          transformation parameter, no observables will change
+          whatsoever.}
 \end{itemize}
-Given these transformations, we can reduce the number of parameters for fit to 7: $\left(\gamma, n, \theta_0, k, \alpha, z, \chi\right)$.
-
-Also, some of these parameters we know from observations:
+Given these transformations, one can reduce the number of parameters
+for fit to 7: $\left(\gamma, n, \theta_0, k, \alpha, z, \chi\right)$.
+
+Also, some of these parameters are known from observations:
 \begin{itemize}
-	\item{Redshifts $z$ of many bursts have been measured \cite{Ackermann:2013zfa}.}
-	\item{We know that relativistic factors $\gamma$ are on the order of magnitude\footnote{Our minimization procedure can find similar minimums for other values of $\gamma$, at least from $100$ to $1000$} of $\gamma = 300$ \cite{Ghirlanda:2011ux}.}
+	\item{Redshifts $z$ of many bursts have been measured
+          \cite{Ackermann:2013zfa}.}
+	\item{We know that relativistic factors $\gamma$ are on the
+          order of magnitude\footnote{Our minimization procedure can
+            find similar minimums for other values of $\gamma$, at
+            least from $100$ to $1000$} of $\gamma = 300$
+          \cite{Ghirlanda:2011ux}.}
 \end{itemize}
@@
 
-We want our cost function $C\left(n, \theta_0, k, \alpha, \chi\right)$ to satisfy the following objectives:
+The best fit parameters are found by minimizing the cost function
+$C\left(n, \theta_0, k, \alpha, \chi\right)$ which should
+satisfy the following objectives:
 \begin{itemize}
-	\item {
-		Total energy of the burst is finite (that is $k+\frac{\alpha}{2}+1 < 0$) and smaller than $\unit[6 \times 10^{53}]{GeV} \approx \unit[10^{51}]{erg}$ \cite{Gehrels:2013xd}.
-	}
-	\item {
-		The stretching factor of the burst should be compatible with the value for the GRB 090926A.
-	}
-	\item {
-		We require all bursts to have the same burst parameters $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k, \alpha\right)$.
-And we know that bursts with small stretching factors like GRB 090902B should be possible to observe.
-So we require that bursts with the stretching factor of GRB 090902B or lower appear in random samples by $z$ and $\chi$ (sampling is done the same way as in section \ref{sec:distribution}).
-	}
-	\item {
-		The ratio of photons with high energies and low energies should be compatible with the value for the GRB 090926A, $\frac{p_{\infty} \left(z,\chi;\unit[1]{GeV},\infty\right)}{p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV}\right)} = 0.057$.
-	}
-	\item {
-		Finally, the burst should not be too faint compared to other bursts from the sample. In other words, the total number of observed photons should approximately equal to the median among the sample.
-	}
+	\item { Total energy of the burst is finite (that is
+          $k+\frac{\alpha}{2}+1 < 0$) and smaller than $\unit[6 \times
+            10^{53}]{GeV} \approx \unit[10^{51}]{erg}$
+          \cite{Gehrels:2013xd}.  }
+	\item { The stretching factor of the burst should be
+          compatible with the value for the GRB 090926A.  }
+	\item { The ratio of photons with high energies and low
+          energies should be compatible with the value for the GRB
+          090926A, $\frac{p_{\infty}
+            \left(z,\chi;\unit[1]{GeV},\infty\right)}{p_{\infty}
+            \left(z,\chi;\unit[0.1]{GeV},\unit[1]{GeV}\right)} =
+          0.057$.  }
+	\item { The burst should not be too faint compared to
+          other bursts from the sample. In other words, the total
+          number of observed photons should approximately equal to the
+          median among the sample.  }
+	\item { We assume that all bursts have the same burst
+          parameters $\left(\gamma, \eta_0, r_0, n, \omega_0,
+          \theta_0, k, \alpha\right)$ and require that bursts with
+          small stretching factors like GRB 090902B should be possible
+          to observe. So we require that bursts with the stretching
+          factor of GRB 090902B or lower appear in random samples by
+          varying $z$ and $\chi$ (sampling is done the same way as in
+          section \ref{sec:distribution}). Note that this requirement
+          unlike others applies to an ensemble of simulated bursts. }
 \end{itemize}
@@
 		Here $\log\kappa_{\text{GRB090926A}} = \frac{1}{2}\left(\log\left(6.62\right) + \log\left(1.99\right)\right)$ and $\Delta\log\kappa_{\text{GRB090926A}} = \frac{1}{2}\left(\log\left(6.62\right) - \log\left(1.99\right)\right)$.
-	}
-	\item{
-		Compute the cost due to the minimal stretching factor from the sample:
-		\begin{equation}
-			C_{\kappa\text{min}} = \max\left(0, 
-				\frac{
-					\log{\kappa_\text{min}} - \log{\kappa_{\text{GRB090902B}}}
-				}{\Delta\log\kappa_{\text{GRB090902B}}}\right)
-		\end{equation}
-		Here $\kappa_\text{min}$ is the minimal stretching factor from the sample computed in the step \ref{item:sample}, $\log\kappa_{\text{GRB090902B}} = \frac{1}{2}\left(\log\left(0.89\right) + \log\left(0.35\right)\right)$ and $\Delta\log\kappa_{\text{GRB090902B}} = \frac{1}{2}\left(\log\left(0.89\right) - \log\left(0.35\right)\right)$.
 	}
@@
 	\item{
+		Compute the cost due to the minimal stretching factor from the sample:
+		\begin{equation}
+			C_{\kappa\text{min}} = \max\left(0, 
+				\frac{
+					\log{\kappa_\text{min}} - \log{\kappa_{\text{GRB090902B}}}
+				}{\Delta\log\kappa_{\text{GRB090902B}}}\right)
+		\end{equation}
+		Here $\kappa_\text{min}$ is the minimal stretching factor from the sample computed in the step \ref{item:sample}, $\log\kappa_{\text{GRB090902B}} = \frac{1}{2}\left(\log\left(0.89\right) + \log\left(0.35\right)\right)$ and $\Delta\log\kappa_{\text{GRB090902B}} = \frac{1}{2}\left(\log\left(0.89\right) - \log\left(0.35\right)\right)$.
+	}
+	\item{
 		Finally, the value of the cost function is the sum of squares of the four:
 		\begin{equation}
-			C = C_\kappa^2 + C_{\kappa\text{min}}^2 + C_f^2 + C_b^2
+			C = C_\kappa^2 + C_f^2 + C_b^2 + C_{\kappa\text{min}}^2
 		\end{equation}
@@
 
-We use the Nelder-Mead (downhill simplex) method for the minimization procedure.
-Initial points were chosen to include the areas in parameter space where each cost is close to $0$ (first $4$ points correspondingly), and to cover the large fraction of the parameter space.
-You can see the values in table \ref{tab:fitInitialPoints}.
-
-The minimization then converged to the following parameter values:
+We use the Nelder-Mead (downhill simplex) method for the minimization
+procedure. Initial points were chosen to include the areas in
+parameter space where each cost is close to $0$ (first $4$ points
+correspondingly), and to cover the large fraction of the parameter
+space. You can see the values in table \ref{tab:fitInitialPoints}.
+
+The minimization converged to the following parameter values:
 \begin{multicols}{2}
@@
 
-Note, that these parameter values (except for $z$ and $\chi$) are universal, and they can describe the whole population of observed bursts, as will be shown in the next section.
+Note, that these parameter values (except for $z$ and $\chi$) are
+universal, and they may describe the whole population of observed
+bursts, as will be shown in the next section.
 
@@
 
-	As was discussed in the previous section our model can reproduce both stretching factors smaller than $1$ (like of GRB 090902B) and larger than $1$ (like of GRB 090926A) depending on redshift and observer's off-axis angle (see fig.
-\ref{fig:sampleLightCurves}).
-
-	Here are results of our tests:
+As was discussed in the previous section our model can reproduce both
+stretching factors smaller than $1$ (like of GRB 090902B) and larger
+than $1$ (like of GRB 090926A) depending on redshift and observer's
+off-axis angle (see fig. \ref{fig:sampleLightCurves}).
+
+Below are the results of the tests of Section~\ref{sec:tests}:
 \begin{itemize}
-		\item{
-			Total energy emitted in $\unit[100]{MeV}$ and more energetic gamma rays $E < \unit[5.89 \times 10^{53}]{GeV}$, which is in agreement with \cite{Gehrels:2013xd}.
-		}
-		\item{
-			The distribution of observable bursts' stretching factors is shown on fig. \ref{fig:kappaDistribution}.
-			This distribution does not contradict to the values of stretching factors obtained in section \ref{sec:observations}.
-		}
-		\item{
-			The fraction of bursts observable in low energy band, which are also observed in high energy band $f_m = 0.110$.
-			The Fermi LAT catalog contains $35$ bursts out of which $4$ can be observed in high energy band.
-			Therefore the observed value $f_o = 0.11 \pm 0.05$ (error obtained from binomial distribution) agrees with the model.
-		}
+	\item{ The total energy emitted in $\unit[100]{MeV}$ and more
+          energetic gamma rays $E < \unit[5.89 \times 10^{53}]{GeV}$,
+          which is in agreement with \cite{Gehrels:2013xd}.  }
+	\item{ The distribution of observable bursts' stretching
+          factors is shown in Fig. \ref{fig:kappaDistribution}.  This
+          distribution does not contradict to the values of stretching
+          factors obtained in section \ref{sec:observations}.  }
+	\item{ The fraction of bursts observable in low energy band,
+          which are also observed in high energy band is $f_m = 0.110$.
+          The Fermi LAT catalog contains $35$ bursts out of which $4$
+          can be observed in high energy band.  Therefore the observed
+          value $f_o = 0.11 \pm 0.05$ (error from binomial
+          distribution) agrees with the model.  }
 \end{itemize}
 
-	Also, there is a correlation between the fraction of high energy photons (that is high energy photon count divided by the total photon count), the stretching factor and the observer's off-axis angle (see fig.
-\ref{fig:correlations}).
-	It allows one to predict observer's off-axis angles.
-
-	So our model is able to explain the time stretching effect from section \ref{sec:observations}.
+Therefore the model successfully explains the variety of the
+stretching factors observed by the Fermi LAT.
+
+\section{Discussion}
+
+The are some caveats and possible directions of improvement of the
+model. First of all, as it's seen in Fig. \ref{fig:correlations} the
+stretching factor of GRB 090926A is higher that the model
+prediction. More data are needed to understand whether this requires a
+refinement of the model or is a statistical fluctuation.
+
+Second, the shapes of light curves produced by the model (see
+fig. \ref{fig:sampleLightCurves}) differ considerably from observed
+ones (e.g. fig. \ref{fig:grb090902B}). In particular, the slope of
+light curves produced by the model is zero at $t = 0$, while the slope
+of the observed light curves at $t = 0$ appears to have a maximum.
+
+Third, the assumption of plasma energy dependence on off-axis angle
+should be better justified.  This might probably be done using
+hydrodynamic simulation of the jet.
+
+We note, that even though it appears that bursts are similar in their
+rest frames, this may be considered only as a first approximation.
+The differences in masses, metallicities, angular momenta of
+progenitor stars, and the differences in local conditions around them
+might account for differences in light curves of GRBs (or, even, for
+especially large stretching factor of GRB 090926A).  This means that
+the minor variations of the model parameters might be necessary.
+
+It seems, however, that all the issues above may be addressed by
+refining the model without changing it's main assumptions.
+
+Our study can be summarized with 3 main conclusions.
+
+First, the time stretching of GRB light curves between different VHE
+(in particular $\unit[100]{MeV} < E < \unit[1]{GeV}$ and $E >
+\unit[1]{GeV}$ ) bands is discovered with the statistical significance
+of $3.3\sigma$.  Depending on the burst the stretching factor may be
+higher of lower than 1, that is the higher or the lower-energy light
+curve is stretched.
+
+Second, the time stretching may be explained with curvature effects,
+that is the effects of jet geometry.  There is no need to introduce
+any new spectral components.
+
+Finally, one may assume that all GRBs are the same in their rest
+frames.  The internal burst parameters (such as $\gamma$, $\eta_0$,
+$r_0$, $n$, $\omega_0$, $\theta_0$, $k$ and $\alpha$) might stay the
+same for all bursts, and this assumption is more or less consistent
+with existing observations.
+
+We note, that the model predicts a correlation between
+the fraction of the high energy photons, the stretching factor and the
+observer's off-axis angle (see fig. \ref{fig:correlations}). This may
+be used as a method to estimate the observer's off-axis angle.
 
@@
 
-%\section{Conclusions}
-
-%\label{sec:conclusions}
-%% TODO: remove section
-
-%% TODO: add more descriptive words
-
-	Our study can be summarized with 3 main conclusions.
-
-	First, the time stretching of GRB light curves between different VHE (in particular $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$) bands is discovered with statistic significance of $3.3\sigma$.
-	This time stretching might be positive or negative (that is higher or lower energy light curve is stretched) depending on the burst.
-
-	Second, this time stretching effect can be explained with curvature effects, that is effects of jet geometry.
-	There is no need to introduce any new spectral components.
-
-	Finally, we can assume all GRBs to be the same in their rest frames.
-	The internal burst parameters (such as $\gamma$, $\eta_0$, $r_0$, $n$, $\omega_0$, $\theta_0$, $k$ and $\alpha$) might stay the same for all bursts, and this assumption is more or less consistent with existing observations.
-
-	However, there are some caveats, and directions for improvement.
-
-	First of all, the model prediction and observations are not completely consistent.
-	If you look at fig. \ref{fig:correlations}, you will see that the stretching factor of GRB 090926A is too high compared to the model prediction.
-	It appears that parameters of the model cannot be fitted to account for it (without ruining agreement with other observations).
-	More data should be analyzed to see whether the problem will be exacerbated, and some refinement of the model might also be required.
-
-	Second, the shapes of light curves produced by the model (see fig. \ref{fig:sampleLightCurves}) differ considerably from observed ones (e.g. fig. \ref{fig:grb090902B}).
-	Most importantly, the slope of light curves produced by the model is zero at $t = 0$, even though the slope of observed light curves at $t = 0$ is not only non-zero, but appears to have a maximum.
-
-	Third, the assumption of plasma energy dependence on off-axis angle should be better justified.
-	This might probably be done using hydrodynamic simulation of the jet.
-
-	Finally, even though it appears that bursts are similar in their rest frames, they cannot be exactly the same.
-	The differences in masses, metallicities, angular momenta of progenitor stars, and the differences in local conditions around them might account for differences in light curves of GRBs (or, even, for especially large stretching factor of GRB 090926A).
-	It means that minor variations of model parameters might be necessary on burst per burst basis.
-
-It seems, however, that all these problems are not unresolvable, and
-may be addressed by refining the model without changing its main
-assumptions.
-%% TODO: remove unresolvable, rephrase
-
-{\bf Acknowledgments.} We thank M.~Pshirkov and S.~Troitsky for
-comments and inspiring discussions.  The work was supported by Russian
-Science Foundation grant 14-12-01340. The analysis is based on data
-and software provided by the Fermi Science Support Center (FSSC). The
-numerical part of the work was done at the cluster of the Theoretical
-Division of INR RAS. G.R. acknowledges the fellowship of the Dynasty
-foundation.
+{\bf Acknowledgments.} We thank A.~Gruzinov, M.~Pshirkov and
+S.~Troitsky for comments and inspiring discussions.  The work was
+supported by Russian Science Foundation grant 14-12-01340. The
+analysis is based on data and software provided by the Fermi Science
+Support Center (FSSC). The numerical part of the work was done at the
+cluster of the Theoretical Division of INR RAS. G.R. acknowledges the
+fellowship of the Dynasty foundation.
 
@@
 \label{sec:fermiCode}
-This appendix contains two things: exact values we are passing to a web form on the Fermi LAT data server, and the code we use to invoke the Fermi tools (including full list of options).
+This appendix contains the details on using the Fermi LAT
+data. Namely, the exact values passed to the web form on the Fermi LAT
+data server and the code used to run the Fermi tools including the
+full list of options.
 
@@
 
-For background, all parameters are the same, except for the time ranges, which are the following:
+For background, all the parameters are the same, except for the time ranges, which are the following:
 \begin{align*}
@@
 
-The commands we use to invoke Fermi Science tools are the following:
+The commands we use to run the Fermi Science Tools are the following:
 \begin{lstlisting}
@@
 \end{lstlisting}
-\texttt{[RA]} and \texttt{[DEC]} here are the coordinates of the burst, and \texttt{[IrfName]} is the Instrument Response Function, which depends on photon's event class (transient, source, clean, or ultraclean) and conversion type (back or front).
-The list of IRF names can be obtained with the \texttt{gtirfs} tool.
-We need to compute PSFs for all of them.
+\texttt{[RA]} and \texttt{[DEC]} here are the coordinates of the
+burst, and \texttt{[IrfName]} is the Instrument Response Function,
+which depends on photon's event class (``SOURCE'' in our case) and
+conversion type (back or front). The list of IRF names is obtained
+with the \texttt{gtirfs} tool.
 
@@
 \end{lstlisting}
-Note, that R.A. coordinates in the \texttt{expcube} FITS files are indexed in reverse order and starting from 180. For example, $\texttt{ra[0]}=180$, $\texttt{ra[1]}=179$, and so on.
-
-\section{Cosmology}
-%% TODO: change title
+Note, that R.A. coordinates in the \texttt{expcube} FITS files are
+indexed in reverse order and start from 180. For example,
+$\texttt{ra[0]}=180$, $\texttt{ra[1]}=179$, and so on.
+
+\section{Cosmological distances}
 \label{sec:cosmology}
-This appendix contains a few formulas we need from cosmology, in particular the area of a photon sphere $A_\text{ph}\left(z\right)$, a sphere over which photons emitted in a particular burst at redshift $z$ are spread; and the volume of the infinitesimal shell surrounding a burst sphere $\dif V\left(z\right)$, a sphere over which bursts at redshift $z$ are distributed.
+This appendix contains a few formulas related to the cosmological
+distances. In particular, it provides formula for the area of a photon
+sphere $A_\text{ph}\left(z\right)$, a sphere over which photons
+emitted in a particular burst at redshift $z$ are spread; and the
+volume of the infinitesimal shell surrounding a burst sphere $\dif
+V\left(z\right)$, a sphere over which bursts at redshift $z$ are
+distributed.
 
@@
 \end{equation}
-We define $a\left(t_\text{obs}\right) = 1$, where $t_\text{obs}$ is the observation time.
+We define $a\left(t_\text{obs}\right) = 1$, where $t_\text{obs}$ is
+the observation time.
 
@@
 \end{align*}
-Here $H_\text{obs} = H\left(t_\text{obs}\right) = \frac{\dot{a}\left( t_\text{obs} \right)}{a\left( t_\text{obs} \right)} = \dot{a}\left( t_\text{obs} \right)$ is the Hubble parameter at the observation time.
+Here $H_\text{obs} = H\left(t_\text{obs}\right) = \frac{\dot{a}\left(
+  t_\text{obs} \right)}{a\left( t_\text{obs} \right)} = \dot{a}\left(
+t_\text{obs} \right)$ is the Hubble parameter at the observation time.
 
@@
 centered on the observer. Also, they have different areas, because the
-scale factor differs at locations of burst and observer.
-%% TODO rephrase, mention back in time
+scale factor differs at the time of emission and observation.
 
@@
 \end{align*}
-Here $_2F_1\left(a, b; c; z\right) = \frac{\Gamma\left(c\right)}{\Gamma\left(b\right)\Gamma\left(c-b\right)}\int_0^1 \frac{t^{b-1}\left(1-t\right)^{c-b-1}}{\left(1-t z\right)^a} \dif t$ is a hypergeometric function.
-We can substitute $a\left(0\right) = \frac{1}{1+z}$, so, finally,
+Here $_2F_1\left(a, b; c; z\right) =
+\frac{\Gamma\left(c\right)}{\Gamma\left(b\right)\Gamma\left(c-b\right)}\int_0^1
+\frac{t^{b-1}\left(1-t\right)^{c-b-1}}{\left(1-t z\right)^a} \dif t$
+is a hypergeometric function. We then substitute $a\left(0\right) =
+\frac{1}{1+z}$ and finally arrive at
 \begin{equation}
@@
 sphere. Therefore, its area is:
-%% TODO: mention that this is the sphere in the past
 \begin{equation}
@@
 
-We need to know one more thing about the bursts sphere -- the volume of the infinitesimal shell surrounding it. For that we can again use the metric:
+It's helpful to calculate one more quantity related to the bursts
+sphere -- the volume of the infinitesimal shell surrounding it. For
+that we again use the metric:
 \begin{align}
@@
 &= A_\text{b}\left(z\right) \frac{\frac{1}{\left(1+z\right)^2}\dif z \left(1+z\right)}{H_\text{obs} \sqrt{\Omega_m\frac{1}{a^3\left(t\right)} + \Omega_\Lambda}} \nonumber\\
-&= A_\text{b}\left(z\right) \frac{1}{\left(1+z\right)} \frac{\dif z}{H_\text{obs} \sqrt{\Omega_m\left(1+z\right)^3 + \Omega_\Lambda}} \nonumber\\
+%&= A_\text{b}\left(z\right) \frac{1}{\left(1+z\right)} \frac{\dif z}{H_\text{obs} \sqrt{\Omega_m\left(1+z\right)^3 + \Omega_\Lambda}} \nonumber\\
 &= \frac{4\pi r^2\left(z\right)}{\left(1+z\right)^3} \frac{\dif z}{H_\text{obs} \sqrt{\Omega_m\left(1+z\right)^3 + \Omega_\Lambda}}

### 2014-12-12 Paper/gammaRays.tex
Fixed a few minor typos

@@
 have enough statistics we require that at least 10 photons are
-detected with the energy greater than $\unit[1]{GeV}$. This leaves are
+detected with the energy greater than $\unit[1]{GeV}$. This leaves us
 with the four bursts for the time span of the Fermi-LAT Gamma-Ray
@@
 the burst.  The analysis is performed with the Fermi Science Tools
-v9r32p5
+v9r33p0
 package~\footnote{\url{http://fermi.gsfc.nasa.gov/ssc/data/analysis/scitools/overview.html}}
@@
 Although the details of the exact mechanism of the GRB are not known
-the are arguments in support of our model.
+there are arguments in support of our model.
 
@@
 \subsection{Light curve}
-We now know enough to compute the observable quantity - the number of
+We now know enough to compute the observable quantity -- the number of
 observed photons $p$. For that we need to integrate the radiation
@@
 In the first test we compute the total energy radiated from the burst,
-and to ensure that it doesn't exceed the mass of the star from which
+and ensure that it doesn't exceed the mass of the star from which
 the burst originated.
@@
 positions, but one should keep the range as small as possible in order
-to avoid production of too many points corresponding to an invisible
+to avoid production of too many points corresponding to invisible
 bursts. For simplicity we consider the rectangular region.
@@
 the burst $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k,
-\alpha\right)$, and 2 of the observers position related to the burst
+\alpha\right)$, and 2 of the observer's position related to the burst
 $\left(z, \chi\right)$. However, some of these parameters only change
@@
 (in particular $\unit[100]{MeV} < E < \unit[1]{GeV}$ and $E >
-\unit[1]{GeV}$ ) bands is discovered with the statistical significance
+\unit[1]{GeV}$) bands is discovered with the statistical significance
 of $3.3\sigma$.  Depending on the burst the stretching factor may be
-higher of lower than 1, that is the higher or the lower-energy light
+higher or lower than 1, that is the higher or the lower-energy light
 curve is stretched.

### 2014-12-12 Paper/gammaRays.tex
Spellchecked a typo

@@
 Out of 19 bursts studied, only 4 have at least 10 high energy events
-remaining after the filtering, and thus eligible for computating the
+remaining after the filtering, and thus eligible for computing the
 stretching factors. The results of this computation are shown in

### 2014-12-12 Paper/gammaRays.tex
Capitalized Section, Table and Figure

@@
 			High and low energy light curves produced by the geometrical model.
-			Burst parameter values are the same as discussed in section \ref{sec:parameters}.
+			Burst parameter values are the same as discussed in Section \ref{sec:parameters}.
 		}
@@
 \ref{sec:observations}. The model and the details on calculations are
-explained in Section \ref{sec:model}. The Section \label{sec:tests}
+explained in Section \ref{sec:model}. The Section \ref{sec:tests}
 introduces three phenomenological tests for the model. The parameters
@@
 \end{itemize}
-The data are summarized in table \ref{tab:bursts}. For the background
+The data are summarized in Table \ref{tab:bursts}. For the background
 estimation, we also download observational data for 24 hours before
@@
 Figures \ref{fig:grb080916C}, \ref{fig:grb090510},
-\ref{fig:grb090902B}, \ref{fig:grb090926A} and table
+\ref{fig:grb090902B}, \ref{fig:grb090926A} and Table
 \ref{tab:observationResults}.
@@
             least from $100$ to $1000$} of $\gamma = 300$
-          \cite{Ghirlanda:2011ux}.}
+          \cite{Ghirlanda:2011bn}.}
 \end{itemize}
@@
           varying $z$ and $\chi$ (sampling is done the same way as in
-          section \ref{sec:distribution}). Note that this requirement
+          Section \ref{sec:distribution}). Note that this requirement
           unlike others applies to an ensemble of simulated bursts. }
@@
 		\label{item:sample}
-		Compute the small sample of 10 bursts (we will need stretching factors and total photon counts) with the same fixed burst parameters $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k, \alpha\right)$, and with $z$ and $\chi$ representatively distributed (as discussed in section \ref{sec:distribution}).
+		Compute the small sample of 10 bursts (we will need stretching factors and total photon counts) with the same fixed burst parameters $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k, \alpha\right)$, and with $z$ and $\chi$ representatively distributed (as discussed in Section \ref{sec:distribution}).
 	}
@@
 	\end{tabular}
-	\caption{Initial points used for minimization procedure from section \ref{sec:parameters}.}
+	\caption{Initial points used for minimization procedure from Section \ref{sec:parameters}.}
 	\label{tab:fitInitialPoints}
@@
 correspondingly), and to cover the large fraction of the parameter
-space. You can see the values in table \ref{tab:fitInitialPoints}.
+space. You can see the values in Table \ref{tab:fitInitialPoints}.
 
@@
           distribution does not contradict to the values of stretching
-          factors obtained in section \ref{sec:observations}.  }
+          factors obtained in Section \ref{sec:observations}.  }
 	\item{ The fraction of bursts observable in low energy band,
@@
 		{\bf Object name or coordinates}.
-		Coordinates are filled from the table \ref{tab:bursts}.
+		Coordinates are filled from the Table \ref{tab:bursts}.
 	}
@@
 		$60$.
-		Events are filtered by location separately by using the point spread functions (PSFs) of the LAT (as described in section \ref{sec:photonSelection}).
+		Events are filtered by location separately by using the point spread functions (PSFs) of the LAT (as described in Section \ref{sec:photonSelection}).
 	}
@@
 		{\bf Observation dates}.
-		To have a decent safety margin, we extend the duration of the burst by $50\%$ to both past and future relative to the table \ref{tab:bursts} time ranges.
+		To have a decent safety margin, we extend the duration of the burst by $50\%$ to both past and future relative to the Table \ref{tab:bursts} time ranges.
 		So, we fill in the following values:
@@
 
-\bibliographystyle{ieeetr}
+\bibliographystyle{hieeetr}
 \bibliography{gammaRays.bib}

### 2014-12-12 Paper/gammaRays.tex
Added zmin value to the text

@@
         on energy has not been studied for the high energy bands.  In
-        this paper we consider the bursts observed by Fermi-LAT from
+        this paper we consider the bursts observed by Fermi LAT from
         2008 August 4 to 2011 August 1, for which at least $10$
@@
 low-redshift burst is very small. So instead of selecting the smallest
-redshift to be $0$, we select it to be a small number $z_\text{min}$.
+redshift to be $0$, we select it to be a small number $z_\text{min} = 0.1$.
 

### 2014-12-12 Paper/gammaRays.tex
Corrected Fermi LAT to Fermi-LAT where appropriate

@@
 \title{
-	Time stretching of the GeV emission of GRBs: Fermi LAT data vs geometrical model
+	Time stretching of the GeV emission of GRBs: Fermi-LAT data vs geometrical model
 }
@@
         directions. The model conforms to the total burst energy
-        constraint and matches the Fermi LAT observations of the
+        constraint and matches the Fermi-LAT observations of the
         fraction of GRBs visible in $\unit[100]{MeV} < E <
@@
 
-\subsection{Fermi LAT Photon Selection}
+\subsection{Fermi-LAT Photon Selection}
 \label{sec:photonSelection}
@@
           emission. The beginning of the interval is used as a fixed
-          point of stretching. The Fermi LAT photons are downloaded
+          point of stretching. The Fermi-LAT photons are downloaded
           for the period extended in time by $50\%$ towards both past
@@
           which are also observed in high energy band is $f_m = 0.110$.
-          The Fermi LAT catalog contains $35$ bursts out of which $4$
+          The Fermi-LAT catalog contains $35$ bursts out of which $4$
           can be observed in high energy band.  Therefore the observed
@@
 \label{sec:fermiCode}
-This appendix contains the details on using the Fermi LAT
-data. Namely, the exact values passed to the web form on the Fermi LAT
+This appendix contains the details on using the Fermi-LAT
+data. Namely, the exact values passed to the web form on the Fermi-LAT
 data server and the code used to run the Fermi tools including the
@@
 
-Parameters that are passed to the web form of the Fermi LAT data server to download the bursts data are the following:
+Parameters that are passed to the web form of the Fermi-LAT data server to download the bursts data are the following:
 \begin{itemize}
@@
 \end{lstlisting}
-Here \texttt{[eventFile]} is the file downloaded from the Fermi LAT data server.
+Here \texttt{[eventFile]} is the file downloaded from the Fermi-LAT data server.
 
@@
 \end{lstlisting}
-where \texttt{[spacecraft]} is the FITS file containing the spacecraft data (also downloaded from the Fermi LAT data server).
+where \texttt{[spacecraft]} is the FITS file containing the spacecraft data (also downloaded from the Fermi-LAT data server).
 

### 2015-11-29 APJ/gammaRaysAPJ.tex
Proceedings editing

--- 2014-12-12 Paper/gammaRays.tex
@@
-\documentclass{article}
-
+%%
+%% Beginning of file 'sample.tex'
+%%
+%% Modified 2005 December 5
+%%
+%% This is a sample manuscript marked up using the
+%% AASTeX v5.x LaTeX 2e macros.
+
+%% The first piece of markup in an AASTeX v5.x document
+%% is the \documentclass command. LaTeX will ignore
+%% any data that comes before this command.
+
+%% The command below calls the preprint style
+%% which will produce a one-column, single-spaced document.
+%% Examples of commands for other substyles follow. Use
+%% whichever is most appropriate for your purposes.
+%%
+%%\documentclass[12pt,preprint]{aastex}
+
+%% manuscript produces a one-column, double-spaced document:
+
+\documentclass[manuscript]{aastex}
+
+\let\captionbox\relax
 \usepackage{commath}
 \usepackage{graphicx}
-\usepackage{fullpage}
+\usepackage{caption}
 \usepackage{subcaption}
@@
 \usepackage{hyperref}
-\usepackage{siunitx}
 \usepackage{listings}
@@
 \usepackage{bm}
-\usepackage{appendix}
-\usepackage{textcomp}
 \usepackage{gensymb}
-\usepackage{booktabs}
-
-\lstset{columns=fullflexible,basicstyle=\ttfamily}
-
-\allowdisplaybreaks
 
@@
 
-\title{
-	Time stretching of the GeV emission of GRBs: Fermi-LAT data vs geometrical model
-}
-
-\author{
-	Maxim S. Piskunov$^{1,2}$\thanks{\href{mailto:maxit@ms2.inr.ac.ru}{maxit@ms2.inr.ac.ru}} ~and
-	Grigory I. Rubtsov$^{2}$\thanks{\href{mailto:grisha@ms2.inr.ac.ru}{grisha@ms2.inr.ac.ru}} \\
-	{\small $^{1}$Faculty of Physics, Moscow State University, Moscow, Russia}\\
-	{\small $^{2}$Institute for Nuclear Research of the RAS, Moscow, Russia}
-}
+%% preprint2 produces a double-column, single-spaced document:
+
+%% \documentclass[preprint2]{aastex}
+
+%% Sometimes a paper's abstract is too long to fit on the
+%% title page in preprint2 mode. When that is the case,
+%% use the longabstract style option.
+
+%% \documentclass[preprint2,longabstract]{aastex}
+
+%% If you want to create your own macros, you can do so
+%% using \newcommand. Your macros should appear before
+%% the \begin{document} command.
+%%
+%% If you are submitting to a journal that translates manuscripts
+%% into SGML, you need to follow certain guidelines when preparing
+%% your macros. See the AASTeX v5.x Author Guide
+%% for information.
+
+% \newcommand{\vdag}{(v)^\dagger}
+% \newcommand{\myemail}{skywalker@galaxy.far.far.away}
+
+%% You can insert a short comment on the title page using the command below.
+
+% \slugcomment{Not to appear in Nonlearned J., 45.}
+
+%% If you wish, you may supply running head information, although
+%% this information may be modified by the editorial offices.
+%% The left head contains a list of authors,
+%% usually a maximum of three (otherwise use et al.).  The right
+%% head is a modified title of up to roughly 44 characters.
+%% Running heads will not print in the manuscript style.
+
+\shorttitle{Time stretching of the GeV emission of GRBs}
+\shortauthors{Piskunov and Rubtsov}
+
+%% This is the end of the preamble.  Indicate the beginning of the
+%% paper itself with \begin{document}.
 
 \begin{document}
-\begin{flushright}  
-INR-TH/2014-034
-\end{flushright}
-\vskip -0.9cm
-{\let\newpage\relax \maketitle}
+
+%% LaTeX will automatically break titles if they run longer than
+%% one line. However, you may use \\ to force a line break if
+%% you desire.
+
+\title{Time stretching of the GeV emission of GRBs: \\
+    Fermi-LAT data vs geometrical model}
+
+%% Use \author, \affil, and the \and command to format
+%% author and affiliation information.
+%% Note that \email has replaced the old \authoremail command
+%% from AASTeX v4.0. You can use \email to mark an email address
+%% anywhere in the paper, not just in the front matter.
+%% As in the title, use \\ to force line breaks.
+
+\author{Maxim S. Piskunov\altaffilmark{1,2} and Grigory I. Rubtsov\altaffilmark{2}}
+\affil{Faculty of Physics, Moscow State University, Moscow, Russia}
+\affil{Institute for Nuclear Research of the RAS, Moscow, Russia}
+
+\email{maxit@ms2.inr.ac.ru}
+\email{grisha@ms2.inr.ac.ru}
+
+%% Notice that each of these authors has alternate affiliations, which
+%% are identified by the \altaffilmark after each name.  Specify alternate
+%% affiliation information with \altaffiltext, with one command per each
+%% affiliation.
+
+\altaffiltext{1}{Faculty of Physics, Moscow State University, Moscow, Russia.}
+\altaffiltext{2}{Institute for Nuclear Research of the RAS, Moscow, Russia.}
+
+%% Mark off your abstract in the ``abstract'' environment. In the manuscript
+%% style, abstract will output a Received/Accepted line after the
+%% title and affiliation information. No date will appear since the author
+%% does not have this information. The dates will be filled in by the
+%% editorial office after submission.
 
@@
         photons were observed with the energy greater than
-        $\unit[1]{GeV}$. These include $4$ bursts: GRB 080916C, GRB
-        090510, GRB 090902B, and GRB 090926A.  We use the
+        $\unit[1]{GeV}$. These include $4$ bursts: \objectname{GRB 080916C}, \objectname{GRB 090510},
+        \objectname{GRB 090902B}, and \objectname{GRB 090926A}.  We use the
         Kolmogorov-Smirnov test to compare the light curves in the two
         bands, $\unit[100]{MeV} < E < \unit[1]{GeV}$ and
-        $E > \unit[1]{GeV}$. For GRB 080916C and GRB
-        090510 the light curves in the two bands are statistically
-        compatible. However, for GRB 090926A, the higher-energy light
+        $E > \unit[1]{GeV}$. For \objectname{GRB 080916C} and
+        \objectname{GRB 090510} the light curves in the two bands are statistically
+        compatible. However, for \objectname{GRB 090926A}, the higher-energy light
         curve is stretched compared to the lower-energy one with a
-        statistical significance of $3.3 \sigma$ and, for GRB 090902B,
+        statistical significance of $3.3 \sigma$ and, for \objectname{GRB 090902B},
         on the contrary, the lower-energy curve is stretched with $2.3
@@
 
+%% Keywords should appear after the \end{abstract} command. The uncommented
+%% example has been keyed in ApJ style. See the instructions to authors
+%% for the journal to which you are submitting your paper to determine
+%% what keyword punctuation is appropriate.
+
+\keywords{gamma-ray burst: general --- gamma-ray burst: individual(\objectname{GRB 080916C},
+\objectname{GRB 090510}, \objectname{GRB 090902B}, \objectname{GRB 090926A})}
+
+%% From the front matter, we move on to the body of the paper.
+%% In the first two sections, notice the use of the natbib \citep
+%% and \citet commands to identify citations.  The citations are
+%% tied to the reference list via symbolic KEYs. The KEY corresponds
+%% to the KEY in the \bibitem in the reference list below. We have
+%% chosen the first three characters of the first author's name plus
+%% the last two numeral of the year of publication as our KEY for
+%% each reference.
+
+%% Authors who wish to have the most important objects in their paper
+%% linked in the electronic edition to a data center may do so by tagging
+%% their objects with \objectname{} or \object{}.  Each macro takes the
+%% object name as its required argument. The optional, square-bracket 
+%% argument should be used in cases where the data center identification
+%% differs from what is to be printed in the paper.  The text appearing 
+%% in curly braces is what will appear in print in the published paper. 
+%% If the object name is recognized by the data centers, it will be linked
+%% in the electronic edition to the object data available at the data centers  
+%%
+%% Note that for sources with brackets in their names, e.g. [WEG2004] 14h-090,
+%% the brackets must be escaped with backslashes when used in the first
+%% square-bracket argument, for instance, \object[\[WEG2004\] 14h-090]{90}).
+%%  Otherwise, LaTeX will issue an error.
+
 \section{Introduction and summary}
@@
 Universe. An extensive studies of these explosions led to a number of
-interesting results, see~\cite{Vianello:2013ela,Gehrels:2013xd} for a
+interesting results, see~\citet{Vianello:2013ela,Gehrels:2013xd} for a
 review. In particular, the total energy emitted in gamma-rays during a
 burst was found to be similar for the different GRBs within an order
-of magnitude~\cite{Bloom:2003wy}. This indicates that the most of the
+of magnitude~\citep{Bloom:2003wy}. This indicates that the most of the
 bursts have similar energetics in their rest frames. Moreover, the
 high energy radiation of GRBs is shown to follow the predictions of
-the synchrotron radiation scenario~\cite{Wang:2013ptaw}.  Several
+the synchrotron radiation scenario~\citep{Wang:2013ptaw}.  Several
 observations are related to the temporal variations of spectra. This
 way, the spectral lags were found between different low energy bands
-\cite{Yi:2005ht} and the very-high-energy radiation was discovered to
+\citep{Yi:2005ht} and the very-high-energy radiation was discovered to
 be extended relative to the x-ray emission
-\cite{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}. In this paper
+\citep{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}. In this paper
 we go further and compare the temporal extension of GRB gamma-ray
@@
 bands) using the data from the Large Area Telescope (LAT) of the Fermi
-satellite~\cite{2009ApJ...697.1071A,Ackermann:2012kna}.
+satellite~\citep{2009ApJ...697.1071A,Ackermann:2012kna}.
 
 We start with the location, time and duration of bursts from the
-Fermi-LAT Gamma-Ray Burst Catalog \cite{Ackermann:2013zfa}. The
+Fermi-LAT Gamma-Ray Burst Catalog \citep{Ackermann:2013zfa}. The
 photons during the burst and the photons for 24 hours before the burst
@@
 The latter are used for the background estimation in both energy bands
-with the technique introduced in \cite{Rubtsov:2011qq}. In order to
+with the technique introduced in \citet{Rubtsov:2011qq}. In order to
 have enough statistics we require that at least 10 photons are
@@
 with the four bursts for the time span of the Fermi-LAT Gamma-Ray
-Burst Catalog. Namely, GRB~080916C \cite{Tajima:2009az}, GRB~090510
-\cite{Ackermann:2010us}, GRB~090902B \cite{Abdo:2009pg} and
-GRB~090926A \cite{Bregeon:2011bu}. For these bursts we stretch the
+Burst Catalog. Namely, \objectname{GRB 080916C} \citep{Tajima:2009az}, \objectname{GRB 090510}
+\citep{Ackermann:2010us}, \objectname{GRB 090902B} \citep{Abdo:2009pg} and
+\objectname{GRB 090926A} \citep{Bregeon:2011bu}. For these bursts we stretch the
 high energy light curve by the arbitrary factor and compare with the
@@
 the $95\%$ confidence level the stretching factor is compatible to $1$
-(no stretching) for the GRB 080916C and GRB 090510. For the other two
-bursts the stretching factor of 1 is, however, excluded.  For GRB
-090902B it should be smaller than 1, so the low energy light curve is
+(no stretching) for the \objectname{GRB 080916C} and \objectname{GRB 090510}. For the other two
+bursts the stretching factor of 1 is, however, excluded.  For \objectname{GRB 090902B}
+it should be smaller than 1, so the low energy light curve is
 stretched with respect to the high energy one, see
@@
 stretching factor of 1 in the opposite direction is observed for the
-GRB 090926A. The high energy light curve of GRB 090926A is stretched
+\objectname{GRB 090926A}. The high energy light curve of \objectname{GRB 090926A} is stretched
 with respect to the low energy one by a factor of at least 1.99 (see
-fig. \ref{fig:grb090926A}).  The stretching of the GRB 090926A high
+fig. \ref{fig:grb090926A}).  The stretching of the \objectname{GRB 090926A} high
 energy light is found with $3.3\sigma$ significance.  All the observed
@@
 called curvature effects. These effects were explored by multiple
-authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.  These
+authors \citep{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.  These
 studies concerned x-ray radiation, and considered the homogeneous
@@
 shown in Figure \ref{fig:sampleLightCurves}. The detailed calculations
-show that the stretching factors of GRB 090902B and GRB 090926A appear
+show that the stretching factors of \objectname{GRB 090902B} and \objectname{GRB 090926A} appear
 naturally in the model. The predicted distribution of stretching
@@
 	\small
-	\begin{tabular}{ l | S[table-format=9.3] | S[table-format=3.3] | S[table-format=-2.3] | S[table-format=1.5] | S[table-format=2.1] | S[table-format=3.1] }
+	\begin{tabular}{ l | c | c | c | c | c | c }
 		\multicolumn{1}{ c |}{GRB} & $\mathrm{GBM\ Trigger\ Time}$ & $\mathrm{R.A.}$ & $\mathrm{Dec.}$ & $\mathrm{Location\ Error}$ & $T_{05}$ & $T_{95}$ \\
 		\multicolumn{1}{ c |}{Name}& $\mathrm{MET,\ sec}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{deg}$ & $\mathrm{sec}$ & $\mathrm{sec}$ \\
-		\multicolumn{1}{ c |}{\texttt{name}} & $\mathrm{\texttt{time}}$ & \texttt{location.ra} & \texttt{location.dec} & \texttt{location.err} & \texttt{startOffset} & \texttt{endOffset} \\
-		\hline
+		\tableline
 		080825C	&	241366429.105	&	233.9  	&	 -4.5  	&	0.75   	&	 3.2	&	 29.4	\\
@@
 	\end{tabular}
-	\caption{Burst data used in our study taken from~\cite{Ackermann:2013zfa}. The third row contains
-          the corresponding \texttt{GRBurst} class variable names of
-          the software package \url{https://github.com/maxitg/GammaRays/blob/master/GRObservations/bursts}}
+	\caption{Burst data used in our study taken from~\citet{Ackermann:2013zfa}.}
 	\label{tab:bursts}
@@
 the Fermi-LAT data
-server~\cite{2009ApJ...697.1071A,Ackermann:2012kna}. In order to
+server~\citep{2009ApJ...697.1071A,Ackermann:2012kna}. In order to
 define the regions of interest we use the catalog
-\cite{Ackermann:2013zfa}, which contains all the bright gamma-ray
+\citep{Ackermann:2013zfa}, which contains all the bright gamma-ray
 bursts seen by the Fermi LAT since 2008 August 4 to 2011 August 1.
 Specifically, we take 4 pieces of information from the tables 2 and 4
-of Ref. \cite{Ackermann:2013zfa}:
+of Ref. \citet{Ackermann:2013zfa}:
 \begin{itemize}
@@
 locations. Knowing the exposures, we use the method introduced in the
-appendix of \cite{Rubtsov:2011qq} to estimate the background for low
+appendix of \citet{Rubtsov:2011qq} to estimate the background for low
 and high-energy bands. Note, that the ``TRANSIENT'' class photons are
@@
 Before applying the KS-test we subtract the background, which is
-considered linear for the duration of the burst :
+considered linear for the duration of the burst:
 
@@
 
-Out of these 4 bursts, two (GRB 080916C and GRB 090510) have
+Out of these 4 bursts, two (\objectname{GRB 080916C} and \objectname{GRB 090510}) have
 stretching factors compatible with $\kappa = 1$ within $2\sigma$
@@
 
-GRB 090926A has, however, high energy radiation stretched with respect
+\objectname{GRB 090926A} has, however, high energy radiation stretched with respect
 to low energy radiation (that is $\kappa > 1$). In contrast to that,
-GRB 090902B has low energy radiation stretched with respect to high
+\objectname{GRB 090902B} has low energy radiation stretched with respect to high
 energy radiation ($\kappa < 1$).
@@
 	\small
-	\begin{tabular}{ l | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] | S[table-format=2.3]@{\; -- \;}S[table-format=2.3] }
+	\begin{tabular}{ l | c@{\; -- \;}c | c@{\; -- \;}c | c@{\; -- \;}c | c@{\; -- \;}c | c@{\; -- \;}c }
 		\multicolumn{1}{ c |}{GRB} & \multicolumn{2}{ c |}{$1\sigma$} & \multicolumn{2}{ c |}{$2\sigma$} & \multicolumn{2}{ c |}{$3\sigma$} & \multicolumn{2}{ c |}{$4\sigma$} & \multicolumn{2}{ c }{$5\sigma$} \\
@@
 are in the LAT field of view at the moment of observation
-\cite{Vianello:2013ela}.  However, only about 30 of them were detected
+\citep{Vianello:2013ela}.  However, only about 30 of them were detected
 by the LAT, and only 4 of them were bright in the high energy band.
@@
 not.  These differences are hard to explain given that the burst
-energetics are similar \cite{Bloom:2003wy}.  These differences in
+energetics are similar \citep{Bloom:2003wy}.  These differences in
 burst counts find natural explanation in our model, in which the
@@
 \begin{align}
-p\left( z, \chi; \tau_1, \tau_2; \omega_1, \omega_2 \right) &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)} \int_0^\infty \dif r \int_0^{\pi} r \dif \theta \int_0^{2\pi} r \sin{\theta} \dif \phi \int_0^\infty \dif \omega' \int_{\tau_1}^{\tau_2} \dif \tau \int_{\omega_1}^{\omega_2} \dif \omega \, \eta\left( r, \theta, \omega' \right) \nonumber\\
+p &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)} \int_0^\infty \dif r \int_0^{\pi} r \dif \theta \int_0^{2\pi} r \sin{\theta} \dif \phi \int_0^\infty \dif \omega' \int_{\tau_1}^{\tau_2} \dif \tau \int_{\omega_1}^{\omega_2} \dif \omega \, \eta\left( r, \theta, \omega' \right) \nonumber\\
 &\qquad {} \times \underbrace{\frac{1}{\gamma^2\left( 1 - v \cos\sigma\left(\theta, \phi, \chi\right) \right)^2}}_{\text{aberration}} \underbrace{\frac{1}{\gamma\left( 1 - v \cos\sigma\left(\theta, \phi, \chi\right) \right)}}_{\text{time dilation}} \nonumber\\
@@
 \begin{align}
-p\left( z, \chi; \tau_1, \tau_2; \omega_1, \omega_2 \right) &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)} \int_0^{\pi} \dif \theta \int_0^{2\pi} \dif \phi \int_{\tau_1}^{\tau_2} \dif \tau \int_{\omega_1}^{\omega_2} \dif \omega \nonumber\\
+p &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)} \int_0^{\pi} \dif \theta \int_0^{2\pi} \dif \phi \int_{\tau_1}^{\tau_2} \dif \tau \int_{\omega_1}^{\omega_2} \dif \omega \nonumber\\
 &\qquad {} \times \eta\left( \frac{\tau}{\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)\left(1+z\right)}, \theta, \gamma\left( 1 - v\cos\sigma\left(\theta, \phi, \chi\right) \right)\left(1+z\right)\omega  \right) \nonumber\\
@@
 \begin{align}
-p\left( z, \chi; \tau_1, \tau_2; \omega_1, \omega_2 \right) &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)} \frac{1}{v^2 \gamma^2 \left(1+z\right)^2} \int_0^{\pi} \dif \theta \int_0^{2\pi} \dif \phi \int_{\omega_1}^{\omega_2} \dif \omega \int_{\tau_1}^{\tau_2} \dif \tau \, \tau^2 \frac{\sin\theta}{\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)^5} \nonumber\\
+p &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)} \frac{1}{v^2 \gamma^2 \left(1+z\right)^2} \int_0^{\pi} \dif \theta \int_0^{2\pi} \dif \phi \int_{\omega_1}^{\omega_2} \dif \omega \int_{\tau_1}^{\tau_2} \dif \tau \, \tau^2 \frac{\sin\theta}{\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)^5} \nonumber\\
 &\qquad {} \times \frac{\eta_0}{1 + \left(\frac{\tau}{r_0\left(\frac{1}{v} - \cos\sigma\left(\theta, \phi, \chi\right)\right)\left(1+z\right)}\right)^n} \nonumber\\
@@
 \label{eq:p}
-p\left( z, \chi; \tau_1, \tau_2; \omega_1, \omega_2 \right) &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)}
+p &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)}
 \frac{2\eta_0}{\left(v\gamma\left(1+z\right)\right)^{2-\alpha}}
@@
 &\qquad {} \times \left( I\left(z,\chi,\omega_2; \theta,\phi\right) - I\left(z,\chi,\omega_1; \theta,\phi\right) \right) \left( J\left(z,\chi,\tau_2; \theta,\phi\right) - J\left( z,\chi,\tau_1; \theta,\phi \right) \right) \\
-I\left(z,\chi,\omega;\theta,\phi\right) &= \frac{
+I &= \frac{
 	\omega
@@
 } \\
-J\left(z,\chi,\tau;\theta,\phi\right) &= \frac{\tau^3}{3} \,
+J &= \frac{\tau^3}{3} \,
 _2F_1\left(
@@
 \begin{equation}
-J\left(z,\chi,\infty;\theta,\phi\right) = \lim_{\tau \rightarrow \infty} J\left(z,\chi,\tau;\theta,\phi\right) = \left(r_0 \left( \frac{1}{v} -  1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi \right) \left(1+z\right)\right)^3 \frac{\pi}{n\sin\frac{3\pi}{n}}
+J = \lim_{\tau \rightarrow \infty} J\left(z,\chi,\tau;\theta,\phi\right) = \left(r_0 \left( \frac{1}{v} -  1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi \right) \left(1+z\right)\right)^3 \frac{\pi}{n\sin\frac{3\pi}{n}}
 \end{equation}
@@
 \begin{align*}
-E = \frac{32\pi^2 \eta_0}{3}\left(\gamma^2 - \frac{1}{4}\right) \int_0^\infty \dif \omega \, \omega \left(\frac{\omega}{\omega_0}\right)^\alpha \int_0^\infty \dif r \, \frac{r^2}{1 + \left(\frac{r}{r_0}\right)^n} \int_0^\infty \dif \theta \, \theta \, \exp\left(-\left(\frac{\theta}{\theta_0}\right)^2\left(\frac{\omega}{\omega_0}\right)^{-2k}\right)
+E &= \frac{32\pi^2 \eta_0}{3}\left(\gamma^2 - \frac{1}{4}\right) \\
+&\times \int_0^\infty \dif \omega \, \omega \left(\frac{\omega}{\omega_0}\right)^\alpha \int_0^\infty \dif r \, \frac{r^2}{1 + \left(\frac{r}{r_0}\right)^n} \int_0^\infty \dif \theta \, \theta \, \exp\left(-\left(\frac{\theta}{\theta_0}\right)^2\left(\frac{\omega}{\omega_0}\right)^{-2k}\right)
 \end{align*}
@@
 b\left(\omega_1,\omega_2\right) &= \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} 2\pi \sin\chi \dif\chi \nonumber\\
-&\approx 2\pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} \chi \dif\chi = \pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_2\right)\,,
+&\approx 2\pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \int_0^{\chi_\text{max}\left(z,\omega_1,\omega_2\right)} \chi \dif\chi \nonumber\\
+&= \pi \int_0^{z_\text{max}\left(\omega_1,\omega_2\right)} \dif V\left(z\right) \rho\left(z\right) \chi^2_\text{max}\left(z,\omega_1,\omega_2\right)\,,
 \end{align}
@@
 	\item{Redshifts $z$ of many bursts have been measured
-          \cite{Ackermann:2013zfa}.}
+          \citep{Ackermann:2013zfa}.}
 	\item{We know that relativistic factors $\gamma$ are on the
@@
             least from $100$ to $1000$} of $\gamma = 300$
-          \cite{Ghirlanda:2011bn}.}
+          \citep{Ghirlanda:2011bn}.}
 \end{itemize}
@@
             10^{53}]{GeV} \approx \unit[10^{51}]{erg}$
-          \cite{Gehrels:2013xd}.  }
+          \citep{Gehrels:2013xd}.  }
 	\item { The stretching factor of the burst should be
-          compatible with the value for the GRB 090926A.  }
+          compatible with the value for the \objectname{GRB 090926A}.  }
 	\item { The ratio of photons with high energies and low
@@
           \theta_0, k, \alpha\right)$ and require that bursts with
-          small stretching factors like GRB 090902B should be possible
+          small stretching factors like \objectname{GRB 090902B} should be possible
           to observe. So we require that bursts with the stretching
-          factor of GRB 090902B or lower appear in random samples by
+          factor of \objectname{GRB 090902B} or lower appear in random samples by
           varying $z$ and $\chi$ (sampling is done the same way as in
@@
 	\item{
-		Set $\gamma = 300$, $\omega_0 = \unit[1]{GeV}$, $z = 2.1062$ (which is the redshift of GRB 090926A)
-	}
-	\item{
-		Set $\eta_0$ and $r_0$ such that the duration of the burst, and the total number of observed photons are compatible with GRB 090926A, $T_{0.99}\left(z,\chi;\unit[0.1]{GeV},\infty\right) = \unit[219.5]{sec}$ and $p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\infty\right) = 179.996$
+		Set $\gamma = 300$, $\omega_0 = \unit[1]{GeV}$, $z = 2.1062$ (which is the redshift of \objectname{GRB 090926A})
+	}
+	\item{
+		Set $\eta_0$ and $r_0$ such that the duration of the burst, and the total number of observed photons are compatible with \objectname{GRB 090926A}, $T_{0.99}\left(z,\chi;\unit[0.1]{GeV},\infty\right) = \unit[219.5]{sec}$ and $p_{\infty} \left(z,\chi;\unit[0.1]{GeV},\infty\right) = 179.996$
 	}
@@
 	\small
-	\begin{tabular}{ S[table-format=2.4] | S[table-format=1.5] | S[table-format=-1.6] | S[table-format=-1.7] | S[table-format=1.8] }
+	\begin{tabular}{ c | c | c | c | c }
 		$\mathrm{n}$ & $\mathrm{\theta_0}$ & $\mathrm{k}$ & $\mathrm{\alpha}$ & $\mathrm{\chi}$ \\
@@
 As was discussed in the previous section our model can reproduce both
-stretching factors smaller than $1$ (like of GRB 090902B) and larger
-than $1$ (like of GRB 090926A) depending on redshift and observer's
+stretching factors smaller than $1$ (like of \objectname{GRB 090902B}) and larger
+than $1$ (like of \objectname{GRB 090926A}) depending on redshift and observer's
 off-axis angle (see fig. \ref{fig:sampleLightCurves}).
@@
           energetic gamma rays $E < \unit[5.89 \times 10^{53}]{GeV}$,
-          which is in agreement with \cite{Gehrels:2013xd}.  }
+          which is in agreement with \citet{Gehrels:2013xd}.  }
 	\item{ The distribution of observable bursts' stretching
@@
 model. First of all, as it's seen in Fig. \ref{fig:correlations} the
-stretching factor of GRB 090926A is higher that the model
+stretching factor of \objectname{GRB 090926A} is higher that the model
 prediction. More data are needed to understand whether this requires a
@@
 might account for differences in light curves of GRBs (or, even, for
-especially large stretching factor of GRB 090926A).  This means that
+especially large stretching factor of \objectname{GRB 090926A}).  This means that
 the minor variations of the model parameters might be necessary.
@@
 
+\acknowledgments
+
 {\bf Acknowledgments.} We thank A.~Gruzinov, M.~Pshirkov and
@@
 
-\begin{appendices}
+{\it Facilities:} \facility{Fermi}.
+
+\appendix
 
@@
 gtselect infile=[eventFile] outfile=filtered.fits
-	ra=INDEF dec=INDEF rad=180 tmin=INDEF tmax=INDEF emin=100 emax=300000
+	ra=INDEF dec=INDEF rad=180 tmin=INDEF tmax=INDEF
+	emin=100 emax=300000
 	zmax=100 evclass=2 convtype=-1 evtable=EVENTS
@@
 \begin{lstlisting}
-gtltcube evfile=timed.fits evtable=EVENTS scfile=[spacecraft] sctable=SC_DATA
+gtltcube evfile=timed.fits evtable=EVENTS
+	scfile=[spacecraft] sctable=SC_DATA
 	outfile=ltcube.fits
-	dcostheta=0.025 binsz=1 phibins=0 tmin=0 tmax=0 zmax=100 zmin=0
+	dcostheta=0.025 binsz=1 phibins=0
+	tmin=0 tmax=0 zmax=100 zmin=0
 \end{lstlisting}
@@
 \begin{lstlisting}
-gtpsf expcube=ltcube.fits outfile=psf_[IrfName].fits outtable=PSF irfs=[IrfName]
-	ra=[RA] dec=[DEC] emin=100 emax=300000 nenergies=41 thetamax=30 ntheta=300
+gtpsf expcube=ltcube.fits outfile=psf_[IrfName].fits
+	outtable=PSF irfs=[IrfName]
+	ra=[RA] dec=[DEC] emin=100 emax=300000
+	nenergies=41 thetamax=30 ntheta=300
 \end{lstlisting}
@@
 \begin{lstlisting}
-gtexpcube2 infile=ltcube.fits cmap=none outfile=expcube_[IrfName].fits irfs=[IrfName]
-	nxpix=360 nypix=180 binsz=1 coordsys=CEL xref=0 yref=0 axisrot=0
-	proj=CAR ebinalg=log emin=100 emax=300000 enumbins=40
-	ebinfile=NONE bincalc=EDGE ignorephi=no thmax=180 thmin=0 table=Exposure
+gtexpcube2 infile=ltcube.fits cmap=none
+	outfile=expcube_[IrfName].fits irfs=[IrfName]
+	nxpix=360 nypix=180 binsz=1 coordsys=CEL
+	xref=0 yref=0 axisrot=0 proj=CAR
+	ebinalg=log emin=100 emax=300000 enumbins=40
+	ebinfile=NONE bincalc=EDGE ignorephi=no
+	thmax=180 thmin=0 table=Exposure
 \end{lstlisting}
@@
 
-\end{appendices}
-
-\bibliographystyle{hieeetr}
-\bibliography{gammaRays.bib}
+\begin{thebibliography}{}
+\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
+
+\bibitem[Abdo et al.(2009)]{Abdo:2009pg}
+Abdo, A., {et~al.} 2009, \apj., 706, L138
+
+\bibitem[Ackermann et al.(2013)]{Ackermann:2013zfa}
+Ackermann, M., Ajello, M., Asano, K., {et~al.} 2013, \apjs, 209,
+  11
+
+\bibitem[Ackermann et al.(2010)]{Ackermann:2010us}
+Ackermann, M., {et~al.} 2010, \apj, 716, 1178
+
+\bibitem[Ackermann et al.(2011)]{Bregeon:2011bu}
+---. 2011, \apj, 729, 114
+
+\bibitem[Ackermann et al.(2012)]{Ackermann:2012kna}
+---. 2012, \apjs, 203, 4
+
+\bibitem[Atwood et al.(2009)]{2009ApJ...697.1071A}
+{Atwood}, W.~B., {Abdo}, A.~A., {Ackermann}, M., {et~al.} 2009, \apj,
+  697, 1071
+
+\bibitem[Bloom et al.(2003)]{Bloom:2003wy}
+Bloom, J.~S., Frail, D., \& Kulkarni, S. 2003, \apj, 594, 674
+
+\bibitem[Castignani et al.(2014)]{Castignani:2014gaa}
+Castignani, G., Guetta, D., Pian, E., {et~al.} 2014, \aap, 565,
+  A60
+
+\bibitem[Gehrels \& Razzaque(2013)]{Gehrels:2013xd}
+Gehrels, N., \& Razzaque, S. 2013, Front.Phys.China., 8, 661
+
+\bibitem[Ghirlanda et al.(2012)]{Ghirlanda:2011bn}
+Ghirlanda, G., Nava, L., Ghisellini, G., {et~al.} 2012,
+  \mnras, 420, 483
+
+\bibitem[Lange \& Pohl(2013)]{Lange:2013uh}
+Lange, J., \& Pohl, M. 2013, \aap, 551, A89
+
+\bibitem[Nakamura \& Ioka(2001)]{Nakamura:2001kd}
+Nakamura, T., \& Ioka, K. 2001, \apj, 554, L163
+
+\bibitem[Rubtsov et al.(2012)]{Rubtsov:2011qq}
+Rubtsov, G., Pshirkov, M., \& Tinyakov, P. 2012, Mon.Not.Roy.Astron.Soc.Lett.,
+  421, L14
+
+\bibitem[Shen et al.(2005)]{Shen:2005ea}
+Shen, R.-F., Song, L.-M., \& Li, Z. 2005, \mnras, 362, 59
+
+\bibitem[Shenoy et al.(2013)]{Shenoy:2013cba}
+Shenoy, A., Sonbas, E., Dermer, C., {et~al.} 2013, \apj, 778, 3
+
+\bibitem[Tajima(2009)]{Tajima:2009az}
+Tajima, H. 2009, arXiv:0907.0714
+
+\bibitem[{Vianello(2012)}]{Vianello:2013ela}
+Vianello, G. 2012, eConf, C121028, 325
+
+\bibitem[{Wang {et~al.}(ApJ)Wang, Liu, \& Lemoine}]{Wang:2013ptaw}
+Wang, X.-Y., Liu, R.-Y., \& Lemoine, M. \apj, 2013, 771, L33
+
+\bibitem[{Yi {et~al.}(2006)Yi, Liang, Qin, \& Lu}]{Yi:2005ht}
+Yi, T.-F., Liang, E.-W., Qin, Y.-P., \& Lu, R.-J. 2006,
+  \mnras, 367, 1751
+
+\end{thebibliography}
 

### 2015-11-29 ITEPProceedings/piskunov.tex
Added acknowledgments
Proceedings editing

@@
 
-\title{TIME STRETCHING OF THE GEV EMISSION OF GRBS: FERMI LAT DATA VS GEOMETRICAL MODEL}
+\title{TIME STRETCHING OF THE GEV EMISSION OF GRBS: FERMI-LAT DATA VS GEOMETRICAL MODEL}
 
@@
 
-	In this paper we consider all the bursts observed by Fermi-LAT since 2008 August 4 to 2011 August 1, for which at least $10$ events with energies $\unit[1]{GeV}$ or higher were observed.
+	In this paper we consider all the bursts observed by Fermi Large Area Telescope (LAT) since 2008 August 4 to 2011 August 1, for which at least $10$ events with energies $\unit[1]{GeV}$ or higher were observed.
 	There are $4$ of them: 080916C, 090510, 090902B, and 090926A.
@@
 	In this study we continue this research and explore the spectral variations between energy bands $(\unit[100]{MeV}, \unit[1]{GeV})$ and $(\unit[1]{GeV}, \unit[300]{GeV})$ (we call them low and high energy bands throughout the paper).
-	We use the Kolmogorov-Smirnov test to compute the time stretching of radiation in one of these bands compared to the other.
+	We use the Kolmogorov-Smirnov (KS) test to compute the time stretching of radiation in one of these bands compared to the other.
 	Results obtained with this approach are discussed in section \ref{sec:data}.
@@
 
-	For these 4 bursts we compute both high and low energy distributions of photon times.
-	We subtract the low energy background estimate from the low energy CDF, and add the high energy background estimate to it.
-	This makes the low energy CDF non-monotonous and, rigorously speaking, we cannot use the 2-sample Kolmogorov-Smirnov test on it.
-	However, since there is a much greater number of photons on lower energies, we can think of the low energy CDF as continuous.
-	Therefore, this non-monotonicity will be negligible, and will not harm the KS-test results much.
+	For these 4 bursts we compute both high and low energy distributions of photon observation times, that is the distributions of time intervals from the burst start to the photon detection in observer's reference frame, in which burst starts are taken to be the beginning of the LAT interval from the Table 4 of \cite{Ackermann:2013zfa}.
 
-	Finally, we stretch the high energy CDF by different factors, and compare it to the low energy CDF using the KS-test.
+	Then we subtract the background, which is considered linear for the duration of the burst (light curves on figures \ref{fig:grb090902B} and \ref{fig:grb090926A} are presented with background subtracted):
+	\begin{equation}
+		\Phi_i\left(t\right) = \frac{p_i\left(T_1, t\right) - b_i \frac{t-T_1}{T_2-T_1}}{p_i\left(T_1, T_2\right) - b_i}
+	\end{equation}
+	Here $\left(T_1, T_2\right)$ is the time range of observations, $p_i\left(t_1, t_2\right)$ is the number of photons observed in the time range $t_1$ to $t_2$ in the $i$'s energy band ($i$ is either low or high for the low and high-energy bands respectively), and $b_i$ is the estimated number of background photons for the time range $T_1$ to $T_2$ in the $i$'s energy band.
+The background to signal ratio $\frac{b_i}{p_i\left(T_1, T_2\right)}$ is the largest for GRB 090902B and equals $6.5\%$.
+
+	The number of degrees of freedom used as an input for the KS-test is $p_i\left(T_1, T_2\right) - b_i$.
+This is lower than the number of independent photons and therefore we are conservative when estimating the probability and significance of the effect.
+The KS-probabilities are computed for pairs $\Phi_\text{low}\left(t\right), \Phi_\text{high}\left(\kappa t\right)$ to obtain the allowed ranges for stretching factors $\kappa$ for a given confidence level.
+
+	Finally, we stretch the high energy cumulative distribution function (CDF) by different factors, and compare it to the low energy CDF using the KS-test.
 	If we require $2\sigma$-significant probability to exclude a particular stretching factor, then the stretching factor of $1$ (which means no stretching) is allowed for GRBs 080916C and 090510.
@@
 
-	First of all, there are around 750 GRBs detected by GBM, half of which were in the LAT field of view at the moment of observation \cite{Vianello:2013ela}.
+	First of all, there are around 750 GRBs detected by Gamma-ray Burst Monitor (GBM), half of which were in the LAT field of view at the moment of observation \cite{Vianello:2013ela}.
 	However, only about 30 of them were detected by the LAT, and only 4 of them were bright in the high energy band.
-	If we extrapolate the uniform jet model to very high energies, this observation would mean that these groups of bursts are internally different: some of them produce VHE radiation, and other do not.
+	If we extrapolate the uniform jet model to very high energies, this observation would mean that these groups of bursts are internally different: some of them produce very high energy (VHE) radiation, and other do not.
 	These differences are hard to explain given that burst energetics are similar \cite{Bloom:2003wy}.
@@
 		\item{
-			Time $t = 0$, a spherical shell of plasma is emitted.
+			Time $t = 0$, a spherical (yet anisotropic) shell of plasma is emitted.
 			The center is called the central engine.
 		}
-		\item{The shell points propagate with a constant velocity $v = \frac{\sqrt{\gamma^2-1}}{\gamma} \sim 1$, so at the time $t$ the radius of the shall is $v t$.}
+		\item{The shell points propagate with a constant velocity $v = \frac{\sqrt{\gamma^2-1}}{\gamma} \sim 1$, so at the time $t$ the radius of the shell is $v t$.}
 		\item{Each point of the shell is an isotropic radiator in its rest frame.}
@@
 			The radiation intensity is a function of the radiator position and the radiation frequency:
-			\begin{equation*}
+			\begin{equation}
 				\eta\left(r,\theta,\omega\right) = 
@@
 				\left(\frac{\omega}{\omega_0}\right)^\alpha
-			\end{equation*}
+			\end{equation}
 			$\eta$ is the number of particles emitted per volume per solid angle per frequency.
@@
 
-	The burst is fully specified by the following set of parameters (we assume these parameters to be the same for all bursts, the specified values are obtained by fitting to make possible observed bursts from section \ref{sec:data}):
+	In order to compute parameters of the radiation intensity function (which are assumed to be the same for all bursts), we fit them to minimize the cost function designed to satisfy the following objectives:
+	\begin{itemize}
+		\item {
+			Total energy of the burst is finite (that is $k+\frac{\alpha}{2}+1 < 0$) and smaller than $\unit[6 \times 10^{53}]{GeV} \approx \unit[10^{51}]{erg}$ \cite{Gehrels:2013xd}.
+			Note that this energy is the same for all bursts because they are assumed to be the same in their reference frames.
+		}
+		\item {
+			The stretching factor of the burst should be compatible with the value for the GRB 090926A.
+		}
+		\item {
+			The ratio of photons with high energies and low energies should be compatible with the value $0.057$ for GRB 090926A.
+		}
+		\item {
+			The burst should not be too faint compared to other bursts from the sample.
+			In other words, the total number of observed photons should approximately equal to the median among the sample.
+		}
+		\item {
+			We assume that all bursts have the same burst parameters $\left(\gamma, \eta_0, r_0, n, \omega_0, \theta_0, k, \alpha\right)$ and require that bursts with small stretching factors like GRB 090902B should be possible to observe.
+			So we require that bursts with the stretching factor of GRB 090902B or lower appear in random samples by varying $z$ and $\chi$.
+			Note that this requirement unlike others applies to an ensemble of simulated bursts.
+		}
+	\end{itemize}
+
+	The minimization converged to the following parameter values:
 	\begin{itemize}
@@
 
-	To test the model against observations, we computed a few more things:
+	To test the model against observations, we computed a few properties of the burst sample obtained by selecting redshifts and off-axis angles randomly (respecting the actual spacial distribution of GRBs):
 	\begin{itemize}
 		\item{
-			Total energy emitted in $\unit[100]{MeV}$ and more energetic gamma rays $E < \unit[5.89 \times 10^{53}]{GeV}$, which is in agreement with \cite{Gehrels:2013xd}.
-		}
-		\item{
 			Distribution of stretching factors of observable bursts.
-			You can see the PDF and the CDF of it on fig. \ref{fig:kappaDistribution}.
+			Only those bursts are included in this distribution, for which at least $10$ photons with energies $\unit[1]{GeV}$ or higher are observed.
+			You can see the PDF and the CDF on fig. \ref{fig:kappaDistribution}.
 			This distribution doesn't contradict with the values of stretching factors obtained in section \ref{sec:data}.
@@
 		\item{
-			The fraction of bursts observable in low energy band, which are also observed in high energy band $f_m = 0.116$.
-			The Fermi LAT catalog contains $35$ bursts out of which $4$ can be observed in high energy band.
+			The fraction of bursts observable (that is, at least $10$ particles are observed in a given energy range) in low energy band, which are also observed in high energy band $f_m = 0.116$.
+			The Fermi LAT catalog \cite{Ackermann:2013zfa} contains $35$ bursts out of which $4$ can be observed in high energy band.
 			Therefore the observed value $f_o = 0.11 \pm 0.05$ (error obtained from binomial distribution) agrees with the model.
@@
 
-	So our model, though with some discrepancies, is able to explain the time stretching result from section \ref{sec:data}.
+	So our model, though with some discrepancies (listed in the next section), is able to explain the time stretching result from section \ref{sec:data}.
 
@@
 		\item{
-			The time stretching of GRB light curves between different VHE bands is confirmed with statistical significance of $3.3\sigma$.
+			The time stretching of GRB light curves between different VHE bands is confirmed with statistical significance of $3.3\sigma$ (that is the probability of no stretching is $9.7 \times 10^{-4}$).
 		}
@@
 		\item{
-			All GRBs are assumed to be the same in their rest frames.
-			There is no need to vary internal burst parameters.
+			Stretching factors both larger and smaller than one can be reproduced without varying internal burst parameters (that is assuming all bursts to be the same in their reference frames).
 		}
@@
 		\item{
-			Assumption of energy dependence on off-axis angle should be better justified.
+			Assumption of energy dependence on off-axis angle should be better justified (with a thermodynamic simulation of jet plasma, for example).
 		}
@@
 
-	We believe, however, that these problems are not unresolvable, and may be addressed by refining the model without changing its main assumptions.
-	And in the end, we think the main ideas of this paper will be close to truth.
+	We think, however, that these problems are not unresolvable, and may be addressed by refining the model without changing its main assumptions.
 
-\bibliographystyle{maik}
+	{\small {\it Acknowledgments.} The work was supported by Russian Science Foundation grant 14-12-01340.}
+
+\bibliographystyle{plain}
 \bibliography{piskunov}
@@
 {\bf Fig.~3.}~Model Overview.
-Here the red and the blue cones represent the volumes through which low and high energy plasma propagates.
+Here lighter and darker cones represent the regions through which low and high energy plasma propagates.
 In case depicted the observer's off-axis angle is smaller than the opening angle of the low energy jet, so, due to the relativistic beaming effect, the most of the observable low energy photons will travel along the straight line from the central engine.
@@
 
-{\bf Fig.~5.}~Stretching factors histogram and CDF produced by our model. Bar lines (black) show observed stretching factors. The sample contains $4096$ bursts.
+{\bf Fig.~5.}~Stretching factors histogram and CDF produced by our model.
+The sample contains $4096$ bursts.
+Numbered error bars correspond respectively to GRB 080916C, GRB 090510, GRB 090902B and GRB 090926A.
 
-{\bf Fig.~6.}~Correlation between the high energy photon fraction, the observer's off-axis angle, and the stretching factor found in the sample produces by our model.
-Black crosses show observations.
-This correlation allows one to estimate observer's off-axis angles of real bursts.
+{\bf Fig.~6.}~Correlations between off-axis angles, stretching factors and high to low energy photon count ratios found in the sample produced by our model.
+The sample contains $4096$ bursts.
+Numbered black crosses show data respectively for GRB 080916C, GRB 090510, GRB 090902B and GRB 090926A with $2\sigma$ error bars.
+This correlation allows one to estimate off-axis angles of observed bursts.
 

### 2015-11-29 QuarksProceedings/piskunov.tex
Proceedings editing

@@
 \bibliographystyle{plain}
-\bibliography{piskunov}
+
+\begin{thebibliography}{10}
+
+\bibitem{Abdo:2009pg}
+A.A. Abdo et~al.
+\newblock {Fermi Observations of GRB 090902B: A Distinct Spectral Component in
+  the Prompt and Delayed Emission}.
+\newblock {\em Astrophys.J.}, 706:L138--L144, 2009.
+
+\bibitem{Ackermann:2013zfa}
+M.~Ackermann, M.~Ajello, K.~Asano, M.~Axelsson, L.~Baldini, et~al.
+\newblock {The First Fermi-LAT Gamma-Ray Burst Catalog}.
+\newblock {\em Astrophys.J.Suppl.}, 209:11, 2013.
+
+\bibitem{Ackermann:2010us}
+M.~Ackermann et~al.
+\newblock {Fermi Observations of GRB 090510: A Short Hard Gamma-Ray Burst with
+  an Additional, Hard Power-Law Component from 10 keV to GeV Energies}.
+\newblock {\em Astrophys.J.}, 716:1178--1190, 2010.
+
+\bibitem{Bregeon:2011bu}
+M.~Ackermann et~al.
+\newblock {Detection of a spectral break in the extra hard component of GRB
+  090926A}.
+\newblock {\em Astrophys.J.}, 729:114, 2011.
+
+\bibitem{Ackermann:2012kna}
+M.~Ackermann et~al.
+\newblock {The Fermi Large Area Telescope On Orbit: Event Classification,
+  Instrument Response Functions, and Calibration}.
+\newblock {\em Astrophys.J.Suppl.}, 203:4, 2012.
+
+\bibitem{Bloom:2003wy}
+Joshua~S. Bloom, D.A. Frail, and S.R. Kulkarni.
+\newblock {GRB energetics and the GRB Hubble diagram: Promises and
+  limitations}.
+\newblock {\em Astrophys.J.}, 594:674--683, 2003.
+
+\bibitem{Castignani:2014gaa}
+Gianluca Castignani, Dafne Guetta, Elena Pian, Lorenzo Amati, Simonetta
+  Puccetti, et~al.
+\newblock {Time delays between Fermi LAT and GBM light curves of GRBs}.
+\newblock 2014.
+
+\bibitem{Gehrels:2004gu}
+N.~Gehrels et~al.
+\newblock {The Swift Gamma-Ray Burst Mission}.
+\newblock {\em AIP Conf.Proc.}, 727:637--641, 2004.
+
+\bibitem{Gehrels:2013xd}
+Neil Gehrels and Soebur Razzaque.
+\newblock {Gamma Ray Bursts in the Swift-Fermi Era}.
+\newblock {\em Front.Phys.China.}, 8:661--678, 2013.
+
+\bibitem{Lange:2013uh}
+Johannes Lange and Martin Pohl.
+\newblock {The average GeV-band Emission from Gamma-Ray Bursts}.
+\newblock 2013.
+
+\bibitem{Nakamura:2001kd}
+Takashi Nakamura and Kunihito Ioka.
+\newblock {Peak luminosity-spectral lag relation caused by the viewing angle of
+  the collimated gamma-ray bursts}.
+\newblock 2001.
+
+\bibitem{Rubtsov:2011qq}
+G.I. Rubtsov, M.S. Pshirkov, and P.G. Tinyakov.
+\newblock {GRB observations by Fermi LAT revisited: new candidates found}.
+\newblock {\em Mon.Not.Roy.Astron.Soc.Lett.}, 421:L14--L18, 2012.
+
+\bibitem{Shen:2005ea}
+Rong-Feng Shen, Li-Ming Song, and Zhuo Li.
+\newblock {Spectral lags and the energy dependence of pulse width in gamma-ray
+  bursts: Contributions from the relativistic curvature effect}.
+\newblock {\em Mon.Not.Roy.Astron.Soc.}, 362:59--65, 2005.
+
+\bibitem{Shenoy:2013cba}
+A.~Shenoy, E.~Sonbas, C.~Dermer, L.C. Maximon, K.S. Dhuga, et~al.
+\newblock {Probing Curvature Effects in the Fermi GRB 110920}.
+\newblock {\em Astrophys.J.}, 778:3, 2013.
+
+\bibitem{Tajima:2009az}
+Hiroyasu Tajima.
+\newblock {Fermi Observations of high-energy gamma-ray emissions from GRB
+  080916C}.
+\newblock 2009.
+
+\bibitem{Vianello:2013ela}
+Giacomo Vianello.
+\newblock {Observations of Gamma-ray Bursts in the Fermi era}.
+\newblock 2013.
+
+\bibitem{Yi:2005ht}
+Ting-Feng Yi, En-Wei Liang, Yi-Ping Qin, and Rui-Jing Lu.
+\newblock {On the spectral lags of the short gamma-ray bursts}.
+\newblock {\em Mon.Not.Roy.Astron.Soc.}, 367:1751--1756, 2006.
+
+\end{thebibliography}
 

### 2015-11-29 aastex52/aasguide.tex
Proceedings editing

%%
%% This is file `aasguide.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% aasclass.dtx  (with options: `guide')
%% 
%% This is a generated file;
%% altering it directly is inadvisable;
%% instead, modify the original source file.
%% 
%% Copyright notice.
%% 
%%    These files are distributed
%%    WITHOUT ANY WARRANTY; without even the implied warranty of
%%    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
%% 
%% \CharacterTable
%%  {Upper-case    \A\B\C\D\E\F\G\H\I\J\K\L\M\N\O\P\Q\R\S\T\U\V\W\X\Y\Z
%%   Lower-case    \a\b\c\d\e\f\g\h\i\j\k\l\m\n\o\p\q\r\s\t\u\v\w\x\y\z
%%   Digits        \0\1\2\3\4\5\6\7\8\9
%%   Exclamation   \!     Double quote  \"     Hash (number) \#
%%   Dollar        \$     Percent       \%     Ampersand     \&
%%   Acute accent  \'     Left paren    \(     Right paren   \)
%%   Asterisk      \*     Plus          \+     Comma         \,
%%   Minus         \-     Point         \.     Solidus       \/
%%   Colon         \:     Semicolon     \;     Less than     \<
%%   Equals        \=     Greater than  \>     Question mark \?
%%   Commercial at \@     Left bracket  \[     Backslash     \\
%%   Right bracket \]     Circumflex    \^     Underscore    \_
%%   Grave accent  \`     Left brace    \{     Vertical bar  \|
%%   Right brace   \}     Tilde         \~}%
%%%  @LaTeX-file{
%%%     filename        = "aastex.dtx",
%%%     version         = "5.2",
%%%     date            = "2005/06/22",
%%%     time            = "16:23:00 GMT",
%%%     checksum        = "5964",
%%%     author          = "Arthur Ogawa (mailto:ogawa@teleport.com)",
%%%     revised by      = "SR Nova Private Ltd."
%%%     copyright       = "Copyright (C) 2003 American Astronomical Society
%%%
%%%                        This work may be distributed and/or modified under the
%%%                        conditions of the LaTeX Project Public License, either version 1.3
%%%                        of this license or (at your option) any later version.
%%%                        The latest version of this license is in
%%%                        http://www.latex-project.org/lppl.txt
%%%                        and version 1.3 or later is part of all distributions of LaTeX
%%%                        version 2003/12/01 or later.
%%%
%%%                        This work has the LPPL maintenance status "maintained".
%%%
%%%                        The Current Maintainer of this work is the American Astronomical Society.
%%%
%%%                        This work consists of all files listed in the document README.
%%%
%%%     address         = "American Astronomical Society
%%%                        USA",
%%%     telephone       = "+1 ???",
%%%     FAX             = "",
%%%     email           = "aastex-help@aas.org",
%%%     codetable       = "ISO/ASCII",
%%%     keywords        = "latex, AAS, journal",
%%%     supported       = "yes",
%%%     abstract        = "formatter for AAS journal submissions",
%%%     docstring       = "The checksum field above generated by ltxdoc",
%%%  }
\NeedsTeXFormat{LaTeX2e}[1995/12/01]%
\ProvidesFile{aasguide.tex}%
 [2005/06/22 5.2/AAS markup document class]%
\documentclass[preprint2]{aastex}%
\makeindex
\hyphenation{com-pu-scripts}
\newcommand\aastex{AASTeX}%
\newcommand\aastexhome{http://aastex.aas.org/}%
\renewcommand\LaTeX{LaTeX}%
\renewcommand\TeX{TeX}%
\makeatletter
\renewcommand\ion[2]{#1$\;${\small\rmfamily{#2}}\relax}%
\renewenvironment{theindex}
               {\if@twocolumn
                  \@restonecolfalse
                \else
                  \@restonecoltrue
                \fi
                \columnseprule \z@
                \columnsep 35\p@
                \twocolumn[\section{\indexname}]%
                \thispagestyle{plain}\parindent\z@
                \parskip\z@ \@plus .3\p@\relax
                \let\item\@idxitem}
               {\if@restonecol\onecolumn\else\clearpage\fi}

% This code stolen from ltxguide.cls:

\let\o@verbatim\verbatim
\def\verbatim{%
  \ifhmode\unskip\par\fi
  \ifx\@currsize\normalsize
     \small
  \fi
  \o@verbatim
}

\renewcommand \verbatim@font {%
  \normalfont \ttfamily
  \catcode`\<=\active
  \catcode`\>=\active
}

\RequirePackage{shortvrb}
\MakeShortVerb{\|}

\begingroup
  \catcode`\<=\active
  \catcode`\>=\active
  \gdef<{\@ifnextchar<\@lt\@meta}
  \gdef>{\@ifnextchar>\@gt\@gtr@err}
  \gdef\@meta#1>{\m{#1}}
  \gdef\@lt<{\char`\<}
  \gdef\@gt>{\char`\>}
\endgroup
\def\@gtr@err{%
   \ClassError{ltxguide}{%
      Isolated \protect>%
   }{%
      In this document class, \protect<...\protect>
      is used to indicate a parameter.\MessageBreak
      I've just found a \protect> on its own.
      Perhaps you meant to type \protect>\protect>?
   }%
}
\def\verbatim@nolig@list{\do\`\do\,\do\'\do\-}

\newcommand{\m}[1]{\mbox{\it #1}}
\renewcommand{\arg}[1]{{\tt\string{}\m{#1}{\tt\string}}}
\newcommand{\oarg}[1]{{\tt[%]
}\m{#1}{\tt%[
]}}
\def\cmd#1{\cs{\expandafter\cmd@to@cs\string#1}}
\def\cmd@to@cs#1#2{\char\number`#2\relax}
\DeclareRobustCommand\cs[1]{\texttt{\char`\\#1}}
\newcommand*{\file}[1]{\texttt{#1}}%

 % end of code stolen from ltxguide.cls. Thanks, Alan.

\makeatother
\title{%
 The \aastex\ Package for Manuscript Preparation
}%

\begin{document}

\maketitle
\tableofcontents
\section{Introduction}

The American Astronomical Society (AAS) has developed a markup package to assist authors in preparing manuscripts intended for submission to AAS-affiliated journals, as well as to other journals that will accept \aastex\ manuscripts.

The most important aspect of the \aastex
package is that it defines the set of commands, or markup,
that can be used to identify the structural elements of papers.
When articles are marked up using this set of standard commands,
they may then be submitted electronically to the editorial
offices and fed into the electronic production of the journals.

This guide contains basic instructions for creating
manuscripts using the \aastex\ markup package.
Authors are expected to be familiar with the editorial
requirements of the journals so that they can make
appropriate submissions; they should also have a basic
knowledge of \LaTeX---for instance, knowing
how to set up equations using \LaTeX\ commands.
A number of useful publications about \LaTeX\ (and \TeX) are listed in
the reference section of this guide.

{\em Authors who wish to submit papers electronically to the ApJ,  AJ, or PASP are strongly encouraged use the \aastex\ markup package as described in this guide.}

\section{\aastex\ Article Markup}

This section describes the commands in the \aastex
package that an author might enter in a manuscript that is being
prepared for electronic submission to one of the journals.
The commands will be described in roughly the same order as they
would appear in a manuscript.
The reader will also find it helpful to examine the
sample files that are distributed with the package and to consult the
information available on the AASTeX Web site located at \aastexhome.
Authors are also reminded to review the instructions to authors and electronic
submissions guidelines for the specific journals to which they submit
their papers.

\subsection{Preamble}\label{sec-preamble}

In \LaTeX\ manuscripts, the preamble is that portion of the file before
the \verb"\begin{document}" command.

\subsubsection{Getting Started}

The first piece of markup in the manuscript declares the
overall style of the document.  Any commands that appear before this
markup will be ignored.
\begin{quote}\index{documentclass@\verb`\documentclass`}
\begin{verbatim}
\documentclass{aastex}
\end{verbatim}
\end{quote}
This specifies the document class as \texttt{aastex} with the
default style (\verb"manuscript").
The paper copy produced by this style file will be double spaced.
Any tables included in the main body of the manuscript will also be
double spaced.

Other substyles are available. They are discussed in \S~\ref{styles}.

\subsubsection{Defining New Commands}

\aastex\ allows authors to define their own commands with
\LaTeX's \verb"\newcommand".
(Authors should not use the plain \TeX\ \verb"\def" command in
AAS journal submissions.)
Authors' \verb"\newcommand" definitions must be placed in the
document preamble.

In general, author-defined commands that are
abbreviations or shorthands are acceptable and can be easily handled
by journal offices and publishers during data conversion;
for example:
\begin{quote}
\begin{verbatim}
\newcommand{\grb}{gamma ray burst}
\end{verbatim}
\end{quote}

However, abbreviations that attempt to define new symbols by using
\LaTeX\ commands for repositioning text tend to
cause problems in the publication process and should be avoided.
In particular, author-defined commands that use any of the commands
listed below are apt to cause problems during data conversion.
\begin{quote}
\verb"\hskip", \verb"\vskip",
\verb"\raise", \verb"\raisebox",
\verb"\lower", \verb"\rlap",
\verb"\kern", \verb"\lineskip",
\verb"\char", \verb"\mathchar",
\verb"\mathcode", \verb"\buildref",
\verb"\mathrel", \verb"\baselineskip"
\end{quote}
Consequently, authors are strongly discouraged from
using them.

Extra symbols are defined for \aastex,
some specifically for an astronomical context, others more
broadly used in math and physics.
In particular, the AMS has additional symbol fonts that are
available in a standard \LaTeX\ package (\verb"amssymb").
All of these symbols are depicted in the additional symbosl tables
supplied with the package and on the \aastex\ Web site.

Before defining a new symbol command,
authors are advised to consult these tables to see whether the
symbol they need already exists.  If it does, they should use the
corresponding markup command.
Authors should \emph{not} redefine existing command names.
When one of these commands is encountered in an electronic
manuscript submitted to a journal, an author's redefinition will
be ignored and the originally-defined command used.

\subsubsection{Editorial Information}

A number of markup commands are available for
editorial office use in recording the publication history
for each manuscript.
These commands should be used by
the editorial and production offices only.
\begin{quote}
\begin{verbatim}
\received{<receipt date>}
\revised{<revision date>}
\accepted{<acceptance date>}

\ccc{<code>}
\cpright{<type>}{<year>}
\end{verbatim}
\end{quote}

Copyright information should be specified with the commands \verb"\cpright"
and \verb"\ccc".  The type of copyright and the corresponding year
are given in \verb"\cpright". Valid copyright types are as follows.
\begin{quote}
\begin{tabular}{l@{\quad}l}
\tt AAS & Copyright assigned to the AAS\\
\tt ASP & Copyright assigned to the ASP\\
\tt Crown & Crown copyright applies\\
%\tt PD & The article is in the public domain\\
\tt none & No copyright can be claimed
\end{tabular}
\end{quote}
The copyright type is case sensitive, so the type string must be
entered exactly as given above.

The Copyright Clearing Center code may be given in the \verb"\ccc"
command. The code is taken as regular text, so any special characters,
notably ``\$,'' must be escaped as appropriate.

\subsubsection{Short Comment on Title Page}

Authors who wish to include a short remark on the title page,
such as the name and date of the journal to which an article
has been submitted, may do so with the following command.
\begin{quote}
\begin{verbatim}
\slugcomment{<text>}
\end{verbatim}
\end{quote}
In the
\texttt{manuscript} style,
these comments appear on the title page after the title and authors;
in the
\texttt{preprint} style,
they are placed at the upper right corner
of the title page.

\subsubsection{Running Heads}\label{shorthead}

Authors are invited to supply running head information using the
following commands.
\begin{quote}
\begin{verbatim}
\shorttitle{<text>}
\shortauthors{<text>}
\end{verbatim}
\end{quote}
Two different kinds of data are generally supplied in running heads.
The left head contains an author list, (last
names, possibly truncated as ``et al.''), while the right head
is an abbreviated form of the paper title.  This running head
information will not appear on the \LaTeX-printed page but will be
passed through to  copy editing staff
for inclusion in the published version.

Editors and publishers impose varying requirements
on the brevity of these data.  A good rule of thumb is to limit the list
of authors to three or else use ``et al.,'' and to limit the
short form of the title to 40--45 characters.
The editors may choose to modify the author-supplied
running heads.

\subsection{Starting the Main Body}

The preamble is a control section.
None of the markup that appears in the preamble
actually typesets anything. The author must include a
\begin{quote}
\begin{verbatim}
\begin{document}
\end{verbatim}
\end{quote}
command to identify the beginning of the main, typeset
portion of the manuscript.

\subsection{Title and Author Information}

Authors should use the
\verb"\title"
 and \verb"\author"\  commands to specify title
and author information and the \verb"\affil" command
to indicate the author's primary affiliation.
Each \verb"\author"
 command
should be followed by a corresponding \verb"\affil"
and optional \verb"\email"\  command.
\begin{quote}
\begin{verbatim}
\title{<text>}
\author{<name(s)>}
\affil{<affiliation>}
\affil{<address>}
\email{<e-mail address>}
\and
\end{verbatim}
\end{quote}
 Line breaks may be inserted in the title
with the \verb"\\" command.  (Long titles will
be broken automatically, so the \verb"\\" markup is not required.)
If the title is explicitly broken over several lines, the
preferred style for titles in AAS and ASP journals is the so-called
``inverted pyramid'' style.  In this style, the longest line
is the first (or top) line, and each succeeding line is shorter.
The text of the title should be entered in mixed case;
it will be printed in upper case or mixed case according to
the style of the publication.
Footnotes are permissible in titles. Be careful to ensure that
alternate affiliations (see below) are properly numbered if a
footnote to the title is specified.

Authors' names should be entered in mixed case.
Names that appear together in the author list for authors who
have the same primary affiliation should be specified in a single
\verb"\author"\  command.
Each author group should be followed by
an \verb"\affil"\  command giving the principle
affiliation of those authors.  Physical and postal address information
for the specified institution  may be included with \verb"\affil".
The address can be broken over several lines using the
\verb"\\" command to indicate
the line breaks.
Usually, however, postal information will fit on one line.
When there is more than one \verb"\author"\  command, the final
one should be preceded by the \verb"\and"\ command.

Authors often have affiliations in addition to their principle employer.
These alternate affiliations may be specified with the \verb"\altaffilmark"
 and \verb"\altaffiltext"
 commands.
These behave like the \verb"\footnotemark"
and \verb"footnotetext"  commands of \LaTeX\ except that they do not take
optional arguments. \verb"\altaffilmark"
 is appended to authors' names in the \verb"\author"
 list and generates superscript identification numbers.
The text for the individual alternate affiliations is generated by the
\verb"\altaffiltext"
 command.
\begin{quote}
\begin{verbatim}
\altaffilmark{<key number(s)>}
\altaffiltext{<numerical key>}{<text>}
\end{verbatim}
\end{quote}
It is up to the author to make sure that
 each \m{key number}  in his or her \verb"\altaffilmark"
 matches the \m{numerical key} for
the corresponding \verb"\altaffiltext".

When there is a lengthy author list, all author names may be
specified in a single \verb"\author"
 command with affiliations
specified using the \verb"\altaffilmark"\ me\-chan\-ism.
In these cases, no \verb"\affil"
 commands are used, and in print, the
affiliations would all be listed in a footnote block at the bottom
of the title page.

\subsection{Abstract}

A paper's abstract should be marked with the
\texttt{abstract} environment.
\begin{quote}
\begin{verbatim}
\begin{abstract}
<abstract text>
\end{abstract}
\end{verbatim}
\end{quote}

If an author is using the \texttt{preprint2} style, and the abstract is too long to fit on the title page, the \texttt{longabstract} option may be used in the \texttt{documentclass} to break the abstract to a new page.

\noindent
\begin{verbatim}
\documentclass[preprint2,longabstract]{aastex}
\end{verbatim}

\noindent Note that authors should use the \texttt{longabstract} option only in \texttt{preprint2} mode and only when the abstract is too long to fit on the title page.

\subsection{Keywords}

Keywords, or subject headings, are accommodated
as a single piece of text.
\begin{quote}
\begin{verbatim}
\keywords{<text>}
\end{verbatim}
\end{quote}
If authors supply keywords, they must be delimited by whatever
punctuation is required by the journal.
They should be specified in alphabetical order.
The \verb"\keywords"
 command will print the proper leading text---``Keywords:,'' ``Subject headings:,'' etc.---according to journal style.

\subsection{Comments to Editors}

Authors may make notes or comments to the copy editor with the
\verb`\notetoeditor` command.
\begin{quote}
\begin{verbatim}
\notetoeditor{<text>}
\end{verbatim}
\end{quote}
This command behaves like a
footnote. Output to the printed page is produced only in the
\texttt{manuscript} style.

\subsection{Sections}

\aastex\ supports four levels of section headings.
\begin{quote}
\begin{verbatim}
\section{<heading>}
\subsection{<heading>}
\subsubsection{<heading>}
\paragraph{<heading>}
\end{verbatim}
\end{quote}
Section headings should be given in upper case or mixed case, depending
on the style of the journal.
Note that these commands delimit sections by marking the
\emph{beginning} of each section;
there are no separate commands to mark the ends.

\subsection{Figure and Table Placement}   \label{place}

When preparing a manuscript for submission to an AAS journal, figures and tables do not generally need to be ``placed'' in the text of the
document where an author would like them to appear but
rather may simply follow the main body of the text.
However, authors may indicate to the editors the preferred placement of
these items by use of the \verb"\place*" commands.
\begin{quote}
\begin{verbatim}
\placetable{<key>}
\placefigure{<key>}
\end{verbatim}
\end{quote}
The \verb"\place*{"\m{key}\verb"}" commands are similar to the
\verb"\ref"
 command in \LaTeX
and require corresponding \verb"\label"
 commands to link them to the
proper elements.

When used in the
\texttt{manuscript} style, the \verb"\place*"\  commands will print a short
message to the editor about figure or table placement.
In the other styles, nothing is printed.

\subsection{Acknowledgments}

\aastex\ supports an
\verb"\acknowledgments"%
\index{acknowledgments@\verb`\acknowledgments`} section.
\begin{quote}
\begin{verbatim}
\acknowledgments
 <acknowledgments text>
\end{verbatim}
\end{quote}
In the \aastex\ styles, acknowledgments are set off from the
conclusion of the body with vertical space. Note the acknowledgments
command takes no arguments.

\subsection{Facilities}

To help organizations obtain information on the effectiveness of their
telescopes, the AAS has created a group of keywords for telescope
facilities.  Using a common set of keywords will make searches for
this information significantly easier and more accurate.
In addition, the use of facility
keywords will be useful for linking papers that utilize the
same telescopes together within the framework of the Virtual Observatory.
A facilities keyword list is available through a
link at the AASTeX Web page, \aastexhome.

The facilities list should appear after the
acknowledgments section.
\begin{quote}
\begin{verbatim}
Facilities: \facility{<facility ID>},
   \facility{<facility ID>},
   \facility{<facility ID>}, ...
\end{verbatim}
\end{quote}
As part of the {\it facility ID} argument,
the author may also include
the name of the instrument in parentheses, e.g.
\verb"facility{HST(WFPC2)}" or \verb"facility{MMT(Blue channel spectrograph)}".
There is no limit to the number of facility keywords
that may be included in a paper.

\subsection{Appendices}

When one or more appendices are needed in a paper, the point where the
main body text ends and the appendix begins should be marked
with the \verb"\appendix"
 command.
\begin{quote}
\begin{verbatim}
\section{<body section>}
\appendix
\section{<appendix section>}
\end{verbatim}
\end{quote}
The \verb"\appendix"
 command takes care of a number of internal
housekeeping concerns, such as identifying sections with letters
instead of numerals, and resetting the equation counter.
Note that the \verb"\appendix"
 command takes no arguments.
Sections in the appendix should be headed with \verb"\section"
 commands.

\subsection{Equations}

Display equations can be typeset in \LaTeX\ in a number of ways.
The following three are probably of greatest use in AASTeX.
\begin{quote}
\begin{verbatim}
\begin{displaymath}
\end{displaymath}

\begin{equation}
\end{equation}

\begin{eqnarray}
\end{eqnarray}
\end{verbatim}
\end{quote}
The
\texttt{displaymath}  environment
will break out a single,
unnumbered formula.  The \texttt{equation} environment does the same
thing except that the equation is
autonumbered by \LaTeX.
To set several formul\ae\ in which vertical alignment
is required, or to display a long equation across multiple lines, use the
\texttt{eqnarray} environment. Each line of the \texttt{eqnarray}
will be numbered
unless a \verb"\nonumber"
 command is inserted
before the equation line delimiter
(\verb"\\").
\LaTeX's equation counter is \emph{not} incremented when
\verb"\nonumber"
 is used.

Authors may occasionally wish to group related equations together and
identify them with letters appended to the equation number.
When this is desired, such related equations should still be set
in \texttt{equation} or \texttt{eqnarray} environments, whichever is
appropriate, and then grouped  within
the \texttt{mathletters} environment.
\begin{quote}
\begin{verbatim}
\begin{mathletters}
<equation> or <eqnarray>
\end{mathletters}
\end{verbatim}
\end{quote}

It is possible to override \LaTeX's automatic numbering within the
\texttt{equation} or \texttt{eqnarray} environments using
\begin{quote}
\begin{verbatim}
\eqnum{<text>}
\end{verbatim}
\end{quote}
When \verb"\eqnum"
 is specified inside an \texttt{equation} environment
or on a particular equation within an \texttt{eqnarray}, the text
supplied as an argument to \verb"\eqnum"
 is used as the equation
identifier.
\LaTeX's equation counter is \emph{not} incremented when \verb"\eqnum"
 is used.
\verb"\eqnum"
 must be used \emph{inside} the math environment.

If, as a consequence of the use of \verb"\eqnum"
 or \verb"\nonumber",
\LaTeX's equation counter gets out of synch with the author's
intended sequence,
the counter may be reset to a particular value.
\begin{quote}
\begin{verbatim}
\setcounter{equation}{<number>}
\end{verbatim}
\end{quote}
The equation counter should be set to the number corresponding to the
last equation that was formatted; therefore, it is most appropriate for this
command to appear immediately after an \texttt{equation} or
\texttt{eqnarray} environment.
The command must be used
\emph{outside} the math environments.

The \texttt{eqsecnum} style file can also be used to modify the way equations
are numbered. See \S~\ref{styles} for details.

\subsection{Citations and Bibliography}   \label{bibliography}

Two options are available for marking citations
and formatting reference lists: the standard \LaTeX\ \verb"thebibliography"
environment, and the \aastex\ \verb"references"\  environment.
Authors are strongly encouraged to use \verb"thebibliography"
in their electronic submissions.

Please note that the bibliographic data supplied by the author in
the reference list must conform to the standards of the journal.
Many of the journals that
accept \aastex\ agreed to reduce typographic overhead, bold, italic, etc.,
in reference lists \citep{Abt90}, and the AAS
has elected not to burden authors with tedious markup commands
to delimit the bibliographic fields.  Instead, citations will be
typeset in roman with no size or style changes.
It is the responsibility of the author to
arrange the required bibliographic fields in the
proper order with the correct punctuation, according to journal
style.

\subsubsection{The {\tt thebibliography} Environment}   \label{bib}

The preferred method for reference management is to use \LaTeX's
\texttt{thebibliography} environment, marking citations in the body
of the paper with
\verb"\citep"\ or \verb"\citet"
and associating references with them using \verb"\bibitem".
The \verb"\cite"-\verb"\bibitem"
mechanism associates citations and references symbolically
while maintaining proper citation syntax within the paper.
In the \verb"\bibitem"
command, the author should specify citation data inside
square brackets and a citation key in curly braces
for each reference. (The \verb"bibitem" command is described
in detail in the next section.)

\noindent
\begin{verbatim}
   \begin{thebibliography}{<dummy>}
   \bibitem[<cite data>]{<key>} <bibliographic data>
      .
      .
   \end{thebibliography}
\end{verbatim}

\noindent Note that the argument \m{dummy} to the start command of the
environment is not
used in the \aastex\ package, but it is included to be consistent with the
syntax of standard \LaTeX. It is acceptable to simply insert an
empty pair of curly braces at the end of the \verb"\begin{thebibliography}"
command.

\subsubsection{Specifying Bibliographic and Citation Information}\label{spec-bib-data}

\aastex\ uses the \verb"natbib" package \citep{Daly98} for
citation management.
The \verb"natbib" package re-implements \LaTeX's \verb"cite" command,
offering greater flexibility for managing citations in the
author-year form.

When using \verb"natbib", bibliographic data are defined
in \verb"bibitem" commands.

\noindent
\begin{verbatim}
   \bibitem[<author>(<year>)]{<key>}
         <bibliographic data>
\end{verbatim}

\noindent
The square-bracketed argument of the \verb"bibitem"
contains the \m{author} portion of the citation
followed by the \m{year} set off in parentheses.
The parentheses are important---\verb"natbib" uses them to
determine the year portion of the citation---so be sure to
include them.
The argument \m{key} in curly braces is the code name
by which the citation is referenced in the text.

When placing citations in the text, the author should use
either a \verb"citep" or a \verb"citet" command.
\begin{quote}
\begin{verbatim}
\citep{<key(s)>}
\citet{<key(s)>}
\end{verbatim}
\end{quote}
The \verb"citep" command produces a citation that is entirely
set off by parentheses, e.g.\ ``(Cox 1995),'' while \verb"citet" permits
the author's name to form part of the text, e.g.\ ``Cox (1995).''
The plain \LaTeX\ \verb"cite" command behaves like \verb"citet".

The citation \m{key} must correspond to the \m{key} in a \verb"bibitem" command.
During processing, information from the square-bracketed argument of the
key's \verb"bibitem" is inserted in
the text at the location of the \verb"\cite"\ command.
Multiple citation keys are separated by commas, e.g.,
\verb"citep"\allowbreak
\verb"knuth84,"\allowbreak
\verb"cox95,"\allowbreak
\verb"lamport94}".

\verb"citep" and \verb"citet" each take optional arguments that specify
extra text to be appended to the citation label. Text in the first
set of square brackets will appear before the cite while
text in the second set will appear after it.
For instance, \verb"citep[chap. 2]{jon90}"
would produce the citation ``(Jones et al., 1990, chap. 2),''
\verb"citep[see][]{jon90}" would produce ``(see Jones et al, 1990),'' and
\verb"citep[see][chap. 2]"\allowbreak\verb"jon90}" would print
``(see Jones et al., 1990, chap. 2).''
In addition, the * form of the
cite commands will print the full author list instead of the
abbreviated form.

The syntax discussed above should be sufficient for the
vast majority of cases; however, \aastex
does use the full \verb"natbib" implementation, so many more
syntax options are available. For details on
the full range of \verb"natbib" citation options, see the section
on using \verb"natbib" on the \aastex\ Web
site and the \verb"natbib" package documentation.

It is not possible to use \verb"\bibitem"
within \aastex's \texttt{references} environment
(\S~\ref{refenv}),
nor will \verb"\cite"\ commands work properly in the main body
if \verb"\bibitem" commands are absent.

\subsubsection{The {\tt references} Environment}\label{refenv}

Some authors might prefer to enter citations directly
into the body of an article. If so, the \verb"references" environment may be
used to format the reference list. The
\texttt{references}  environment
simply sets off
the list of references and adjusts spacing parameters.
\begin{quote}
\begin{verbatim}
\begin{references}
\reference{<key>} <bibliographic data>
   .
   .
\end{references}
\end{verbatim}
\end{quote}

While the
\texttt{references}  environment
remains supported in \aastex,
we anticipate that authors will prefer the stronger capabilities of
the standard \LaTeX\ \verb"thebibliography" commands as extended by \verb"natbib".

\subsubsection{Abbreviations for Journal Names}

\aastex\ commands for journal abbreviations have been deprecated.
Authors are encouraged to use standard journal abbreviations defined
by the ADS group; see http://ads.hardvard.edu/journal-abbrevs.

%\aastex\ commands are available for many of the most--frequently-referenced
%journals so that authors may use the markup rather than having
%to look up a particular journal's abbreviation.
%In principle, all the journals should be using the
%same abbreviations, but it is fair to anticipate some changes in
%the specific abbreviations before a system is finally settled on.
%If authors use the journal macros, they need not worry about
%changes to journal style governing abbreviations in citations.
%A listing of the current abbreviation macros appears
%in Table~\ref{journame}.
%\begin{table}
%\footnotesize
%\begin{center}
%\caption{Abbreviations for Journal Names}\label{journame}
%\begin{tabular}{l@{\quad}p{2.5in}}
%\verb"\araa" & Annual Review of Astronomy and Astrophysics\\
%\verb"\actaa" & Acta Astronomica\\
%\verb"\ao" & Applied Optics\\
%\verb"\aj" & Astronomical Journal\\
%\verb"\azh" & Astronomicheskii Zhurnal\\
%\verb"\aap" & Astronomy and Astrophysics\\
%\verb"\aapr" & Astronomy and Astrophysics Reviews\\
%\verb"\apj" & Astrophysical Journal\\
%\verb"\apjl" & \rule[.5ex]{2em}{.4pt}, Letters to the Editor\\
%\verb"\apjs" & \rule[.5ex]{2em}{.4pt}, Supplement Series\\
%\verb"\aplett" & Astrophysics Letters and Communications\\
%\verb"\apspr" & Astrophysics Space Physics Research\\
%\verb"\apss" & Astrophysics and Space Science\\
%\verb"\aaps" & \rule[.5ex]{2em}{.4pt}, Supplement Series\\
%\verb"\baas" & Bulletin of the AAS\\
%\verb"\bain" & Bulletin Astronomical Inst. Netherlands\\
%\verb"\caa" & Chinese Astronomy and Astrophysics\\
%\verb"\cjaa" & Chinese Journal of Astronomy and Astrophysics\\
%\verb"\fcp" & Fundamental Cosmic Physics\\
%\verb"\gca" & Geochimica Cosmochimica Acta\\
%\verb"\grl" & Geophysics Research Letters\\
%\verb"\iaucirc" & IAU Circular\\
%\verb"\icarus" & Icarus\\
%\verb"\jcp" & Journal of Chemical Physics\\
%\verb"\jgr" & Journal of Geophysical Research\\
%\verb"\jcap" & Journal of Cosmology and Astroparticle Physics\\
%\verb"\jqsrt" & Journal of Quantitative Specstroscopy and Radiative Transfer\\
%\verb"\jrasc" & Journal of the RAS of Canada\\
%\verb"\memras" & Memoirs of the RAS\\
%\verb"\memsai" & Mem. Societa Astronomica Italiana\\
%\verb"\mnras" & Monthly Notices of the RAS\\
%\verb"\na" & New Astronomy\\
%\verb"\nar" & New Astronomy Review\\
%\verb"\nat" & Nature\\
%\verb"\nphysa" & Nuclear Physics A\\
%\verb"\physscr" & Physica Scripta\\
%\verb"\pra" & Physical Review A\\
%\verb"\prb" & Physical Review B\\
%\verb"\prc" & Physical Review C\\
%\verb"\prd" & Physical Review D\\
%\verb"\pre" & Physical Review E\\
%\verb"\prl" & Physical Review Letters\\
%\verb"\physrep" & Physics Reports\\
%\verb"\planss" & Planetary Space Science\\
%\verb"\procspie" & Proceedings of the SPIE\\
%\verb"\pasa" & Publications of the ASA\\
%\verb"\pasj" & Publications of the ASJ\\
%\verb"\pasp" & Publications of the ASP\\
%\verb"\qjras" & Quarterly Journal of the RAS\\
%\verb"\rmxaa" & Revista Mexicana de Astronomia y Astrofisica \\
%\verb"\skytel" & Sky and Telescope\\
%\verb"\solphys" & Solar Physics\\
%\verb"\ssr" & Space Science Reviews\\
%\verb"\sovast" & Soviet Astronomy\\
%\verb"\zap" & Zeitschrift fuer Astrophysik\\
%\end{tabular}
%\end{center}
%\end{table}

\subsection{Figures}

\subsubsection{Electronic Art}

If an author wishes
to embed graphics in a manuscript, it is necessary that
the graphics files conform
to the Encapsulated PostScript (EPS) standard \citep{PLRM}. The
author must also have an appropriate DVI translator,
one that targets PostScript output devices. (Detailed information
on preparing and submitting electronic art is
available in the submissions instructions for the journals.)

Several commands are available for including EPS files
in \aastex\ manuscripts. They should be placed within the
 \texttt{figure} environment.
\begin{quote}
\begin{verbatim}
\begin{figure}
\figurenum{<text>}
\epsscale{<num>}
\plotone{<epsfile>}
\plottwo{<epsfile>}{<epsfile>}
\caption{<text>}
\end{figure}
\end{verbatim}
\end{quote}
When \verb"\figurenum" is specified inside the {\tt figure} environment,
the text supplied as an argument to \verb"\figurenum" will be used as the
figure identifier.
\LaTeX's figure counter is \emph{not} incremented when \verb"\figurenum"
is used.
\verb"\figurenum" must be used \emph{inside} the {\tt figure} environment.

\verb"\plotone" inserts the graphic in the named EPS file,
scaled so that the horizontal
dimension fits the width of the body text;
the vertical dimension is scaled to maintain the aspect ratio.
\verb"\plottwo" inserts two plots next to each other.
Scale factors are determined automatically from information in the
EPS file.

The automatic scaling may be overridden with the command
\verb"\epsscale{<num>}", where \m{num} is a scaling factor
 in decimal units, e.g., 0.80.

The \verb"plotone" and \verb"plottwo"
macros are invocations of the \verb"graphicx"
\verb"includegraphics" command. In most instances, using
\verb"plotone" or \verb"plottwo" should work for placing
figures in \aastex\ documents. However, if more
flexibility is needed, the \verb"includegraphics" command
may be invoked directly. For instance, to rotate
an image by 90 degrees, use
\begin{quote}
\begin{verbatim}
\includegraphics[angle=90]{<epsfile>}
\end{verbatim}
\end{quote}
See the \verb"graphicx" package documentation or the \aastex\ Web site for a
complete list of the available arguments to \verb"includegraphics".

Authors are strongly encouraged to use \verb"plotone", \verb"plottwo", or \verb"includegraphics" to place their EPS figures. However, as of the current release, the old \verb"plotfiddle" macro from v4.0 has been reintroduced and may be used if the desired figure placement cannot be achieved with one of the methods above. The syntax of the command is
\begin{quote}
\begin{verbatim}
\plotfiddle{<epsfile>}{<vsize>}{<rot>}{<hsf>}
       {<vsf>}{<htrans>}{<vtrans>}
\end{verbatim}
\end{quote}
where the arguments are
\begin{center}
\begin{tabular}{l@{\quad}p{2in}}
\tt epsfile & name of the EPS file \\
\tt vsize & vertical white space to allow for plot (\LaTeX\ dimension)\\
\tt rot & rotation angle (degrees)\\
\tt hsf & horizontal width of scaled figure (PS points)\\
\tt vsf & vertical height of scaled figure (PS points)\\
\tt htrans & horizontal translation (PS points)\\
\tt vtrans & vertical translation (PS points)\\
\end{tabular}
\end{center}
PostScript points are 1/72 inches, so an \m{htrans} of 72 moves
the graphic 1 inch to the right. Note that the \m{vtrans} argument is
discarded in the reimplented macro and is included only for
backward compatibility with \aastex\ v4.0.

\subsubsection{Figure Captions} \label{legends}

Regardless of whether an author includes electronic art in a manuscript, figure captions, or legends, should be provided. If art is included in the document, use the \verb"\caption"\ command within the \texttt{figure} environment.  If electronic art is not included, figure captions may be grouped together at the end of the document using \verb"\figcaption".
\begin{quote}
\begin{verbatim}
\figcaption[<filename>]{<text>\label{<key>}}
\end{verbatim}
\end{quote}
The optional argument, \m{filename}, can be used to iden\-ti\-fy
the file for the corresponding figure;
\m{text} refers to the caption for that figure.  The author
may provide a \verb"\label"\ with a unique \m{key} for cross-referencing
purposes.

When the \verb"\figcaption"\  command is used, the figure
identification label,
e.g., ``Figure 1,'' is generated automatically by the command itself, so
there is no need to key this information.
 There is an upper limit of seven figure captions per page.
Footnotes are \emph{not} supported for figures.

\subsection{Tables}  \label{tables}

There is support in the \aastex\ package for tables via two mechanisms:
\LaTeX's standard \texttt{table} environment,
and the
\texttt{deluxetable} environment, which allows for the formatting
of lengthy tabular material.  Tables may be
marked up using either mechanism, although use of
\texttt{deluxetable} is preferred.
Authors should \emph{not} use the \LaTeX
\texttt{tabbing} environment when preparing electronic
submissions.

\subsubsection{The \texttt{deluxetable} Environment}  \label{dte}

Authors are encouraged to use the
\texttt{deluxetable} environment to format their tables since it automatically
handles many formatting tasks, including table numbering and insertion
of horizontal rules. It also provides mechanisms for breaking tables and
controlling width and vertical spacing that are unavailable in the \LaTeX
\texttt{tabular} environment.

The \texttt{deluxetable} environment
is delimited by \LaTeX's familiar
\verb"\begin"\  and \verb"\end"\  constructs.
The content consists of preamble commands and table data,
the latter delimited by \verb"\startdata"
 and \verb"\enddata".
\begin{quote}
\begin{verbatim}
\begin{deluxetable}{<cols>}
<preamble commands>
\startdata
<table data>
\enddata
\end{deluxetable}
\end{verbatim}
\end{quote}
The argument \m{cols} specifies the justification for each col\-umn.
An alignment token, ``l,'' ``c,'' or ``r,'' is given for each column,
indicating flush left, centered, or flush right.

\subsubsection{Preamble to the \texttt{deluxetable}}%

There are several commands in the
\texttt{deluxetable} environment
that
must be given in the preamble.
\begin{quote}
\begin{verbatim}
\tabletypesize{<font size command>}
\rotate
\tablewidth{<dimen>}
\tablenum{<text>}
\tablecolumns{<num>}
\tablecaption{<text>\label{<key>}}
\tablehead{<text>}
\end{verbatim}
\end{quote}

If a table is too wide for the printed page, the font size of the table
can be cahnged with the \verb"\tabletypesize" command, which takes as
an argument one of the the font size change commands:
\verb"\small"
 (11pt), \verb"\footnotesize"
 (10pt), or
\verb"\scriptsize"
 (8pt).

To force a table to be set in landscape orientation,
use the \verb"\rotate"\ \label{cmd-rotate} command. Note that
most DVI previewers will not properly render rotated \verb"deluxetable"
output, so in order to see what the table looks like,
it must be output to PostScript and viewed there.

The width of a deluxetable is defined by \verb"\tablewidth".
If this command is omitted, the default width is the width of the page.
The table can be set to its natural width by specifying
a dimension of \verb"0pt".
Long tables may have a natural width that is
different for each page.  The natural width for each page will be
printed to the log file during processing.
Authors may then use this log information to
define a fixed table width in order to give the
table a more uniform appearance across pages.

It is possible to override \LaTeX's automatic numbering within the
\texttt{deluxetable} environment.
When \verb"\tablenum"
 is specified inside a \texttt{deluxetable}
preamble,
the text supplied as an argument to \verb"\tablenum"
 will be used as the
table identifier.
\LaTeX's equation counter is \emph{not} incremented when
\verb"\tablenum"
 is used.

The caption (actually, the title) of the table is specified
in \verb"\tablecaption".
The text of \verb"\tablecaption"
 should be brief;
explanatory notes should be specified in the end notes to the table
(see \S \ref{tabendnotes}
 below).  If the caption
does not appear
centered above the table after processing, then specify the width of
the table explicitly with the \verb"\tablewidth"
 command and rerun
\LaTeX\ on the file.   If an author supplies a \verb"\label"\ for cross-referencing purposes, this, too, should be included in the \verb"\tablecaption".

Column headings are specified with \verb"\tablehead".
Within the \verb"\tablehead", each column heading should be given
in a \verb"\colhead", which will ensure that the heading
is centered
on the natural width of the column.
There should be a heading for each column so that there are as
many \verb"\colhead"
 commands in the \verb"\tablehead"
 as there
are data columns.

\begin{quote}
\begin{verbatim}
\tablehead{
\colhead{<heading>} & \colhead{<heading>}}
\end{verbatim}
\end{quote}

If more complicated column headings are required,
any valid \texttt{tabular} command that constitutes a proper
head line in a \LaTeX\ table may be used. For example, the
\verb"multicolumn" command below would create a table head with
text centered over five columns.
\begin{quote}
\begin{verbatim}
\multicolumn{5}{c}{<text>}
\end{verbatim}
\end{quote}
Consult \citet{Lamport} or \citet{Kopka99} for further details
on the available table commands.

The \verb"\tablecolumns{"\m{num}\verb"}" command
is necessary if an author has
multi-line column headings produced by \verb"\tablehead"
 or other \LaTeX
commands and is using either the \verb"\cutinhead"
 or \verb"\sidehead"
markup (see below).  The \m{num} argument should be
set to the true number of columns in the
table.  The command must come before the \verb"\startdata"
 command.

\subsubsection{Content of the \texttt{deluxetable}}%

After the table title and column headings have been specified,
data rows can be entered.
Data rows are delimited with the \verb"\startdata"
 and \verb"\enddata"\  commands.
The end of each row is indicated with the standard \LaTeX\ \verb"\\" command.
Data cells within a row are separated with \& (ampersand) characters.
\begin{quote}
\begin{verbatim}
\startdata
<data line>\\
<cell>&<cell>&<cell>\\
<more data lines>\\
\enddata
\end{verbatim}
\end{quote}

Column alignment within the data columns can be adjusted with the \TeX
\verb"\phantom{"\m{string}\verb"}" command,
where \m{string} can be any character, e.g., \verb"\phantom{$\arcmin$}".
A blank character of width \m{string} is then inserted in the table.
Four commands have been predefined for this purpose.
\begin{quote}
\begin{tabular}{l@{\quad}p{2in}}
{\verb"\phn"%
} & {phantom numeral 0-9}\\
{\verb"\phd"%
} & {phantom decimal point}\\
{\verb"\phs"%
} & {phantom $\pm$ sign}\\
{\verb"\phm{"\m{string}\verb"}"} & {generic phantom}\\
\end{tabular}
\end{quote}

Extra vertical space can be inserted between rows with an optional argument to the
\verb"\\" command.
\begin{quote}
\begin{verbatim}
\\[<dimen>]
\end{verbatim}
\end{quote}
The argument is a dimension
and may be specified in any units that are legitimate in \LaTeX.

In a table, it may happen that several rows of data are
associated with a single object or item.
Such logical groupings should not be broken across pages.
In these cases, the \verb"tablebreak" command may be used to force
a page break at the desired point.
\begin{quote}
\begin{verbatim}
<table row>\\
\tablebreak
<next table row>\\
\end{verbatim}
\end{quote}
This command can be used any time that the default \verb"deluxetable" page
breaks need to be overridden.

Journals often require that table cells that contain no data
be explicitly marked.  This is to differentiate such cells from
blank cells, which are frequently interpreted as implicitly
repeating the entry in the corresponding cell in the row preceding.
Table cells for which there are no data should contain
a \verb"\nodata"
 command.
\begin{quote}
\begin{verbatim}
\nodata
\end{verbatim}
\end{quote}

Within the \verb"deluxetable" body, two kinds of special heads are
allowed, \verb"cutinhead" and \verb"sidehead".
A cut-in head is a piece of text centered across the width
of the table. It is spaced above and below
from the data rows that precede and
follow it and will appear set off by rules in the \LaTeX\ output.
Similarly, the command for a side head produces a row spanning
the width of the table but with the text left justified.
\begin{quote}
\begin{verbatim}
\cutinhead{<text>}
\sidehead{<text>}
\end{verbatim}
\end{quote}

Table footnotes (more properly, table \emph{end notes})
may also be used in the \texttt{deluxetable} environment.
Their use is described in detail in \S \ref{tabendnotes}.

\subsubsection{The \texttt{table} Environment}

Authors may also compose tables using the \texttt{table} environment.
\begin{quote}
\begin{verbatim}
\begin{table}
\end{table}
\end{verbatim}
\end{quote}
The \texttt{table}\ environment
encloses not only the tabular
material but also any title or footnote information
associated with the table.

Titles, or captions, for tables are indicated with a \verb"caption"
command
\begin{quote}
\begin{verbatim}
\caption{<text>\label{<key>}}
\end{verbatim}
\end{quote}
A table label, e.g. ``Table 2,'' is generated automatically
by \verb"\caption".
 The author
may provide a \verb"\label"\ in the caption with a unique \m{key}
for cross-referencing
purposes.

The table body should appear within the \texttt{tabular} environment.
\begin{quote}
\begin{verbatim}
\begin{tabular}{<cols>}
\end{tabular}
\end{verbatim}
\end{quote}
The alignment tokens in \m{cols} specify the justification for each column. The letters ``l,'' ``c,'' or ``r'' is given for each column,
indicating left, center, or right justification.
Consult \citet{Lamport} for details about using
the
\texttt{tabular} environment
to prepare tables.

Each \texttt{tabular} table must appear within a \texttt{table}
environment. There should be only one \texttt{tabular} table per
\texttt{table}
environment.
If the journal requests manuscripts with only one table per page,
the author may need to insert a \verb"\clearpage"
 command after especially short tables.

Use the \verb"tableline" command to insert horizontal rules in the
\verb"tabular" environment.
\begin{quote}
\verb"\tableline"
 \end{quote}
The use of vertical rules should be avoided.

As with the \texttt{deluxetable} environment, it
 is possible to override \LaTeX's automatic numbering within the
\texttt{table} environment using \verb"\tablenum".
\verb"\tablenum"
 must be used \emph{inside} the \texttt{table}
environment.

\subsubsection{Table End Notes} \label{tabendnotes}

\aastex\ supports footnotes and end notes within tables;
this support applies to both the
\texttt{deluxetable} environment
and the standard \LaTeX\ \texttt{table} environment.

Footnotes for tables are usually identified by lowercase letters
rather than numbers. Use the \verb"tablenotemark" and \verb"tablenotetext"
commands to supply table footnotes.
As with \verb"\altaffilmark" and \verb"\altaffiltext", a note label,
usually a letter, is required.
\begin{quote}
\begin{verbatim}
\tablenotemark{<key letter(s)>}
\tablenotetext{<alpha key>}{<text>}
\end{verbatim}
\end{quote}
The \m{key letter} of the \verb"tablenotemark" should be the same as the
\m{alpha key} for the corresponding \m{text}.
It is the responsibility of the author to make the correspondence
correct.

Sometimes authors tabulate materials that have corresponding
references and may want to associate these references with the table.
Authors may also wish to append a short paragraph of explanatory
notes that pertain to the entire table. These elements should be
specified with the commands below.
\begin{quote}
\begin{verbatim}
\tablerefs{<reference list>}
\tablecomments{<text>}
\end{verbatim}
\end{quote}
The \verb"\tablenotetext", \verb"\tablecomments", and
\verb"\tablerefs" commands \emph{must}
 be specified after the
\verb"\end{tabular}" or \verb"\enddata" and before
the closing \verb"\end{table}" or \verb"\end{deluxetable}".

\subsection{Supplemental Materials}

For many years now, authors have been taking advantage of the AAS journals'
ability to post supplemental materials with their papers in the electronic
editions.  Even though each paper must stand on its own
scientifically without the supplements, these materials are reviewed in the
peer review process and should be included in initial manuscript
submissions. As with regular figures and tables, papers with online only
data must reference each electronic object in the main text
and include an explanation of what the reader will
find in the electronic edition.

The four most popular types of supplemental materials are
machine-readable tables, online color figures, online-only figures, and
animations.  Please see each journal's Web site for details on
what types of supplemental materials are acceptable, how to
submit these materials, author tools for preparing them,
and their associated costs.

\subsubsection{Machine-readable Tables}

Online-only tables submitted to AAS journals are converted
to a machine-readable
format for presentation in the electronic edition.
Machine-readable tables have two
parts: the formatted ASCII data and a metadata header that provides
format, units, and short explanations of each column of data.  This
structure is designed to provide maximum flexibility and ease of use
for readers who wish to further manipulate the data with their own
computer programs or with software like Excel.

For each machine-readable table, the author should include a short sample
version of the table in his or her \LaTeX\ submission. This sample version
will appear in the print edition as well as
in HTML in the electronic edition.  The sample table should be
5 to 15 lines long.  It should include a table note at the end
 with text indicating that a machine-readable
version will be available in the electronic edition. For instance,
\begin{quote}
\begin{verbatim}
\tablecomments{Table 1 is
published in its entirety in
the electronic edition of
the Astrophysical Journal.
A portion is shown here
for guidance regarding its
form and content.}
\end{verbatim}
\end{quote}
Each example table must
be cited and numbered as if it were a fully printed table.

\subsubsection{Online Color Figures}

For some journals, authors may submit figures
that will appear in black and white in the
print edition of a journal but in color in the electronic edition.
In these cases, authors must submit separate black and white and RGB color
versions of each figure labelled according to the
file-naming conventions required for the publication.
For the benefit of print journal readers, figure captions should be written
with the color information placed inside parentheses,
for instance, ``The dotted
line (colored blue in electronic edition)
is the H$_o$ = 75 km s-1 Mpc-1 model,'' and should
include a note directing the reader to
``See the electronic edition of the Journal
for a color version of this figure.''

\subsubsection{Online-only Figures}

This feature is mainly useful for articles that contain large compendia
of identification charts and other supplemental graphic material that need not
be printed in full in the paper journal. Online-only figures are intended to
provide supplemental information that is not critical to the scientific
content of the article but that might nonetheless be of interest to the reader.

Online-only figures must be mentioned explicitly by number and appear
in correct numerical order in the body of the text.  At least one figure
in a series must be displayed as an example figure for the print edition.
The caption should carry a message indicating that
more figures are available in the electronic edition---for instance,
``Plots for all sources are available in the electronic edition of the journal.''
Enough information must be included in the figure caption for readers of the
print edition to determine what is contained in the online-only figures.

\subsubsection{Animations}

Currently, only animations in the MPEG format are accepted by the AAS journals.
Authors must supply a still frame from the animation in EPS format marked
up like a regular figure that will serve as an example for the reader.
They should include text in the caption for the
still frame indicating that an animation is
available electronically. For instance,
``This figure is also available as an mpeg
animation in the electronic edition of the {\it Astrophysical Journal}.''
As with online-only figures, authors must include enough information in the
figure caption for readers of the print edition to determine what the
animation illustrates.

\subsection{Miscellaneous}

\subsubsection{Celestial Objects and Data Sets}

Authors who wish to have the most important objects in their paper
linked to a data center in the electronic edition may do so using the
\begin{quote}
\begin{verbatim}
\objectname[<catalog ID>]{<text>}
\end{verbatim}
\end{quote}
macro, or its alias \verb"object".
The text contained in the required argument
will be printed in the paper and will serve as a link anchor
in the electronic edition. The catalog ID
given as an optional argument will be carried through as the identification
key in the link to a data center. Note that links will only
be activated if the name provided in the argument is recognized by a
participating data center.  It is the author's responsibility to use the
correct identifier.

Similar markup is available for linking to data sets hosted
at participating data centers.
\begin{quote}
\begin{verbatim}
\dataset[<catalog ID>]{<text>}
\end{verbatim}
\end{quote}
In the paper, the text in the required argument will be
printed while the the catalog ID value will be passed through to form
links to data centers.

When an article contains \verb"object" or \verb"dataset" commands, the
publisher will be able to use the markup to
pass along a list of objects and data sets used in the
paper to database personnel. Software can then be used to construct links to
those databases.
Please check with each journal's Web site for instructions on how to
determine the object and data set identifiers, the location of
verifications tools, and information on where in the paper these
macros can best be used.

\subsubsection{Ionic Species and Chemical Bonds}

When discussing atomic species, ionization levels can be indicated
with the following command.
\begin{quote}
\begin{verbatim}
\ion{<element>}{<level>}
\end{verbatim}
\end{quote}
The ionization state is specified as the second argument
and should be given as a numeral.
For example, ``\ion{Ca}{III}'' would be marked up as \verb"\ion{Ca}{3}".

For single, double, and triple chemical bonds, use the following macros.
\begin{quote}
\begin{verbatim}
\sbond
\dbond
\tbond
\end{verbatim}
\end{quote}

\subsubsection{Fractions}

\aastex\ contains commands that permit authors to specify alternate
forms for fractions.
\LaTeX\ will set fractions in displayed math as built-up fractions;
however, it is sometimes desirable to use case fractions in
displayed equations.
In such instances, one should use \verb"\case"
rather than \verb"\frac". Note, however, that authors submitting
manuscripts electronically to AAS journals
should generally find it unnecessary to use any markup other than the
standard \LaTeX\ \verb"\frac".

\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{@{}llc@{}}
Built-up & \verb"\frac{1}{2}" & $\displaystyle\frac{1}{2}$ \\
Case     & \verb"\case{1}{2}" & $\case{1}{2}$ \\
Shilled  & \verb"1/2" & $1/2$ \\
\end{tabular}
\end{center}

\subsubsection{Astronomical Symbols}

As mentioned earlier, the \aastex\ package
contains a collection of assorted macros
for symbols and abbreviations specific to an astronomical context.
These are commonly useful and also somewhat difficult for authors
to produce themselves because fussy kerning is required.
See the symbols pages provided with the package distribution.
Most of these commands can be used in both running text and math. However,
\verb"\lesssim"\  and \verb"\gtrsim"\  can only be used in math mode.

\subsubsection{Hypertext Constructs}

The \verb"\anchor"\ \label{cmd-anchor} command is a general-purpose
hypertext link tag, associating \verb"<text>" in the manuscript with
the specified resource (\verb"<href>").
\begin{quote}
\begin{verbatim}
\anchor{<href>}{<text>}
\end{verbatim}
\end{quote}
\verb"<href>"\ should be specified as a \emph{full} URI, including the
\verb"scheme:"\  designator (\verb"http:", \verb"ftp:", etc.).

The \verb"\url" command supports the special case where an author
wishes to express a URL in the text.
\begin{quote}
\begin{verbatim}
\url{<text>}
\end{verbatim}
\end{quote}

The \verb"\email"\  command is used to identify e-mail addresses
anywhere in the manuscript.
The text of the argument is the e-mail address.
Please do \emph{not} prepend the \verb"mailto:" part.
\begin{quote}
\begin{verbatim}
\email{<address>}
\end{verbatim}
\end{quote}
This command should be used to indicate authors' e-mail addresses
in author lists at the beginning of manuscripts.

\subsection{Concluding the File}

The last command in the electronic manuscript file should be the
\begin{quote}
\begin{verbatim}
\end{document}
\end{verbatim}
\end{quote}
command, which must appear after all the material in the paper.
This command directs the formatter to finish processing the manuscript.

\section{Style Options}   \label{styles}

\subsection{Manuscript Style}

The default style option is the \verb"manuscript" style. This style will
produce double-spaced pages printed in a single column at the width
of the page.

\subsection{Preprints}

Two preprint styles are available.
The \texttt{preprint} style is similar to the \verb"manuscript" style,
but it is single-spaced and fully justified.
The \texttt{preprint2} style produces a two-column preprint.

\subsubsection{Single-column Preprint}

A single-column preprint format may be specified with the
\texttt{preprint} style option.
\begin{quote}
\begin{verbatim}
\documentclass[preprint]{aastex}
\end{verbatim}
\end{quote}
The size of the typeface used is under author control by way of
\LaTeX's \m{nn}\texttt{pt} class options
(where \m{nn} is 10, 11, or 12).
Use of 10 point type is not recommended with the
\texttt{preprint} style.

\subsubsection{Two-column Preprint}

The
\texttt{preprint2} style
has the principle function of providing two-column formatting.
\begin{quote}
\begin{verbatim}
\documentclass[preprint2]{aastex}
\end{verbatim}
\end{quote}
It is important to remember
that text lines are considerably shorter when two columns are typeset
side by side on a page.  Long equations, wide tables and figures, and
the like, may not typeset in this format without some adjustments.
The expenditure of great effort to adapt copy and markup for
two-column pages is counterproductive.
Remember that the main
goal of this package is to produce draft-, or referee-, format pages.
It is the responsibility of the editors and publishers to
produce publication-format papers for the journals.

The \texttt{preprint2} style
sets the article's front matter---the title, author, abstract,
and keyword material---on a separate page at full text width.
The body of the article is set in a two-column page grid,
the appendices in a one-column page grid,
and the bibliography in a two-column page grid.
(This manual was prepared using the \texttt{preprint2} style.)

The author may supply \LaTeX's \verb"\twocolumn"\  or \verb"onecolumn"
commands whenever desired.
Be aware, however, that using explicit column-switching commands can
cause formatting problems.

\subsection{The {\tt eqsecnum} Style}

The \texttt{eqsecnum} style file can be used to modify the way equations
are numbered.
\begin{quote}
\begin{verbatim}
\documentclass[eqsecnum]{aastex}
\end{verbatim}
\end{quote}
Normally, equations are numbered sequentially through the
entire paper, starting over at ``(A1)'' if there is an appendix.
If \texttt{eqsecnum} appears in the \texttt{documentclass} command,
equation num\-bers will be sequential through each section and will be
formatted ``({\it sec-eqn}),'' where ``{\it sec}'' is the current section number and
``{\it eqn}'' is the number of the equation within that section.

\subsection{The {\tt flushrt} Style }

A \verb"flushrt"\ style option is available for authors that prefer to
have their margins left and right justified.
\begin{quote}
\begin{verbatim}
\documentclass[eqsecnum,flushrt]{aastex}
\end{verbatim}
\end{quote}
Note the \texttt{preprint} and \texttt{preprint2} styles
are already flush right by default.

\section{Additional Documentation}  \label{docs}

The preceding explanation of the markup commands in the
\aastex\ package has merit for defining syntax, but many
authors will prefer to examine the sample papers that are
included with the style files.
The files of interest are described below.

A comprehensive example employing nearly all of the capabilities
of the package (in terms of markup as well as formatting)
is in \verb"sample.tex".
This file is annotated with comments that describe
the purpose of most of the markup.
\verb"sample.tex"\ includes three tables: two marked up using the
\texttt{deluxetable} environment
and another table using the \LaTeX\ \texttt{table}
environment.

In \verb"table.tex", a complex but short example of the
\texttt{deluxetable} environment
demonstrates some of the techniques
that can be used to generate complex column headings and to align
variable-width columns.
Here the \LaTeX\ \verb"\multicolumn" command is used to span a heading
over several columns.  When \verb"\multicolumn" is used along with the
\verb"\cutinhead" command, the \verb"\tablecolumns" command must be used to specify
the number of columns in the table---otherwise the \verb"\cutinhead" command
will not work properly.
This table also makes use of the \verb"\phn" command to better align some of the
columns.

This user guide (\verb"aasguide.tex")
is also marked up with the \aastex\ package,
although it is not exemplary as a scientific paper.

Many of the markup commands described in the preceding
sections are standard \LaTeX\ commands. The reader who is
unfamiliar with their syntax is referred to the \LaTeX
references works cited in the bibliography, in particular
\citet{Kopka99} and \citet{Lamport}.

Authors who wish to know the ins and outs of \TeX\ itself
should read the \emph{\TeX book} \citep{Knuth}.
This resource contains a good deal of information about
typography in general.  Many details of mathematical typography are
discussed in \emph{Mathematics into Type} by \citet{Swanson}.

\section{Acknowledgments}

\aastex\ was designed and written by Chris Biemesderfer in 1988.
Substantial revisions were made by Lee Brotzman and Pierre
Landau when the package was updated to v4.0.
\aastex\ was rewritten as a \LaTeX 2$\epsilon$ class by Arthur Ogawa
for the  v5.0 release. It was updated to v5.2 by
SR Nova Private Ltd.
The documentation has benefited from revisions by
Jeannette Barnes, Sara Zimmerman, and Greg Schwarz.

\begin{thebibliography}{}

\bibitem[Abt(1990)]{Abt90}
  Abt, H.  1990, ApJ, 357, 1 (editorial)

\bibitem[Adobe(1999)]{PLRM}
  Adobe Systems, Inc.  1999,
  \anchor{http://www.adobe.com/prodindex/postscript/}
  {PostScript Language Reference Manual} (Reading, MA: Addison-Wesley)

\bibitem[Daly(1998)]{Daly98}
  Daly, P. 1998, \emph{Natural Sciences Citations and References}
  (\texttt{natbib} package documentation)

\bibitem[Goossens, Mittelbach, and Samarin(1994)]{Goossens94}
  Goossens, M., Mittelbach, F., and Samarin, A. 1994,
  \emph{The \LaTeX\ Companion} (Reading, MA: Addison-Wesley)

\bibitem[Hahn(1993)]{Hahn93}
  Hahn, J. 1993,  \emph{\LaTeX\ for Everyone}
  (Englewood Cliffs, NJ: Prentice-Hall)

\bibitem[Knuth(1984)]{Knuth}
  Knuth, D. 1984, \emph{The \TeX book} (Reading, MA: Addison-Wesley).
  \newblock Revised to cover \TeX3, 1991.

\bibitem[Kopka and Daly(1999)]{Kopka99}
  Kopka, H. and Daly, P. 1999,  \emph{A Guide to \LaTeX},
  3rd edition, (Reading, MA: Addison-Wesley)

\bibitem[Lamport(1994)]{Lamport}
  Lamport, L. 1994, \emph{\LaTeX: A Document Preparation System\/},
  2nd edition, (Reading, MA: Addison-Wesley)

\bibitem[Swanson(1979)]{Swanson}
  Swanson, E. 1979, \emph{Mathematics into Type}
  (Providence, RI: American Mathematical Society)

\end{thebibliography}
\end{document}
\endinput
%%
%% End of file `aasguide.tex'.

### 2015-11-29 Paper/gammaRays.tex
Proceedings editing
Add more explanation to the section 5.1
Transition between general description of a fit, to the fit of GRB
090926A is explained.

@@
         \centering
-        \includegraphics[width=0.7\textwidth]{modelOverview}
+        \includegraphics[width=0.7\textwidth]{modelOverview.pdf}
         \caption{ Model Overview.  The red and blue cones represent
@@
 	\item{If we make a simultaneous transformation of $r_0
-          \rightarrow \lambda r_0$ and $\eta_0 \rightarrow
-          \frac{1}{\lambda^3}\eta_0$, where $\lambda$ is a parameter
+          \rightarrow \lambda_1 r_0$ and $\eta_0 \rightarrow
+          \frac{1}{\lambda_1^3}\eta_0$, where $\lambda_1$ is a parameter
           of transformation, only duration of the burst will
@@
 	\item{Finally, if we make a following transformation:
-          $\omega_0 \rightarrow \lambda \omega_0$, $\theta_0
-          \rightarrow \lambda^k \theta_0$, $\eta_0 \rightarrow
-          \lambda^\alpha \eta_0$, where $\lambda$ is another
+          $\omega_0 \rightarrow \lambda_2 \omega_0$, $\theta_0
+          \rightarrow \lambda_2^k \theta_0$, $\eta_0 \rightarrow
+          \lambda_2^\alpha \eta_0$, where $\lambda_2$ is another
           transformation parameter, no observables will change
@@
 
-The best fit parameters are found by minimizing the cost function
-$C\left(n, \theta_0, k, \alpha, \chi\right)$ which should
-satisfy the following objectives:
+We are going to fit parameters specifically for GRB 090926A for its stretching factor is the most significant.
+While we assume burst related parameters $\left(n, \theta_0, k, \alpha\right)$ to be the same for all bursts, $\chi$ here considered to be the off-axis angle of GRB 090926A.
+At the same time, we need to make sure that burst related universal parameters $\left(n, \theta_0, k, \alpha\right)$ are consistent with observation of other bursts such as GRB 090902B.
+
+In particular, the best fit parameters are found by minimizing the cost function
+$C\left(n, \theta_0, k, \alpha, \chi\right)$ which
+satisfies the following objectives:
 \begin{itemize}

### 2015-11-30 Paper/gammaRays.tex
Add references from the 2nd physics reviewer

@@
 of magnitude~\cite{Bloom:2003wy}. This indicates that the most of the
-bursts have similar energetics in their rest frames. Moreover, the
+bursts have similar energetics in their rest frames. {\bf Also, it is generally accepted that} the
 high energy radiation of GRBs is shown to follow the predictions of
-the synchrotron radiation scenario~\cite{Wang:2013ptaw}.  Several
+the synchrotron radiation scenario~\cite{Wang:2013ptaw}, {\bf although there are problems with that model~\cite{Daigne:2010fb} and there are bursts in which additional radiation component clearly appears~\cite{Abdo:2009pg} }.
+Several
 observations are related to the temporal variations of spectra. This
-way, the spectral lags were found between different low energy bands
-\cite{Yi:2005ht} and the very-high-energy radiation was discovered to
+way, the spectral lags were found between different low energy bands {\bf 
+\cite{cheng1995time,Wu:1999ms} }and the very-high-energy radiation was discovered to
 be extended relative to the x-ray emission
@@
 called curvature effects. These effects were explored by multiple
-authors \cite{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.  These
+authors{\bf \cite{Meszaros:1997je,Soderberg:2001mk,Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba} }.  These
 studies concerned x-ray radiation, and considered the homogeneous
@@
 $\left(1+z\right)^3$. It is clear, however, that since no stars
-existed at very small redshifts, the burst density should decline
+existed at very {\bf large} redshifts, the burst density should decline
 there, so we add an exponential cutoff to it:
@@
           the observed photon count.}
-	\item{If we make a simultaneous transformation of $r_0
+	\item{{\bf If we make a simultaneous transformation of $r_0
           \rightarrow \lambda_1 r_0$ and $\eta_0 \rightarrow
-          \frac{1}{\lambda_1^3}\eta_0$, where $\lambda_1$ is a parameter
+          \frac{1}{\lambda_1^3}\eta_0$, where $\lambda_1$} is a parameter
           of transformation, only duration of the burst will
@@
           or stretching factor, will not be affected.}
-	\item{Finally, if we make a following transformation:
+	\item{{\bf Finally, if we make a following transformation:
           $\omega_0 \rightarrow \lambda_2 \omega_0$, $\theta_0
           \rightarrow \lambda_2^k \theta_0$, $\eta_0 \rightarrow
-          \lambda_2^\alpha \eta_0$, where $\lambda_2$ is another
+          \lambda_2^\alpha \eta_0$, where $\lambda_2$ } is another
           transformation parameter, no observables will change
@@
 
-We are going to fit parameters specifically for GRB 090926A for its stretching factor is the most significant.
+{\bf We are going to fit parameters specifically for GRB 090926A for its stretching factor is the most significant.
 While we assume burst related parameters $\left(n, \theta_0, k, \alpha\right)$ to be the same for all bursts, $\chi$ here considered to be the off-axis angle of GRB 090926A.
@@
 $C\left(n, \theta_0, k, \alpha, \chi\right)$ which
-satisfies the following objectives:
+satisfies the following objectives:}
 \begin{itemize}

### 2015-12-17 APJ/gammaRaysAPJ.tex
Ported recent changes to APJ

@@
 of magnitude~\citep{Bloom:2003wy}. This indicates that the most of the
-bursts have similar energetics in their rest frames. Moreover, the
-high energy radiation of GRBs is shown to follow the predictions of
-the synchrotron radiation scenario~\citep{Wang:2013ptaw}.  Several
+bursts have similar energetics in their rest frames. {\bf Also, it is generally accepted that} the
+high energy radiation of GRBs follow{\bf s} the predictions of
+the synchrotron radiation scenario~\citep{Wang:2013ptaw}, {\bf although there are problems with that model~\citep{Daigne:2010fb} and there are bursts in which additional radiation component clearly appears~\citep{Abdo:2009pg} }.
+Several
 observations are related to the temporal variations of spectra. This
-way, the spectral lags were found between different low energy bands
-\citep{Yi:2005ht} and the very-high-energy radiation was discovered to
+way, the spectral lags were found between different low energy bands {\bf \citep{cheng1995time,Wu:1999ms} }and the very-high-energy radiation was discovered to
 be extended relative to the x-ray emission
@@
 called curvature effects. These effects were explored by multiple
-authors \citep{Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}.  These
+authors {\bf \citep{Meszaros:1997je,Soderberg:2001mk,Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}}.  These
 studies concerned x-ray radiation, and considered the homogeneous
@@
 background on the spacecraft position.
+{\bf Therefore, the background we estimate is mostly sky background (for the background from misidentification of protons for ``SOURCE'' class events is subdominant), and it remains stable over the period of $24$ hours, making our background estimate valid during the time of the burst.}
 
@@
           the observed photon count.}
-	\item{If we make a simultaneous transformation of $r_0
-          \rightarrow \lambda r_0$ and $\eta_0 \rightarrow
-          \frac{1}{\lambda^3}\eta_0$, where $\lambda$ is a parameter
+	\item{{\bf If we make a simultaneous transformation of $r_0
+          \rightarrow \lambda_1 r_0$ and $\eta_0 \rightarrow
+          \frac{1}{\lambda_1^3}\eta_0$, where $\lambda_1$} is a parameter
           of transformation, only duration of the burst will
@@
           or stretching factor, will not be affected.}
-	\item{Finally, if we make a following transformation:
-          $\omega_0 \rightarrow \lambda \omega_0$, $\theta_0
-          \rightarrow \lambda^k \theta_0$, $\eta_0 \rightarrow
-          \lambda^\alpha \eta_0$, where $\lambda$ is another
+	\item{{\bf Finally, if we make a following transformation:
+          $\omega_0 \rightarrow \lambda_2 \omega_0$, $\theta_0
+          \rightarrow \lambda_2^k \theta_0$, $\eta_0 \rightarrow
+          \lambda_2^\alpha \eta_0$, where $\lambda_2$ } is another
           transformation parameter, no observables will change
@@
 
-The best fit parameters are found by minimizing the cost function
-$C\left(n, \theta_0, k, \alpha, \chi\right)$ which should
-satisfy the following objectives:
+{\bf We are going to fit parameters specifically for GRB 090926A for its stretching factor is the most significant.
+While we assume burst related parameters $\left(n, \theta_0, k, \alpha\right)$ to be the same for all bursts, $\chi$ here considered to be the off-axis angle of GRB 090926A.
+At the same time, we need to make sure that burst related universal parameters $\left(n, \theta_0, k, \alpha\right)$ are consistent with observation of other bursts such as GRB 090902B.
+
+In particular, the best fit parameters are found by minimizing the cost function
+$C\left(n, \theta_0, k, \alpha, \chi\right)$ which
+satisfies the following objectives:}
 \begin{itemize}
@@
 
+\bibitem[Cheng et al.(1995)]{cheng1995time}
+Cheng, L.-X., Ma, Y.-Q., Cheng, K.-S., Lu, T., Zhou, Y.-Y. 1995, \aap, 300, 746
+
+\bibitem[Daigne {et~al.}(2011)]{Daigne:2010fb}
+Daigne, F., Bosnjak, Z., Dubus, G. 2011, \aap, 526, A110
+
 \bibitem[Gehrels \& Razzaque(2013)]{Gehrels:2013xd}
@@
 
+\bibitem[Meszaros et al.(1998)]{Meszaros:1997je}
+Meszaros, P., Rees, M.-J., Wijers, R.-A.-M.-J. 1998, \apj, 499, 301
+
 \bibitem[Nakamura \& Ioka(2001)]{Nakamura:2001kd}
@@
 
+\bibitem[Soderberg(2001)]{Soderberg:2001mk} 
+Soderberg, A.-M. \& Fenimore, E.-E. 2001, astro-ph/0103277
+
 \bibitem[Tajima(2009)]{Tajima:2009az}
@@
 Wang, X.-Y., Liu, R.-Y., \& Lemoine, M. \apj, 2013, 771, L33
+
+\bibitem[Wu \& Fenimore(2000)]{Wu:1999ms}
+Wu, B., Fenimore, E. 2000, \apj, 535, L29
 

### 2015-12-18 APJ/gammaRaysAPJ.tex
Added references from first reviewer and Postnov

@@
 interesting results, see~\citet{Vianello:2013ela,Gehrels:2013xd} for a
-review. In particular, the total energy emitted in gamma-rays during a
-burst was found to be similar for the different GRBs within an order
-of magnitude~\citep{Bloom:2003wy}. This indicates that the most of the
-bursts have similar energetics in their rest frames. {\bf Also, it is generally accepted that} the
+review. In particular, there are indications that the total energy emitted in gamma-rays during a
+burst is similar for the different GRBs within an order
+of magnitude~\citep{postnov1999grb,Bloom:2003wy}.
+One should keep in mind, though, that the width of the distribution differs for multiple authors, and, in fact, is model dependent.
+More recent results, for example, suggest that energetics of different bursts might be different~\citep{racusin2009jet, chandra2012radio}.
+In this paper, however, we explore the possibility that most of the
+bursts have similar energetics in their rest frames to simplify analysis.
+When more precise energy distribution becomes available, the analysis in this paper may be improved.
+{\bf Also, it is generally accepted that} the
 high energy radiation of GRBs follow{\bf s} the predictions of
@@
 called curvature effects. These effects were explored by multiple
-authors {\bf \citep{Meszaros:1997je,Soderberg:2001mk,Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba}}.  These
+authors {\bf \citep{Meszaros:1997je,Soderberg:2001mk,Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba,granot2002off}}.  These
 studies concerned x-ray radiation, and considered the homogeneous
@@
 
+\bibitem[Chandra \& Frail(2012)]{chandra2012radio}
+Chandra, P., Frail, D.-A. 2012, \apj, 746, 156
+
 \bibitem[Ghirlanda et al.(2012)]{Ghirlanda:2011bn}
@@
 
+\bibitem[Granot et al.(2002)]{granot2002off}
+Granot, J., Panaitescu, A., Kumar, P., Woosley, S.-E. 2002, \apjl, 570, L61
+
 \bibitem[Lange \& Pohl(2013)]{Lange:2013uh}
@@
 Nakamura, T., \& Ioka, K. 2001, \apj, 554, L163
+
+\bibitem[Postnov et. al.(1999)]{postnov1999grb}
+Postnov, K.-A., Prokhorov, M.-E., Lipunov, V.-M. 1999, astro-ph/9908136
+
+\bibitem[Racusin et. al.(2009)]{racusin2009jet}
+Racusin, J.-L., Liang, E.-W., Burrows, D.-N., Falcone, A., Sakamoto, T., Zhang, B.-B., Zhang, B., Evans, P., Osborne, J 2009, \apj, 698, 43
 

### 2015-12-18 APJ/gammaRaysAPJ.tex
Added comments on curvature effects

@@
 interesting results, see~\citet{Vianello:2013ela,Gehrels:2013xd} for a
-review. In particular, there are indications that the total energy emitted in gamma-rays during a
+review. {\bf In particular, there are indications that the total energy emitted in gamma-rays during a
 burst is similar for the different GRBs within an order
 of magnitude~\citep{postnov1999grb,Bloom:2003wy}.
-One should keep in mind, though, that the width of the distribution differs for multiple authors, and, in fact, is model dependent.
+One should keep in mind, though, that there is no agreement on the width of the distribution, and it is in fact model dependent.
 More recent results, for example, suggest that energetics of different bursts might be different~\citep{racusin2009jet, chandra2012radio}.
@@
 high energy radiation of GRBs follow{\bf s} the predictions of
-the synchrotron radiation scenario~\citep{Wang:2013ptaw}, {\bf although there are problems with that model~\citep{Daigne:2010fb} and there are bursts in which additional radiation component clearly appears~\citep{Abdo:2009pg} }.
+the synchrotron radiation scenario~\citep{Wang:2013ptaw}}, {\bf although there are problems with that model~\citep{Daigne:2010fb} and there are bursts in which additional radiation component clearly appears~\citep{Abdo:2009pg} }.
 Several
@@
 called curvature effects. These effects were explored by multiple
-authors {\bf \citep{Meszaros:1997je,Soderberg:2001mk,Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba,granot2002off}}.  These
-studies concerned x-ray radiation, and considered the homogeneous
-distribution of radiation sources throughout the jet. Our model is
-based on the reverse assumption that the highest energy radiators are
+authors {\bf \citep{Meszaros:1997je,Soderberg:2001mk,Nakamura:2001kd,Shen:2005ea,Shenoy:2013cba,granot2002off}}. {\bf These
+studies either considered the homogeneous
+distribution of radiation sources throughout the jet, or assumed dependence of energy and Lorentz factor on the opening angle. Our model is
+based on a similar assumption that the highest energy radiators are
 concentrated near the axis of the jet and consequently the jet opening
@@
 the small fraction of GRBs visible above GeV and the hard
-spectrum.
+spectrum. We assume the Lorentz factor to be constant throughout the jet for simplicity.}
 

### 2015-12-18 APJ/gammaRaysAPJ.tex
Added reference to Goldstein 2015

@@
 burst is similar for the different GRBs within an order
-of magnitude~\citep{postnov1999grb,Bloom:2003wy}.
+of magnitude~\citep{postnov1999grb,Bloom:2003wy,Goldstein:2015fib}.
 One should keep in mind, though, that there is no agreement on the width of the distribution, and it is in fact model dependent.
@@
 
+\bibitem[Goldstein et al.(2015)]{Goldstein:2015fib}
+Goldstein, A., Connaughton, V., Briggs, M.-S., Burns, E. 2015, arXiv:1512.04464
+
 \bibitem[Granot et al.(2002)]{granot2002off}

### 2015-12-18 APJ/referee_reply.txt
Referee reply draft added. Minor modification to the paper.

Dear Editor,
Dear Referee,

We thank you strongly for the comprehensive review of the manuscript which helped us to improve the paper with the account of your comments. We are sorry for the vast delay with the reply due to personal reasons. Please find below answers for the second Referee's comments and comments of the expert on the statistical aspects. We have also studied the literature suggested by the first Referee and referenced when relevant.

Answers to the second Referee report

Q: I think that the consideration of relevant observations and previous papers is weak and needs to be improved. 
A: We agree that many relevant works were indeed missed and now it's fixed in an updated version (see below)

Q: Section 1: "In particular, the total energy emitted in gamma-rays during a burst was found to be similar for the different GRBs within an order of magnitude (Bloom et al. 2003)." More recent papers, with a larger sample and post Fermi, find a wider range of total energy. For example, Frail et al., ApJ, 2012, 746, 156.
"Moreover, the high energy radiation of GRBs is shown to follow the predictions of the synchrotron radiation scenario." This is statement is seriously incomplete and misleading. Many in the field think it likely that the keV-MeV emission is from synchrotron emission, but there there are serious conflicts between at least simple version of synchrotron predictions and observations. There is a vast literature on this topic and numerous papers attempting to resolve the problems. Furthermore, it is unquestionable that multiple emission components are observed in some Fermi GRBs. For example, in addition to the likely-synchrotron component, GRB 090902B has a power-law component (Abdo, A. A. et al. 2009, ApJ, 706, L138).
Yi et al. seems a peculiar reference for spectral lags ... its about short versus long BATSE GRBs. There were numerous papers about BATSE lags during the 1990s. Perhaps find a review article?
"curvature effects. These effects were explored by multiple authors..." There are earlier papers on this topic. e.g., Rees. Fenimore.

A: An introduction is revised according to the comments. References to the recent results are included.

Q: section 2.1 I'm doubtful that the background is constant in time for 24 hours. Perhaps the authors should consult with a member of the LAT team, or justify this assumption.

A: The requirement of the stability of the background is the reason why the "TRANSIENT" event class was not used. For the "SOURCE" class events, the sky background dominates over the misidentification of protons. The corresponding statement is added to section 2.2.

Q: Section 3.
"However, only about 30 of them were detected by the LAT, and only 4 of them were bright in the high energy band. If one extrapolates the uniform jet model to the very high energies, this observation would mean that these groups of bursts are internally different: some of them produce VHE radiation, while the others do not. These differences are hard to explain given that the burst energetics are similar"
There are numerous ways to explain this in addition to jet viewing effects: 1) current evidence is that burst energetics are not similar, 2) some synchrotron spectra might have a cutoff, 3) there might be a range of jet Lorentz factions, leading to a range of cutoffs due to gamma-gamma opacity, 4) the high-energy emission may mostly be due to an additional component (e.g., inverse Compton). The ratio between the nominally synchrotron component and the additional component may vary.
How does this analysis of curvature effects compare to previous published papers?

A: The statement is revised in Section 3. We argue that since the particular reason is non known, one may consider explanation which are based or not based on the internal difference of GRBs.
The difference between our model and previous published papers is now given in the Introduction.

Q: section 4.2
There are many papers on the GRB rate versus z, relating to the star formation rate. Generally there is thought to be a bias towards Pop II and low metallicity stars -- and thus a deficit near z = 0. One of the star formation rate curves could be used.
A: Unfortunately this requires to rerun all the calculation. We will include the star formation data in the next applications of the model.

Q: "since no stars existed at very small redshifts" is "large" meant?
A: corrected

Q: section 5.1
I found this section confusing. It seems to start as a general discussion of how to fit the model to the data, then to implicitly change to fitting the the data of GRB 090926A.
A: Additional explanation is added to Section 5.1. GRB 090926A has the most significant stretching factor and it's explanation is a key to fit all the observations.

Answer to the comments of the expert on the statistical aspects

Q: I see two problems with the use of `KS Probability' in Figures 3-5.

a) The Anderson-Darling statistic is similar to KS but always more sensitive. See https://asaip.psu.edu/Articles/beware-the-kolmogorov-smirnov-test.
b) The KS and AD tests are hypothesis tests, and are not designed for multiple applications for parameter estimation (Stretching Factor). There is no problem with using these statistics as measures of proximity between data and model, but the tabulated significance levels are not applicable (see Web page above). As the original data are time series, simple bootstrapping cannot be performed, so I am not sure how significance levels can be estimated. Some careful thought is needed, or the study should proceed without estimation of the accuracy of the best-fit Stretching Factors.

A: 
We thank the expert for the issues rised. 
a) Specially thanks for interesting reference and suggestion. It's intriguing to test whether AD is stronger in our particular case. Unfortunately, this would require redesign of the code which is optimized for performance with KS. We'll try AD within the next application of our model. 
b) A special bootstrapping procedure should be designed for the precise statistical meaning of the accuracy. Still, the accuracy is given as an estimate, while the main result is the statistical significance of the deviation from stretching factor 1.0. This is estimated by just comparing the data with the expected distribution.

Sincerely yours,
Maxim Piskunov, Grigory Rubtsov

### 2015-12-18 APJ/gammaRaysAPJ.tex
Referee reply draft added. Minor modification to the paper.

@@
 interesting results, see~\citet{Vianello:2013ela,Gehrels:2013xd} for a
-review. {\bf In particular, there are indications that the total energy emitted in gamma-rays during a
-burst is similar for the different GRBs within an order
-of magnitude~\citep{postnov1999grb,Bloom:2003wy,Goldstein:2015fib}.
-One should keep in mind, though, that there is no agreement on the width of the distribution, and it is in fact model dependent.
-More recent results, for example, suggest that energetics of different bursts might be different~\citep{racusin2009jet, chandra2012radio}.
-In this paper, however, we explore the possibility that most of the
-bursts have similar energetics in their rest frames to simplify analysis.
-When more precise energy distribution becomes available, the analysis in this paper may be improved.
-{\bf Also, it is generally accepted that} the
-high energy radiation of GRBs follow{\bf s} the predictions of
-the synchrotron radiation scenario~\citep{Wang:2013ptaw}}, {\bf although there are problems with that model~\citep{Daigne:2010fb} and there are bursts in which additional radiation component clearly appears~\citep{Abdo:2009pg} }.
-Several
-observations are related to the temporal variations of spectra. This
-way, the spectral lags were found between different low energy bands {\bf \citep{cheng1995time,Wu:1999ms} }and the very-high-energy radiation was discovered to
-be extended relative to the x-ray emission
-\citep{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}. In this paper
-we go further and compare the temporal extension of GRB gamma-ray
-radiation in the two energy bands $\unit[100]{MeV} < E <
+review. {\bf In particular, there are indications that the total
+  energy emitted in gamma-rays during a burst is similar for the
+  different GRBs within an order of
+  magnitude~\citep{postnov1999grb,Bloom:2003wy,Goldstein:2015fib}.
+  One should keep in mind, though, that there is no agreement on the
+  width of the distribution, and it is in fact model dependent. More
+  recent results, for example, suggest that energetics of different
+  bursts might be different~\citep{racusin2009jet,
+    chandra2012radio}. In this paper, however, we explore the
+  possibility that most of the bursts have similar energetics in their
+  rest frames to simplify analysis. When more precise energy
+  distribution becomes available, the analysis in this paper may be
+  improved. {\bf Also, it is generally accepted that} the high energy
+  radiation of GRBs follow{\bf s} the predictions of the synchrotron
+  radiation scenario~\citep{Wang:2013ptaw}}, {\bf although there are
+  problems with that model~\citep{Daigne:2010fb} and there are bursts
+  in which additional radiation component clearly
+  appears~\citep{Abdo:2009pg}. } Several observations are related to
+the temporal variations of spectra. This way, the spectral lags were
+found between different low energy bands {\bf
+  \citep{cheng1995time,Wu:1999ms} }and the very-high-energy radiation
+was discovered to be extended relative to the x-ray emission
+\citep{Castignani:2014gaa,Lange:2013uh,Vianello:2013ela}. In this
+paper we go further and compare the temporal extension of GRB
+gamma-ray radiation in the two energy bands $\unit[100]{MeV} < E <
 \unit[1]{GeV}$ and $E > \unit[1]{GeV}$ (hereafter, low and high-energy
@@
 studies either considered the homogeneous
-distribution of radiation sources throughout the jet, or assumed dependence of energy and Lorentz factor on the opening angle. Our model is
+distribution of radiation sources throughout the jet, or assumed
+dependence of energy and Lorentz factor on the opening angle. The model of this paper is
 based on a similar assumption that the highest energy radiators are
@@
 background on the spacecraft position.
-{\bf Therefore, the background we estimate is mostly sky background (for the background from misidentification of protons for ``SOURCE'' class events is subdominant), and it remains stable over the period of $24$ hours, making our background estimate valid during the time of the burst.}
+{\bf Therefore, the background we estimate is mostly sky background
+  for the background from misidentification of protons for ``SOURCE''
+  class events is subdominant. Sky background remains stable over the period of $24$ hours, making our estimate valid during the time of the burst.}
 
@@
 different: some of them produce VHE radiation, while the others do
-not.  These differences are hard to explain given that the burst
-energetics are similar \citep{Bloom:2003wy}.  These differences in
+not. {\bf The particular reason for these differences is not
+  known. Explanations may include additional spectral components and
+  variations of the spectra or jet Lorentz factors within GRB
+  population. In this paper we elaborate on explanation, which is not
+  based on the difference of GRB. 
+}  
+The differences in
 burst counts find natural explanation in our model, in which the
@@
 
-{\it Facilities:} \facility{Fermi}.
+{\it Facilities:} \facility{Fermi LAT}.
 
